---
title: "oct_31_analysis_memo"
format: pdf
editor: visual
---

---
title: "Memo: Contact → Deservingness Diagnostics"
format: pdf
---

## Purpose

This memo documents the diagnostic analyses evaluating how distinct forms of contact with the criminal legal system—**incarceration experience** (`incarc`), **proximate contact** (`knowincarc`), and **victimization** (`vict`)—relate to perceived deservingness of incarcerated people (**DV:** `d_incppl_pris_abc`, 0–100).\
Models include party ID, ideology, punitiveness, FIRE items, demographics, and a **generalized deservingness** control (`desor_core`). Analyses are unweighted and emphasize *model behavior*, *explanatory power*, and *unique variance*, not publication-ready estimates.

------------------------------------------------------------------------

## Data Snapshot

-   N (rows used across sections): \~2,222; all variables numeric.\
-   DV distribution: mean ≈ **46**, SD ≈ **28**, modest right skew (0–100).\
-   Contact prevalence: **8%** `incarc`, **31%** `knowincarc`, **12%** `vict` (expected imbalance).\
-   Missingness: ideology has more missing than other covariates; models use listwise deletion.

------------------------------------------------------------------------

## Tests and What Each Tells Us

### 1. Base and Full Models (with HC1 robust SEs)

```{r}
lm(d_incppl_pris_abc ~ contact + controls)
coeftest(model, vcov = vcovHC(model, type = "HC1"))
```

**Purpose:** Estimate direction and magnitude of associations; inference robust to heteroskedasticity.

**Key coefficients (full models):**

| Predictor | Estimate | SE | p-value | Interpretation |
|----|----|----|----|----|
| `incarc` | +8.70 | 2.31 | \<.001 | Direct incarceration → higher deservingness |
| `knowincarc` | +5.35 | 1.25 | \<.001 | Knowing someone incarcerated → higher deservingness |
| `vict` | +1.83 | 1.84 | n.s. | Victimization effect small / unreliable |
| `desor_core` | +0.74 | 0.035 | ≪ .001 | Strong positive relationship |
| `social_ideology` | −1.96 | 0.85 | .02 | Conservative orientation → lower deservingness |
| `penalpoint` | −1.18 | 0.35 | \<.001 | Punitiveness → lower deservingness |

------------------------------------------------------------------------

### 2. Univariate Explanatory Power (Adj. R² and AIC)

Writing

------------------------------------------------------------------------

lm(d_incppl_pris_abc \~ each_variable)

Text

**Purpose:** Evaluate how much outcome variance each variable explains alone.\

**Metric:** Adjusted R² (fit adjusted for model size) and AIC (parsimony).

**Findings (sorted by Adj. R²):**

| Variable          | Adj. R²     | Interpretation                   |
|-------------------|-------------|----------------------------------|
| `desor_core`      | 0.30        | Explains 30% of outcome variance |
| `social_ideology` | 0.08        | Moderate relationship            |
| `fiscal_ideology` | 0.076       | Moderate relationship            |
| FIRE items        | 0.045–0.057 | Smaller                          |
| `penalpoint`      | 0.036       | Smaller                          |
| Partisanship      | 0.034–0.035 | Smaller                          |
| Contact vars      | \< 0.02     | Minimal alone                    |

------------------------------------------------------------------------

### 3. Incremental Δ Adjusted R² Tests

Writing

------------------------------------------------------------------------

ΔR2 \<- summary(expanded)\$adj.r.squared - summary(base)\$adj.r.squared

Text

**Purpose:** Quantify added variance from each control beyond a contact-only model.\

**Finding:** Adding `desor_core` raises Adj. R² by ≈ **+0.24**; other controls negligible.\

**Note:** Sensitive to listwise N differences—use descriptively.

------------------------------------------------------------------------

### 4. Semi-Partial R² (Unique Variance in Full Model)

Writing

------------------------------------------------------------------------

library(rsq)\
rsq.partial(full_model, type = "sp")

Text

**Purpose:** Measure each predictor’s unique contribution within the full model.

**Unique Variance (Semi-partial R²):**

| Variable          | Incarcerated | Victimized | Proximate |
|-------------------|--------------|------------|-----------|
| `desor_core`      | 0.2266       | 0.2263     | 0.2252    |
| Contact           | 0.0090       | 0.0006     | 0.0115    |
| `penalpoint`      | 0.0073       | 0.0066     | 0.0070    |
| `social_ideology` | 0.0036       | 0.0033     | 0.0036    |

**Takeaway:** Attitudes ≫ contact ≫ ideology/punitiveness.

------------------------------------------------------------------------

### 5. Standardized Coefficients (β\*)

Writing

------------------------------------------------------------------------

β\* = β \* sd(X) / sd(Y)

Text

**Purpose:** Put predictors on a common SD scale to compare magnitudes.\

(Intercept not standardized.)

| Predictor         | β\*   | Interpretation      |
|-------------------|-------|---------------------|
| `desor_core`      | 0.51  | Strongest predictor |
| `incarc`          | 0.08  | Modest, positive    |
| `knowincarc`      | 0.06  | Modest, positive    |
| `vict`            | 0.02  | Small, not sig.     |
| `social_ideology` | −0.11 | Negative            |
| `penalpoint`      | −0.08 | Negative            |

------------------------------------------------------------------------

## Core Conclusions

1.  

2.  

    **Generalized deservingness (`desor_core`) dominates** perceived deservingness of incarcerated people (≈30% total, ≈23% unique, β\*≈0.51).

3.  

4.  

    **Contact variables** have positive, independent, but modest effects: incarceration and proximate contact are significant; victimization is not.

5.  

6.  

    **Attitudinal and ideological controls** (punitiveness, social ideology) are consistent but smaller.

7.  

8.  

    **Party ID, FIRE items, and demographics** are negligible once attitudes are controlled.

## Modeling Decisions

-   

-   

    Retain `desor_core` as a control in all subsequent models (major confounder).

-   

-   

    Continue to policy outcomes (`prisonhelp`, `prisonpen`) using the same control set.

-   

-   

    Test **mediation:** `contact → desor_core → policy attitude`.

-   

-   

    Add robustness: weights, block-restricted, and alternative racial-attitudes batteries.

-   

------------------------------------------------------------------------

## Figures and Tables (Committee-Facing)

| Element         | Description                                         |
|-----------------|-----------------------------------------------------|
| **Figure A1**   | DV density (mean ≈ 46, SD ≈ 28)                     |
| **Table A2**    | Cross-tabs of contact types                         |
| **Table B1**    | Univariate Adj. R² (sorted)                         |
| **Tables 1–12** | Model coefficients, semi-partial R², standardized β |
| **Figure C1**   | Semi-partial R² bars: `desor_core` ≫ others         |
| **Figure C2**   | ΔAdj. R² gain when adding `desor_core`              |

stored objects for reference

```{r}
# Descriptives
summary_table; dv_density_data; contact_crosstab

# Model Coefficients
coef_table_incarc; coef_table_vict; coef_table_prox

# Unique Variance
sp_table_inc; sp_table_v; sp_table_prox

# Standardized Betas
table_std_incarc; table_std_vict; table_std_prox

# Bivariate Diagnostics
results_inc; results_vict; results_prox

# Incremental Fit
minc_incr2; mvict_incr2; mprox_incr2

# N Checks
ncheck_inc; ncheck_vict; ncheck_prox

# Cross-Model Summary
contact_model_summary

```

## Bridge to Next Phase: Policy + Mediation

Next analyses will shift to **policy outcomes** (`prisonhelp`, `prisonpen`) to examine whether contact influences support for criminal-justice policies directly and/or indirectly through deservingness perceptions.

**Planned analyses:**

1.  

2.  

    **Direct effects:**

    ```         
    lm(prisonhelp ~ contact + desor_core + controls)
    lm(prisonpen  ~ contact + desor_core + controls)
    ```

3.  

4.  

    **Mediation:**

    ```         
    mediation::mediate(model_m, model_y, treat = "contact", mediator = "desor_core")
    ```

5.  

6.  

    **Robustness checks:**

    -   

    -   

        Weighted models

    -   

    -   

        Block-restricted subsets

    -   

    -   

        Alternative racial-attitudes controls

    -   

7.  

8.  

    **Outputs:**\

    Report standardized β and semi-partial R² for all policy and mediation models to parallel this stage.

9.  

------------------------------------------------------------------------

## Quick Glossary

| Statistic | What It Tests | Use |
|----|----|----|
| **Adj. R²** | Overall model fit | Assess explanatory power |
| **ΔAdj. R²** | Incremental fit from added variable | Descriptive variable importance |
| **Semi-partial R²** | Unique variance | Primary “importance” metric |
| **Standardized β** | Scaled effect size | Compare across predictors |
| **HC1 SEs** | Robust inference | Validate significance under heteroskedasticity |
