---
title: "analysis"
format: html
editor: visual
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,       # Hide code by default
  warning = FALSE,
  message = FALSE
)

# Load packages
library(dplyr)
library(tidyr)
library(readr)
library(janitor)   # for clean_names()
library(tibble)
library(stringr)
library(haven)
library(gt)
library(marginaleffects)
library(mediation)
library(naniar)
library(rsq)
library(patchwork)
library(modelsummary)
library(rlang)
library(forcats)
library(ggplot2)
library(broom)
library(knitr)

source("variable_labels.R")
```

```{r}

here::here()

sample <- read_csv("data/cleaned_control_data.csv")

not_full <- read_csv("data/cleaned_sans_control_data.csv")


#sample <- readr::read_csv("data/cleaned/...")
#sample <- readr::read_csv("data/cleaned/sept_14.csv") 
#sample <- readr::read_csv("data/cleaned_control_data.csv")


#excluded no control responders: 

excluded <- readr::read_csv("data/cleaned_sans_control_data.csv")

nrow(sample)
nrow(excluded)
#sample
#names(sample)
```

# Descriptive Analyses

# 1. Policy

### 2.1A Police Policy

```{r}

source("variable_labels.R")
cor_matrix <- cor(
  sample[police_items_1_to_5],
  use = "pairwise.complete.obs"
)

round(cor_matrix, 2)

library(psych)


alpha_result <- psych::alpha(sample[police_items_1_to_5])

print(alpha_result)

names(alpha_result)

alpha_result$alpha.drop

alpha_result

as.data.frame(alpha_result$alpha.drop)

View(alpha_result$alpha.drop)

print(sample %>%
  pivot_longer(cols = all_of(police_items_1_to_5)) %>%
  ggplot(aes(x = value)) +
  geom_histogram(binwidth = 1) +
  facet_wrap(~ name))

#avg check
sample <- sample %>%
  mutate(police_index = rowMeans(across(all_of(police_items_1_to_5)), na.rm = TRUE))

summary(sample$police_index)

# police index 1: index of all the police items
sample <- sample %>%
  mutate(
    police_reform_index = rowMeans(across(all_of(police_items_1_to_5)), na.rm = TRUE)
  )

# police index 2: index of the items that hang together best

police_reform_core <- c(
  "no_strangle",
  "police_tracker",
  "police_training",
  "charge_police"
)

sample <- sample %>%
  mutate(
    police_reform_core_index = rowMeans(across(all_of(police_reform_core)), na.rm = TRUE)
  )

# last - 3 groups 
sample <- sample %>%
  mutate(
    police_violence_prevention = rowMeans(across(c("q15_2_1_to_5", "q15_4_1_to_5")), na.rm = TRUE),
    police_accountability = rowMeans(across(c("q15_3_1_to_5", "q15_5_1_to_5")), na.rm = TRUE),
    police_local_residency = q15_1_1_to_5
  )

## POLICE MEANS ANALYSIS 

# index 1: police_reform_index

## baseline: no impact
all_baseline_mean <- mean(
  sample$police_reform_index[sample$impact == 0],
  na.rm = TRUE
)

## compute means for each impact group
all_police_means <- tibble(
  var = impact_vars
) %>%
  mutate(
    label = impact_labels[var],
    mean = sapply(var, function(v) {
      mean(
        sample$police_reform_index[
          sample[[v]] == 1
        ],
        na.rm = TRUE
      )
    })
  ) %>%
  bind_rows(
    tibble(
      var = "no_impact",
      label = "No impact",
      mean = all_baseline_mean
    ),
    .
  ) %>%
  mutate(
    diff_from_no_impact = mean - all_baseline_mean
  )

all_police_means

##
# index 2: 
##

# baseline: no impact
baseline_mean <- mean(
  sample$police_reform_core_index[sample$impact == 0],
  na.rm = TRUE
)

# compute means for each impact group
police_means <- tibble(
  var = impact_vars
) %>%
  mutate(
    label = impact_labels[var],
    mean = sapply(var, function(v) {
      mean(
        sample$police_reform_core_index[
          sample[[v]] == 1
        ],
        na.rm = TRUE
      )
    })
  ) %>%
  bind_rows(
    tibble(
      var = "no_impact",
      label = "No impact",
      mean = baseline_mean
    ),
    .
  ) %>%
  mutate(
    diff_from_no_impact = mean - baseline_mean
  )

print(police_means) # index 2
print(all_police_means)
## index 3?
##

```

**Write up on coding police variables decision**

all - participants' opposition/ support (coded 1-5, unsure = 3)

\(1\) Require that police officers live in the cities or towns they police

\(2\) Make it a crime for police to use chokeholds or strangleholds

\(3\) Create a federal government database to track police officers who have been accused of misconduct

\(4\) Require police to be trained in nonviolent alternatives to deadly force

\(5\) Allow civilians to file lawsuits against police officers that use excessive force or misconduct

My loose overall theory is that experiences with crime victimization may influence trust in government and policy learning. Ie, people who have been victimized may view the CJ system as differently legitimate and trustworthy, potentially extend that to the state; they will also learn things about policy. That's the context in which I'm thinking about how to use these survey items.

My initial view is – theoretically – these mostly might fall into two categories. The first category is policing police, or police oversight – questions 3 and 5. The second is about police violence, or police violence prevention - so questions 2 and 4. My initial view is that question 1 is different, or might not as clearly fit into the buckets. but it could also fit into some idea of constraining police?

I want to make this decision in an appropriate way from the research design perspective – not select on my (xyz - forgot what this is called).

I'm not entirely clear if grouping these variables in a single index improves analyses – more power, especially given that my victimized sample is only 277? One option is use as a whole index, but then also break out separate analyses. Another is to do 3 (keep local as a standalone – this might look like:

police_map \<- tibble::tribble( \~var, \~label, \~section, "q15_1_1_to_5", "Local Police", "Misc. Police", "q15_2_1_to_5", "No Strangle", "Police Violence", "q15_3_1_to_5", "Police Tracker", "Police Police", "q15_4_1_to_5", "Police Training", "Police Violence", "q15_5_1_to_5", "Charge Police", "Police Police", )

) – odds are I would still want to run analyses with them as distinct variables. Last option is all 3?

Relationships between police variables

- Police should be local

  - & no strangle = 0.41

  - & track misconduct = 0.45

  - & require alternatives training = 0.39

  - & allow civilian lawsuits against = 0.39

- Strangleholds & chokeholds illegal

  - & police local = 0.41

  - & track misconduct = 0.58

  - & require alternatives training = 0.55

  - & allow civilian lawsuits against = 0.58

- Federal misconduct database

  - & police local = 0.45

  - & no strangle = 0.48

  - & track misconduct = 0.67

  - & require alternatives training = 0.70

- Require police to be trained in nonviolent alternatives to deadly force

  - & police local = 0.39

  - & no strangle = 0.55

  - & track misconduct = 0.67

  - & allow civilian lawsuits against = 0.65

- Allow civilians to file lawsuits against police officers that use excessive force or misconduct

  - & police local = 0.39

  - & no strangle = 0.58

  - & track misconduct = 0.70

  - & require alternatives training = 0.65

Alpha for all = 0.85 – reason to use as one.

ADD DECISION WRITE UP

Decision: two layers. First (index), second (ones that hang together best), then last (3) created above.

**Basic Means Write-Up: Index 1**

```{r}


```

```{r}

```

Police Means Tables

```{r}


```

# 2. Good Controls?

## Y ~ X, Z ~ X, y ~ z

```{r}

m_a <- lm(min_wage ~ pid_3, data = sample) # given someone's PID, how much min wage support would we predict



m_b <- lm(impact ~ pid_3, data = sample)  # given unit's pid, how much carceral impact would we predict? 

y_resid <- resid(m_a)
x_resid <- resid(m_b)

m_c <- lm(y_resid ~ x_resid) # among the variation in impact that is unrelated to PID, is that variation associated with the variation in minimum-wage support that is also unrelated to PID?

summary(m_c)
```
```{r}

m_a1 <- lm(min_wage ~ self_ideo, data = sample) # given someone's PID, how much min wage support would we predict



m_b1 <- lm(impact ~ self_ideo, data = sample)  # given unit's pid, how much carceral impact would we predict? 

y_resid_a1 <- resid(m_a1)
x_resid_a2 <- resid(m_b1)

m_c_a1 <- lm(y_resid_a1 ~ x_resid_a2) # among the variation in impact that is unrelated to PID, is that variation associated with the variation in minimum-wage support that is also unrelated to PID?

summary(m_c_a1)
```


all 
```{r}

covs_min_wage <- lm(min_wage ~ White + Black + Hisp_Latino + Asian_Am + Native_Am + OtherRace, data = sample)

summary(covs_min_wage)

covs_min_wage_1 <- lm(min_wage ~ pid_3 + self_ideo + desor_core + White + Black + Hisp_Latino + Asian_Am + Native_Am + OtherRace + gov_attn, data = sample)

summary(covs_min_wage_1)

```


# 3. Policy Regressions

This section continues to use variable_labels.R. The goal of this section was to avoid and clean up extensive seperated regressions, and instead incorporate use loop that iteratively work through outcome variables and incorporate controls.

PRE 08-26, it had two sets of regression models.

Set 1 are asks "across the full sample, how is each form of impact associated with each policy outcome before adding controls (ie, prisonhelp \~ incarc) and then after/as we add controls (ie, prisonhelp \~incarc + Black + Hispanic + OtherRace + gender). (policy_dummies_waves)

Set 2 asks instead how each impacted subgroup compares specifically compares with respondents who are fully not-impacted. ie, I throw out respondents who are impacted in some other way, but not directly incarcerated; I compare impact == 0 to incarc == 1).

As of now, it (policy_vs_nonimp_waves) compares those groups (ie incarc == 1 and impact == 0) and then does so with controls.

Then this chunk builds tables.

\*\*\*08-26 I HAVE NOW RECODED incarc, knowincarc, and vict so that 1 = impacted in that particular way and 0 = NOT IMPACTED at all (ie, impact == 0).

This means that I don't need the same 2 goal situation. I am now cleaning this up so that the main models will say:

For every policy DV:

1.  Any impact vs no impact
2.  Direct impact vs no impact
3.  Proximal impact vs no impact
4.  Victim impact vs no impact

Run each through: Model 1: no controls Model 2: demographics Model 3: demographics + party/ideology Model 4: full attitudes

### 3.1A: Framework for model sequence

control_waves \| \|-- nocontrol \|-- demo \|-- party_ideo \|-- full_attitudes

```{r}

names(sample)
# ============================================================
# 3. TO START - big picture here is defining the model "waves." The first is bivariate o no controls ("nocontrol"), then demo is models with just demographic controls (8-26 this is just race I think, then demographic controls and part ID and fiscal ideology and social ideology (party_ideo) then with all controls, so aforementioned plus FIRE and *eventually, need to fix* deservingness orientation. 
# The following controls are grouped in vriable_labels, but prepped in variable_prep_script:
#   demo_controls
#   party_ideo_controls
#   desor_core***
#   fire_controls
# ============================================================

demo_controls <- c(
  "Black", "Hisp_Latino", "Asian_Am", "Native_Am", "OtherRace", "gender"
)
problematic_race_vars <- c(
  "White",
  "Black",
  "Hisp_Latino",
  "AllOthers"
)
problematic_race_labels <- c(
  White = "White",
  Black = "Black",
  Hisp_Latino = "Hisp_Latino",
  OtherRaces = "AllOthers"
)

party_ideo_controls <- c(
  "pid_3",
  "fiscal_ideology",
  "social_ideology"
)

party_ideo_controls_labels <- c(
  pid_3 = "Party ID",
  # Republican = "Republican",
  # Democrat = "Democrat",
  fiscal_ideology = "Fiscal ideology",
  social_ideology = "Social ideology"
)

fire_controls <- c(
  "fire_rare",
  "fire_privilege",
  "fire_angry",
  "fire_fear"
)

fire_controls_labels <- c(
  fire_rare = "Racism rare",
  fire_privilege = "White privilege",
  fire_angry = "Angry about racism",
  fire_fear = "Fear other races"
)

control_waves <- list(
  nocontrol = character(0),
  demo = demo_controls,
  party_ideo = c(problematic_race_vars, party_ideo_controls),
  full_attitudes = c(problematic_race_vars, party_ideo_controls,
                     fire_controls, desor_core)
)
```


### 3.2 Define lm function that will waves will operate thru

```{r}
# ============================================================
# This starts by defining my reusable function. That function says:

#"whenever (later) I give this function a dataset (df), a dependent variable (dv), an independent variable (iv), and some controls (controls), R should:

##(a) start with a default of assuming there are no controls, if no controls are given (character (0)); 
## (b) in the right-hand side (rhs) or the equation, use the supplied iv and supplied controls to make a vector with all vraibles that should appear to the right of ~
## (c) (fml) construct an ordinary regression syntax with the supplied dv on the left of ~, then supplied rhs info and turn it into a signle string (collapse = "+"), and 
## (d) treat this as a statistical formula rather than a string (as.formula())
## (e) run the regression
# ============================================================

make_policy_model <- function(df, dv, iv, controls = character(0)) { #character(0) just means empty character vector so this function works for uncontrolled model then for controlled model
  rhs <- c(iv, controls)
  fml <- as.formula(paste(dv, "~", paste(rhs, collapse = " + ")))
  lm(fml, data = df)
}

# ============================================================
# MAIN MODEL GENERATING LOOOP Now, I need to line up buckets to put the models 
## that place is "policy_dummies_waves"
#
# Then, I need to say how I will fill it. I will fill it with waves of models that loop through dvs and iteratively add controls 
#for each dv, generates 4 substantive comparisons:
### impact versus non impact
### incarc versus non impact
### knowincarc versus non impact
### victimized vs non impact
# ============================================================

policy_dummies_waves <- list() #this is actually now main model generating loop

for (wave_name in names(control_waves)) {#calls list of waves in control waves

  model_list <- list() #fresh model list within each wave

  for (dv in policy_vars) {
    for (iv in impact_vars) { #loop thru dvs, ivs, labels file gives the 4 ivs; fits 4 models for each outcome

      model_name <- paste0("m_", dv, "_", iv) #ie model using prisonhelp and incarc will be named m_prisonhelp_incarc

      # NOW, actually telling it to run and store the regressions
      model_list[[model_name]] <- make_policy_model(
        df = sample,
        dv = dv, 
        iv = iv,
        controls = control_waves[[wave_name]]
      )
    }
  }

  policy_dummies_waves[[wave_name]] <- model_list
}

##after the two inner loops finish organization for storing them. 
## object is roughly: policy_dummies_waves : A no control -> A1 no control m_prisonhelp_impact, A2 no control m_prisonhelp_incarc. . ..etc; B1 demographic controls m_prisonhelp_impact....(see below)
#policy_dummies_waves
# |
# |-- nocontrol
# |     |-- m_prisonhelp_impact
# |     |-- m_prisonhelp_incarc
# |     |-- m_prisonhelp_knowincarc
# |     |-- m_prisonhelp_vict
# |     |-- ...
# |
# |-- demo
# |     |-- m_prisonhelp_impact
# |     |-- m_prisonhelp_incarc
# |     |-- ...
# |
# |-- party_ideo
# |
# |-- full_attitudes
 
# ============================================================
# NEXT -- making hopefully more convenient references so that  instead of typing out the full "policy_dummies_waves[["nocontrol"]][["m_prisonhelp_incarc"]] I can type "policy_dummies_models[{"m_prisonhelp_incarc"}] (really tho, the table making code came first and this lets me adopt above outputs to )
## 
#
# T 
# ============================================================



policy_dummies_models <- policy_dummies_waves[["nocontrol"]]
policy_dummies_demo_models <- policy_dummies_waves[["demo"]]
policy_dummies_partyideo_models <- policy_dummies_waves[["party_ideo"]]
policy_dummies_full_models <- policy_dummies_waves[["full_attitudes"]]

```



## 3.3 -- PRE 08-26 RECODING "restricted" model for when incarc, knowincarc, vict were coded such that 0 meant not incarc, not vict, etc.. (08-2026 no longer necessary for main reported relationships having recoded incarc, knowincarc, and vict)

```{r}
# goal 1 wave loop above
# 
# make_restricted_model <- function(df, dv, subgroup_var, controls = character(0)) {
# 
#   if (subgroup_var == "impact") {
#     rhs <- c("impact", controls)
#     fml <- as.formula(paste(dv, "~", paste(rhs, collapse = " + ")))
#     return(lm(fml, data = df))
#   }
# 
#   df_sub <- df %>%
#     dplyr::filter(impact == 0 | .data[[subgroup_var]] == 1) %>%
#     dplyr::mutate(treat = ifelse(.data[[subgroup_var]] == 1, 1, 0))
# 
#   rhs <- c("treat", controls)
#   fml <- as.formula(paste(dv, "~", paste(rhs, collapse = " + ")))
#   lm(fml, data = df_sub)
# }
```

### 3.3 Appendix Tables - Each IV Model Grouped around DV

```{r}
# ============================================================
# Build one LaTeX regression-results table per DV
# Assumes  variable-label script has already been sourced,
# so the following objects already exist:
#   - impact_vars
#   - impact_labels
#   - policy_vars
#   - policy_map
# Also assumes policy_dummies_models already exists.
# ============================================================


# ------------------------------------------------------------
# sanity checks: make sure the variables referenced in 
# helper objects actually exist in the data
# -----------------------------------------------------------

missing_policy_vars <- setdiff(policy_vars, names(sample))
missing_impact_vars <- setdiff(impact_vars, names(sample))

if (length(missing_policy_vars) > 0) {
  stop("These policy_vars are missing from sample: ",
       paste(missing_policy_vars, collapse = ", "))
}

if (length(missing_impact_vars) > 0) {
  stop("These impact_vars are missing from sample: ",
       paste(missing_impact_vars, collapse = ", "))
}

# ------------------------------------------------------------
# create a named vector of DV labels from policy_map
# ------------------------------------------------------------

policy_labels <- setNames(policy_map$label, policy_map$var)

# ------------------------------------------------------------
# helper function: convert p-values to significance stars
# ------------------------------------------------------------

p_to_stars <- function(p) {
  dplyr::case_when(
    is.na(p)  ~ "",
    p < .001  ~ "***",
    p < .01   ~ "**",
    p < .05   ~ "*",
    TRUE      ~ ""
  )
}

# ------------------------------------------------------------
# function: build a clean one-DV results table
# ------------------------------------------------------------

make_dv_results_table <- function(dv_name, model_list) {

  rows <- lapply(impact_vars, function(iv_name) {

    model_name <- paste0("m_", dv_name, "_", iv_name)

    if (!model_name %in% names(model_list)) {
      stop("Model not found in policy_dummies_models: ", model_name)
    }

    mod <- model_list[[model_name]]

    coef_row <- broom::tidy(mod) %>%
      dplyr::filter(term != "(Intercept)") %>%
      dplyr::mutate(
        iv_name = iv_name,
        iv_label = impact_labels[[iv_name]],
        coef = round(estimate, 3),
        se = round(std.error, 3),
        p = round(p.value, 3),
        stars = p_to_stars(p.value),
        coef_with_stars = paste0(format(coef, nsmall = 3), stars),
        n = broom::glance(mod)$nobs,
        r2 = round(broom::glance(mod)$r.squared, 3)
      ) %>%
      dplyr::select(iv_label, coef_with_stars, se, p, n, r2)

    coef_row
  })

  dplyr::bind_rows(rows) %>%
    dplyr::rename(
      `Independent Variable` = iv_label,
      `Coefficient` = coef_with_stars,
      `SE` = se,
      `p-value` = p,
      `N` = n,
      `R^2` = r2
    )
}

# ------------------------------------------------------------
# function: convert one DV results table into LaTeX code
# ------------------------------------------------------------

make_dv_latex_table <- function(dv_name, model_list) {

  results_df <- make_dv_results_table(dv_name, model_list)

  caption_text <- paste0("Bivariate Models Predicting ", policy_labels[[dv_name]])
  label_text <- paste0("tab:", dv_name, "_bivariate")

  knitr::kable(
    results_df,
    format = "latex",
    booktabs = TRUE,
    caption = caption_text,
    label = label_text,
    align = c("l", "c", "c", "c", "c", "c"),
    escape = TRUE
  )
}

# ------------------------------------------------------------
# create one LaTeX table per DV and store them in a named list
# ------------------------------------------------------------

dv_latex_tables <- lapply(policy_vars, function(dv_name) {
  make_dv_latex_table(dv_name, policy_dummies_models)
})

names(dv_latex_tables) <- policy_vars

# ------------------------------------------------------------
# example: print one table to the console so you can copy/paste
# into Overleaf
# ------------------------------------------------------------

#cat(dv_latex_tables[["prisonhelp"]])

# ------------------------------------------------------------
# write all tables to separate .tex files
# ------------------------------------------------------------

for (dv_name in policy_vars) {
  writeLines(
    dv_latex_tables[[dv_name]],
    con = paste0(dv_name, "_bivariate_table.tex")
  )
}
```

### 3.4 Compare Each Impacted Subgroup to Non-Impacted

Restricted sample comparison logic (not impacted vs incarcerated) ultimate goal: no-controls to controls main results table built from the restricted-sample non-impacted-versus-subgroup models. then extend the restricted-sample models rather than rebuilding everything from scratch when adding controls.

Why doing this instead of a factor variable for all impact: -restricted-sample regressions keep the logic consistent with everything I are already doing and make it much easier to scale to later models with controls. A t-test table and a restricted-sample dummy regression are telling the same story in the no-controls case. But the regression route gives me a single framework you can preserve when you later add demographic controls, then political controls, then attitudinal controls. If I switch to t-tests for model 2 and regressions for later models, I create an unnecessary break in the workflow and probably more formatting work.

A single categorical variable like 0 = not impacted, 1 = direct only, 2 = proximal only, 3 = victim only, 4 = multiple forms would answer a different question. It would force you to partition the sample into mutually exclusive bins, which is useful for some appendix descriptives and maybe later robustness checks, but it would also collapse substantively important overlap into a catch-all “multiple forms” group and make it harder to estimate the specific contrast you currently care about: whether each exposure type, as experienced in the real world, is associated with different outcomes relative to the fully non-impacted. For your main table, the restricted-sample approach is cleaner and more faithful to the theory.

```{r}

# ============================================================
# Goal #2: compare each impacted subgroup to the fully
# non-impacted group
# ============================================================

policy_vs_nonimp_waves <- list()

for (wave_name in names(control_waves)) {

  model_list <- list()

  for (dv in policy_vars) {
    for (iv in impact_vars) {

      model_name <- paste0("comp_", dv, "_", iv)

      model_list[[model_name]] <- make_restricted_model(
        df = sample,
        dv = dv,
        subgroup_var = iv,
        controls = control_waves[[wave_name]]
      )
    }
  }

  policy_vs_nonimp_waves[[wave_name]] <- model_list
}

policy_vs_nonimp_nocontrol_models <- policy_vs_nonimp_waves[["nocontrol"]]
policy_vs_nonimp_demo_models <- policy_vs_nonimp_waves[["demo"]]
policy_vs_nonimp_partyideo_models <- policy_vs_nonimp_waves[["party_ideo"]]
policy_vs_nonimp_full_models <- policy_vs_nonimp_waves[["full_attitudes"]]


summary(policy_vs_nonimp_nocontrol_models$comp_prisonhelp_impact)
summary(policy_vs_nonimp_demo_models$comp_prisonhelp_impact)

```

reusable system for:

a results-section table for prisonhelp and prisonpen,

appendix tables for the remaining DVs in the same architecture,

a blank second model block for later controlled models,

placeholder control rows so the table skeleton already matches where you are headed.

### 3.5 Tables Loop

```{r}
# ============================================================
# Advisor-style restricted-sample tables
# Goal: compare each impacted subgroup to non-impacted only
# Model 2 = no-controls restricted-sample models
# Model 3 = placeholder for later controlled models
#
# Packages used:
# dplyr, stringr, purrr, lmtest, sandwich
# ============================================================

library(dplyr)
library(stringr)
library(purrr)
library(lmtest)
library(sandwich)

# ------------------------------------------------------------
# Labels from your variable script
# ------------------------------------------------------------
policy_labels <- setNames(policy_map$label, policy_map$var)

# ------------------------------------------------------------
# Formatting helpers
# ------------------------------------------------------------
p_to_stars <- function(p) {
  dplyr::case_when(
    is.na(p) ~ "",
    p < .001 ~ "***",
    p < .01  ~ "**",
    p < .05  ~ "*",
    TRUE     ~ ""
  )
}

fmt_num <- function(x, digits = 3) {
  ifelse(is.na(x), "", sprintf(paste0("%.", digits, "f"), x))
}

latex_escape <- function(x) {
  x %>%
    str_replace_all("\\\\", "\\\\\\\\") %>%
    str_replace_all("([#$%&_{}])", "\\\\\\1") %>%
    str_replace_all("~", "\\\\textasciitilde{}") %>%
    str_replace_all("\\^", "\\\\textasciicircum{}")
}

# ------------------------------------------------------------
# Extract one non-intercept coefficient from a model
# robust = HC1 by default
# ------------------------------------------------------------
get_single_term_stats <- function(mod, robust = TRUE) {
  if (robust) {
    ct <- lmtest::coeftest(mod, vcov. = sandwich::vcovHC(mod, type = "HC1"))
    out <- data.frame(
      term      = rownames(ct),
      estimate  = ct[, 1],
      std.error = ct[, 2],
      p.value   = ct[, 4],
      stringsAsFactors = FALSE
    )
  } else {
    sm <- summary(mod)$coefficients
    out <- data.frame(
      term      = rownames(sm),
      estimate  = sm[, 1],
      std.error = sm[, 2],
      p.value   = sm[, 4],
      stringsAsFactors = FALSE
    )
  }

  out %>%
    filter(term != "(Intercept)") %>%
    slice(1) %>%
    mutate(
      n  = stats::nobs(mod),
      r2 = summary(mod)$r.squared
    )
}

# ------------------------------------------------------------
# Build one cell for one DV x subgroup x model-list
# Returns something like:
# \makecell[c]{0.210*** \\ (0.030)}
#
# If model_list is NULL, returns blank cell
# ------------------------------------------------------------
make_comp_cell <- function(model_list, dv, iv, robust = TRUE) {

  if (is.null(model_list)) return("")

  model_name <- paste0("comp_", dv, "_", iv)

  if (!model_name %in% names(model_list)) return("")

  stats_row <- get_single_term_stats(model_list[[model_name]], robust = robust)

  coef_txt <- paste0(
    fmt_num(stats_row$estimate, 3),
    p_to_stars(stats_row$p.value)
  )

  se_txt <- paste0("(", fmt_num(stats_row$std.error, 3), ")")

  paste0("\\makecell[c]{", coef_txt, " \\\\ ", se_txt, "}")
}

# ------------------------------------------------------------
# Split a vector into chunks for appendix tables
# ------------------------------------------------------------
chunk_vector <- function(x, chunk_size = 2) {
  split(x, ceiling(seq_along(x) / chunk_size))
}

# ------------------------------------------------------------
# Main table builder
#
# dv_vec: which DVs go in the columns for each model block
# model2_list: required (your no-controls restricted-sample list)
# model3_list: optional, NULL for now
# row_vars: rows to display under carceral contact
# include_placeholder_control_rows: include blank future rows
# ------------------------------------------------------------
make_restricted_panel_latex_table <- function(
    dv_vec,
    model2_list,
    model3_list = NULL,
    caption_text,
    label_text,
    model2_title = "Model 2",
    model3_title = "Model 3",
    row_vars = c("impact", "incarc", "knowincarc", "vict"),
    control_vars = c("Black", "Hisp_Latino", "Asian_Am", "Native_Am", "OtherRace", "gender"),
    control_labels = c(
      Black = "Black",
      Hisp_Latino = "Hispanic/Latino",
      Asian_Am    = "Asian American",
      Native_Am   = "Native American",
      OtherRace = "Other Race",
      gender = "Male"
    ),
    robust = TRUE
) {

  # ----- column setup -----
  n_dv <- length(dv_vec)
  n_model_blocks <- 2
  total_data_cols <- n_dv * n_model_blocks
  align_string <- paste0("l", paste(rep("c", total_data_cols), collapse = ""))

  # ----- header rows -----
  header_top <- c(
    "",
    paste0("\\multicolumn{", n_dv, "}{c}{", latex_escape(model2_title), "}"),
    paste0("\\multicolumn{", n_dv, "}{c}{", latex_escape(model3_title), "}")
  )

  header_second <- c(
    "",
    latex_escape(policy_labels[dv_vec]),
    latex_escape(policy_labels[dv_vec])
  )

  cmid_model2 <- paste0("\\cmidrule(lr){2-", 1 + n_dv, "}")
  cmid_model3 <- paste0("\\cmidrule(lr){", 2 + n_dv, "-", 1 + (2 * n_dv), "}")
  cmid_line <- paste(cmid_model2, cmid_model3, collapse = " ")

  # ----- helper: extract a specific term from a model -----
  get_term_cell <- function(model_list, dv, iv, term_name, robust = TRUE) {

    if (is.null(model_list)) return("")

    model_name <- paste0("comp_", dv, "_", iv)
    if (!model_name %in% names(model_list)) return("")

    mod <- model_list[[model_name]]

    if (robust) {
      ct <- lmtest::coeftest(mod, vcov. = sandwich::vcovHC(mod, type = "HC1"))
      out <- data.frame(
        term      = rownames(ct),
        estimate  = ct[, 1],
        std.error = ct[, 2],
        p.value   = ct[, 4],
        stringsAsFactors = FALSE
      )
    } else {
      sm <- summary(mod)$coefficients
      out <- data.frame(
        term      = rownames(sm),
        estimate  = sm[, 1],
        std.error = sm[, 2],
        p.value   = sm[, 4],
        stringsAsFactors = FALSE
      )
    }

    if (!term_name %in% out$term) return("")

    row <- out[out$term == term_name, ]

    coef_txt <- paste0(
      fmt_num(row$estimate, 3),
      p_to_stars(row$p.value)
    )

    se_txt <- paste0("(", fmt_num(row$std.error, 3), ")")

    paste0("\\makecell[c]{", coef_txt, " \\\\ ", se_txt, "}")
  }

  # ----- body rows -----
  body_lines <- c()

  # Section header
  body_lines <- c(
    body_lines,
    paste(
      c(
        "\\textit{Carceral Contact (ref. Not impacted)}",
        rep("", total_data_cols)
      ),
      collapse = " & "
    ),
    "\\\\"
  )

  # For each subgroup, print coefficient row + repeated control rows
  for (iv in row_vars) {

    row_label <- impact_labels[[iv]]
    main_term <- ifelse(iv == "impact", "impact", "treat")

    # main subgroup row
    model2_cells <- sapply(dv_vec, function(dv) {
      get_term_cell(model2_list, dv, iv, main_term, robust = robust)
    })

    model3_cells <- sapply(dv_vec, function(dv) {
      get_term_cell(model3_list, dv, iv, main_term, robust = robust)
    })

    subgroup_line <- paste(
      c(latex_escape(row_label), model2_cells, model3_cells),
      collapse = " & "
    )

    body_lines <- c(body_lines, paste0(subgroup_line, " \\\\"))

    # repeated control rows under this subgroup
    for (ctrl in control_vars) {

      model2_ctrl_cells <- sapply(dv_vec, function(dv) {
        get_term_cell(model2_list, dv, iv, ctrl, robust = robust)
      })

      model3_ctrl_cells <- sapply(dv_vec, function(dv) {
        get_term_cell(model3_list, dv, iv, ctrl, robust = robust)
      })

      ctrl_line <- paste(
        c(latex_escape(control_labels[[ctrl]]), model2_ctrl_cells, model3_ctrl_cells),
        collapse = " & "
      )

      body_lines <- c(body_lines, paste0(ctrl_line, " \\\\"))
    }

    body_lines <- c(body_lines, "\\addlinespace")
  }

  # ----- assemble LaTeX -----
  out <- c(
    "\\begin{table}[!htbp]",
    "\\centering",
    paste0("\\caption{", latex_escape(caption_text), "}"),
    paste0("\\label{", latex_escape(label_text), "}"),
    "\\begin{threeparttable}",
    paste0("\\begin{tabular}{", align_string, "}"),
    "\\toprule",
    paste(header_top, collapse = " & "),
    " \\\\",
    cmid_line,
    paste(header_second, collapse = " & "),
    " \\\\",
    "\\midrule",
    body_lines,
    "\\bottomrule",
    "\\end{tabular}",
    "\\begin{tablenotes}[flushleft]",
    "\\footnotesize",
    "\\item Entries are OLS coefficients with HC1 robust standard errors in parentheses.",
    "\\item Each cell is estimated from a separate restricted-sample regression comparing the named subgroup to the non-impacted group.",
    "\\item Because rows come from different restricted samples, observation counts and $R^2$ are not shown in this table format.",
    "\\item * p$<$0.05, ** p$<$0.01, *** p$<$0.001.",
    "\\end{tablenotes}",
    "\\end{threeparttable}",
    "\\end{table}"
  )

  out
}
```

```{r}
results_dvs <- c("prisonhelp", "prisonpen")

results_table_tex <- make_restricted_panel_latex_table(
  dv_vec = results_dvs,
  model2_list = policy_vs_nonimp_nocontrol_models,
  model3_list = policy_vs_nonimp_demo_models,
  caption_text = "OLS Models of Carceral Policy Orientations - No controls to Demographic Controls",
  label_text = "tab:carceral_policy_orientations_nocontrol_demo",
  model2_title = "Model 2",
  model3_title = "Model 3",
  row_vars = c("impact", "incarc", "knowincarc", "vict"),
  control_vars = c("Black", "Hispanic/Latino", "Asian_Am",
                   "Native_Am", "OtherRace", "gender"),
  control_labels = c(
    Black = "Black",
    Hisp_Latino = "Hispanic/Latino",
    Asian_Am    = "Asian American",
    Native_Am   = "Native American",
    OtherRace = "Other Race",
    gender = "Male"
  ),
  robust = TRUE
)

writeLines(results_table_tex, "results_policy_orientations_nocontrol_demo_table.tex")


##model for just desor

dep_variable <-(sample$knowincarc)
desor_model <- lm(desor_core ~ dep_variable + Black+Hisp_Latino+OtherRace+fiscal_ideology+social_ideology, data = sample)
summary(desor_model)

writeLines(results_table_tex, "desor_controls_table.tex")
```

```{r}
# ------------------------------------------------------------
# 1. RESULTS-SECTION TABLE
# ------------------------------------------------------------
results_dvs <- c("prisonhelp", "prisonpen")

results_table_tex <- make_restricted_panel_latex_table(
  dv_vec = results_dvs,
  model2_list = policy_vs_nonimp_nocontrol_models,
  model3_list = policy_vs_nonimp_demo_models,
  caption_text = "OLS Models of Carceral Policy Orientations - No controls to Demographic Controls",
  label_text = "tab:carceral_policy_orientations_nocontrol_demo",
  model2_title = "Model 2",
  model3_title = "Model 3",
  row_vars = c("impact", "incarc", "knowincarc", "vict"),
  control_vars = c("Black", "Hispanic/Latino", "Asian_Am",
                   "Native_Am", "OtherRace", "gender"),
  control_labels = c(
    Black = "Black",
    Hisp_Latino = "Hispanic/Latino",
    Asian_Am    = "Asian American",
    Native_Am   = "Native American",
    OtherRace = "Other race",
    gender = "Male"
  ),
  robust = TRUE
)

writeLines(results_table_tex, "results_policy_orientations_nocontrol_demo_table.tex")

# Optional preview in console
# cat(results_table_tex, sep = "\n")


# ------------------------------------------------------------
# 2. APPENDIX TABLES FOR REMAINING POLICY DVs
#    Same architecture, chunked two DVs at a time
# ------------------------------------------------------------
appendix_dvs <- setdiff(policy_vars, results_dvs)

# You can keep help/punish separated if you want cleaner appendix organization
appendix_help_dvs <- appendix_dvs[policy_map$section[match(appendix_dvs, policy_map$var)] == "Help"]
appendix_punish_dvs <- appendix_dvs[policy_map$section[match(appendix_dvs, policy_map$var)] == "Punish"]

appendix_help_chunks <- chunk_vector(appendix_help_dvs, chunk_size = 2)
appendix_punish_chunks <- chunk_vector(appendix_punish_dvs, chunk_size = 2)

appendix_tables <- list()

# Help tables
for (i in seq_along(appendix_help_chunks)) {
  dv_chunk <- appendix_help_chunks[[i]]

  tex <- make_restricted_panel_latex_table(
    dv_vec = dv_chunk,
    model2_list = policy_vs_nonimp_nocontrol_models,
    model3_list = NULL,
    caption_text = paste0("Restricted-Sample OLS Models of Help-Oriented Policy Outcomes, Appendix Table ", i),
    label_text = paste0("tab:appendix_help_restricted_", i),
    model2_title = "Model 2",
    model3_title = "Model 3",
    row_vars = c("impact", "incarc", "knowincarc", "vict"),
    include_placeholder_control_rows = TRUE,
    robust = TRUE
  )

  appendix_tables[[paste0("help_", i)]] <- tex
  writeLines(tex, paste0("appendix_help_restricted_table_", i, ".tex"))
}

# Punish tables
for (i in seq_along(appendix_punish_chunks)) {
  dv_chunk <- appendix_punish_chunks[[i]]

  tex <- make_restricted_panel_latex_table(
    dv_vec = dv_chunk,
    model2_list = policy_vs_nonimp_nocontrol_models,
    model3_list = NULL,
    caption_text = paste0("Restricted-Sample OLS Models of Punitive Policy Outcomes, Appendix Table ", i),
    label_text = paste0("tab:appendix_punish_restricted_", i),
    model2_title = "Model 2",
    model3_title = "Model 3",
    row_vars = c("impact", "incarc", "knowincarc", "vict"),
    include_placeholder_control_rows = TRUE,
    robust = TRUE
  )

  appendix_tables[[paste0("punish_", i)]] <- tex
  writeLines(tex, paste0("appendix_punish_restricted_table_", i, ".tex"))
}
```

Looking just at prison help and prison pun after adding in demographic factors

```{r}
summary(policy_vs_nonimp_demo_models$comp_prisonhelp_impact)
summary(policy_vs_nonimp_demo_models$comp_prisonpen_impact)

summary(policy_vs_nonimp_demo_models$comp_prisonhelp_incarc)
summary(policy_vs_nonimp_demo_models$comp_prisonpen_incarc)

summary(policy_vs_nonimp_demo_models$comp_prisonhelp_knowincarc)
summary(policy_vs_nonimp_demo_models$comp_prisonpen_knowincarc)

summary(policy_vs_nonimp_demo_models$comp_prisonhelp_vict)
summary(policy_vs_nonimp_demo_models$comp_prisonpen_vict)


comp_prisonhelp_incarc     
comp_prisonhelp_knowincarc  
comp_prisonhelp_vict       
comp_prisonpen_impact       
comp_prisonpen_incarc      
comp_prisonpen_knowincarc  
comp_prisonpen_vict        


results_dvs <- c("prisonhelp", "prisonpen")

results_table_tex <- make_restricted_panel_latex_table(
  dv_vec = results_dvs,
  model2_list = policy_vs_nonimp_nocontrol_models,
  model3_list = policy_vs_nonimp_demo_models,
  caption_text = "OLS Models of Carceral Policy Orientations - No controls to Demographic Controls",
  label_text = "tab:carceral_policy_orientations_nocontrol_demo",
  model2_title = "Model 2",
  model3_title = "Model 3",
  row_vars = c("impact", "incarc", "knowincarc", "vict"),
  include_placeholder_control_rows = TRUE,
  robust = TRUE
)

writeLines(results_table_tex, "results_policy_orientations_nocontrol_demo_table.tex")
```

##. ##. \## 4-8 Policy MV Regressions (non-restricted)

```{r}
names(sample)
#checkin that reported restricted table matches
#restricted_incarc <- sample %>%
#  filter(impact == 0 | incarc == 1)

#m_check <- lm(prisonhelp ~ incarc + Black + Hispanic + OtherRace + #gender + Republican + Democrat + fiscal_ideology + social_ideology #+desor_core + fire_rare + fire_privilege + fire_angry + fire_fear, data #= restricted_incarc)

#summary(m_check)


#impact - help 
test_impact_full <- lm(prisonhelp ~ impact + Black + Hisp_Latino + OtherRace + gender + Republican + Democrat + fiscal_ideology + social_ideology +desor_core + fire_rare + fire_privilege + fire_angry + fire_fear, data = sample)

test_incarc_full <- lm(prisonhelp ~ incarc + Black + Hisp_Latino + OtherRace + gender + Republican + Democrat + fiscal_ideology + social_ideology +desor_core + fire_rare + fire_privilege + fire_angry + fire_fear, data = sample)

#knowincarc 
test_knowincarc_full <- lm(prisonhelp ~ knowincarc + Black + Hisp_Latino + OtherRace + gender + Republican + Democrat + fiscal_ideology + social_ideology +desor_core + fire_rare + fire_privilege + fire_angry + fire_fear, data = sample)

test_vict_full <- lm(prisonhelp ~ vict + Black + Hisp_Latino + OtherRace + gender +  Republican + Democrat + fiscal_ideology + social_ideology +desor_core + fire_rare + fire_privilege + fire_angry + fire_fear, data = sample)

test_impact_full
test_knowincarc_full
test_incarc_full
test_vict_full

#impact - pun 
test_impact_full_pen <- lm(prisonpen ~ impact + Black + Hisp_Latino + OtherRace + gender + Republican + Democrat + fiscal_ideology + social_ideology +desor_core + fire_rare + fire_privilege + fire_angry + fire_fear, data = sample)

test_incarc_full_pen <- lm(prisonpen ~ incarc + Black + Hisp_Latino + OtherRace + gender + Republican + Democrat + fiscal_ideology + social_ideology +desor_core + fire_rare + fire_privilege + fire_angry + fire_fear, data = sample)

#knowincarc 
test_knowincarc_full_pen <- lm(prisonpen ~ knowincarc + Black + Hisp_Latino + OtherRace + gender + Republican + Democrat + fiscal_ideology + social_ideology +desor_core + fire_rare + fire_privilege + fire_angry + fire_fear, data = sample)

test_vict_full_pen <- lm(prisonpen ~ vict + Black + Hisp_Latino + OtherRace + gender +  Republican + Democrat + fiscal_ideology + social_ideology +desor_core + fire_rare + fire_privilege + fire_angry + fire_fear, data = sample)

test_impact_full_pen
test_knowincarc_full_pen
test_incarc_full_pen
test_vict_full_pen

```

Same as above but for policies

```{r}
policy_dvs <- policy_map %>%
  dplyr::filter(!var %in% c("prisonhelp", "prisonpen")) %>%
  dplyr::pull(var)

impact_vars <- c("impact", "incarc", "knowincarc", "vict")

policy_models <- list()

for (dv in policy_dvs) {
  for (iv in impact_vars) {
    
    model_name <- paste0("m_", dv, "_", iv)
    
    formula <- as.formula(paste(dv, "~", iv))
    
    policy_models[[model_name]] <- lm(formula, data = sample)
  }
}

policy_models_by_dv <- split(
  policy_models,
  gsub("m_(.*)_(impact|incarc|knowincarc|vict)", "\\1", names(policy_models))
)

length(policy_models_by_dv)        # should be 32
names(policy_models_by_dv)         # all model names

# print one
summary(policy_models[["m_q23_2_rev_impact"]])

# or loop through all
lapply(policy_models, summary)

#table for appendix
library(dplyr)
library(purrr)
library(broom)
library(stringr)
library(tidyr)

impact_labels <- c(
  impact = "Any impact",
  incarc = "Direct incarceration",
  knowincarc = "Proximal incarceration contact",
  vict = "Victimization"
)

comparison_table <- imap_dfr(policy_models, function(model, model_name) {
  
  tidied <- broom::tidy(model)
  glanced <- broom::glance(model)
  
  dv <- str_match(model_name, "^m_(.*)_(impact|incarc|knowincarc|vict)$")[,2]
  iv <- str_match(model_name, "^m_(.*)_(impact|incarc|knowincarc|vict)$")[,3]
  
  coef_row <- tidied %>%
    filter(term == iv)
  
  tibble(
    model_name = model_name,
    dv = dv,
    impact_var = iv,
    impact_label = impact_labels[iv],
    estimate = coef_row$estimate,
    std.error = coef_row$std.error,
    statistic = coef_row$statistic,
    p.value = coef_row$p.value,
    n = glanced$nobs,
    r.squared = glanced$r.squared,
    adj.r.squared = glanced$adj.r.squared
  )
})

#add readable dv labels and significance stars
comparison_table <- comparison_table %>%
  left_join(
    policy_map %>% dplyr::select(var, label, section),
    by = c("dv" = "var")
  ) %>%
  mutate(
    stars = case_when(
      p.value < .001 ~ "***",
      p.value < .01  ~ "**",
      p.value < .05  ~ "*",
      p.value < .1   ~ "†",
      TRUE ~ ""
    ),
    coef_with_stars = paste0(sprintf("%.3f", estimate), stars),
    se_fmt = paste0("(", sprintf("%.3f", std.error), ")"),
    p_fmt = sprintf("%.3f", p.value),
    r2_fmt = sprintf("%.3f", r.squared)
  ) %>%
  dplyr::select(
    section, label, dv, impact_label, impact_var,
    coef_with_stars, se_fmt, p_fmt, n, r2_fmt,
    estimate, std.error, p.value, r.squared, adj.r.squared, model_name
  ) %>%
  arrange(section, label, factor(impact_var, levels = c("impact", "incarc", "knowincarc", "vict")))

appendix_table_wide <- comparison_table %>%
  dplyr::select(label, section, impact_var, coef_with_stars, se_fmt, n, r2_fmt) %>%
  mutate(result = paste0(coef_with_stars, " ", se_fmt)) %>%
  dplyr::select(label, section, impact_var, result, n, r2_fmt) %>%
  pivot_wider(
    names_from = impact_var,
    values_from = c(result, n, r2_fmt),
    names_glue = "{impact_var}_{.value}"
  ) %>%
  relocate(section, label)

View(appendix_table_wide)
```

1.  

# 4. DV: Deservingness

### 4.1 All Des Means Loop

```{r}
source("variable_labels.R")

# Optional: prettier target labels
pretty_target <- function(x) {
  x %>%
    str_remove("^d_") %>%
    str_remove("_abc$|_a$|_b$|_c$") %>%
    str_replace_all("_", " ") %>%
    str_to_title()
}

# -----------------------------
# 2. Reshape data long
# -----------------------------

long_df_heatmap <- sample %>%
  dplyr::select(all_of(impact_vars), all_of(all_deserving_vars)) %>%
  pivot_longer(
    cols = all_of(all_deserving_vars),
    names_to = "target",
    values_to = "deservingness"
  ) %>%
  pivot_longer(
    cols = all_of(impact_vars),
    names_to = "impact_group",
    values_to = "in_group"
  ) %>%
  mutate(
    target_label = pretty_target(target),
    impact_group = recode(impact_group, !!!impact_labels)
  )
# -----------------------------
# 3. Summarize means/sd/n
#    among subgroup == 1
# -----------------------------

summary_df <- long_df_heatmap %>%
  filter(in_group == 1, !is.na(deservingness)) %>%
  group_by(impact_group, target, target_label) %>%
  summarise(
    n = n(),
    mean_score = mean(deservingness, na.rm = TRUE),
    sd_score = sd(deservingness, na.rm = TRUE),
    var_score = var(deservingness, na.rm = TRUE),
    .groups = "drop"
  )

# View or export
summary_df
# write.csv(summary_df, "deservingness_by_impact_summary.csv", row.names = FALSE)

# Visual 1: heatmaps of means
ggplot(summary_df, aes(
  x = impact_group,
  y = fct_reorder(target_label, mean_score, .fun = base::mean),
  fill = mean_score
)) +
  geom_tile(color = "white") +
  labs(
    x = NULL,
    y = NULL,
    fill = "Mean\nDeservingness",
    title = "Mean Deservingness by Impact Subgroup Across All Targets"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 7),
    legend.position = "right"
  )
```

4-8 looking at means for all by impact

```{r}
overall_target_means <- sample %>%
  dplyr::select(all_of(all_deserving_vars)) %>%
  pivot_longer(
    cols = everything(),
    names_to = "target",
    values_to = "deservingness"
  ) %>%
  mutate(target_label = pretty_target(target)) %>%
  group_by(target, target_label) %>%
  summarise(
    n = sum(!is.na(deservingness)),
    mean_score = mean(deservingness, na.rm = TRUE),
    sd_score = sd(deservingness, na.rm = TRUE),
    var_score = var(deservingness, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(mean_score))

overall_target_means
View(overall_target_means)

#lookup table
target_means_wide <- summary_df %>%
  dplyr::select(target, target_label, impact_group, mean_score) %>%
  pivot_wider(
    names_from = impact_group,
    values_from = mean_score
  )

View(target_means_wide)
write.csv(target_means_wide, "target_means_by_impact_group_wide.csv", row.names = FALSE)

write.csv(overall_target_means, "overall_target_means.csv", row.names = FALSE)
```

Difference from overall heat map

```{r}
# -----------------------------
# 4. Overall means for each target
# -----------------------------

overall_means <- long_df_heatmap %>%
  filter(!is.na(deservingness)) %>%
  group_by(target, target_label) %>%
  summarise(
    overall_mean_score = mean(deservingness, na.rm = TRUE),
    .groups = "drop"
  )

# -----------------------------
# 5. Merge with subgroup summary and calculate differences
# -----------------------------

diff_df <- summary_df %>%
  left_join(overall_means, by = c("target", "target_label")) %>%
  mutate(
    diff_from_overall = mean_score - overall_mean_score
  )

# -----------------------------
# 6. Order targets by average subgroup deviation
# -----------------------------

diff_df <- diff_df %>%
  group_by(target_label) %>%
  mutate(avg_diff = mean(diff_from_overall, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(
    target_label = forcats::fct_reorder(target_label, avg_diff)
  )

# -----------------------------
# 7. Plot difference-from-overall heatmap
# -----------------------------

ggplot(diff_df, aes(
  x = impact_group,
  y = target_label,
  fill = diff_from_overall
)) +
  geom_tile(color = "white") +
  labs(
    x = NULL,
    y = NULL,
    fill = "Difference\nfrom overall mean",
    title = "How Each Impact Subgroup Differs from the Full Sample"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 7),
    legend.position = "right"
  )

heatmap_plot <- ggplot(diff_df, aes(
  x = impact_group,
  y = target_label,
  fill = diff_from_overall
)) +
  geom_tile(color = "white") +
  labs(
    x = NULL,
    y = NULL,
    fill = "Difference\nfrom overall mean",
    title = "How Each Impact Subgroup Differs from the Full Sample"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 6),
    axis.text.x = element_text(size = 10),
    plot.title = element_text(size = 12, face = "bold"),
    legend.position = "right",
    plot.margin = margin(10, 10, 10, 10)
  )

heatmap_plot

ggsave(
  filename = "deservingness_heatmap_diff.png",
  plot = heatmap_plot,
  width = 8,
  height = 24,
  dpi = 300,
  bg = "white"
)

ggsave(
  filename = "deservingness_heatmap_diff.png",
  plot = heatmap_plot,
  width = 9,
  height = 30,
  dpi = 300,
  bg = "white"
)
```

### 4.2 Incarc Des Means

-means -density plots

```{r}
#DVs
source("variable_labels.R")

#separate object from above for just incarcerated des groups

long_df_table <- sample %>%
  dplyr::select(all_of(impact_vars), all_of(deserving_vars)) %>%
  pivot_longer(
    cols = all_of(deserving_vars),
    names_to = "target",
    values_to = "deservingness"
  )

get_diff_table <- function(df, group_var) {
  df %>%
    dplyr::filter(!is.na(deservingness)) %>%
    dplyr::group_by(target, !!rlang::sym(group_var)) %>%
    dplyr::summarise(mean = mean(deservingness, na.rm = TRUE), .groups = "drop") %>%
    tidyr::pivot_wider(names_from = !!rlang::sym(group_var), values_from = mean) %>%
    dplyr::mutate(
      baseline = `0`,
      diff = `1` - `0`,
      pct_diff = 100 * (diff / `0`)
    ) %>%
    dplyr::select(target, baseline, pct_diff)
}

names(long_df)

impact_tab <- get_diff_table(long_df_table, "impact")
incarc_tab <- get_diff_table(long_df_table, "incarc")
know_tab   <- get_diff_table(long_df_table, "knowincarc")
vict_tab   <- get_diff_table(long_df_table, "vict")

get_pvals_impact <- function(df) {
  df %>%
    dplyr::filter(!is.na(deservingness), !is.na(impact)) %>%
    dplyr::group_by(target) %>%
    dplyr::summarise(
      pval = t.test(deservingness ~ impact)$p.value,
      .groups = "drop"
    )
}

add_stars <- function(p) {
  dplyr::case_when(
    p < 0.001 ~ "***",
    p < 0.01  ~ "**",
    p < 0.05  ~ "*",
    TRUE ~ ""
  )
}

names(long_df_table)

impact_pvals <- get_pvals_impact(long_df_table)

impact_tab <- impact_tab %>%
  dplyr::left_join(impact_pvals, by = "target") %>%
  dplyr::mutate(
    pct_diff = paste0(round(pct_diff, 1), add_stars(pval))
  ) %>%
  dplyr::select(-pval)

density_compare <- function(df, subgroup_var, label, xvar) {

  plot_df <- df %>%
    dplyr::transmute(
      d = .data[[xvar]],
      group = dplyr::case_when(
        impact == 0 ~ "Not impacted",
        .data[[subgroup_var]] == 1 ~ label,
        TRUE ~ NA_character_
      )
    ) %>%
    dplyr::filter(!is.na(group), !is.na(d))

  means_df <- plot_df %>%
    dplyr::group_by(group) %>%
    dplyr::summarise(mean_val = mean(d, na.rm = TRUE), .groups = "drop")

  ggplot2::ggplot(plot_df, ggplot2::aes(x = d, color = group)) +
    ggplot2::geom_density(linewidth = 1.2) +
    ggplot2::geom_vline(
      data = means_df,
      ggplot2::aes(xintercept = mean_val, color = group),
      linetype = "dashed"
    ) +
    ggplot2::labs(x = "Deservingness", y = "Density", color = NULL) +
    ggplot2::theme_minimal()
}

## TABLES

impact_tab <- impact_tab %>%
  dplyr::rename(pct_impact = pct_diff)

incarc_tab <- incarc_tab %>%
  dplyr::rename(pct_incarc = pct_diff)

know_tab <- know_tab %>%
  dplyr::rename(pct_knowincarc = pct_diff)

vict_tab <- vict_tab %>%
  dplyr::rename(pct_vict = pct_diff)

table_out <- impact_tab %>%
  dplyr::select(target, baseline, pct_impact) %>%
  dplyr::left_join(
    incarc_tab %>% dplyr::select(target, pct_incarc),
    by = "target"
  ) %>%
  dplyr::left_join(
    know_tab %>% dplyr::select(target, pct_knowincarc),
    by = "target"
  ) %>%
  dplyr::left_join(
    vict_tab %>% dplyr::select(target, pct_vict),
    by = "target"
  )

# readable row labels from variable_labels.R
table_out <- table_out %>%
  dplyr::mutate(
    target = unname(targets[target])
  )

table_out <- table_out %>%
  dplyr::mutate(
    baseline = round(baseline, 1),
    pct_incarc = round(pct_incarc, 1),
    pct_knowincarc = round(pct_knowincarc, 1),
    pct_vict = round(pct_vict, 1)
  )

table_out_print <- table_out %>%
  dplyr::mutate(
    pct_impact = paste0(pct_impact, "\\%"),
    pct_incarc = paste0(pct_incarc, "\\%"),
    pct_knowincarc = paste0(pct_knowincarc, "\\%"),
    pct_vict = paste0(pct_vict, "\\%")
  )

latex_tab <- knitr::kable(
  table_out_print,
  format = "latex",
  booktabs = TRUE,
  escape = FALSE,
  col.names = c(
    "Group",
    "Not impacted mean",
    "\\% change: Impacted",
    "\\% change: Incarcerated",
    "\\% change: Knows incarcerated",
    "\\% change: Victim"
  ),
  align = c("l", "c", "c", "c", "c", "c")
)

cat(latex_tab)
```

#### into plots

```{r}

# -----------------------------
# 2. helper to build one comparison
# -----------------------------
make_plot_df <- function(df, xvar, subgroup_var, subgroup_label, targets) {
  df %>%
    transmute(
      deserving = .data[[xvar]],
      group = case_when(
        impact == 0 ~ "Not impacted",
        .data[[subgroup_var]] == 1 ~ subgroup_label,
        TRUE ~ NA_character_
      )
    ) %>%
    filter(!is.na(deserving), !is.na(group)) %>%
    mutate(
      target = targets,
      comparison = subgroup_label
    )
}

# -----------------------------
# 3. build full plotting data
# -----------------------------
plot_df <- bind_rows(
  make_plot_df(sample, "d_incppl_pris_abc", "impact", "Impacted", "Incarcerated people"),
  make_plot_df(sample, "d_incppl_pris_abc", "incarc", "Ever incarcerated", "Incarcerated people"),
  make_plot_df(sample, "d_incppl_pris_abc", "knowincarc", "Knows incarcerated", "Incarcerated people"),
  make_plot_df(sample, "d_incppl_pris_abc", "vict", "Victim", "Incarcerated people"),

  make_plot_df(sample, "d_crim_abc", "impact", "Impacted", "Criminals"),
  make_plot_df(sample, "d_crim_abc", "incarc", "Ever incarcerated", "Criminals"),
  make_plot_df(sample, "d_crim_abc", "knowincarc", "Knows incarcerated", "Criminals"),
  make_plot_df(sample, "d_crim_abc", "vict", "Victim", "Criminals"),

  make_plot_df(sample, "d_incwomen_abc", "impact", "Impacted", "Incarcerated women"),
  make_plot_df(sample, "d_incwomen_abc", "incarc", "Ever incarcerated", "Incarcerated women"),
  make_plot_df(sample, "d_incwomen_abc", "knowincarc", "Knows incarcerated", "Incarcerated women"),
  make_plot_df(sample, "d_incwomen_abc", "vict", "Victim", "Incarcerated women")
)

# order facets nicely
plot_df <- plot_df %>%
  mutate(
    target = factor(
      target,
      levels = c("Incarcerated people", "Criminals", "Incarcerated women")
    ),
    comparison = factor(
      comparison,
      levels = c("Impacted", "Ever incarcerated", "Knows incarcerated", "Victim")
    ),
    group = factor(
      group,
      levels = c("Not impacted", "Impacted", "Ever incarcerated", "Knows incarcerated", "Victim")
    )
  )

# -----------------------------
# 4. means for dashed lines
# -----------------------------
means_df <- plot_df %>%
  group_by(target, comparison, group) %>%
  summarise(mean_val = mean(deserving, na.rm = TRUE), .groups = "drop")

# -----------------------------
# 5. plot
# -----------------------------
p <- ggplot(plot_df, aes(x = deserving, color = group)) +
  geom_density(linewidth = 1) +
  geom_vline(
    data = means_df,
    aes(xintercept = mean_val, color = group),
    linetype = "dashed",
    linewidth = 0.8,
    show.legend = FALSE
  ) +
  facet_grid(target ~ comparison) +
  coord_cartesian(xlim = c(0, 100)) +
  labs(
    x = "Deservingness score",
    y = "Density",
    color = NULL
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    strip.text = element_text(face = "bold")
  )



ggsave(
  "deservingness_3x4_density.png",
  plot = p,
  width = 14,
  height = 9,
  dpi = 300,
  bg = "white"
)

ggsave(
  "deservingness_3x4_density.pdf",
  plot = p,
  width = 14,
  height = 9
)
```

4-5 the tables aren't working. I just am pivoting to making sure I can print out all the info I need to fix the writing in my analysis.

```{r}
summary(policy_vs_nonimp_nocontrol_models$comp_prisonhelp_incarc)
summary(policy_vs_nonimp_demo_models$comp_prisonhelp_incarc)
summary(policy_vs_nonimp_partyideo_models$comp_prisonhelp_incarc)
summary(policy_vs_nonimp_full_models$comp_prisonhelp_incarc)


summary(policy_vs_nonimp_nocontrol_models$comp_prisonhelp_i)
summary(policy_vs_nonimp_demo_models$comp_prisonhelp_incarc)
summary(policy_vs_nonimp_partyideo_models$comp_prisonhelp_incarc)
summary(policy_vs_nonimp_full_models$comp_prisonhelp_incarc)

```

```{r}
# ============================================================
# EXPORT ALL MODEL SUMMARIES AND ROBUST COEFFICIENT TABLES
# ============================================================

library(lmtest)
library(sandwich)

# -----------------------------
# 1. Collect all model lists
# -----------------------------
all_model_lists <- list(
  goal1_nocontrol   = policy_dummies_models,
  goal1_demo        = policy_dummies_demo_models,
  goal1_partyideo   = policy_dummies_partyideo_models,
  goal1_full        = policy_dummies_full_models,
  goal2_nocontrol   = policy_vs_nonimp_nocontrol_models,
  goal2_demo        = policy_vs_nonimp_demo_models,
  goal2_partyideo   = policy_vs_nonimp_partyideo_models,
  goal2_full        = policy_vs_nonimp_full_models
)

# -----------------------------
# 2. Helper: write one model list
# -----------------------------
write_model_list_outputs <- function(model_list, list_name, out_dir = "model_outputs") {

  if (!dir.exists(out_dir)) dir.create(out_dir, recursive = TRUE)

  summary_file <- file.path(out_dir, paste0(list_name, "_summary_output.txt"))
  robust_file  <- file.path(out_dir, paste0(list_name, "_robust_output.txt"))

  # -------- summary() output --------
  con_sum <- file(summary_file, open = "wt")
  for (model_name in names(model_list)) {
    mod <- model_list[[model_name]]

    writeLines(paste0(
      "\n",
      strrep("=", 90), "\n",
      "MODEL: ", model_name, "\n",
      strrep("=", 90), "\n"
    ), con_sum)

    capture.output(summary(mod), file = con_sum, append = TRUE)
    writeLines("\n\n", con_sum)
  }
  close(con_sum)

  # -------- robust coefficient output --------
  con_rob <- file(robust_file, open = "wt")
  for (model_name in names(model_list)) {
    mod <- model_list[[model_name]]

    writeLines(paste0(
      "\n",
      strrep("=", 90), "\n",
      "MODEL: ", model_name, "\n",
      strrep("=", 90), "\n"
    ), con_rob)

    robust_tab <- lmtest::coeftest(mod, vcov. = sandwich::vcovHC(mod, type = "HC1"))
    capture.output(print(robust_tab), file = con_rob, append = TRUE)

    writeLines("\n--- Model fit ---\n", con_rob)
    fit_info <- c(
      paste("N   =", nobs(mod)),
      paste("R2  =", round(summary(mod)$r.squared, 4)),
      paste("Adj =", round(summary(mod)$adj.r.squared, 4))
    )
    writeLines(fit_info, con_rob)

    writeLines("\n\n", con_rob)
  }
  close(con_rob)

  invisible(list(
    summary_file = summary_file,
    robust_file = robust_file
  ))
}

# -----------------------------
# 3. Run for all model lists
# -----------------------------
output_files <- lapply(names(all_model_lists), function(list_name) {
  write_model_list_outputs(
    model_list = all_model_lists[[list_name]],
    list_name = list_name,
    out_dir = "model_outputs"
  )
})

names(output_files) <- names(all_model_lists)

# -----------------------------
# 4. Optional: make one master file
# -----------------------------
master_summary_file <- file.path("model_outputs", "ALL_MODEL_SUMMARIES.txt")
master_robust_file  <- file.path("model_outputs", "ALL_MODEL_ROBUST.txt")

con_master_sum <- file(master_summary_file, open = "wt")
for (list_name in names(all_model_lists)) {
  writeLines(paste0(
    "\n\n",
    strrep("#", 100), "\n",
    "MODEL LIST: ", list_name, "\n",
    strrep("#", 100), "\n"
  ), con_master_sum)

  for (model_name in names(all_model_lists[[list_name]])) {
    mod <- all_model_lists[[list_name]][[model_name]]

    writeLines(paste0(
      "\n",
      strrep("=", 90), "\n",
      "MODEL: ", model_name, "\n",
      strrep("=", 90), "\n"
    ), con_master_sum)

    capture.output(summary(mod), file = con_master_sum, append = TRUE)
    writeLines("\n\n", con_master_sum)
  }
}
close(con_master_sum)

con_master_rob <- file(master_robust_file, open = "wt")
for (list_name in names(all_model_lists)) {
  writeLines(paste0(
    "\n\n",
    strrep("#", 100), "\n",
    "MODEL LIST: ", list_name, "\n",
    strrep("#", 100), "\n"
  ), con_master_rob)

  for (model_name in names(all_model_lists[[list_name]])) {
    mod <- all_model_lists[[list_name]][[model_name]]

    writeLines(paste0(
      "\n",
      strrep("=", 90), "\n",
      "MODEL: ", model_name, "\n",
      strrep("=", 90), "\n"
    ), con_master_rob)

    robust_tab <- lmtest::coeftest(mod, vcov. = sandwich::vcovHC(mod, type = "HC1"))
    capture.output(print(robust_tab), file = con_master_rob, append = TRUE)

    writeLines("\n--- Model fit ---\n", con_master_rob)
    fit_info <- c(
      paste("N   =", nobs(mod)),
      paste("R2  =", round(summary(mod)$r.squared, 4)),
      paste("Adj =", round(summary(mod)$adj.r.squared, 4))
    )
    writeLines(fit_info, con_master_rob)

    writeLines("\n\n", con_master_rob)
  }
}
close(con_master_rob)

# -----------------------------
# 5. Confirm output locations
# -----------------------------
print(output_files)
cat("\nMaster files written to:\n",
    master_summary_file, "\n",
    master_robust_file, "\n")
```

## deservingness means by target by impact

```{r}
library(dplyr)
library(tidyr)

# -----------------------------
# 1. Define the three targets
# -----------------------------
deserving_vars <- c("d_incppl_pris_abc", "d_crim_abc", "d_incwomen_abc")

targets <- c(
  d_incppl_pris_abc = "Incarcerated people",
  d_crim_abc = "Criminals",
  d_incwomen_abc = "Incarcerated women"
)

# -----------------------------
# 2. Create simple group flags
# -----------------------------
df <- sample %>%
  mutate(
    not_impacted = if_else(impact == 0, 1, 0, missing = 0),
    impacted_grp = if_else(impact == 1, 1, 0, missing = 0),
    incarcerated_grp = if_else(incarc == 1, 1, 0, missing = 0),
    knowincarc_grp = if_else(knowincarc == 1, 1, 0, missing = 0),
    victim_grp = if_else(vict == 1, 1, 0, missing = 0)
  )

# -----------------------------
# 3. Put the data in long format
# -----------------------------
long_df <- df %>%
  dplyr::select(
    not_impacted, impacted_grp, incarcerated_grp, knowincarc_grp, victim_grp,
    all_of(deserving_vars)
  ) %>%
  pivot_longer(
    cols = all_of(deserving_vars),
    names_to = "target",
    values_to = "score"
  ) %>%
  mutate(
    target = recode(target, !!!targets)
  )

# -----------------------------
# 4. Compute means for each group x target
# -----------------------------
group_means <- bind_rows(
  long_df %>%
    filter(not_impacted == 1) %>%
    group_by(target) %>%
    summarise(mean_score = mean(score, na.rm = TRUE), .groups = "drop") %>%
    mutate(group = "Not impacted"),

  long_df %>%
    filter(impacted_grp == 1) %>%
    group_by(target) %>%
    summarise(mean_score = mean(score, na.rm = TRUE), .groups = "drop") %>%
    mutate(group = "Impacted"),

  long_df %>%
    filter(incarcerated_grp == 1) %>%
    group_by(target) %>%
    summarise(mean_score = mean(score, na.rm = TRUE), .groups = "drop") %>%
    mutate(group = "Incarcerated"),

  long_df %>%
    filter(knowincarc_grp == 1) %>%
    group_by(target) %>%
    summarise(mean_score = mean(score, na.rm = TRUE), .groups = "drop") %>%
    mutate(group = "Knows incarcerated"),

  long_df %>%
    filter(victim_grp == 1) %>%
    group_by(target) %>%
    summarise(mean_score = mean(score, na.rm = TRUE), .groups = "drop") %>%
    mutate(group = "Victim")
) %>%
  dplyr::select(group, target, mean_score)

# View the long table
group_means
group_means_wide <- group_means %>%
  pivot_wider(
    names_from = target,
    values_from = mean_score
  )

group_means_wide

group_differences <- group_means_wide %>%
  mutate(
    `Incarcerated people - Criminals` =
      `Incarcerated people` - `Criminals`,

    `Incarcerated women - Criminals` =
      `Incarcerated women` - `Criminals`,

    `Incarcerated women - Incarcerated people` =
      `Incarcerated women` - `Incarcerated people`
  )

group_differences

final_deservingness_table <- group_differences %>%
  mutate(across(where(is.numeric), ~ round(.x, 2)))

View(final_deservingness_table)
```

to overleaf

```{r}

kable(
  final_deservingness_table,
  format = "latex",
  booktabs = TRUE,
  digits = 2,
  align = c("l", rep("r", ncol(final_deservingness_table) - 1)),
  caption = "Mean deservingness scores and within-group target differences"
)

library(knitr)

kable(
  final_deservingness_table,
  format = "latex",
  booktabs = TRUE,
  digits = 2,
  align = c("l", rep("r", ncol(final_deservingness_table) - 1)),
  caption = "Mean deservingness scores and within-group target differences"
) %>%
  cat()
```

## deservingness straight up models with controls

```{r}

m_des_inc_impact <- lm(d_incppl_pris_abc ~ impact + Black + Hisp_Latino + OtherRace + gender + Republican + Democrat + fiscal_ideology + social_ideology + desor_core, data = sample)

m_des_inc_incarc <- lm(d_incppl_pris_abc ~ incarc + Black + Hisp_Latino + OtherRace + gender + Republican + Democrat + fiscal_ideology + social_ideology + desor_core, data = sample)

m_des_inc_knowincarc <- lm(d_incppl_pris_abc ~ knowincarc + Black + Hisp_Latino + OtherRace + gender + Republican + Democrat + fiscal_ideology + social_ideology + desor_core, data = sample)

m_des_inc_vict <- lm(d_incppl_pris_abc ~ vict + Black + Hisp_Latino + OtherRace + gender + Republican + Democrat + fiscal_ideology + social_ideology + desor_core, data = sample)

##

m_des_crim_impact <- lm(d_crim_abc ~ impact + Black + Hisp_Latino + OtherRace + gender + Republican + Democrat + fiscal_ideology + social_ideology + desor_core, data = sample)

m_des_crim_incarc <- lm(d_crim_abc ~ incarc + Black + Hisp_Latino + OtherRace + gender + Republican + Democrat + fiscal_ideology + social_ideology + desor_core, data = sample)

m_des_crim_knowincarc <- lm(d_crim_abc ~ knowincarc + Black + Hisp_Latino + OtherRace + gender + Republican + Democrat + fiscal_ideology + social_ideology + desor_core, data = sample)

m_des_crim_vict <- lm(d_crim_abc ~ vict + Black + Hisp_Latino + OtherRace + gender + Republican + Democrat + fiscal_ideology + social_ideology + desor_core, data = sample)

##

m_des_incwomen_imp <- lm(d_incwomen_abc ~ impact + Black + Hisp_Latino + OtherRace + gender + Republican + Democrat + fiscal_ideology + social_ideology + desor_core, data = sample)

m_des_incwomen_incarc <- lm(d_incwomen_abc ~ incarc + Black + Hisp_Latino + OtherRace + gender + Republican + Democrat + fiscal_ideology + social_ideology + desor_core, data = sample)

m_des_incwomen_knowinc <- lm(d_incwomen_abc ~ knowincarc + Black + Hisp_Latino + OtherRace + gender + Republican + Democrat + fiscal_ideology + social_ideology + desor_core, data = sample)

m_des_incwomen_vict <- lm(d_incwomen_abc ~ vict + Black + Hisp_Latino + OtherRace + gender + Republican + Democrat + fiscal_ideology + social_ideology + desor_core, data = sample)


##


summary(m_des_inc_impact)
summary(m_des_inc_incarc)
summary(m_des_inc_knowincarc)
summary(m_des_inc_vict)

summary(m_des_crim_impact)
summary(m_des_crim_incarc)
summary(m_des_crim_knowincarc)
summary(m_des_crim_vict)

summary(m_des_incwomen_imp)
summary(m_des_incwomen_incarc)
summary(m_des_incwomen_knowinc)
summary(m_des_incwomen_vict)
```

## Des Restricted Sample

```{r}
library(sandwich)
library(lmtest)
library(broom)
##
# model generation code
##
# DVs


deserving_vars <- c("d_incppl_pris_abc", "d_crim_abc", "d_incwomen_abc")
impact_vars <- c("impact", "incarc", "knowincarc", "vict")

model_specs <- list(
  nocontrols = NULL,
  demo = demo_controls,
  full = c(demo_controls, all_attitudinal_controls)
)

get_restricted_data <- function(data, iv) {
  if (iv == "impact") {
    data %>% filter(!is.na(impact))
  } else {
    data %>% filter((.data[[iv]] == 1) | (impact == 0))
  }
}

restricted_model_list <- list()
results_df_restricted <- list()

for (dv in deserving_vars) {
  for (spec_name in names(model_specs)) {
    controls <- model_specs[[spec_name]]
    
    for (iv in impact_vars) {
      dat_iv <- get_restricted_data(sample, iv)
      vars_needed <- c(dv, iv, "impact", controls)
      dat_iv <- dat_iv %>% filter(complete.cases(across(all_of(vars_needed))))
      
      form <- if (is.null(controls)) {
        as.formula(paste(dv, "~", iv))
      } else {
        as.formula(paste(dv, "~", iv, "+", paste(controls, collapse = " + ")))
      }
      
      model <- lm(form, data = dat_iv)
      robust <- coeftest(model, vcov = vcovHC(model, type = "HC1"))
      
      model_name <- paste(dv, spec_name, iv, sep = "_")
      restricted_model_list[[model_name]] <- list(
        model = model,
        robust = robust,
        data = dat_iv
      )
      
      d0 <- dat_iv
      d1 <- dat_iv
      d0[[iv]] <- 0
      d1[[iv]] <- 1
      
      baseline <- mean(predict(model, newdata = d0), na.rm = TRUE)
      predicted <- mean(predict(model, newdata = d1), na.rm = TRUE)
      coef_iv <- predicted - baseline
      pct_change <- ((predicted - baseline) / baseline) * 100
      
      results_df_restricted[[model_name]] <- data.frame(
        baseline = baseline,
        coef = coef_iv,
        predicted = predicted,
        pct_change = pct_change,
        se = robust[iv, "Std. Error"],
        p = robust[iv, "Pr(>|t|)"],
        dv = dv,
        spec = spec_name,
        iv = iv,
        n = nrow(dat_iv),
        r2 = summary(model)$r.squared,
        adj_r2 = summary(model)$adj.r.squared
      )
    }
  }
}


results_df_restricted <- do.call(rbind, results_df_restricted)
View(results_df_restricted)

restricted_model_list[["d_incppl_pris_abc_demo_incarc"]]
coef(restricted_model_list[["d_incppl_pris_abc_demo_incarc"]]$model)
restricted_model_list[["d_incppl_pris_abc_demo_incarc"]]$robust
restricted_model_list[["d_incppl_pris_abc_demo_incarc"]]$robust


get_model_table <- function(model_entry) {
  robust_mat <- model_entry$robust
  
  data.frame(
    term = rownames(robust_mat),
    estimate = robust_mat[, "Estimate"],
    se = robust_mat[, "Std. Error"],
    p = robust_mat[, "Pr(>|t|)"],
    stringsAsFactors = FALSE,
    row.names = NULL
  )
}


#incarc, demo controls
print(get_model_table(restricted_model_list[["d_incppl_pris_abc_demo_impact"]]))
print(get_model_table(restricted_model_list[["d_incppl_pris_abc_demo_incarc"]]))
print(get_model_table(restricted_model_list[["d_incppl_pris_abc_demo_knowincarc"]]))
print(get_model_table(restricted_model_list[["d_incppl_pris_abc_demo_vict"]]))

#crim, demo controls
print(get_model_table(restricted_model_list[["d_crim_abc_demo_impact"]]))
print(get_model_table(restricted_model_list[["d_crim_abc_demo_incarc"]]))
print(get_model_table(restricted_model_list[["d_crim_abc_demo_knowincarc"]]))
print(get_model_table(restricted_model_list[["d_crim_abc_demo_vict"]]))

#inc women, demo controls
print(get_model_table(restricted_model_list[["d_incwomen_abc_demo_impact"]]))
print(get_model_table(restricted_model_list[["d_incwomen_abc_demo_incarc"]]))
print(get_model_table(restricted_model_list[["d_incwomen_abc_demo_knowincarc"]]))
print(get_model_table(restricted_model_list[["d_incwomen_abc_demo_vict"]]))


#incarc, demo controls
print(get_model_table(restricted_model_list[["d_incppl_pris_abc_full_impact"]]))
print(get_model_table(restricted_model_list[["d_incppl_pris_abc_full_incarc"]]))
print(get_model_table(restricted_model_list[["d_incppl_pris_abc_full_knowincarc"]]))
print(get_model_table(restricted_model_list[["d_incppl_pris_abc_full_vict"]]))

#crim, demo controls
print(get_model_table(restricted_model_list[["d_crim_abc_full_impact"]]))
print(get_model_table(restricted_model_list[["d_crim_abc_full_incarc"]]))
print(get_model_table(restricted_model_list[["d_crim_abc_full_knowincarc"]]))
print(get_model_table(restricted_model_list[["d_crim_abc_full_vict"]]))

#inc women, demo controls
print(get_model_table(restricted_model_list[["d_incwomen_abc_full_impact"]]))
print(get_model_table(restricted_model_list[["d_incwomen_abc_full_incarc"]]))
print(get_model_table(restricted_model_list[["d_incwomen_abc_full_knowincarc"]]))
print(get_model_table(restricted_model_list[["d_incwomen_abc_full_vict"]]))




summary(d_incppl_pris_abc_demo_impact)
d_incppl_pris_abc_demo_incarc 
d_incppl_pris_abc_demo_knowincarc
d_incppl_pris_abc_demo_vict

```

Interpretation: The baseline is the average predicted deservingness score for the comparison group within each restricted sample when the focal IV is set to 0. The comparison group is people with no experience (impact = 0). So the baseline now means: what is the expected deservingness score among non-impacted respondents (within the relevant sample and model specification). The baseline is NOT the regression intercept.

-Predicted: the average predicted score if everyone in that same sample had the focal experience (ie incarc = 1). The coef is the difference between those two quantities, and pct_change is the percent increase relative to the non-impacted baseline. Each row then is answering the questionL how much higher or lower is the predicted deservingness score for people with this type of contact compared to people with no contact.

Substantively, the patterns are quite strong and very consistent across DVs. For incarcerated people, the baseline sits around 43–44 across specifications, and all forms of contact increase deservingness relative to that group. Direct incarceration produces the largest shifts (roughly 22–27% depending on controls), proximal contact is somewhat smaller, and victimization is consistently the smallest and sometimes not statistically significant once you include full controls. That ordering is doing real theoretical work for you: it suggests that experiential proximity to incarceration increases generosity in a graded way, but that victimization is a qualitatively different kind of experience rather than just another form of “contact.”

You see the same structure for “criminals,” but with larger effect sizes. The baseline is lower (around 29–30), and the increases for direct incarceration are much larger (often \~29–33% even with controls). That’s important because “criminals” are the most negatively constructed target in your set, so the fact that contact moves those evaluations substantially is strong evidence that experience is reshaping underlying deservingness judgments, not just marginal preferences. This lines up cleanly with your broader argument about contact altering how people interpret stigmatized groups rather than simply shifting policy opinions at the margin.

For incarcerated women, the baseline is higher (around 50), and the effects are more compressed. You still see positive differences for all forms of contact, but they are smaller and attenuate more with controls. That pattern is also theoretically coherent: respondents already view this group as more deserving, so there is less room for contact to shift evaluations upward. In other words, contact effects are conditional on the prior social construction of the group, which is exactly the kind of mechanism your framework is trying to get at.

The role of controls is also now interpretable in a clean way. Moving from no controls to demo to full controls generally attenuates the percent differences, but does not eliminate them—especially for direct and proximal incarceration. That gives you a defensible claim that these differences are not simply artifacts of race, gender, ideology, or racial attitudes. In your language, it supports the idea that carceral contact is associated with distinct political orientations, rather than just sorting into pre-existing ideological camps. This connects directly to the literature you’re drawing on about how policy experience shapes political views and engagement .

So the bottom line is that this table now does what you need it to do. The baseline is a meaningful “no experience” reference point, the percent changes are interpretable group differences, and the patterns across DVs and specifications are theoretically informative rather than artifacts of model construction.

1.  

2.  

3.  

4.  

5.  

6.  

7.  

8.  

9.  


