---
title: "analysis"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,       # Hide code by default
  warning = FALSE,
  message = FALSE
)

# Load packages
library(dplyr)
library(tidyr)
library(readr)
library(janitor)   # for clean_names()
library(tibble)
library(stringr)
library(haven)
library(tibble)
library(gt)
library(marginaleffects)
library(mediation)
library(naniar)
library(rsq)

```

```{r}
here::here()
raw_data_cleaned <- readr::read_csv("data/cleaned/coded_data_may9.csv")

#view(raw_data_cleaned)
```

## Preliminary Analyses

| Decision                | Notes                                        |
|-------------------------|----------------------------------------------|
| which version of table? | prisonerdeserves (block aware and cleaned)   |
| NAs?                    | drop silently for plot but log               |
| scale?                  | raw 0-100 for now                            |
| compare subgroups?      | for now, overall distribution of just incppl |
| summary overlays?       | mean line, kernal density, histogram bars    |
| plot aesthetics?        | clear x-axis label                           |
| Output                  |                                              |

### 1.1. Main DV: Deservingness of Incarcerated People/Prisoners

About

-   mean: 46.13

-   median: 50

-   SD: 27.96

-   variance: 782.03

```{r, prechecks}
#check counts: 

#raw_data_cleaned %>% count(d_incppl_pris_abc)

#summary:
#summary(raw_data_cleaned$d_incppl_pris_abc)



#histogram
incarc_ppl_hist <- hist(raw_data_cleaned$d_incppl_pris_abc, breaks = 20,
     main = "Deservingness of Incarcerated People/Prisoners",
     xlab = "Deservingness (0–100)", col = "gray80", border = "white")


#mean(raw_data_cleaned$d_incppl_pris_abc, na.rm = TRUE)

#sd(raw_data_cleaned$d_incppl_pris_abc, na.rm = TRUE)

#var(raw_data_cleaned$d_incppl_pris_abc, na.rm = TRUE)

boxplot(raw_data_cleaned$d_incppl_pris_abc)

#density plot
dv_density_data <- raw_data_cleaned %>%
  dplyr::select(d_incppl_pris_abc) %>%
  filter(!is.na(d_incppl_pris_abc))

```

### 1.2. Control: Generalized Deservingness (desor)

```{r}

# Histogram for generalized deservingness control variable
desor_core_hist <- hist(raw_data_cleaned$desor_core, breaks = 20,
     main = "'Needy' Deservingness Orientation",
     xlab = "Average Deservingness of Help-Target Groups (0–100)",
     col = "gray80", border = "white")

summary(raw_data_cleaned$desor_core)
sd(raw_data_cleaned$desor_core, na.rm = TRUE)

```

### 1.3. Generalized Deservingness (desor) x

### 1.4 Cross-Tabs of Contact Types

```{r}
library(dplyr)
library(gt)
library(stringr)

#basic
contact_crosstab <- raw_data_cleaned %>%
  summarise(
    incarc_yes = sum(incarc == 1, na.rm=TRUE),
    knowincarc_yes = sum(knowincarc == 1, na.rm=TRUE),
    vict_yes = sum(vict == 1, na.rm=TRUE)
  )


# Optional: define custom labels for key combinations
pattern_labels <- tibble::tibble(
  contact_pattern = c(
    "001001", "000000", "100000", "010000", "100100",
    "010100", "110110", "110100", "011010", "001100"
  ),
  label = c(
    "No or unsure only",
    "Skipped all Q4 items",
    "Been incarcerated only",
    "Knows someone incarcerated only",
    "Been incarcerated & victimized",
    "Knows incarcerated & victimized",
    "All four: incarcerated, knows, victim, knows victim",
    "Incarcerated, knows, victim",
    "Knows incarcerated & knows victim",
    "Victimized only"
  )
)

# Join with summary table
combo_summary_labeled <- combo_summary %>%
  left_join(pattern_labels, by = "contact_pattern") %>%
  mutate(label = ifelse(is.na(label), "Other", label))
```

#### A. All Contact Combos

pull from 2B

```{r}

```

## 2. M1: Predicting deservingness

#### 2.1 Without Controls

```{r}
m1 <- lm(d_incppl_pris_abc ~ impact, data = raw_data_cleaned)
summary(m1)

```

#### 2.2 M1 Interpretation:\*

-   Q: does any carceral contact or victimization change perceived deservingness of incarcerated people (`d_incppl_pris_abc`, 0–100 scale) compared to those without any such contact (`impact = 0`).

-   Coefficients:

    -   Intercept = 43.31

        -   The average deservingness score of incarcerated people, for respondents with no contact (`impact = 0`), is on average 43.31 out of 100 on the deservingness scale. This is higher than I might have thought.

    -   impact = 7.17

        -   People who have had contact (direct, indirect, or victimization) rate incarcerated people 7.17 points higher on the 0-100 deservingness scale — on average – than those without contact.

        -   Significance: The t-val is 5.881 and p is less than 0.001, suggesting this effect is significantly significant and the probability that it is due to chance is very low.

        -   Model fit: The R-squared is 0.0156, meaning only 1.6% of the variation in deservingness is explained by this binary variable alone.

```{r, m1plot}
library(ggplot2)
# Summarize means and confidence intervals
plot_data <- raw_data_cleaned %>%
  filter(!is.na(impact)) %>%  # <- filter out NA values
  group_by(impact) %>%
  summarize(
    mean_deserving = mean(d_incppl_pris_abc, na.rm = TRUE),
    se = sd(d_incppl_pris_abc, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) %>%
  mutate(
    impact_label = ifelse(impact == 1, "Any Contact", "No Contact"),
    ci_lower = mean_deserving - 1.96 * se,
    ci_upper = mean_deserving + 1.96 * se
  )

# Plot
ggplot(plot_data, aes(x = impact_label, y = mean_deserving)) +
  geom_col(fill = "gray60") +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.2) +
  labs(
    x = NULL,
    y = "Mean Deservingness Rating (0–100)",
    title = "Deservingness of Incarcerated People by Contact Experience"
  ) +
  theme_minimal(base_size = 14)


```

#### 2.3 (M1c) With Controls

```{r}
m1c <- lm(d_incppl_pris_abc ~ impact +
    Republican + Democrat +
    fiscal_ideology + social_ideology +
    penalpoint + desor_core +
    fire_rare + fire_privilege + fire_angry + fire_fear +
    gender + Black + Hispanic + OtherRace,
  data = raw_data_cleaned
)
summary(m1c)
```

#### 2.4 Diagnostics

**Variance inflation factors – multi-colinearity**

```{r}

library(car)
# Run VIFs on  existing model object (replace 'm1' with your model name)
print(vif(m1c))

```

--\> shows everything 1-2, predictors mostly independent, EXCEPT:

-   fiscal_ideology 4.4

-   social_ideology 4.7

-   --\> moderate correlation

**Inspect for heteroskedasticity & nonlinearity**

```{r}

plot(m1c, which = 1)

```

```{r}
# Normal Q-Q plot
plot(m1c, which = 2)
```

```{r}
library(ggplot2)

ggplot(m1c, aes(x = .fitted, y = .resid)) +
  geom_point(alpha = 0.4) +
  geom_smooth(se = FALSE, color = "red") +
  labs(x = "Fitted Values", y = "Residuals",
       title = "Residuals vs Fitted (Check for Nonlinearity/Heteroskedasticity)") +
  theme_minimal()

```

```{r}
library(lmtest)
bptest(m1c)
```

BP Test suggests residuals DO NOT have constant variance – there is evidence of heteroskedasticy. Some subsets of observations have larger/smaller residual spread than others. Coefficient estimates are not biased, but **standard errors are biased.**

--\> I should use robust standard errors

#### 2.5 ROBUST SE's

```{r}

library(lmtest)
library(sandwich)

coeftest(m1c, vcov = vcovHC(m1c, type = "HC1"))
```

***Using standardized betas – show relative explantory strength, not total explained variance***

```{r, m1cs}

m1cs <- lm(d_incppl_pris_abc ~ impact + Republican + Democrat + fiscal_ideology +
         social_ideology + penalpoint + desor_core + fire_rare + fire_privilege +
         fire_angry + fire_fear + gender + Black + Hispanic + OtherRace,
         data = raw_data_cleaned)

# 1. Standardized coefficients
library(broom)
y_sd <- sd(raw_data_cleaned$d_incppl_pris_abc, na.rm = TRUE)

# compute sd for each predictor (binary and continuous)
sds <- sapply(model.frame(m1cs)[-1], function(x) sd(x, na.rm=TRUE))  # excludes response
coefs <- coef(m1cs)
# standardized coef = b * sd(x) / sd(y).  For factor dummies sd(x) is the std dev of 0/1 vector.
std_coefs <- coefs * c(1, sds / y_sd)   # note: first coef is intercept so alignment matters
# better to use broom::tidy:
tidy(m1cs) |> 
  mutate(std_est = estimate * c(0, sds / y_sd)) 
```

#### 2.6 Interpretation

-   N = 1,577 observations used – additional 645 deleted for missingness

-   Coefficients:

    -   Intercept = 2.97 (p = 0.63)

        -   The predicted deservingness score when all predictors are 0, [so not particularly meaningful.]{.underline} Means respondent:
            -   is independent or other
            -   has the most conservative ideology
            -   ...is not impacted
            -   is
            -   (add remaining)

    -   impact = +4.60\* (p \< 0.001)

        -   Being impacted by the CJ system (incarcerated, knows, victimized) increases the predicted deservingness rating by 4.6 points on average out of 100 (4.6%), holding all else constant.

            -   Remains statistically significant, particularly because the outcome range is 0-100

            -   supports H1

    -   Republican = -0.25 (p = 0.90)

        -   Compared to Independents/Others, Republicans rate incarcerated people just 0.39 points higher, and this is not statistically significant.

        -   **No evidence** that Republican identification alone predicts harsher or more generous deservingness scores in this model (when other factors are controlled).

        -   May reflect that **ideology and punitiveness** absorb more of the variance that would otherwise be attributed to party label.

    -   Democrat = -0.72 (p = 0.61)

        -   Compared to Independents/Others, Democrats score incarcerated people 0.72 points lower on deservingness, but this is not statistically significant.

        -   There is **no reliable difference** in prisoner deservingness between Democrats and Independents/Others once we account for ideology, contact, and other controls.

        -   Coupled with the Republican coefficient, this suggests **very weak direct effects of party ID per se**, which may be surprising given elite cues—but again, could reflect ideological and experiential absorption.

    -   fiscal_ideology = +0.13 (p = 0.88)

        -   This is the effect of being more fiscally conservative (scale reversed: higher = more conservative).

        -   Not statistically significant, and the coefficient is small.

        -   **Fiscal ideology appears unrelated** (ONCE SOCIAL IDEOLOGY INCLUDED) to how respondents rate the deservingness of incarcerated people, net of party, social ideology, punitiveness, and racial attitudes.

        -   This aligns with your expectation that deservingness is more influenced by *social/cultural* orientations than by economic/fiscal ideology.

    -   social_ideology = -2.03\* (0.017)

        -   each unit increase in social conservativism predicts a 2.03 point decrease in deservingness of incarcerated people (on a 0-100 scale)

        -   marginally significant (p\~0.014) and in the expected direction

        -   some evidence social conservativism is associated with lower perceived deservingness

        -   effect size is moderate – but this all goes into the pile of "do further testing to see which controls are meaningful

    -   PenalPunitiveness = -1.15 (p \<0.0009)

        -   participant view on whether purpose of incarceration is to punish or rehabilitate.

        -   a one unit increase in belief that incarceration is primarily about punishment is associated with a 1.15-point decrease in the perceived deservingness of incarcerated people.

        -   **substantive significance:** a full-scale shift (from 1 to 7) predicts a 6.8 point drop in deservingness on a 0-100 scale – modest but meaningful, especially in the context of other predictors.

        -   this supports the claim that punitive orientations towards the purpose of incarceration shape how people evaluate the deservingness of targets.

        -   **modeling implication**: it's not entirely endogenous to deservingness ratings, bc it comes from a distinct question about the penal telos. I think this tells us that deservingness ratings don't exclusively reflect beliefs about what prison is for.

            -   However, I'm not sure this variable captures "punitivenness." In part, it bleeds into my questions about affective reactions versus judgement.

    -   desor_core = +0.73 (p \< 0.001)

        -   Interpretation:

            -   For each one-point increase in the average deservingness rating of *help-target groups*, perceived deservingness of incarcerated people increases by 0.73

        -   [**Substantive scale:**]{.underline}

            -   ranges from 0-100, so a 10-pt increase corresponds to a 7.6 point increase in prisoner deservingness – comperable to but in opp direction of penalpun effect

        -   conceptual interpretation:

            -   **so far, strongest predictor in the model**. those who generally see "helping" groups as deserving also extend that view to incarcerated people.

            -   captures individual-level variation in generalized generosity twoard marginalized/help needing populations.

            -   inclusion seems important because it helps isolate target-specific effects, shows that prisoner deservingness is partially embedded in broader affective/moral schemas about who deserves help.

        -   Modeling implication: controlling for desor might let me argue that the effect of contact on prisoner deservingness isn't just a byproduct of being generally generous – it's targeted.

    -   penalpoint = -1.147 (SE 0.345, p \<0.001) – the more someone views as prison's purpose being punishment, rather than rehabilitation, the slightly less deserving (1.15 points lower) they see incarcerated people.

    -   fire_rare = 0.39734 (p = 0.60)

        -   agreement that racial problems are rare and isolated, so more rejection of racial minimization (racial empathy) predicts lightly higher deservingness perceptions of incarcerated people.

        -   takeaway – not statistically significant, so can't rule out the possibility the relationship is due to chance. But direction aligns with theory.

    -   fire_privilege = 0.05353 (p = 0.94)

        -   Meaning: Agreement that white people benefit from unearned racial advantages

        -   Interpretation: The near-zero, nonsignificant coefficient indicates this belief does not meaningfully predict differences in perceived deservingness of incarcerated people in this model.

        -   Note: This item may be less emotionally activated than the others or more subject to social desirability bias, dampening its predictive power

    -   fire_angry = 1.18110 (p = 0.16)

        -   Meaning: Agreement with *“I am angry that racism exists.”*

        -   Interpretation: The positive direction suggests that respondents who are more emotionally angry about racism tend to see incarcerated people as more deserving, consistent with frameworks linking racial empathy to support for equity-enhancing policies.

            -   Note: While not significant, this coefficient is substantively large and may become more stable with a larger sample or better scaling.

    -   fire_fear (0.30112) (P = 0.73)

        -   Meaning: Agreement with *“I am fearful of people of other races.”*

        -   Interpretation: Respondents who express more racial fear are (slightly) more likely to rate incarcerated people as less deserving, though the effect is small and nonsignificant

        -   This aligns with theories linking racial threat or fear to punitive or exclusionary attitudes.

    -   gender = -1.02 (p=0.38 )

        -   men estimated to rate incarcerated people as slightly less deserving, but difference is small and not significant

    -   Black = rate incarcerated people as slightly lower (3.3) than do white respondents, modest but significant group-level difference (0.045)

    -   Hispanic = -3.25 (p= 0.057)

        -   rate incarcerated people 3 points less deserving than non-hispanic whites, holding all else constant

    -   OtherRace = dropped

Residual SE: 22.6 –\> moderate dispersion, consistent with attitude data

R squared: explains 34% of the variance in deservingness

F = 54.1, p \< 0.001 –\> overal model highly significant.

SUMMARY:

Individuals with personal or proximate contact with the carceral system perceive incarcerated people as significantly more deserving of help or compassion (β = 4.60, p \< 0.001). This relationship remains robust after adjusting for partisanship, ideology, punitiveness, racial attitudes, and demographics. In contrast, more socially conservative and punitive respondents evaluate incarcerated people as substantially less deserving. General deservingness orientations exert the largest effect, suggesting that judgments about incarcerated people reflect broader moral schemas about who merits social support. Race and gender do not systematically differentiate deservingness perceptions, although Hispanic respondents express somewhat lower ratings than White respondents.

BEFORE MOVING ONTO INTERPRETATION

### 2.7 With Weights

```{r}
# weights don't exist yet -- would be on 

# m1_w <- lm(prisonerdeserves ~ impact, data = raw_data_cleaned, weights = weight_var)
# summary(m1_w)
```

### 2.8 Variable Check -

Confounder, mediator, collider?

-   missingness patterns

-   variable types

-   distributions

-   outliers

Univariate explanatory power / how much does each control explain on its own?

```{r}

controls <- c("impact", "Republican","Democrat","fiscal_ideology","social_ideology","penalpoint","desor_core","fire_rare", "fire_privilege","fire_angry","fire_fear","gender","Black","Hispanic","OtherRace")

for (v in controls) {
  f <- as.formula(paste("d_incppl_pris_abc ~", v))
  fm <- lm(f, data = raw_data_cleaned)
  results <- rbind(results, 
                   data.frame(var = v, 
                              adjR2 = summary(fm)$adj.r.squared, 
                              AIC = AIC(fm)))
}

results <- results[order(-results$adjR2), ]
print(results)

#NEXT -- testing when adding controls in to m1 sequentially


for(v in controls[-1])
{
  f <- as.formula(paste("d_incppl_pris_abc ~ impact +", v))
  fm <- lm(f, data = raw_data_cleaned)
  cat(v, "adjR2 =", summary(fm)$adj.r.squared-summary(m1)$adj.r.squared, "\n")
}

#AIC - mirrors fit mostly, suggests to check counts for vars

for (v in controls) {
  f <- as.formula(paste("d_incppl_pris_abc ~", v))
  fm <- lm(f, data = raw_data_cleaned)
  cat(v, "N =", nobs(fm), "\n")
```

Find from adjusted R2 (explaining spread) that:

-   desor_core explains 30% of the variation in deservingness scores.

-   ideology measures explain 7-8%

-   FIRE variables explain 3-6%

-   PID 3-4%

-   impact explains very little – 1.5%

## 3. M2: by Impact

### 3.1 Incarcerated

```{r, minc}

minc <- lm(d_incppl_pris_abc ~ incarc +
    Republican + Democrat +
    fiscal_ideology + social_ideology +
    penalpoint + desor_core +
    fire_rare + fire_privilege + fire_angry + fire_fear +
    gender + Black + Hispanic + OtherRace,
  data = raw_data_cleaned
)
summary(minc)

```

#### 3.1a Robust SEs

```{r}
coef_table_incarc <- coeftest(minc, vcov = vcovHC(minc, type = "HC1", cluster = NULL)) %>% broom::tidy()
```

#### 3.1b. Standardized betas

```{r}

minc_s <- lm(d_incppl_pris_abc ~ incarc + Republican + Democrat + fiscal_ideology +
         social_ideology + penalpoint + desor_core + fire_rare + fire_privilege +
         fire_angry + fire_fear + gender + Black + Hispanic + OtherRace,
         data = raw_data_cleaned)

# 1. Standardized coefficients

inc_sd <- sd(raw_data_cleaned$d_incppl_pris_abc, na.rm = TRUE)

# compute sd for each predictor (binary and continuous)
inc_sds <- sapply(model.frame(minc_s)[-1], function(x) sd(x, na.rm=TRUE))  # excludes response
coefs <- coef(minc_s)

# standardized coef = b * sd(x) / sd(y).  For factor dummies sd(x) is the std dev of 0/1 vector.
inc_std_coefs <- coefs * c(NA, inc_sds / inc_sd)   # note: first coef is intercept so alignment matters
# better to use broom::tidy:
tidy(minc_s) |> 
  mutate(std_est = estimate * c(NA, inc_sds / inc_sd)) 

table_std_incarc <- tidy(minc_s) |> mutate(std_est = estimate * c(NA, inc_sds / inc_sd))
```

#### 3.1c Control Variable Check

```{r}

controls_inc <- c("incarc", "Republican","Democrat","fiscal_ideology","social_ideology","penalpoint","desor_core","fire_rare", "fire_privilege","fire_angry","fire_fear","gender","Black","Hispanic","OtherRace")
  
results_inc <- data.frame() #to reset

for (v_inc in controls_inc) {
  f_inc <- as.formula(paste("d_incppl_pris_abc ~", v_inc))
  fm_inc <- lm(f_inc, data = raw_data_cleaned)
  results_inc <- rbind(results_inc, 
                   data.frame(var = v_inc, 
                              adjR2 = summary(fm_inc)$adj.r.squared, 
                              AIC = AIC(fm_inc)))
}

results_inc <- results_inc[order(-results_inc$adjR2), ]
print(results_inc)

#NEXT -- testing when adding controls in to minc sequentially

#first make no control model version
minc_base <- lm(d_incppl_pris_abc ~ incarc, data = raw_data_cleaned)

minc_incr2 <- for(v_inc in controls_inc[-1])
{
  f_inc <- as.formula(paste("d_incppl_pris_abc ~ incarc +", v_inc))
  fm_inc <- lm(f_inc, data = raw_data_cleaned)
  cat(v_inc, "adjR2 =", summary(fm_inc)$adj.r.squared-summary(minc_base)$adj.r.squared, "\n")
}

#semi-partial R2 for each model predictor (how much unique variance each predictor explains in the full model, controlling for all other variables)

sp_r2 <- rsq.partial(minc, type = "sp")
sp_r2

#in table
sp_table_inc <- data.frame(
  variable = sp_r2$variable,
  semi_partial_R2 = sp_r2$partial.rsq
)

# largest to smallest
sp_table_inc <- sp_table %>%
  arrange(desc(semi_partial_R2))

# round for display
sp_table_inc$semi_partial_R2 <- round(sp_table_inc$semi_partial_R2, 4)

print(sp_table)

```

Interpretation (Control Variable Check)

Find from adjusted R2 (explaining spread) that:

-   desor_core explains 30% of the variation in deservingness scores.

-   ideology measures explain 7-8%

-   FIRE variables explain 3-6%

-   PID 3-4%

-   incarceration explains very little – 0.008%

AIC - mirrors fit mostly, suggests to check counts for vars

```{r}
for (v in controls_inc) {
  f <- as.formula(paste("d_incppl_pris_abc ~", v))
  fm <- lm(f, data = raw_data_cleaned)
  cat(v, "N =", nobs(fm), "\n")
}
```

### 3.2 Victimized

```{r, mv}

#10-31 AGAIN checking that I know what I'm talking about re: counts -- YES VICT, 277

mv <- lm(d_incppl_pris_abc ~ vict +
    Republican + Democrat +
    fiscal_ideology + social_ideology +
    penalpoint + desor_core +
    fire_rare + fire_privilege + fire_angry + fire_fear +
    gender + Black + Hispanic + OtherRace,
  data = raw_data_cleaned
)
summary(mv)
```

#### 3.2a Robust SEs

```{r}
coef_table_vict <- coeftest(mv, vcov = vcovHC(mv, type = "HC1"))%>% broom::tidy()
```

#### 3.2b Standardized Betas

```{r}

mvs <- lm(d_incppl_pris_abc ~ vict + Republican + Democrat + fiscal_ideology +
         social_ideology + penalpoint + desor_core + fire_rare + fire_privilege +
         fire_angry + fire_fear + gender + Black + Hispanic + OtherRace,
         data = raw_data_cleaned)

# 1. Standardized coefficients
library(broom)
y_sd_v <- sd(raw_data_cleaned$d_incppl_pris_abc, na.rm = TRUE)

# compute sd for each predictor (binary and continuous)
sds_v <- sapply(model.frame(mvs)[-1], function(x) sd(x, na.rm=TRUE))  # excludes response
coefs_v <- coef(mvs)

# standardized coef = b * sd(x) / sd(y).  For factor dummies sd(x) is the std dev of 0/1 vector.
std_coefs_v <- coefs_v * c(NA, sds_v / y_sd_v)   

# better to use broom::tidy:
tidy(mvs) |> 
  mutate(std_est_v = estimate * c(NA, sds_v / y_sd_v)) 

table_std_vict   <- tidy(mvs)    |> mutate(std_est_v = estimate * c(NA, sds_v / y_sd_v))
```

#### 3.2c Control Variable Check

Univariate explanatory power / how much does each control explain on its own?

```{r}

controls_vict <- c("vict", "Republican","Democrat","fiscal_ideology","social_ideology","penalpoint","desor_core","fire_rare", "fire_privilege","fire_angry","fire_fear","gender","Black","Hispanic","OtherRace")

results_vict <- data.frame()

for (v_vict in controls_vict) {
  f_vict <- as.formula(paste("d_incppl_pris_abc ~", v_vict))
  fm_vict <- lm(f_vict, data = raw_data_cleaned)
  results_vict <- rbind(results_vict, 
                   data.frame(var = v_vict, 
                              adjR2 = summary(fm_vict)$adj.r.squared, 
                              AIC = AIC(fm_vict)))
}

results_vict <- results_vict[order(-results_vict$adjR2), ]

print(results_vict)

#NEXT -- testing when adding controls in to m1 sequentially--needs fix
#first make base
mv_base <- lm(d_incppl_pris_abc ~ vict, data=raw_data_cleaned)

mvict_incr2 <-for(v_v in controls_vict[-1])
{
  f_v <- as.formula(paste("d_incppl_pris_abc ~ vict +", v_v))
  fm_v <- lm(f_v, data = raw_data_cleaned)
  cat(v_v, "adjR2 =", summary(fm_v)$adj.r.squared-summary(mv_base)$adj.r.squared, "\n")
}


#semi-partial R2 for each model predictor (how much unique variance each predictor explains in the full model, controlling for all other variables)

sp_r2_v <- rsq.partial(mv, type = "sp")
sp_r2_v

#in table
sp_table_v <- data.frame(
  variable_v = sp_r2_v$variable,
  semi_partial_R2_v = sp_r2_v$partial.rsq
)

# largest to smallest
sp_table_v <- sp_table_v %>%
  arrange(desc(semi_partial_R2_v))

# round for display
sp_table_v$semi_partial_R2_v <- round(sp_table_v$semi_partial_R2_v, 4)

print(sp_table_v)


```

### 3.3 Proximate

```{r}

mki <- lm(d_incppl_pris_abc ~ knowincarc +
    Republican + Democrat +
    fiscal_ideology + social_ideology +
    penalpoint + desor_core +
    fire_rare + fire_privilege + fire_angry + fire_fear +
    gender + Black + Hispanic + OtherRace,
  data = raw_data_cleaned
)
summary(mki)
```

#### 3.3a Robust SEs

```{r}
coef_table_prox <- coeftest(mki, vcov = vcovHC(mki, type = "HC1"))%>% broom::tidy()
```

#### 3.3b Standardized Betas

```{r}
mkis <- lm(d_incppl_pris_abc ~ knowincarc + Republican + Democrat + fiscal_ideology +
         social_ideology + penalpoint + desor_core + fire_rare + fire_privilege +
         fire_angry + fire_fear + gender + Black + Hispanic + OtherRace,
         data = raw_data_cleaned)

# 1. Standardized coefficients
library(broom)
y_sd_prox <- sd(raw_data_cleaned$d_incppl_pris_abc, na.rm = TRUE)

# compute sd for each predictor (binary and continuous)
sds_prox <- sapply(model.frame(mkis)[-1], function(x) sd(x, na.rm=TRUE))  # excludes response
coefs_prox <- coef(mkis)
# standardized coef = b * sd(x) / sd(y).  For factor dummies sd(x) is the std dev of 0/1 vector.
std_coefs_prox <- coefs_prox * c(NA, sds_prox / y_sd_prox)   # note: first coef is intercept so alignment matters
# better to use broom::tidy:
tidy(mkis) |> 
  mutate(std_est_prox = estimate * c(NA, sds_prox / y_sd_prox))

table_std_prox   <- tidy(mkis)   |> mutate(std_est_prox = estimate * c(NA, sds_prox / y_sd_prox))
```

#### 3.3c Control Variable Check

Univariate explanatory power / how much does each control explain on its own?

```{r}

controls_prox <- c("knowincarc","Republican","Democrat","fiscal_ideology","social_ideology","penalpoint","desor_core","fire_rare", "fire_privilege","fire_angry","fire_fear","gender","Black","Hispanic","OtherRace")

results_prox <- data.frame()

for (v_prox in controls_prox) {
  f_prox <- as.formula(paste("d_incppl_pris_abc ~", v_prox))
  fm_prox <- lm(f_prox, data = raw_data_cleaned)
  results_prox <- rbind(results_prox, 
                   data.frame(var = v_prox, 
                              adjR2 = summary(fm_prox)$adj.r.squared, 
                              AIC = AIC(fm_prox)))
}

results_prox <- results_prox[order(-results_prox$adjR2), ]

print(results_prox)

#NEXT -- testing when adding controls in to m1 sequentially--needs fix
#first make base
mki_base <- lm(d_incppl_pris_abc ~ knowincarc, data=raw_data_cleaned)

mprox_incr2<-for(v_prox in controls_prox[-1])
{
  f_prox <- as.formula(paste("d_incppl_pris_abc ~ knowincarc +", v_prox))
  fm_prox <- lm(f_prox, data = raw_data_cleaned)
  cat(v_prox, "adjR2 =", summary(fm_prox)$adj.r.squared-summary(mki_base)$adj.r.squared, "\n")
}


#semi-partial R2 for each model predictor (how much unique variance each predictor explains in the full model, controlling for all other variables)

sp_r2_prox <- rsq.partial(mki, type = "sp")
sp_r2_prox

#in table
sp_table_prox <- data.frame(
  variable_prox = sp_r2_prox$variable,
  semi_partial_R2_prox = sp_r2_prox$partial.rsq
)

# largest to smallest
sp_table_prox <- sp_table_prox %>%
  arrange(desc(semi_partial_R2_prox))

# round for display
sp_table_prox$semi_partial_R2_prox <- round(sp_table_prox$semi_partial_R2_prox, 4)

print(sp_table_prox)


```

### M2 Tables: Impact

```{r}

key_vars <- c("d_incppl_pris_abc","incarc","knowincarc","vict","desor_core",
              "fiscal_ideology","social_ideology","penalpoint",
              "Republican","Democrat","fire_rare","fire_privilege",
              "fire_angry","fire_fear","gender","Black","Hispanic","OtherRace")

summary_table <- raw_data_cleaned %>%
  dplyr::select(all_of(key_vars)) %>%
  skimr::skim()


#count N in each
# M2 - Incarcerated variable Ns
ncheck_inc <- sapply(controls_inc, function(v) 
  nobs(lm(as.formula(paste("d_incppl_pris_abc ~", v)), data = raw_data_cleaned))
) |> 
  tibble::enframe(name = "variable", value = "N")

# M2 - Victimized variable Ns
ncheck_vict <- sapply(controls_vict, function(v)
  nobs(lm(as.formula(paste("d_incppl_pris_abc ~", v)), data = raw_data_cleaned))
) |>
  tibble::enframe(name = "variable", value = "N")

# M2 - Proximate variable Ns
ncheck_prox <- sapply(controls_prox, function(v)
  nobs(lm(as.formula(paste("d_incppl_pris_abc ~", v)), data = raw_data_cleaned))
) |>
  tibble::enframe(name = "variable", value = "N")




#model comparison
contact_model_summary <- bind_rows(
  table_std_incarc %>% dplyr::select(term, std_est) %>% mutate(model = "Incarcerated"),
  table_std_vict   %>% dplyr::select(term, std_est_v) %>% mutate(model = "Victimized"),
  table_std_prox   %>% dplyr::select(term, std_est_prox) %>% mutate(model = "Proximate")
)


save(
  #summary
  summary_table, dv_density_data, contact_crosstab, incarc_ppl_hist,
  desor_core_hist,
  #model coefs
     coef_table_incarc, coef_table_vict, coef_table_prox,
  #semi-partial r2
     sp_table_inc, sp_table_v, sp_table_prox,
  #standardized betas
     table_std_incarc, table_std_vict, table_std_prox, 
  #univar and incremental fit diagnostics
  results_inc,results_vict,results_prox,
  minc_incr2,mvict_incr2, mprox_incr2,
  #N for each
  ncheck_inc, ncheck_vict, ncheck_prox,
  #model comparison table
  contact_model_summary, 
     file = "data/analysis_objects.RData")

summary_table
dv_density_data
contact_crosstab
incarc_ppl_hist
desor_core_hist
  #model coefs
  coef_table_incarc
coef_table_vict
coef_table_prox
  sp_table_inc
  sp_table_v
  sp_table_prox
  #standardized betas
    table_std_incarc
    table_std_vict
    table_std_prox
  #univar and incremental fit diagnostics
  results_inc
  results_vict
  results_prox
  minc_incr2
  mvict_incr2
  mprox_incr2
  #N for each
  ncheck_inc
  ncheck_vict
  ncheck_prox
  #model comparison table
  contact_model_summary

```

## 4. M3 & M4 - Deservingness on Carceral Policy

My second core hypothesis (**H2)**:

more generous perceptions of carceral citizens' deservingness will predict greater and more consistent support for assistance-oriented criminal legal policies and reduced support for punitive criminal legal policies.

To test this, I estimate outcome models for two constructed policy indices: prisonhelp and prisonpen. In each model, I include carceral contact (impact) and deservingness (prisonerdeserves) as key independent variables, along with the same controls used in M1.

In each model, $delta_1$ captures the association between carceral contact and policy preferences, controlling for deservingness and covariates. $delta_2$ estimates the relationship between deservingness and policy support, which directly tests H2. A positive and statistically significant $delta_2$ in the help model—and a negative $delta_2$ in the punitive model—would support the hypothesis that warmer evaluations of incarcerated people correspond to greater support for decarceral policies.

### M3 Deservingness on Help Policy

$$
\texttt{prisonhelp}_{i} = \delta_0 + \delta_1 \cdot \texttt{impact}_{i} + \delta_2 \cdot \texttt{prisonerdeserves}_{i} + X^{\prime}_{i} \cdot \texttt{controlsvector} + \epsilon_{i}
$$

ie: the more someone views incarcerated people as deserving:

-   the more they will support assistance-orientated CJ policy

-   the more they will oppose punitive policy

Variables needed:

-   Mediator: deservingness scores: d_incppl_pris_abc

-   any CJ contact vs no CJ contact

    -   impact

-   types of contact vs no contact

    -   knowincarc

    -   vict

    -   incarc

-   outcome indices:

    -   prisonhelp

        -   

    -   prisonpen

-   controls vector

```{r, controls vector}
var(raw_data_cleaned$prisonhelp, na.rm=TRUE)

names(raw_data_cleaned)

m3_0 <- lm(prisonhelp ~ Republican + Democrat , data = raw_data_cleaned)

summary(m3_0)

```

**Outcome variable: prisonhelp**

Average respondent support for policy items (1-5, 5 is strongly support):

-   requiring prison laborers to be paid minimum wage

-   eliminating prison phone call costs

-   increasing funding for GED/college in prisons

-   alternative-to-prison sentencing for convicted parents

-   restoring voting rights post-sentence (OG 1-3)

-   for bureaucratic requirements enabling voting access while incarcerated (OG 1-3)

Basic Stats:

-   variance = 0.488

**Variable logic check**

First, checking how respondents who identify as Republican vs Democrat differ on average support for help policies (Republican and Democrat are both dummy variables, with Independent/Other as the reference group). Confirmed that Democrats are more supportive of help policies than Republicans on average.

```{r}
prisonhelp_by_party <- lm(prisonhelp ~ Republican + Democrat, data = raw_data_cleaned)

summary(prisonhelp_by_party)
```

### 4.\_ Preliminary M3 & M4 Models with Controls

```{r}
#character vector of controls vector for m3 m4
 m3m4controls <- c("d_incppl_pris_abc",
  "impact","Republican", "Democrat", "fiscal_ideology", "social_ideology",
  "penalpoint", "desor_core", "fire_rare", "fire_privilege", "fire_angry",
  "fire_fear", "gender", "Black", "Hispanic", "OtherRace"
)

#M3 M4 DVs
m3dv <- "prisonhelp"

m4dv <- "prisonpen"

#Formulas where paste constructs a formula string of controls and turns it into an actual formula object
m3_controls_formula <- as.formula(
  paste(m3dv, "~", paste(m3m4controls, collapse = " + "))
)

m4_controls_formula <- as.formula(
  paste(m4dv, "~", paste(m3m4controls, collapse = " + "))
)

#M3 M4 OLS Models
m3_basic <- lm(m3_controls_formula, data = raw_data_cleaned)
m4_basic <- lm(m4_controls_formula, data = raw_data_cleaned)
```

### 4.\_ Interpretation (Preliminary M3 & M4 Models with Controls)

#### M3 Basic Model Interpretation

```{r}
summary(m3_basic)
```

The intercept indicates that, holding all controls at 0, the average support for help policies is 2.9 (on a 1-5 scale). This is not super substantive because many predictors are scaled starting at 1. Still, this reflects respondents who: -give incarcerated people either 0 or 50?? deservingness scores (are on a 0-100 scale) -are not impacted (0= no carceral contact) -who are political independents -have 0 fiscal ideology scores so most conservative fiscal ideology I think) - have 0 social ideology scores (also I think most conservative) - have 0 penalpoint scores (see prison as primarily for rehabilitation -- should I un-reverse code this so it is in same expected direction as conservative to liberal?) - have 0? desor_core scores (general orientation is to see help-needing groups as undeserving) (is constructed by average of deservingness scores that are 0-100) - strongly disagree that racial problems are rare (1-4 scale) - strongly agree that white people benefit from unearned racial advantages (1-4 scale) - strongly agree that they are angry that racism exists (1-4 scale) - strongly disagree that they are fearful of other races (1-4) - are women or other? codebook says 1= man - are white (Black, Hispanic, OtherRace are all 0)

...are 2.9 out of 5 (most supportive) so medium support or are slightly supportive of prison help policies. I'm not sure what the t and p value here mean for the intercept.

Coefficients: -having one unit warmer deservingness perceptions of incarcerated people increases support for help policies by 0.004 points, holding all else constant. So a 10-point increase in deservingness corresponds to a 0.04 point increase in help policy support. This is statistically significant (p\<0.001) but substantively small. -being impacted in some way decreases support for help policies by 0.037 points, holding all else constant. P=0.25 means this is statistically indistinguishable from 0 or likely due to chance. -being republican or democrat instead of independent both decrease support for help policies (by 0.12 points for democrats and 0.13 points for republicans. These effects are significant. Ideology is insignificant. - penalpoint: each one unit increase (1-7 scale, sees the purpose of prison more about punishment) decreases support for prison help policies decreases support for help policies by 0.039 points, holding all else constant. This is significant (p\<0.001) -on desor_core: each one unit increase (0-100 scale, more general orientation to see help-needing groups as deserving) decreases support for help policies by 0.007 points, holding all else constant. This is not significant.

Fit Statistics -- the model explains about 13% of variance in help policy support (R2 is 0.129) after adjusting for the number of predictors.

The F-stat and p-val confirm that, collectively, the predictors explain more variance than an intercept only model.

```{r}
summary(m4_basic)
```

### 4.\_\_ Diagnostic checks

Variance Inflation/multico. with vif --rule: 1 = no colinearity, less than 5 is sometimes concerning, sometimes stricter threshold is 2.5 - for m3_basic

-for m4_basic Neither model shows concerning multicollinearity according to VIFs. However, fiscal and social ideology are almost 5 for both models.

```{r}

vif(m3_basic)

vif(m4_basic)

```

Influence / leverage check

```{r}
plot(m3_basic, which=4) # cook's distance
plot(m4_basic, which=4) # cook's distance
```

Pairwise correlations and partial check (suppression?)

```{r}
# Focus on the key attitudinal predictors
vars_to_check <- c("desor_core", "d_incppl_pris_abc", "penalpoint", "social_ideology", "fiscal_ideology")

# Pairwise correlations
cor_mat <- raw_data_cleaned %>%
  dplyr::select(all_of(vars_to_check)) %>%
  cor(use = "pairwise.complete.obs")

print(round(cor_mat, 3))

car::avPlots(m3_basic, terms = ~ desor_core + d_incppl_pris_abc + penalpoint + social_ideology)

car::avPlots(m4_basic, terms = ~ desor_core + d_incppl_pris_abc + penalpoint + social_ideology)


```

Pairwise interpretation:

| Relationship | Interpretation |   |
|----|----|----|
| desor_core & inc ppl | strong shared variance, likely suppression |  |
| desorcore & ideo -0.37 | expected liberal/conservative |  |
| desorcore & penalpoint: -0.19 | mild relationship |  |
| fiscal & social ideology | redundant, reason to combine or drop fiscal? |  |

**Takeaway on desorcore negative coefficient:** once I partial out the incarcerated-specific deservingness, `desor_core`’s remaining variance is probably identifying respondents who are generally generous but not especially supportive of carceral “help” (e.g., people who prioritize social welfare but are skeptical when it comes to prison).

Confirmed above with regressing on residual gets 0.07, a small positive correlation (r\~0..07) between the residualized desor_core (removed its overlap with incarcerated ppl deservingness) and prisonhelp. Suggests there is probably a stronger correlation in the bivariate relationship that collapses after residualizing which suggests all (or nearly) the positive relationship between this orientation towards help-needing groups and help-oriented prison policy is shared with deservingness of incarcerated people.

```{r}

# Get residuals within a complete-data framedesor_core on the highly correlated predictor (d_incppl_pris_abc)
tmp <- raw_data_cleaned %>%
  dplyr::select(prisonhelp, desor_core, d_incppl_pris_abc) %>%
  dplyr::filter(complete.cases(.))

desor_resid <- lm(desor_core ~ d_incppl_pris_abc, data = tmp)$residuals


# 2. Correlate those residuals with the outcome (prisonhelp)
cor(tmp$prisonhelp, desor_resid)


```

Partial/semi partial R2

```{r}
rsq::rsq.partial(m3_basic, type = "sp")
rsq::rsq.partial(m4_basic, type = "sp")
```

Adjusted R2

```{r}

base_m3 <- lm(prisonhelp ~ d_incppl_pris_abc + impact, data = raw_data_cleaned)
full <- update(base_m3, . ~ . + desor_core)
summary(full)$adj.r.squared - summary(base_m3)$adj.r.squared

base_m4 <- lm(prisonpen ~ d_incppl_pris_abc + impact, data = raw_data_cleaned)
full_m4 <- update(base_m4, . ~ . + desor_core)
summary(full_m4)$adj.r.squared - summary(base_m4)$adj.r.squared
```

**First Without Controls -- Help**

```{r, m3_0}

m3_0 <- lm(prisonhelp ~ impact)

hist(raw_data_cleaned$prisonhelp)

hist(raw_data_cleaned$prisonpen)

```

```{r, m3}

m3 <- lm(prisonhelp ~ impact + d_incppl_pris_abc +
    Republican + Democrat + fiscal_ideology + social_ideology +penalpoint + desor_core + fire_rare + fire_privilege + fire_angry + fire_fear + gender + Black + Hispanic + OtherRace,

    data = raw_data_cleaned
)

summary(m3)

```

### M4 Deservingness on Punitive Policy

**First Without Controls -- Punitive Policy** First, check how respondents who identify as Republican vs Democrat differ on average support for punitive policies (Republican and Democrat are both dummy variables, with Independent/Other as the reference group).

```{r}
prisonpen_by_party <- lm(prisonhelp ~ Republican + Democrat, data = raw_data_cleaned)
summary(prisonpen_by_party)
```

$$
\texttt{prisonpen}_{i} = \delta_0 + \delta_1 \cdot \texttt{impact}_{i} + \delta_2 \cdot \texttt{prisonerdeserves}_{i} + X^{\prime}_{i} \cdot \texttt{controlsvector} + \epsilon_{i}
$$

In each model, $delta_1$ captures the association between carceral contact and policy preferences, controlling for deservingness and covariates. $delta_2$ estimates the relationship between deservingness and policy support, which directly tests H2. A positive and statistically significant $delta_2$ in the help model—and a negative $delta_2$ in the punitive model—would support the hypothesis that warmer evaluations of incarcerated people correspond to greater support for decarceral policies.

```{r, m4}

m4 <- lm(prisonpen ~ impact + d_incppl_pris_abc +
    Republican + Democrat + fiscal_ideology + social_ideology + penalpoint + desor_core + fire_rare + fire_privilege + fire_angry + fire_fear + gender + Black + Hispanic + OtherRace,

    data = raw_data_cleaned
)


summary(m4)



```

## 5 M5 & M6

To formally test whether perceptions of deservingness mediate the effect of carceral contact on policy preferences, I first estimate a set of interaction models to examine whether the relationship between deservingness and policy views differs depending on whether respondents have been directly or proximally impacted by the criminal legal system. This approach allows me to test whether contact moderates how deservingness evaluations translate into policy preferences—i.e., whether deservingness matters more (or less) for those with lived experience.

### M5: Interaction for Support of Carceral Policy Help

$$
\texttt{prisonhelp}_{i} = \theta_0 + \theta_1 \cdot \texttt{Deservingness}_{i} + \theta_2 \cdot \texttt{Contact}_{i} + \theta_3 (\texttt{Deservingness}_{i} \times \texttt{Contact}_{i}) + X^{\prime}_{i} \cdot \texttt{controlsvector} + \epsilon_{i}
$$

A significant interaction term (\$\\theta_3\$) would indicate that the effect of deservingness on policy support varies depending on whether the respondent has had contact with the criminal legal system. These models help determine whether deservingness operates similarly across groups—or if lived experience conditions how moral evaluations shape policy preferences.

```{r, m5}

m5a<- lm(prisonhelp ~ d_incppl_pris_abc * impact,
            data = raw_data_cleaned)
summary(m5a)

m5b <- lm(prisonhelp ~ d_incppl_pris_abc * impact,
            data = raw_data_cleaned)

m5 <- lm(prisonhelp ~ d_incppl_pris_abc * impact + Republican + Democrat + fiscal_ideology + social_ideology + penalpoint + desor_core + fire_rare + fire_privilege + fire_angry + fire_fear + gender + Black + Hispanic + OtherRace, 
            data = raw_data_cleaned)

summary(m5)

```

#### M5 Interpretation

The intercept is still n

The coefficient on the interaction term is -0.0003483

### M5_C : Interaction for Support of Carceral Policy Help with Mean-Centered Deservingness

### Model 6: Interaction for Support of Carceral Policy Punishment

$$
\texttt{prisonpen}_{i} = \theta_0 + \theta_1 \cdot \texttt{Deservingness}_{i} + \theta_2 \cdot \texttt{Contact}_{i} + \theta_3 (\texttt{Deservingness}_{i} \times \texttt{Contact}_{i}) + X^{\prime}_{i} \cdot \texttt{controlsvector} + \epsilon_{i}
$$

A significant interaction term (\$\\theta_3\$) would indicate that the effect of deservingness on policy support varies depending on whether the respondent has had contact with the criminal legal system. These models help determine whether deservingness operates similarly across groups—or if lived experience conditions how moral evaluations shape policy preferences.

```{r, m6}

```

### M6_C : Interaction for Support of Carceral Policy Punishment with Mean-Centered Deservingness
