---
title: "Correlations"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(tidyr)
library(readr)
library(janitor)   # for clean_names()
library(tibble)
library(stringr)
library(haven)
library(dplyr)
library(tidyr)
library(readr)
library(janitor)   # for clean_names()
library(tibble)
library(stringr)
library(haven)
library(tibble)
library(gt)
library(psych)
library(ggplot2)


source("variable_labels.R")
```


```{r}
sample <- read_csv("data/cleaned_control_data.csv")

not_full <- read_csv("data/cleaned_sans_control_data.csv")
```

```{r, desor}
#desor_core is respondent level composite

summary(sample$desor_core)

hist(sample$desor_core)

desors <- lm(desor_core ~ impact, sample)
  
summary(desors)

```


Correlations when full, incarc, knowincarc, vict, impacted, not impacted

```{r, desor0}
# ------------------------------------------------------------
# Deservingness / "sympathy" orientation: CORE items
# ------------------------------------------------------------

desor_cont_vars <- c(
   "d_welfare_tanf_abc","d_fstamps_snap_abc","d_medicaid_abc","d_medicare_abc",
   "d_unemployed_abc","d_poorfam_abc","d_homeless_abc","d_immigrants_abc","d_unauth_undoc_abc")

# ------------------------------------------------------------
# 1. Define analysis groups
# ------------------------------------------------------------

desor_groups <- list(
  
  full_sample = sample,
  
  incarc_1 = sample %>%
    filter(incarc == 1),
  
  knowincarc_1 = sample %>%
    filter(knowincarc == 1),
  
  vict_1 = sample %>%
    filter(vict == 1),
  
  impact_1 = sample %>%
    filter(impact == 1),
  
  impact_0 = sample %>%
    filter(impact == 0)
)


# Check N in each group
sapply(desor_groups, nrow)

# ------------------------------------------------------------
# 2. Correlations within each group
# ------------------------------------------------------------

desor_cont_corrs <- purrr::map(
  desor_groups,
  ~ psych::corr.test(
      .x %>% select(all_of(desor_cont_vars)),
      use = "pairwise",
      method = "pearson",
      adjust = "none"
    )
)
```

# ------------------------------------------------------------
# Inspect
# ------------------------------------------------------------

```{r, desor1}
# Full sample correlations
round(desor_cont_corrs$full_sample$r, 2)

# Incarcerated respondents
round(desor_cont_corrs$incarc_1$r, 2)

# Know someone incarcerated
round(desor_cont_corrs$knowincarc_1$r, 2)

# Victimized respondents
round(desor_cont_corrs$vict_1$r, 2)

# Any impact
round(desor_cont_corrs$impact_1$r, 2)

# Not impacted
round(desor_cont_corrs$impact_0$r, 2)

desor_cont_corrs$full_sample$n

desor_cont_corrs$incarc_1$n

desor_cont_corrs$knowincarc_1$n

desor_cont_corrs$vict_1$n

desor_cont_corrs$impact_1$n

desor_cont_corrs$impact_0$n

```



```{r, desor2}
# ------------------------------------------------------------
# 3. Number of observed deservingness items, by group
# ------------------------------------------------------------

desor_item_counts <- purrr::map(
  desor_groups,
  ~ {
    cont_mat <- .x %>%
      select(all_of(desor_cont_vars))
    
    cont_n <- rowSums(!is.na(cont_mat))
    
    table(cont_n, useNA = "ifany")
  }
)

desor_item_counts


# ------------------------------------------------------------
# 4. Reliability / internal consistency by group
# ------------------------------------------------------------

desor_cont_alpha <- purrr::map(
  desor_groups,
  ~ {
    
    cont_mat <- .x %>%
      select(all_of(desor_cont_vars))
    
    cont_n <- rowSums(!is.na(cont_mat))
    
    cont_keep <- cont_n >= 5
    
    psych::alpha(
      cont_mat[cont_keep, ],
      check.keys = TRUE
    )
  }
)

# Full sample
desor_cont_alpha$full_sample$total$raw_alpha
desor_cont_alpha$full_sample$total$std.alpha
desor_cont_alpha$full_sample$total$average_r
desor_cont_alpha$full_sample$n.obs

# Incarcerated
desor_cont_alpha$incarc_1$total$raw_alpha
desor_cont_alpha$incarc_1$total$std.alpha
desor_cont_alpha$incarc_1$total$average_r
desor_cont_alpha$incarc_1$n.obs

# vict
desor_cont_alpha$vict_1$total$raw_alpha

# Know Incarcerated
desor_cont_alpha$knowincarc_1$total$raw_alpha

# Impacted
desor_cont_alpha$impact_1$total$raw_alpha


# Not impacted
desor_cont_alpha$impact_0$total$raw_alpha
```


```{r}
# Item Stats
View(desor_cont_alpha$full_sample$item.stats)
```


```{r}
View(desor_cont_alpha$incarc_1$item.stats)
```


```{r}
View(desor_cont_alpha$vict_1$item.stats)
```


```{r}
View(desor_cont_alpha$knowincarc_1$item.stats)
```


```{r}
View(desor_cont_alpha$impact_1$item.stats)
```


```{r}
View(desor_cont_alpha$impact_0$item.stats)

```

# Help Outcome Correlations

Correlations when full, incarc, knowincarc, vict, impacted, not impacted

```{r, desor0}
# ------------------------------------------------------------
# Carceral state help policies
# ------------------------------------------------------------

q25_help_vars <- c("min_wage", "free_calls", "ed_funds", "alt_sentencing")

# ------------------------------------------------------------
# 1. Define analysis groups
# ------------------------------------------------------------

help_groups <- list(
  
  full_sample = sample,
  
  incarc_1 = sample %>%
    filter(incarc == 1),
  
  knowincarc_1 = sample %>%
    filter(knowincarc == 1),
  
  vict_1 = sample %>%
    filter(vict == 1),
  
  impact_1 = sample %>%
    filter(impact == 1),
  
  impact_0 = sample %>%
    filter(impact == 0)
)


# Check N in each group
sapply(help_groups, nrow)

# ------------------------------------------------------------
# 2. Correlations within each group
# ------------------------------------------------------------

help_cont_corrs <- purrr::map(
  help_groups,
  ~ psych::corr.test(
      .x %>% select(all_of(q25_help_vars)),
      use = "pairwise",
      method = "pearson",
      adjust = "none"
    )
)
```

# ------------------------------------------------------------
# Inspect
# ------------------------------------------------------------

```{r, desor1}
# Full sample correlations
round(help_cont_corrs$full_sample$r, 2)

# Incarcerated respondents
round(help_cont_corrs$incarc_1$r, 2)

# Know someone incarcerated
round(help_cont_corrs$knowincarc_1$r, 2)

# Victimized respondents
round(help_cont_corrs$vict_1$r, 2)

# Any impact
round(help_cont_corrs$impact_1$r, 2)

# Not impacted
round(help_cont_corrs$impact_0$r, 2)

help_cont_corrs$full_sample$n

help_cont_corrs$incarc_1$n

help_cont_corrs$knowincarc_1$n

help_cont_corrs$vict_1$n

help_cont_corrs$impact_1$n

help_cont_corrs$impact_0$n

```



```{r, desor2}
# ------------------------------------------------------------
# 3. Number of observed deservingness items, by group
# ------------------------------------------------------------

help_item_counts <- purrr::map(
  help_groups,
  ~ {
    help_mat <- .x %>%
      select(all_of(q25_help_vars))
    
    help_n <- rowSums(!is.na(help_mat))
    
    table(help_n, useNA = "ifany")
  }
)

help_item_counts


# ------------------------------------------------------------
# 4. Reliability / internal consistency by group
# ------------------------------------------------------------

help_alpha <- purrr::map(
  help_groups,
  ~ {
    
    help_mat <- .x %>%
      select(all_of(q25_help_vars))
    
    help_n <- rowSums(!is.na(help_mat))
    
    help_keep <- help_n >= 4
    
    psych::alpha(
      help_mat[help_keep, ],
      check.keys = TRUE
    )
  }
)

# Full sample
help_alpha$full_sample$total$raw_alpha
help_alpha$full_sample$total$std.alpha
help_alpha$full_sample$total$average_r
help_alpha$full_sample$n.obs

# Incarcerated
help_alpha$incarc_1$total$raw_alpha
help_alpha$incarc_1$total$std.alpha
help_alpha$incarc_1$total$average_r
help_alpha$incarc_1$n.obs

# vict
help_alpha$vict_1$total$raw_alpha

# Know Incarcerated
help_alpha$knowincarc_1$total$raw_alpha

# Impacted
help_alpha$impact_1$total$raw_alpha


# Not impacted
help_alpha$impact_0$total$raw_alpha
```


```{r}
# Item Stats
View(help_alpha$full_sample$item.stats)
```


```{r}
View(help_alpha$incarc_1$item.stats)
```


```{r}
View(help_alpha$vict_1$item.stats)
```


```{r}
View(help_alpha$knowincarc_1$item.stats)
```


```{r}
View(help_alpha$impact_1$item.stats)
```


```{r}
View(help_alpha$impact_0$item.stats)

```


# Police Outcome Correlations

Correlations when full, incarc, knowincarc, vict, impacted, not impacted

```{r, desor0}
# ------------------------------------------------------------
# Carceral state help policies
# ------------------------------------------------------------

police_items <- c(
  "local_police",
  "no_strangle",
  "police_tracker",
  "police_training",
  "charge_police"
)

# ------------------------------------------------------------
# 1. Define analysis groups
# ------------------------------------------------------------

police_groups <- list(
  
  full_sample = sample,
  
  incarc_1 = sample %>%
    filter(incarc == 1),
  
  knowincarc_1 = sample %>%
    filter(knowincarc == 1),
  
  vict_1 = sample %>%
    filter(vict == 1),
  
  impact_1 = sample %>%
    filter(impact == 1),
  
  impact_0 = sample %>%
    filter(impact == 0)
)


# Check N in each group
sapply(police_groups, nrow)

# ------------------------------------------------------------
# 2. Correlations within each group
# ------------------------------------------------------------

police_corrs <- purrr::map(
  police_groups,
  ~ psych::corr.test(
      .x %>% select(all_of(police_items)),
      use = "pairwise",
      method = "pearson",
      adjust = "none"
    )
)

```

# ------------------------------------------------------------
# Inspect
# ------------------------------------------------------------

```{r, desor1}
# Full sample correlations
round(police_corrs$full_sample$r, 2)

# Incarcerated respondents
round(police_corrs$incarc_1$r, 2)

# Know someone incarcerated
round(police_corrs$knowincarc_1$r, 2)

# Victimized respondents
round(police_corrs$vict_1$r, 2)

# Any impact
round(police_corrs$impact_1$r, 2)

# Not impacted
round(police_corrs$impact_0$r, 2)

police_corrs$full_sample$n

police_corrs$incarc_1$n

police_corrs$knowincarc_1$n

police_corrs$vict_1$n

police_corrs$impact_1$n

police_corrs$impact_0$n

```



```{r, desor2}
# ------------------------------------------------------------
# 3. Number of observed deservingness items, by group
# ------------------------------------------------------------

police_item_counts <- purrr::map(
  police_groups,
  ~ {
    police_mat <- .x %>%
      select(all_of(police_items))
    
    police_n <- rowSums(!is.na(police_mat))
    
    table(police_n, useNA = "ifany")
  }
)

police_item_counts


# ------------------------------------------------------------
# 4. Reliability / internal consistency by group
# ------------------------------------------------------------

police_alpha <- purrr::map(
  police_groups,
  ~ {
    
    police_mat <- .x %>%
      select(all_of(police_items))
    
    police_n <- rowSums(!is.na(police_mat))
    
    police_keep <- police_n >= 5
    
    psych::alpha(
      police_mat[police_keep, ],
      check.keys = TRUE
    )
  }
)

# Full sample
police_alpha$full_sample$total$raw_alpha
police_alpha$full_sample$total$std.alpha
police_alpha$full_sample$total$average_r
police_alpha$full_sample$n.obs

# Incarcerated
police_alpha$incarc_1$total$raw_alpha
police_alpha$incarc_1$total$std.alpha
police_alpha$incarc_1$total$average_r
police_alpha$incarc_1$n.obs

# vict
police_alpha$vict_1$total$raw_alpha

# Know Incarcerated
police_alpha$knowincarc_1$total$raw_alpha

# Impacted
police_alpha$impact_1$total$raw_alpha


# Not impacted
police_alpha$impact_0$total$raw_alpha
```


```{r}
# Item Stats
View(police_alpha$full_sample$item.stats)
```


```{r}
View(police_alpha$incarc_1$item.stats)
```


```{r}
View(police_alpha$vict_1$item.stats)
```


```{r}
View(police_alpha$knowincarc_1$item.stats)
```


```{r}
View(police_alpha$impact_1$item.stats)
```


```{r}
View(police_alpha$impact_0$item.stats)

```


# Duplicative police correlations?

```{r}
cor_matrix <- cor(
  sample[police_items],
  use = "pairwise.complete.obs"
)

round(cor_matrix, 2)

library(psych)


alpha_result <- psych::alpha(sample[police_items])

print(alpha_result)

names(alpha_result)

alpha_result$alpha.drop

alpha_result

as.data.frame(alpha_result$alpha.drop)

View(alpha_result$alpha.drop)

print(sample %>%
  pivot_longer(cols = all_of(police_items)) %>%
  ggplot(aes(x = value)) +
  geom_histogram(binwidth = 1) +
  facet_wrap(~ name))

#avg check
sample <- sample %>%
  mutate(police_index = rowMeans(across(all_of(police_items)), na.rm = TRUE))

summary(sample$police_index)

# police index 1: index of all the police items
sample <- sample %>%
  mutate(
    police_reform_index = rowMeans(across(all_of(police_items)), na.rm = TRUE)
  )

# police index 2: index of the items that hang together best

police_reform_core <- c(
  "no_strangle",
  "police_tracker",
  "police_training",
  "charge_police"
)

sample <- sample %>%
  mutate(
    police_reform_core_index = rowMeans(across(all_of(police_reform_core)), na.rm = TRUE)
  )

# last - 3 groups 
# sample <- sample %>%
#   mutate(
#     police_violence_prevention = rowMeans(across(c("q15_2_1_to_5", "q15_4_1_to_5")), na.rm = TRUE),
#     police_accountability = rowMeans(across(c("q15_3_1_to_5", "q15_5_1_to_5")), na.rm = TRUE),
#     police_local_residency = q15_1_1_to_5
#   )

## POLICE MEANS ANALYSIS 

# index 1: police_reform_index

## baseline: no impact
all_baseline_mean <- mean(
  sample$police_reform_index[sample$impact == 0],
  na.rm = TRUE
)

## compute means for each impact group
all_police_means <- tibble(
  var = impact_vars
) %>%
  mutate(
    label = impact_labels[var],
    mean = sapply(var, function(v) {
      mean(
        sample$police_reform_index[
          sample[[v]] == 1
        ],
        na.rm = TRUE
      )
    })
  ) %>%
  bind_rows(
    tibble(
      var = "no_impact",
      label = "No impact",
      mean = all_baseline_mean
    ),
    .
  ) %>%
  mutate(
    diff_from_no_impact = mean - all_baseline_mean
  )

all_police_means

##
# index 2: 
##

# baseline: no impact
baseline_mean <- mean(
  sample$police_reform_core_index[sample$impact == 0],
  na.rm = TRUE
)

# compute means for each impact group
police_means <- tibble(
  var = impact_vars
) %>%
  mutate(
    label = impact_labels[var],
    mean = sapply(var, function(v) {
      mean(
        sample$police_reform_core_index[
          sample[[v]] == 1
        ],
        na.rm = TRUE
      )
    })
  ) %>%
  bind_rows(
    tibble(
      var = "no_impact",
      label = "No impact",
      mean = baseline_mean
    ),
    .
  ) %>%
  mutate(
    diff_from_no_impact = mean - baseline_mean
  )

print(police_means) # index 2
print(all_police_means)
## index 3?
##
```

### 2.1A Police Policy

```{r}

source("variable_labels.R")



```

**Write up on coding police variables decision**

all - participants' opposition/ support (coded 1-5, unsure = 3)

\(1\) Require that police officers live in the cities or towns they police

\(2\) Make it a crime for police to use chokeholds or strangleholds

\(3\) Create a federal government database to track police officers who have been accused of misconduct

\(4\) Require police to be trained in nonviolent alternatives to deadly force

\(5\) Allow civilians to file lawsuits against police officers that use excessive force or misconduct

My loose overall theory is that experiences with crime victimization may influence trust in government and policy learning. Ie, people who have been victimized may view the CJ system as differently legitimate and trustworthy, potentially extend that to the state; they will also learn things about policy. That's the context in which I'm thinking about how to use these survey items.

My initial view is – theoretically – these mostly might fall into two categories. The first category is policing police, or police oversight – questions 3 and 5. The second is about police violence, or police violence prevention - so questions 2 and 4. My initial view is that question 1 is different, or might not as clearly fit into the buckets. but it could also fit into some idea of constraining police?

I want to make this decision in an appropriate way from the research design perspective – not select on my (xyz - forgot what this is called).

I'm not entirely clear if grouping these variables in a single index improves analyses – more power, especially given that my victimized sample is only 277? One option is use as a whole index, but then also break out separate analyses. Another is to do 3 (keep local as a standalone – this might look like:

police_map \<- tibble::tribble( \~var, \~label, \~section, "q15_1_1_to_5", "Local Police", "Misc. Police", "q15_2_1_to_5", "No Strangle", "Police Violence", "q15_3_1_to_5", "Police Tracker", "Police Police", "q15_4_1_to_5", "Police Training", "Police Violence", "q15_5_1_to_5", "Charge Police", "Police Police", )

) – odds are I would still want to run analyses with them as distinct variables. Last option is all 3?

Relationships between police variables

- Police should be local

  - & no strangle = 0.41

  - & track misconduct = 0.45

  - & require alternatives training = 0.39

  - & allow civilian lawsuits against = 0.39

- Strangleholds & chokeholds illegal

  - & police local = 0.41

  - & track misconduct = 0.58

  - & require alternatives training = 0.55

  - & allow civilian lawsuits against = 0.58

- Federal misconduct database

  - & police local = 0.45

  - & no strangle = 0.48

  - & track misconduct = 0.67

  - & require alternatives training = 0.70

- Require police to be trained in nonviolent alternatives to deadly force

  - & police local = 0.39

  - & no strangle = 0.55

  - & track misconduct = 0.67

  - & allow civilian lawsuits against = 0.65

- Allow civilians to file lawsuits against police officers that use excessive force or misconduct

  - & police local = 0.39

  - & no strangle = 0.58

  - & track misconduct = 0.70

  - & require alternatives training = 0.65

Alpha for all = 0.85 – reason to use as one.

ADD DECISION WRITE UP

Decision: two layers. First (index), second (ones that hang together best), then last (3) created above.
