---
title: "data_prep"
format: html
editor: visual
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(
  echo = FALSE,       # Hide code by default
  warning = FALSE,
  message = FALSE
)

# Load packages
library(dplyr)
library(tidyr)
library(readr)
library(janitor)   # for clean_names()
library(tibble)
library(stringr)
library(haven)
library(ggplot2)
library(ggbeeswarm)
library(readxl)
library(stringr)
library(dplyr)


```

OVERALL:
I start by constructing the relevant info first (with the exception of excluding respondents who don't consent), then describing what that info says, then I make exclusion (potentially additional flag) decisions last. 

# 1. Basic Prep

Load data

```{r}

#| include: true
# Unzip and load data

raw_data <- read_csv("./data/raw/scotg_results.csv")

raw_data <- clean_names(raw_data)
 
#names(raw_data)

```

Standardized and cleaned column names by (1) converting to lowerase, (2) replaces spaces with characters, removes accents, ensures uniqueness by adding numbers to duplicates, removing leading/trailing whitespace, creates syntatically valid names

Checked for changes, main change is made lowercase.

#### a. D&P Labels

This happens first because it involves a column I want to drop in the next cleaning step.

Assign Deserving & Power Blocks – redefine block using column names. Used the below to create block assignments for participants in A (n=900), B (n=898), and C (N=898) (to ultimately consolidate).

Explanation: Participants were sorted into 3 sections (ABC). For the deservingness and power questions, all 3 participant sections evaluated many of the same groups, but two of the three evaluated extra groups that participants in the other sections did not.

This shows up in the raw data:

- there are 3 versions of the same variables – one corresponding to participants in section a, another in section b, another in section c.

Here, I am creating a block variable to keep track of what block each participant was sorted into, so that when I merge the disparate variables, it is still easy to know which group each participant was assigned to (thus, what groups they would have been prompted to evaluated and also important because it appears that the deserving questions for participants in block A were forced choice and snapped to increments of 10: not the case for B and C).

```{r}
#| include: true
# DeservingPower Blocks

#FL_8_DO_Deserving&PowerBlockA %in% names(raw_data)
#View(raw_data)

# Add participant block variable (group a, b, c)
raw_data <- raw_data %>%
  mutate(
    block = case_when(
      !is.na(`fl_8_do_deserving_power_block_a`) ~ "A",
      !is.na(`fl_8_do_deserving_power_block_b`) ~ "B",
      !is.na(`fl_8_do_deserving_power_block_c`) ~ "C",
      TRUE ~ NA_character_
    )
  )

#count(raw_data, block)

# Save labeled but otherwise unmodified version
write_csv(raw_data, "./data/cleaned/block_labeled_raw_data.csv")

```


#### b. Cleaned data frame, deleting un-needed vars

```{r}
#| include: true

columns_to_drop <- c(
  "status", "start_date", "end_date", "ip_address", "recipient_last_name", "recipient_first_name", 
  "recipient_email", "external_reference", "recorded_date", "location_latitude", 
  "location_longitude", "distribution_channel", "user_language"
)

#confirming (07-08-26) these all exist still because I'm getting an error when I try to run the drop line

# columns_to_drop %in% names(raw_data)

# Aggregate all columns containing "_DO" (Display Order)

grep("_do_", names(raw_data), value = TRUE)

do_columns <- grep("_do", names(raw_data), value = TRUE)


#confirming (07-08-26) these all exist still because I'm getting an error when I try to run the drop line

# do_columns %in% names(raw_data)


# Drop do columns and others
raw_data <- raw_data %>%
  select(-all_of(c(columns_to_drop, do_columns)))

#09-08-26 above drop line ran

#view(raw_data)
```

### c. Drop non-consenters

```{r}
#| include: true
# Filter to those who consented
raw_data <- raw_data %>%
  filter(consent == 1)

#remove consent col

raw_data <- raw_data %>%
  select(-consent)

# Save again

write_csv(raw_data, "data/cleaned/raw_data_consent.csv")
```

Still using "raw data"

# 2. Initial Clean & Quality

### a. attn checks

uncleaned: q9, character, can be 1 2 3 4 NA 3 -99 --see 2095 passed, 599 did not q17, character, lots of different responses -2063 passed, 631 not q114_4, character can be 6 5 NA 4 -99

```{r}
# summary(raw_data$q114_4)
# 
# unique(raw_data$q17)
# 
# table(raw_data$q9, useNA = "ifany")
# table(raw_data$q17, useNA = "ifany")
# table(raw_data$q114_4, useNA = "ifany")
# 
# unique(raw_data$q114_4)
# 
# raw_data %>%
#   count(q114_4, sort = TRUE)

```

- create flags for passed/not each check and people who failed all

```{r}
#| include: true

# 1. create logical variable (true/false) that marks whether respondent answered Q9 with 4

raw_data$attn_q9_pass = !is.na(raw_data$q9) & raw_data$q9 == "4"

raw_data$attn_q9_typo = !is.na(raw_data$q9) & raw_data$q9 == "1"


#2. attention flag -  text responses to Q17, remove leading/trailing spaces, and convert everything to lowercase for standardization)

attn_text <- tolower(str_trim(raw_data$q17))
#table(attn_text)

#2a. create list of acceptable answers for Q17
accepted_variants <- c(
  "purple",
  "\"purple\"",
  "pruple",
  "purpke",
  "purplee",
  "purplre",
  "💜",
  "🟣",
  "purple🟣",
  "purple6"
)

#2b. check if each cleaned Q17 response is in the list of accepted variants. if yes, respondent passed attn check
raw_data$attn_check_q17 <- attn_text %in% accepted_variants
table(raw_data$attn_check_q17, useNA = "ifany")

#3. attn flag -- Create a flag for whether the respondent answered Q114_4 with a 6 (i.e., passed the third attention check

# table(raw_data$q114_4, useNA = "ifany")

raw_data$attn_q114_pass = !is.na(raw_data$q114_4) & raw_data$q114_4 == "6"


# older version counted NA as passed. lots of NAs likely because late in the survey. For now, only care is that I have passes recorded and in next code chunk will filter out those who failed all 3 checks.
# attn_check_q1114 <- raw_data$q114_4 == 6
# 
# #If Q114_4 equals -99 (an invalid code), treat that as a failed check.
# attn_check_q1114[raw_data$q114_4 == -99] <- FALSE
# 
# #If Q114_4 is NA (they didn’t see the question), count that as a pass because they weren’t shown the question
# attn_check_q1114[is.na(raw_data$q114_4)] <- TRUE  # not shown → pass
```

attn_q9_pass, attn_check_q17, attn_q114_pass
```{r}
raw_data <- raw_data %>%
  mutate(
    # Was each attention check actually answered?
    attn_q9_answered =
      !is.na(q9) & q9 != "-99",
    
    attn_q17_answered =
      !is.na(q17) & str_trim(q17) != "",
    
    attn_q114_answered =
      !is.na(q114_4) & q114_4 != "-99"
  )
```

```{r}
raw_data <- raw_data %>%
  mutate(
    attn_pass_count = rowSums(
      cbind(
        as.integer(attn_q9_pass),
        as.integer(attn_check_q17),
        as.integer(attn_q114_pass)
      ),
      na.rm = TRUE
    ),
    attn_fail_count = 3 - attn_pass_count
  )

raw_data <- raw_data %>%
  mutate(
    attn_answered_count =
      attn_q9_answered +
      attn_q17_answered +
      attn_q114_answered
  )
```

1,604 respondents passed all 3 attn checks 599 did not supply correct answers none of them 32 passed 1 459 passed 2


### b. survey duration

```{r}
#| include: true
raw_data <- raw_data %>%
 mutate(duration_in_seconds = as.numeric(duration_in_seconds))

raw_data <- raw_data %>%
  mutate(duration_minutes = duration_in_seconds / 60)

#get pre-cleaned survey duration info: will want this to justify cleaning cutoffs

#summary(raw_data_cleaned$duration_minutes)
```


Grouping ppl into speed categories

```{r}
summary(raw_data$duration_minutes)
raw_data <- raw_data %>%
  mutate(
    sub_8  = if_else(duration_minutes <= 8, 1, 0),
    sub_12 = if_else(duration_minutes <= 12, 1, 0),
    sub_20 = if_else(duration_minutes <= 20, 1, 0),
    sub_30 = if_else(duration_minutes <= 30, 1, 0),
    over_52 = if_else(duration_minutes >= 52, 1, 0),
    over_180 = if_else(duration_minutes >= 180, 1, 0)
  )

```

### c. progress/finished 

```{r}

raw_data <- raw_data %>%
 mutate(progress = as.numeric(progress))

raw_data <- raw_data %>%
  mutate(
    nonfinisher = if_else(finished == 0, 1, 0)
  )



```

visualization of respondent progress

```{r}
# visual of response times

uncleaned_progress <-
ggplot(raw_data,
       aes(x = progress, y = 0)) +
  geom_quasirandom(
    groupOnX = FALSE,
    width = .35,
    alpha = .35,
    size = 1.2
  ) +
  coord_cartesian(xlim = c(0, 120)) +
  labs(
    x = "Survey progress (number)",
    y = NULL,
    title = "Raw Survey Progress"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )


uncleaned_progress # save uncleaned_response_times as image


ggsave(
  "uncleaned_progress.png",
  plot = uncleaned_progress,
  width = 14,
  height = 9,
  dpi = 300,
)
```

### d. NAs & completion of vars -> end

### ***Des and pwr consolidation

Objective: consolidate blocks (i.e. so only 1 Incarcerated People/Prisoners)

ABC loop to consolidate groups. Verified. After below, dataset in memory has unified variables.

```{r consolidation, eval=TRUE}

# Load the mapping
mapping <- read_excel("scratch for consolidation.xlsx")

# Goal: standardize variable names in the mapping sheet so they match the names in raw_data_cleaned
# Clean column names and values in mapping to match raw_data_cleaned
mapping <- mapping %>%
  rename( #rename simplifies the column name for use in this loop
    var_a = `Deserving Group A`,
    var_b = `Deserving Group B`,
    var_c = `Deserving Group C`
  ) %>%
  mutate(
    var_a = str_replace_all(tolower(trimws(var_a)), " ", "_"),# trim removes whitespace
    var_b = str_replace_all(tolower(trimws(var_b)), " ", "_"),#lowercases
    var_c = str_replace_all(tolower(trimws(var_c)), " ", "_"),
    target_slug = trimws(target_slug)
  )

#undiagnostic

for (i in 1:nrow(mapping)) {

  new_var <- mapping$target_slug[i]
  var_a   <- mapping$var_a[i]
  var_b   <- mapping$var_b[i]
  var_c   <- mapping$var_c[i]

  raw_data[[new_var]] <- NA_real_

  if (!is.na(var_a) && var_a %in% names(raw_data)) {
    x <- as.numeric(raw_data[[var_a]])
    x[x == -99] <- NA_real_

    raw_data[[new_var]][raw_data$block == "A"] <-
      x[raw_data$block == "A"]
  }

  if (!is.na(var_b) && var_b %in% names(raw_data)) {
    x <- as.numeric(raw_data[[var_b]])
    x[x == -99] <- NA_real_

    raw_data[[new_var]][raw_data$block == "B"] <-
      x[raw_data$block == "B"]
  }

  if (!is.na(var_c) && var_c %in% names(raw_data)) {
    x <- as.numeric(raw_data[[var_c]])
    x[x == -99] <- NA_real_

    raw_data[[new_var]][raw_data$block == "C"] <-
      x[raw_data$block == "C"]
  }
}

mapping$target_slug %in% names(raw_data)
summary(raw_data[mapping$target_slug])


```

LIMITED Deserving/power for 1 group only: recode groups directly

```{r}
#| include: true
# Load the split mapping
split_map <- read.csv("./data/deserving_split.csv")

split_map <- split_map %>%
  rename(
    old_var = var.name,
    new_var = target_slug
  ) %>%
  mutate(
    old_var = str_replace_all(tolower(trimws(old_var)), " ", "_"),
    new_var = trimws(new_var)
  )

skipped <- list()

for (i in 1:nrow(split_map)) {

  old_var <- split_map$old_var[i]
  new_var <- split_map$new_var[i]

  if (!(old_var %in% names(raw_data))) {
    skipped[[new_var]] <- paste("Missing column:", old_var)
    next
  }

  x <- as.numeric(raw_data[[old_var]])
  x[x == -99] <- NA_real_

  if (all(is.na(x))) {
    skipped[[new_var]] <- paste("No valid data in", old_var)
    next
  }

  raw_data[[new_var]] <- x
}

if (length(skipped) > 0) {
  print("Skipped variables:")
  print(skipped)
} else {
  print("All group-specific variables created successfully.")
}

names(raw_data)

# OLD
# Loop through and rename each variable
#for (i in 1:nrow(split_map)) {
  
#  old_var <- split_map$var.name[i]     # e.g., "Deserving Group A_43"
#  new_var <- split_map$target_slug[i]  # e.g., "d_noncit_parent_a"
  
  # Just copy the values over
#  raw_data[[new_var]] <- raw_data[[old_var]]
#}

```

Checking this worked before removing the old variables

```{r}
#CHECKS
# 1. Did every mapped ABC target get created?
setdiff(mapping$target_slug, names(raw_data))

# 2. Are there any -99s left in consolidated ABC variables?

sapply(
  raw_data[mapping$target_slug],
  function(x) sum(x == -99, na.rm = TRUE)
)

# 3. Checking q26 (inc ppl) as ex to be used target directly
raw_data %>%
  select(
    block,
    deserving_group_a_26,
    d_incppl_pris_abc
  ) %>%
  filter(block == "A") %>%
  slice_head(n = 20)

```

Variables I can now prepare to drop

```{r}
block_vars_to_drop <- grep(
  "^(deserving_group|power_group)_[abc]_",
  names(raw_data),
  value = TRUE
)
```

Check/inspect to make sure right ones to drop

```{r}
length(block_vars_to_drop)
head(block_vars_to_drop, 20)
tail(block_vars_to_drop, 20)

```

Save pre-dropped version

```{r}
unclean_with_block_spec_vars <- raw_data
write_csv(unclean_with_block_spec_vars, "data/cleaned/unclean_with_block_spec_vars.csv")
```

Save for analysis dataset without block-spec/deserv vars (uncleaned_data_light, later to be cleaned_data_light)

```{r}
raw_data <- raw_data %>%
  dplyr::select(-all_of(block_vars_to_drop))

```


### 2d. NAs & completion of vars
I want to understand basic distribution of missingness/NAs, but because of the block specific variables I need to do so seperately for the trhee. 
```{r}
# Block-specific variables -----------------------------------------------

block_a_vars <- c(
  "d_noncit_parent_a",
  "d_incparents_a",
  "d_hsteen_a",
  "d_athiest_a",
  "d_cath_a",
  "d_evan_a",
  "d_jews_a",
  "d_pal_a",
  "d_muslim_a",
  "d_blackwom_a",
  "d_blackmen_a",
  "d_lds_a",
  "d_longcov_a",
  "d_exon_a",
  "d_weed_a",
  "d_opioid_a",
  "d_whitecol_a",
  "d_tech_a",
  "d_smokers_a",
  "d_genx_a",
  "d_boomer_a",
  "d_mill_a",
  "d_genz_a"
)

block_b_vars <- c(
  "d_metoo_b",
  "d_christnat_b",
  "d_ruralam_b",
  "d_urbanam_b",
  "d_suburbam_b",
  "d_billion_b",
  "d_blm_b",
  "d_maga_b",
  "d_progs_b",
  "d_soc_b",
  "d_proudboys_b",
  "d_attnys_b",
  "d_military_b",
  "d_vets_b",
  "d_solds_b",
  "d_essentials_b",
  "d_nurses_b",
  "d_drs_b",
  "d_dems_b",
  "d_reps_b",
  "d_media_b"
)

block_c_vars <- c(
  "d_bluecol_c",
  "d_working_c",
  "d_midclass_c",
  "d_migrantwork_c",
  "d_smallbiz_c",
  "d_taxpayers_c",
  "d_elderly_c",
  "d_moms_c",
  "d_retires_c",
  "d_sexoffenders_c",
  "d_sexhar_vict_c",
  "d_sexharrasser_c",
  "d_formerlyinc_c",
  "d_homeown_c",
  "d_regufee_c",
  "d_gunvi_vict_c",
  "d_polbrut_vict_c",
  "d_unins_c",
  "d_mentill_c",
  "d_aids_c",
  "d_gay_c",
  "d_lesbians_c"
)

all_block_vars <- c(
  block_a_vars,
  block_b_vars,
  block_c_vars
)

quality_derived_vars <- c(
  "attn_q9_pass",
  "attn_q9_typo",
  "attn_check_q17",
  "attn_q114_pass",
  "attn_pass_count",
  "attn_nonpass_count",
  "attn_answered_count",
  "duration_minutes",
  "sub_8",
  "sub_12",
  "sub_20",
  "sub_30",
  "sub_52",
  "sub_180",
  "nonfinisher",
  "common_na_count",
  "block_na_count",
  "na_count"
)

non_question_vars <- c(
  "response_id",
  "respondent_id",
  "block",
  "duration_in_seconds",
  "progress",
  "finished"
)

common_vars <- setdiff(
  names(raw_data),
  c(
    all_block_vars,
    quality_derived_vars,
    non_question_vars
  )
)
```

```{r}
raw_data <- raw_data %>%
  mutate(
    
    # NAs among variables available to everyone
    common_na_count = rowSums(
      is.na(pick(all_of(common_vars)))
    ),
    
    # NAs only within the block the respondent was assigned
    block_na_count = case_when(
      block == "A" ~ rowSums(is.na(pick(all_of(block_a_vars)))),
      block == "B" ~ rowSums(is.na(pick(all_of(block_b_vars)))),
      block == "C" ~ rowSums(is.na(pick(all_of(block_c_vars)))),
      TRUE ~ NA_real_
    ),
    
    # Total meaningful NAs
    na_count = common_na_count + block_na_count
  )
```


# 5. Description of Consented, pre-screening sample



```{r}
ncol(raw_data)
nrow(raw_data)
count(raw_data, block)
```
## attention checks

How'd they do on attention checks?
```{r}
count(raw_data, attn_pass_count)
count(raw_data, attn_answered_count)
```

Overall, 1,604 respondents passed all 3 checks. 462 passed 2, 29 passed 1, and 599 respondents "failed" all 3 checks -- or they did not supply correct answers. 

table:
```{r}
raw_data %>%
  summarise(
    n = n(),

    q9_pass_n = sum(attn_q9_pass),
    q9_pass_pct = mean(attn_q9_pass) * 100,
    
    attn_q9_typo_n = sum(attn_q9_typo),
    attn_q9_typo_pct = mean(attn_q9_typo) * 100,

    q17_pass_n = sum(attn_check_q17),
    q17_pass_pct = mean(attn_check_q17) * 100,

    q114_pass_n = sum(attn_q114_pass),
    q114_pass_pct = mean(attn_q114_pass) * 100
  )
```
Among pre-cleaned, consenting respondents (n = 2694), 78% (n=2095) passed the first check, 6.8% (n = 182) at least selected "strongly" (selected strongly disagree instead of strongly agree); 76% (n = 2066) passed the second attention check by typing purple or a very mild mispelling of it (ommitting a single letter, for example); and 60% (n = 1604) passed the last attention check or indicated that they never use a fake social media platform.

table factoring in answered
```{r}
attn_check_summary <- tibble(
  check = c("Q9", "Q17", "Q114"),
  
  total_n = nrow(raw_data),
  
  answered_n = c(
    sum(raw_data$attn_q9_answered),
    sum(raw_data$attn_q17_answered),
    sum(raw_data$attn_q114_answered)
  ),
  
  nonanswer_n = c(
    sum(!raw_data$attn_q9_answered),
    sum(!raw_data$attn_q17_answered),
    sum(!raw_data$attn_q114_answered)
  ),
  
  pass_n = c(
    sum(raw_data$attn_q9_pass),
    sum(raw_data$attn_check_q17),
    sum(raw_data$attn_q114_pass)
  )
) %>%
  mutate(
    # Answered, but did not pass
    wrong_n = answered_n - pass_n,
    
    # Everyone who did not pass: wrong answer + nonanswer
    nonpass_n = total_n - pass_n,
    
    answered_pct = answered_n / total_n * 100,
    
    # Your current percentages
    pass_pct_all = pass_n / total_n * 100,
    
    # More useful for actual attention performance:
    # Of people who ANSWERED, how many passed?
    pass_pct_answered = pass_n / answered_n * 100,
    
    # Of all apparent non-passes, what % are just nonanswers?
    nonanswer_pct_of_nonpass =
      nonanswer_n / nonpass_n * 100,
    
    # Of people who answered, what % actually got it wrong?
    wrong_pct_answered =
      wrong_n / answered_n * 100
  )

attn_check_summary
```
Attention checks themselves actually show respondent drop-off. The first check was answered by 2,298 respondents, the second was answered by 2,075 respondents, and the third was only answered by 1,663 respondents. For the first, of those who answered, 203 answered incorrectly and 2,095 answered correctly. For the second, 9 answered incorrectly and 2,066 passed. For the third, 59 answered incorrectly and 1,604 passed. 

**save that table for overleaf. 

```{r}
raw_data %>%
  count(attn_pass_count, attn_answered_count)
```

## duration
what does survey duration actually look like? 

```{r}
# visual of response times

uncleaned_response_times <-
ggplot(raw_data,
       aes(x = duration_minutes, y = 0)) +
  geom_quasirandom(
    groupOnX = FALSE,
    width = .35,
    alpha = .35,
    size = 1.2
  ) +
  coord_cartesian(xlim = c(0, 120)) +
  labs(
    x = "Survey duration (minutes)",
    y = NULL,
    title = "Survey Duration"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )


uncleaned_response_times # save uncleaned_response_times as image


ggsave(
  "uncleaned_response_times.png",
  plot = uncleaned_response_times,
  width = 14,
  height = 9,
  dpi = 300,
)

```

```{r}
summary(raw_data$duration_minutes)
```
Took respondents from less than a second to days to "complete" the survey. 
flags:
-sub_8
-sub_12
-sub_20
-sub_30
-over_52
-over_180

```{r}
raw_data %>%
  summarise(
    n = n(),

    sub_8_n = sum(sub_8),
    sub_8_pct = mean(sub_8) * 100,
    
    sub_12_n = sum(sub_12),
    sub_12_pct = mean(sub_12) * 100,

    sub_20_n = sum(sub_20),
    sub_20_pct = mean(sub_20) * 100,

    sub_30_n = sum(sub_30),
    sub_30_pct = mean(sub_30) * 100,
    
    over_52_n = sum(over_52),
    over_52_pct = mean(over_52) * 100,
    
    over_180_n = sum(over_180),
    over_180_pct = mean(over_180) * 100
  )
```
Of the 2,694 consenting respondents, 13% (n = 352) reached the end of the survey in less or equal to 8 minutes; 20% (n = 539) in less than or equal to 12 minutes; 38% (n = 1021) in less than or equal to 20 minutes, 59% (n = 1592) in or less than 30, 21% (n = 563) in 52 minutes or more, and 7% or 200 respondents took two hours or more. 

## finishers

nonfinisher

```{r}

raw_data %>%
  summarise(
    n = n(),
    
    finished_n = sum(finished == 1),
    finished_pct = finished_n / n,
    
    nonfinished_n = sum(nonfinisher),
    nonfinished_pct = nonfinished_n / n
  )


table(raw_data$finished, useNA = "ifany")
```
Of the 2,694 consenting respondents, 83% finished (as in, reached the end of the survey, irregardless of skipped questions) (n = 2,231) and 17% did not (n = 463).

## missingness? 

```{r}
summary(raw_data$na_count)
summary(raw_data$common_na_count)
summary(raw_data$block_na_count)

```


```{r}
# visual of nas/missingness

uncleaned_missingness <-
ggplot(raw_data,
       aes(x = na_count, y = 0)) +
  geom_quasirandom(
    groupOnX = FALSE,
    width = .35,
    alpha = .35,
    size = 1.2
  ) +
  coord_cartesian(xlim = c(0, 120)) +
  labs(
    x = "na count",
    y = NULL,
    title = "na count"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )


uncleaned_missingness 


ggsave(
  "uncleaned_missingness.png",
  plot = uncleaned_missingness,
  width = 14,
  height = 9,
  dpi = 300,
)
```


```{r}
raw_data %>%
  group_by(finished) %>%
  summarise(
    n = n(),
    min_na = min(na_count, na.rm = TRUE),
    q1_na = quantile(na_count, .25, na.rm = TRUE),
    median_na = median(na_count, na.rm = TRUE),
    mean_na = mean(na_count, na.rm = TRUE),
    q3_na = quantile(na_count, .75, na.rm = TRUE),
    max_na = max(na_count, na.rm = TRUE)
  )
```


```{r}
raw_data %>%
  group_by(block) %>%
  summarise(
    n = n(),
    median_na = median(na_count, na.rm = TRUE),
    mean_na = mean(na_count, na.rm = TRUE),
    q75_na = quantile(na_count, .75, na.rm = TRUE),
    q90_na = quantile(na_count, .90, na.rm = TRUE),
    max_na = max(na_count, na.rm = TRUE)
  )
```

```{r}
raw_data %>%
  mutate(
    progress_group = case_when(
      progress < 25  ~ "<25%",
      progress < 50  ~ "25–<50%",
      progress < 75  ~ "50–<75%",
      progress < 100 ~ "75–<100%",
      progress == 100 ~ "100%",
      TRUE ~ NA_character_
    )
  ) %>%
  group_by(progress_group) %>%
  summarise(
    n = n(),
    median_na = median(na_count, na.rm = TRUE),
    mean_na = mean(na_count, na.rm = TRUE),
    min_na = min(na_count, na.rm = TRUE),
    max_na = max(na_count, na.rm = TRUE)
  )
```

```{r}
raw_data %>%
  group_by(attn_answered_count) %>%
  summarise(
    n = n(),
    median_na = median(na_count, na.rm = TRUE),
    mean_na = mean(na_count, na.rm = TRUE),
    median_progress = median(progress, na.rm = TRUE),
    pct_finished = mean(finished == 1, na.rm = TRUE) * 100
  )
```
```{r}
raw_data %>%
  group_by(finished) %>%
  summarise(
    n = n(),
    
    median_progress =
      median(progress, na.rm = TRUE),
    
    median_na =
      median(na_count, na.rm = TRUE),
    
    median_common_na =
      median(common_na_count, na.rm = TRUE),
    
    median_block_na =
      median(block_na_count, na.rm = TRUE),
    
    median_attn_answered =
      median(attn_answered_count, na.rm = TRUE),
    
    median_attn_passed =
      median(attn_pass_count, na.rm = TRUE)
  )
```



## clustering of indicators?

overlap summary
```{r}
overlap_summary <- tibble(
  check = c("Q9", "Q17", "Q114"),
  
  total_n = nrow(raw_data),
  
  answered_n = c(
    sum(raw_data$attn_q9_answered),
    sum(raw_data$attn_q17_answered),
    sum(raw_data$attn_q114_answered)
  ),
  
  nonanswer_n = c(
    sum(!raw_data$attn_q9_answered),
    sum(!raw_data$attn_q17_answered),
    sum(!raw_data$attn_q114_answered)
  ),
  
  pass_n = c(
    sum(raw_data$attn_q9_pass),
    sum(raw_data$attn_check_q17),
    sum(raw_data$attn_q114_pass)
  )
) %>%
  mutate(
    # Answered, but did not pass
    wrong_n = answered_n - pass_n,
    
    # Everyone who did not pass: wrong answer + nonanswer
    nonpass_n = total_n - pass_n,
    
    answered_pct = answered_n / total_n * 100,
    
    # Your current percentages
    pass_pct_all = pass_n / total_n * 100,
    
    # More useful for actual attention performance:
    # Of people who either didn't answer or got all 3 , how many passed?
    pass_pct_answered = pass_n / answered_n * 100,
    
    # Of all apparent non-passes, what % are just nonanswers?
    nonanswer_pct_of_nonpass =
      nonanswer_n / nonpass_n * 100,
    
    # Of people who answered, what % actually got it wrong?
    wrong_pct_answered =
      wrong_n / answered_n * 100
  )

overlap_summary

```

### enough usable measurements?
Ie, did respondents supply enough of the questionnaire to furnish usable measurements?

Among people who broke off—or among finishers with substantial missingness—what analytically important information is actually present?

q15_1 q15_2 q15_3 q15_4 q15_5 q15 q25_1 q25_2 q25_3 q25_4 q25_1 q4_1_1
q4_1_2 q4_1_3 q4_2_1 q4_2_2 q4_2_3

q15_1 q15_2 q15_3 q15_4 q15_5 q15 q25_1 q25_2 q25_3 q25_4 q25_1 q4_1_1
q4_1_2 q4_1_3 q4_2_1 q4_2_2 q4_2_3, q62_1,q51_1, q51_2, q51_3, q51_4, q51_5, q51_6, q51_7, q51_8, dem101,

q32, q33, q34, q35, q36, q37, q38, q111,

used variables
```{r}

used_vars <- c("d_crim_abc", "d_incppl_pris_abc","d_incwomen_abc","q15_1", "q15_2", "q15_3", "q15_4", "q15_5",
"q25_1", "q25_2", "q25_3", "q25_4", "q4_1_1", "q4_1_2", "q4_1_3", "q4_2_1", "q4_2_2", "q4_2_3","q15_1", "q15_2", "q15_3", "q15_4", "q15_5", "q25_1", "q25_2", "q25_3", "q25_4", "q25_1", "q62_1", "q23_2","q23_7",
"q4_1_1", "q4_1_2", "q4_1_3", "q4_2_1", "q4_2_2", "q4_2_3",
"q62_1","q51_1", "q51_2", "q51_3", "q51_4", "q51_5", "q51_6", "q51_7", "q51_8","dem101","q32", "q33", "q34", "q35", "q36", "q37", "q38", "q111"
)

used_var_missing <- raw_data %>% 
  summarise(
    across(
      all_of(used_vars),
      ~sum(is.na(.))
    )
  ) %>%
  pivot_longer(
    everything(),
    names_to = "variable",
    values_to = "n_missing"
  ) %>%
  mutate(
    n_answered = nrow(raw_data)- n_missing,
    pct_missing = n_missing/nrow(raw_data) * 100,
    pct_answered = n_answered/nrow(raw_data) *100
  ) %>%
  arrange(desc(pct_missing))

View(used_var_missing)
```

# 4. SAVE: cleaned_light.csv

```{r}
#light csv
write_csv(cleaned_data_light, "data/cleaned/cleaned_light.csv") # main file to use going forward

testing <- readr::read_csv("data/cleaned/cleaned_light.csv")

```
