---
title: "data_prep"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,       # Hide code by default
  warning = FALSE,
  message = FALSE
)

# Load packages
library(dplyr)
library(tidyr)
library(readr)
library(janitor)   # for clean_names()
library(tibble)
library(stringr)
library(haven)
library(ggplot2)
library(ggbeeswarm)
library(readxl)
library(stringr)
library(dplyr)

```

## 1. Clean & Basic Prep


Load data
```{r}
#| eval: false
#| include: false
# Unzip and load data

raw_data <- read_csv("./data/raw/scotg_results.csv")

raw_data <- clean_names(raw_data)
 
#names(raw_data)

```
Standardized and cleaned column names by (1) converting to lowerase, (2) replaces spaces with characters, removes accents, ensures uniqueness by adding numbers to duplicates, removing leading/trailing whitespace, creates syntatically valid names

Checked for changes, main change is made lowercase.

#### a. Cleaned data frame, deleting un-needed vars

```{r}
#| include: false

columns_to_drop <- c(
  "status", "start_date", "end_date", "ip_address", "recipient_last_name", "recipient_first_name", 
  "recipient_email", "external_reference", "recorded_date", "location_latitude", 
  "location_longitude", "distribution_channel", "user_language"
)

# Aggregate all columns containing "_DO" (Display Order)

grep("_do_", names(raw_data), value = TRUE)

do_columns <- grep("_do", names(raw_data), value = TRUE)

# Drop do columns and others
raw_data <- raw_data %>%
  select(-all_of(c(columns_to_drop, do_columns)))

```


### b. Drop non-consenters

```{r}
#| eval: false
#| include: false

# Filter to those who consented
raw_data <- raw_data %>%
  filter(consent == 1)

#remove consent col

raw_data_cleaned <- raw_data_cleaned %>%
  select(-consent)

# Save again

write_csv(raw_data_cleaned, "data/cleaned/cleaned_data_consent.csv")
```


#### b. D&P Labels

Assign Deserving & Power Blocks – redefine block using column names. Used the below to create block assignments for participants in A (n=900), B (n=898), and C (N=898) (to ultimately consolidate).

Explanation: Participants were sorted into 3 sections (ABC). For the deservingness and power questions, all 3 participant sections evaluated many of the same groups, but two of the three evaluated extra groups that participants in the other sections did not.

This shows up in the raw data:

-   there are 3 versions of the same variables – one corresponding to participants in section a, another in section b, another in section c.

Here, I am creating a block variable to keep track of what block each participant was sorted into, so that when I merge the disparate variables, it is still easy to know which group each participant was assigned to (thus, what groups they would have been prompted to evaluated and also important because it appears that the deserving questions for participants in block A were forced choice and snapped to increments of 10: not the case for B and C).

```{r}
#| include: false
# DeservingPower Blocks

#FL_8_DO_Deserving&PowerBlockA %in% names(raw_data)
#View(raw_data)

# Add participant block variable (group a, b, c)
raw_data <- raw_data %>%
  mutate(
    block = case_when(
      !is.na(`fl_8_do_deserving_power_block_a`) ~ "A",
      !is.na(`fl_8_do_deserving_power_block_b`) ~ "B",
      !is.na(`fl_8_do_deserving_power_block_c`) ~ "C",
      TRUE ~ NA_character_
    )
  )
#count(raw_data, block)
# Save labeled but otherwise unmodified version
write_csv(raw_data, "./data/cleaned/labeled_raw_data.csv")

```


## 2. Clean & Quality
###  attn checks
uncleaned:
q9, character, can be 1 2 3 4 NA 3 -99
--see 2095 passed, 599 did not
q17, character, lots of different responses
-2063 passed, 631 not
q114_4, character can be 6 5 NA 4 -99

```{r}
# summary(raw_data_cleaned$q114_4)
# 
# unique(raw_data_cleaned$q17)
# 
# table(raw_data_cleaned$q9, useNA = "ifany")
# table(raw_data_cleaned$q17, useNA = "ifany")
# table(raw_data_cleaned$q114_4, useNA = "ifany")
# 
# unique(raw_data_cleaned$q114_4)
# 
# raw_data_cleaned %>%
#   count(q114_4, sort = TRUE)

```

-   create flags for passed/not each check and people who failed all

```{r}
#| eval: false
#| include: false

# 1. create logical variable (true/false) that marks whether respondent answered Q9 with 4

raw_data_cleaned$attn_q9_pass = !is.na(raw_data_cleaned$q9) & raw_data_cleaned$q9 == "4"

#true if 4

# table(raw_data_cleaned$q9, useNA = "ifany")
# table(raw_data_cleaned$attn_q9_pass, useNA = "ifany")


#2. attention flag -  text responses to Q17, remove leading/trailing spaces, and convert everything to lowercase for standardization)
attn_text <- tolower(str_trim(raw_data_cleaned$q17))

#2a. create list of acceptable answers for Q17
accepted_variants <- c("purple", "urple", "puple","purpel", "\"purple\"", "purpke", "purplee", "prurprle", "purpl", "purp", "💜", "🟣")

#2b. chheck if each cleaned Q17 response is in thhe list of accepted variants. if yes, respondent passed attn check
raw_data_cleaned$attn_check_q17 <- attn_text %in% accepted_variants
table(attn_check_q17, useNA = "ifany")

#3. attn flag -- Create a flag for whether the respondent answered Q114_4 with a 6 (i.e., passed the third attention check

# table(raw_data_cleaned$q114_4, useNA = "ifany")

raw_data_cleaned$attn_q114_pass = !is.na(raw_data_cleaned$q114_4) & raw_data_cleaned$q114_4 == "6"

# older version counted NA as passed. lots of NAs likely because late in the survey. For now, only care is that I have passes recorded and in next code chunk will filter out those who failed all 3 checks.
# attn_check_q1114 <- raw_data_cleaned$q114_4 == 6
# 
# #If Q114_4 equals -99 (an invalid code), treat that as a failed check.
# attn_check_q1114[raw_data_cleaned$q114_4 == -99] <- FALSE
# 
# #If Q114_4 is NA (they didn’t see the question), count that as a pass because they weren’t shown the question
# attn_check_q1114[is.na(raw_data_cleaned$q114_4)] <- TRUE  # not shown → pass
```

attn_q9_pass, attn_check_q17, attn_q114_pass

```{r}
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    attn_pass_count = rowSums(
      cbind(
        as.integer(attn_q9_pass),
        as.integer(attn_check_q17),
        as.integer(attn_q114_pass)
      ),
      na.rm = TRUE
    ),
    attn_fail_count = 3 - attn_pass_count
  )

count(raw_data_cleaned, attn_pass_count)

raw

```
1,604 respondents passed all 3 attn checks
599 passed none of them
32 passed 1
459 passed 2

```{r}
write_csv(raw_data_cleaned, "data/cleaned/cleaned_data_consent_pre_attn_rest.csv")

data_pre_attn_exclusion <- raw_data_cleaned

raw_data_cleaned <- raw_data_cleaned %>%
  filter(attn_pass_count > 0)

count(raw_data_cleaned, attn_pass_count)

```

With dropped respondents (failed all 3), I have 32 respondents who passed 1, 459 who passed 2, and 1,604 who passed all 3. Total = 2,095 respondents retained.

To check specific attention check pass/fail counts:
attn_q9_pass,
attn_check_q17,
attn_q114_pass
attn_pass_count

### e. Clean: via survey duration

```{r}
#| eval: false
#| include: false

raw_data_cleaned <- raw_data_cleaned %>%
 mutate(duration_in_seconds = as.numeric(duration_in_seconds))

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(duration_minutes = duration_in_seconds / 60)

#get pre-cleaned survey duration info: will want this to justify cleaning cutoffs

summary(raw_data_cleaned$duration_minutes)


quantile(
  raw_data_cleaned$duration_minutes,
  probs = c(.01, .05, .25, .5, .75, .95, .99),
  na.rm = TRUE
)

quantile(
  raw_data_cleaned$duration_minutes,
  probs = c(.50, .60, .70, .75, .80, .85, .90, .95, .97, .99),
  na.rm = TRUE
)
raw_data_cleaned %>%
  mutate(
    duration_group = case_when(
      duration_minutes < 15 ~ "< 15 min",
      duration_minutes < 30 ~ "15–29 min",
      duration_minutes < 45 ~ "30–44 min",
      duration_minutes < 60 ~ "45–59 min",
      duration_minutes < 90 ~ "60–89 min",
      duration_minutes < 120 ~ "90–119 min",
      duration_minutes < 360 ~ "2–6 hours",
      TRUE ~ "6+ hours"
    )
  ) %>%
  count(duration_group)

# visual of response times

uncleaned_response_times <-
ggplot(raw_data_cleaned,
       aes(x = duration_minutes, y = 0)) +
  geom_quasirandom(
    groupOnX = FALSE,
    width = .35,
    alpha = .35,
    size = 1.2
  ) +
  coord_cartesian(xlim = c(0, 120)) +
  labs(
    x = "Survey duration (minutes)",
    y = NULL,
    title = "Survey Duration"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )


uncleaned_response_times
# save uncleaned_response_times as image


ggsave(
  "uncleaned_response_times.png",
  plot = uncleaned_response_times,
  width = 14,
  height = 9,
  dpi = 300,
)

```
Grouping ppl into speed
```{r}

raw_data_cleaned <- raw_data_cleaned %>%
  filter(duration_minutes >= 8)

raw_data_cleaned$speeder <- ifelse(raw_data_cleaned$duration_minutes < 12, 1, 0) 

```

```{r}
speed_review <- raw_data_cleaned %>%
  filter(duration_minutes >= 8) %>%
  mutate(
    speed_group = case_when(
      duration_minutes < 12 ~ "8–<12 min",
      duration_minutes < 20 ~ "12–<20 min",
      TRUE ~ "20+ min"
    ),
    
    attn_fail_count = 3 - attn_pass_count,
    
    any_attn_fail = attn_fail_count >= 1,
    two_attn_fails = attn_fail_count >= 2
  )

speed_review %>%
  count(speed_group)
```
```{r}
speed_review %>%
  count(speed_group, attn_fail_count)
```
 I should probably remove people who were too fast two slow and only passed one attention check. or flag. 

remember: 
attn_q9_pass,
attn_check_q17,
attn_q114_pass
attn_pass_count
```{r}
speed_review %>%
  group_by(speed_group) %>%
  summarise(
    n = n(),
    
    q9_fail_n = sum(attn_q9_pass == FALSE, na.rm = TRUE),
    q9_fail_pct = mean(attn_q9_pass == FALSE, na.rm = TRUE) * 100,
    
    q17_fail_n = sum(attn_check_q17 == FALSE, na.rm = TRUE),
    q17_fail_pct = mean(attn_check_q17 == FALSE, na.rm = TRUE) * 100,
    
    q1114_fail_n = sum(attn_q114_pass == FALSE, na.rm = TRUE),
    q1114_fail_pct = mean(attn_q114_pass == FALSE, na.rm = TRUE) * 100
  )
```
fast respondents who had another issue
```{r}
speed_review %>%
  filter(
    duration_minutes < 12,
    attn_fail_count >= 1
  ) %>%
  select(
    response_id,
    duration_minutes,
    attn_pass_count,
    attn_fail_count,
    attn_q9_pass,
    attn_check_q17,
    attn_q114_pass
  ) %>%
  arrange(duration_minutes)
```
```{r}
speed_review %>%
  filter(
    duration_minutes < 20,
    attn_fail_count >= 1
  ) %>%
  select(
    response_id,
    duration_minutes,
    attn_pass_count,
    attn_fail_count,
    attn_q9_pass,
    attn_check_q17,
    attn_q114_pass
  ) %>%
  arrange(duration_minutes) %>%
  View()
```

```{r}
speed_review %>%
  summarise(
    # 8 to <12 minutes
    n_under_12 = sum(duration_minutes < 12),
    n_under_12_attn_fail =
      sum(duration_minutes < 12 & any_attn_fail),
    pct_under_12_attn_fail =
      mean(any_attn_fail[duration_minutes < 12]) * 100,
    
    # 8 to <20 minutes
    n_under_20 = sum(duration_minutes < 20),
    n_under_20_attn_fail =
      sum(duration_minutes < 20 & any_attn_fail),
    pct_under_20_attn_fail =
      mean(any_attn_fail[duration_minutes < 20]) * 100
  )
```

### e. completed vars
for data that has survived screening so far . . .
-which analytically important questions were they eligible to see
-which did they answer
-what pattens of missingness exist
-what consequences could those patterns have for analysis

q15_1 q15_2 q15_3 q15_4 q15_5 q15 q25_1 q25_2 q25_3 q25_4 q25_1 q4_1_1 q4_1_2 q4_1_3 q4_2_1 q4_2_2 q4_2_3, q62_1,q51_1 through q51_8, dem101, q32-38, q111, 

### f. completion (all and vars)

Did these people answer variables I use? 
AAPOR explicitly distinguishes survey breakoff from item nonresponse. Someone can reach the end but skip questions; someone can break off before the end but still provide usable answers to the crucial variables. AAPOR recommends coding unanswered questions distinctly from substantive answers such as “don't know” and “prefer not to answer,” and emphasizes documenting partial responses rather than automatically treating them as invalid cases.

And the breakoff literature gives you another reason not to equate finished == 0 with poor-quality respondent. Breakoff can be related to questionnaire design, burden, where a question occurs, and respondent characteristics. Peytchev found that breakoffs did not simply look like inattentive respondents.

Need to look for:
-loss of information/precision
-selection

Selection: do people who don't answer policy questions differ systematically in incarceratione xperience, ideology, education, survey engagement, race, underlying policy view? 

Missing data lit says percent missing does not tell by itself whether missingness is dangerous/problematic. Ordinary regressions usually perform complete-case/listwise deletion for the variables appearing in that particular model. That means that adding a control variable with missing observations can change the regression sample. Coeficcient can then change partyly becuase added the control but also partialyl because the model is being estimated with a different sample. 


## Retained Respondents who did not finish

```{r}

table(raw_data_cleaned$finished, useNA = "ifany")

finishers_df <- raw_data_cleaned %>% filter(finished==1)
write_csv(finishers_df, "data/cleaned/finishers.csv")

nonfinishers_df <- raw_data_cleaned %>% filter(finished == 0)
write_csv(nonfinishers_df, "data/cleaned/nonfinishers.csv")

```

```{r}
count(nonfinishers_df, block)
count(finishers_df, block)
```

what block are they in?

-   18 in A

-   28 in B

-   20 in C

| Block   | nonfinishers | finishers |
|---------|--------------|-----------|
| Block A | 20           | 680       |
| Block B | 29           | 685       |
| Block C | 20           | 661       |

```{r}
count(nonfinishers_df, block)
```

| Experiences       | nonfinishers | finishers |
|-------------------|--------------|-----------|
| Incarcerated      | 8            | 162       |
| Know incarcerated | 21           | 668       |
| Victims           | 60           | 267       |
|                   |              |           |

-   8 incarcerated

-   21 know incarcerated

-   60 victims

```{r}

count(nonfinishers_df, q23_1)

count(finishers_df, q15_1)
```

| NAs in                   | nonfinishers | finishers                |
|--------------------------|--------------|--------------------------|
| d_crim_abc               |              |                          |
| d_incppl_pris_abc        |              |                          |
| d_incwomen_abc           |              |                          |
| all Police Questions (5) | 16           |                          |
| all prison questions     | all 16 NAs   | 170 (171 for 25_4 and 6) |
| q25_1                    | 16           |                          |
| q25_2                    | ...          |                          |
| re enfranchisement       | 35           | 497                      |
| jail voting              | 35           | 497                      |


deservingness scores aside, variables I use/care about (in names at THIS stage): 

q15_1 q15_2 q15_3 q15_4 q15_5 q15 q25_1 q25_2 q25_3 q25_4 q25_1 q4_1_1 q4_1_2 q4_1_3 q4_2_1 q4_2_2 q4_2_3

```{r}
names(raw_data_cleaned)
```


## 2. Deservingness

### A. Deservingness and Power consolidation

Objective: consolidate blocks (i.e. so only 1 Incarcerated People/Prisoners)

```{r}
#| eval: false
#| include: false

mapping <- read_csv("./data/Unified_Deserving_Power_Target_Mapping.csv")
#view(mapping)
```

Prep by converting to numeric

```{r}
#| eval: false
#| include: false
# Identify all Deserving and Power columns
deserving_power_cols <- names(raw_data_cleaned)[str_detect(names(raw_data_cleaned), "^(Deserving|Power) Group")]

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(
    all_of(deserving_power_cols),
    as.character
  ))

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(
    all_of(deserving_power_cols),
    ~ as.numeric(na_if(.x, "-99"))
  ))

#table(raw_data_cleaned$`Deserving Group A_1`, useNA = "always")

```

ALL Evals: Create consolidated variable names for deservingness — OLD(?) can probably remove below because covered by excel loop.

```{r}
#| eval: false
#| include: false

new_vars <- c(
 "d_black_afam_abc", "d_native_ind_abc", "d_asianam_abc", "d_latino_his_abc", "d_whiteam_abc",
 "d_mideastam_abc", "d_incwomen_abc", "d_collegewom_abc", "d_enviro_abc", "d_feminists_abc",
 "d_proabort_choice_abc", "d_abortprov_abc", "d_welfare_tanf_abc", "d_fstamps_snap_abc",
 "d_unions_abc", "d_medicaid_abc", "d_medicare_abc", "d_unemployed_abc", "d_police_abc",
 "d_poorfam_abc", "d_wealthy_abc", "d_teachers_abc", "d_sci_experts_abc", "d_child_abc",
 "d_crim_abc", "d_incppl_abc", "d_homeless_abc", "d_immigrants_abc", "d_unauth_undoc_abc",
 "d_gunowners_abc", "d_voters_abc", "d_lgbt_abc", "d_trans_abc", "d_transkids_abc", "d_gunmans_abc",
 "d_ice_abc", "d_socmed_abc", "d_nra_abc", "d_bigbank_abc", "d_bigcorp_abc", "d_uni_abc",
 "d_comcoll_abc", "d_noncit_parent_a", "d_incparents_a", "d_hsteen_a", "d_athiest_a", "d_cath_a",
 "d_evan_a", "d_jews_a", "d_pal_a", "d_muslim_a", "d_blackwom_a", "d_blackmen_a", "d_lds_a",
 "d_longcov_a", "d_exon_a", "d_weed_a", "d_opioid_a", "d_whitecol_a", "d_tech_a", "d_smokers_a",
 "d_genx_a", "d_boomer_a", "d_mill_a", "d_genz_a"
)

for (var in new_vars) {
 raw_data_cleaned[[var]] <- NA #to initialize
}
```

Ran new version of ABC loop to consolidate groups. Verified. After below, dataset in memory has unified variables.

```{r consolidation, eval=TRUE}



# Load the mapping
mapping <- read_excel("scratch for consolidation.xlsx")
view(mapping)

# Goal: standardize variable names in the mapping sheet so they match the names in raw_data_cleaned
# Clean column names and values in mapping to match raw_data_cleaned
mapping <- mapping %>%
  rename( #rename simplifies the column name for use in this loop
    var_a = `Deserving Group A`,
    var_b = `Deserving Group B`,
    var_c = `Deserving Group C`
  ) %>%
  mutate(
    var_a = str_replace_all(tolower(trimws(var_a)), " ", "_"),# trim removes whitespace
    var_b = str_replace_all(tolower(trimws(var_b)), " ", "_"),#lowercases
    var_c = str_replace_all(tolower(trimws(var_c)), " ", "_")
  )


view(mapping)
#undiagnostic

for (i in 1:nrow(mapping)) {
  new_var <- mapping$target_slug[i]
  var_a <- mapping$var_a[i]
  var_b <- mapping$var_b[i]
  var_c <- mapping$var_c[i]

  # Initialize with NAs
  raw_data_cleaned[[new_var]] <- NA_real_

  # Assign A group values
  if (!is.na(var_a) && var_a %in% names(raw_data_cleaned)) {
    raw_data_cleaned[[new_var]][raw_data_cleaned$block == "A"] <- 
      as.numeric(raw_data_cleaned[[var_a]][raw_data_cleaned$block == "A"])
  }

  # Assign B group values
  if (!is.na(var_b) && var_b %in% names(raw_data_cleaned)) {
    raw_data_cleaned[[new_var]][raw_data_cleaned$block == "B"] <- 
      as.numeric(raw_data_cleaned[[var_b]][raw_data_cleaned$block == "B"])
  }

  # Assign C group values
  if (!is.na(var_c) && var_c %in% names(raw_data_cleaned)) {
    raw_data_cleaned[[new_var]][raw_data_cleaned$block == "C"] <- 
      as.numeric(raw_data_cleaned[[var_c]][raw_data_cleaned$block == "C"])
  }
}

mapping$target_slug %in% names(raw_data_cleaned)
summary(raw_data_cleaned[mapping$target_slug])
view(raw_data_cleaned)


#NEW VERSION OFdrop block specific variables -- this didn't seem to work -- nothing appears when I look at block_vars_to_drop

#block_vars_to_drop <- grep("^(deserving_group|power_group)_[abc]_", names(raw_data_cleaned), value = TRUE)

#cleaned_data_light <- raw_data_cleaned %>%
  select(-all_of(block_vars_to_drop))


```

**Old (pre Oct 27 6pm, and pre Oct 27)**

```{r}
# DELETE FROM HERE?

#full new oct 27
for (i in 1:nrow(mapping)) {
  new_var <- mapping$target_slug[i]
  var_a <- mapping$var_a[i]
  var_b <- mapping$var_b[i]
  var_c <- mapping$var_c[i]
  
  # Optional: diagnostic check
  missing_vars <- c(var_a, var_b, var_c)[!c(var_a, var_b, var_c) %in% names(raw_data_cleaned)]
  if (length(missing_vars) > 0) {
    message("⚠️ Missing column(s) for ", new_var, ": ", paste(missing_vars, collapse = ", "))
  }
  
  # Main: consolidation
  raw_data_cleaned[[new_var]] <- with(raw_data_cleaned, case_when(
    block == "A" & var_a %in% names(raw_data_cleaned) ~ as.numeric(get(var_a)),
    block == "B" & var_b %in% names(raw_data_cleaned) ~ as.numeric(get(var_b)),
    block == "C" & var_c %in% names(raw_data_cleaned) ~ as.numeric(get(var_c)),
    TRUE ~ NA_real_
  ))
}



# Create consolidated variables by starting to loop through each row in mapping spreadsheet. For each target group, new_var will be consolidated name, var_a etc will be the claned name of the A-block specific var
for (i in 1:nrow(mapping)) {
  new_var <- mapping$target_slug[i]
  var_a <- mapping$var_a[i]
  var_b <- mapping$var_b[i]
  var_c <- mapping$var_c[i]
  
  #  check for whether any source variable names are missing from dataset
  if (!all(c(var_a, var_b, var_c) %in% names(raw_data_cleaned))) {
    warning(paste("Missing column(s) for:", new_var))
  }


if (!all(c(var_a, var_b, var_c) %in% names(raw_data_cleaned))) {
  print(paste("❌ Missing column(s) for:", new_var))
}

#verify: 

raw_data_cleaned %>%
  select(block, deserving_group_a_26, d_incppl_pris_abc) %>%
  filter(block == "A") %>%
  slice(1:10)

#to here?

```

LIMITED Deserving/power for 1 group only: recode groups directly

```{r}
#| eval: false
#| include: false

# Load the split mapping
split_map <- read.csv("./data/deserving_split.csv")

# Clean and normalize names to match raw_data_cleaned
split_map <- split_map %>%
  rename(old_var = `var.name`, new_var = target_slug) %>%
  mutate(
    old_var = str_replace_all(tolower(trimws(old_var)), " ", "_"),
    new_var = trimws(new_var)
  )

skipped <- list()

for (i in 1:nrow(split_map)) {
  old_var <- split_map$old_var[i]
  new_var <- split_map$new_var[i]

  # Only proceed if old_var is in dataset
  if (!(old_var %in% names(raw_data_cleaned))) {
    skipped[[new_var]] <- paste("Missing column:", old_var)
    next
  }

  # Check if there is *any valid data* (not -99)
  non_missing <- raw_data_cleaned[[old_var]] != -99 & !is.na(raw_data_cleaned[[old_var]])
  if (!any(non_missing, na.rm = TRUE)) {
    skipped[[new_var]] <- paste("No valid data in", old_var)
    next
  }

  # If we get here: create the new variable
  raw_data_cleaned[[new_var]] <- raw_data_cleaned[[old_var]]
}

if (length(skipped) > 0) {
  print("⚠️ Skipped variables:")
  print(skipped)
} else {
  print("✅ All group-specific variables renamed successfully.")
}

names(raw_data_cleaned)

# OLD
# Loop through and rename each variable
#for (i in 1:nrow(split_map)) {
  
#  old_var <- split_map$var.name[i]     # e.g., "Deserving Group A_43"
#  new_var <- split_map$target_slug[i]  # e.g., "d_noncit_parent_a"
  
  # Just copy the values over
#  raw_data_cleaned[[new_var]] <- raw_data_cleaned[[old_var]]
#}

```

##### Clean -99s

```{r}
# Combine both sets of variables to clean
vars_to_clean <- c(mapping$target_slug, split_map$new_var)

# Only keep variables that exist in the dataset
vars_to_clean <- vars_to_clean[vars_to_clean %in% names(raw_data_cleaned)]

# Recode -99 to NA
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(all_of(vars_to_clean), ~ {
    if (is.character(.x)) {
      na_if(.x, "-99")  # character version
    } else if (is.numeric(.x)) {
      na_if(.x, -99)    # numeric version
    } else {
      .x  # leave unchanged
    }
  }))

sapply(raw_data_cleaned[vars_to_clean], class)

```

```{r}



```

### FOR NOW - Oct 27 7pm – skip dropping to create the light version (intended to prep for "light" version of csv without old deservingness vars (in lieu of consoldiated and renamed)

```{r}

#TO SKIP: 
  cleaned_data_light <- raw_data_cleaned
  
  
#vector of old variable names
block_vars_to_drop <- grep("^(deserving_group|power_group)_[abc]_", names(raw_data_cleaned), value = TRUE)

#create version with them dropped
#cleaned_data_light <- raw_data_cleaned %>%
#  select(-all_of(block_vars_to_drop))

view(cleaned_data_light)

```


# SAVE: cleaned_light.csv

```{r}


write_csv(raw_data_cleaned, "data/cleaned/cleaned_data_consent.csv")

#light csv
write_csv(cleaned_data_light, "data/cleaned/cleaned_light.csv") # main file to use going forward

testing <- readr::read_csv("data/cleaned/cleaned_light.csv")

sum(is.na(raw_data_cleaned$q51_1))

table(raw_data_cleaned$q51_1, useNA = "ifany")
```

# ARCHIVED CODE


OLD: consolidating ABCs

```{r blockspecifictargets}
#| eval: false
#| include: false
# Load the spreadsheet with the mappings

mapping <- read_excel("scratch for consolidation.xlsx")


#CHECKING
#view(mapping)
#view(raw_data_cleaned)
#colnames(raw_data_cleaned)

#clean mapping col names to match raw_data_cleaned
mapping <- mapping %>%
  mutate(
    `Deserving Group A` = str_replace_all(tolower(trimws(`Deserving Group A`)), " ", "_"),
    `Deserving Group B` = str_replace_all(tolower(trimws(`Deserving Group B`)), " ", "_"),
    `Deserving Group C` = str_replace_all(tolower(trimws(`Deserving Group C`)), " ", "_")
  )

#preview mapping to confirm fix
mapping %>%
  select(target_slug, `Deserving Group A`, `Deserving Group B`, `Deserving Group C`) %>%
  head(10)
print(
  mapping %>%
    select(target_slug, `Deserving Group A`, `Deserving Group B`, `Deserving Group C`) %>%
    head(10)
)

dim(mapping)  # shows number of rows and columns
print(mapping)
names(mapping)

# Create composite deservingness variables from the correct block column only
for (i in 1:nrow(mapping)) {
  new_var <- mapping$target_slug[i]
  var_a <- mapping$`Deserving Group A`[i]
  var_b <- mapping$`Deserving Group B`[i]
  var_c <- mapping$`Deserving Group C`[i]

  raw_data_cleaned[[new_var]] <- case_when(
    raw_data_cleaned$block == "A" ~ raw_data_cleaned[[var_a]],
    raw_data_cleaned$block == "B" ~ raw_data_cleaned[[var_b]],
    raw_data_cleaned$block == "C" ~ raw_data_cleaned[[var_c]],
    TRUE ~ NA_real_
  )
}

names(raw_data_cleaned)

```
