---
title: "data_prep"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,       # Hide code by default
  warning = FALSE,
  message = FALSE
)

# Load packages
library(dplyr)
library(tidyr)
library(readr)
library(janitor)   # for clean_names()
library(tibble)
library(stringr)
library(haven)
library(ggplot2)
library(ggbeeswarm)
library(readxl)
library(stringr)
library(dplyr)


```

# 1. Clean & Basic Prep

Load data

```{r}
#| eval: false
#| include: false
# Unzip and load data

raw_data <- read_csv("./data/raw/scotg_results.csv")

raw_data <- clean_names(raw_data)
 
#names(raw_data)

```

Standardized and cleaned column names by (1) converting to lowerase, (2) replaces spaces with characters, removes accents, ensures uniqueness by adding numbers to duplicates, removing leading/trailing whitespace, creates syntatically valid names

Checked for changes, main change is made lowercase.

#### a. D&P Labels
This happens first because it involves a column I want to drop in the next cleaning step.

Assign Deserving & Power Blocks – redefine block using column names. Used the below to create block assignments for participants in A (n=900), B (n=898), and C (N=898) (to ultimately consolidate).

Explanation: Participants were sorted into 3 sections (ABC). For the deservingness and power questions, all 3 participant sections evaluated many of the same groups, but two of the three evaluated extra groups that participants in the other sections did not.

This shows up in the raw data:

- there are 3 versions of the same variables – one corresponding to participants in section a, another in section b, another in section c.

Here, I am creating a block variable to keep track of what block each participant was sorted into, so that when I merge the disparate variables, it is still easy to know which group each participant was assigned to (thus, what groups they would have been prompted to evaluated and also important because it appears that the deserving questions for participants in block A were forced choice and snapped to increments of 10: not the case for B and C).

```{r}
#| include: false
# DeservingPower Blocks

FL_8_DO_Deserving&PowerBlockA %in% names(raw_data)
#View(raw_data)

# Add participant block variable (group a, b, c)
raw_data <- raw_data %>%
  mutate(
    block = case_when(
      !is.na(`fl_8_do_deserving_power_block_a`) ~ "A",
      !is.na(`fl_8_do_deserving_power_block_b`) ~ "B",
      !is.na(`fl_8_do_deserving_power_block_c`) ~ "C",
      TRUE ~ NA_character_
    )
  )

#count(raw_data, block)

# Save labeled but otherwise unmodified version
write_csv(raw_data, "./data/cleaned/block_labeled_raw_data.csv")

```
#### b. Cleaned data frame, deleting un-needed vars

```{r}
#| include: false

columns_to_drop <- c(
  "status", "start_date", "end_date", "ip_address", "recipient_last_name", "recipient_first_name", 
  "recipient_email", "external_reference", "recorded_date", "location_latitude", 
  "location_longitude", "distribution_channel", "user_language"
)

# Aggregate all columns containing "_DO" (Display Order)

grep("_do_", names(raw_data), value = TRUE)

do_columns <- grep("_do", names(raw_data), value = TRUE)

# Drop do columns and others
raw_data <- raw_data %>%
  select(-all_of(c(columns_to_drop, do_columns)))

```

### c. Drop non-consenters

```{r}
#| eval: false
#| include: false

# Filter to those who consented
raw_data <- raw_data %>%
  filter(consent == 1)

#remove consent col

raw_data <- raw_data %>%
  select(-consent)

# Save again

write_csv(raw_data, "data/cleaned/raw_data_consent.csv")
```
Still using "raw data"

# 2. Clean & Quality

### a. attn checks

uncleaned: q9, character, can be 1 2 3 4 NA 3 -99 --see 2095 passed, 599 did not q17, character, lots of different responses -2063 passed, 631 not q114_4, character can be 6 5 NA 4 -99

```{r}
# summary(raw_data$q114_4)
# 
# unique(raw_data$q17)
# 
# table(raw_data$q9, useNA = "ifany")
# table(raw_data$q17, useNA = "ifany")
# table(raw_data$q114_4, useNA = "ifany")
# 
# unique(raw_data$q114_4)
# 
# raw_data %>%
#   count(q114_4, sort = TRUE)

```

- create flags for passed/not each check and people who failed all

```{r}
#| eval: false
#| include: false

# 1. create logical variable (true/false) that marks whether respondent answered Q9 with 4

raw_data$attn_q9_pass = !is.na(raw_data$q9) & raw_data$q9 == "4"

#true if 4

# table(raw_data$q9, useNA = "ifany")
# table(raw_data$attn_q9_pass, useNA = "ifany")


#2. attention flag -  text responses to Q17, remove leading/trailing spaces, and convert everything to lowercase for standardization)
attn_text <- tolower(str_trim(raw_data$q17))

#2a. create list of acceptable answers for Q17
accepted_variants <- c("purple", "urple", "puple","purpel", "\"purple\"", "purpke", "purplee", "prurprle", "purpl", "purp", "💜", "🟣")

#2b. chheck if each cleaned Q17 response is in the list of accepted variants. if yes, respondent passed attn check
raw_data$attn_check_q17 <- attn_text %in% accepted_variants
table(raw_data$attn_check_q17, useNA = "ifany")

#3. attn flag -- Create a flag for whether the respondent answered Q114_4 with a 6 (i.e., passed the third attention check

# table(raw_data$q114_4, useNA = "ifany")

raw_data$attn_q114_pass = !is.na(raw_data$q114_4) & raw_data$q114_4 == "6"

# older version counted NA as passed. lots of NAs likely because late in the survey. For now, only care is that I have passes recorded and in next code chunk will filter out those who failed all 3 checks.
# attn_check_q1114 <- raw_data$q114_4 == 6
# 
# #If Q114_4 equals -99 (an invalid code), treat that as a failed check.
# attn_check_q1114[raw_data$q114_4 == -99] <- FALSE
# 
# #If Q114_4 is NA (they didn’t see the question), count that as a pass because they weren’t shown the question
# attn_check_q1114[is.na(raw_data$q114_4)] <- TRUE  # not shown → pass
```

attn_q9_pass, attn_check_q17, attn_q114_pass

```{r}
raw_data <- raw_data %>%
  mutate(
    attn_pass_count = rowSums(
      cbind(
        as.integer(attn_q9_pass),
        as.integer(attn_check_q17),
        as.integer(attn_q114_pass)
      ),
      na.rm = TRUE
    ),
    attn_fail_count = 3 - attn_pass_count
  )

count(raw_data, attn_pass_count)

```

1,604 respondents passed all 3 attn checks 599 passed none of them 32 passed 1 459 passed 2

```{r}
write_csv(raw_data, "data/cleaned/raw_data_consent_pre_attn_rest.csv")

data_pre_attn_exclusion <- raw_data # saving pre exclusion, or just use raw_data

raw_data_cleaned <- raw_data %>%
  filter(attn_pass_count > 0)#switched to raw_data_cleaned

count(raw_data_cleaned, attn_pass_count)

```

With dropped respondents (failed all 3), I have 32 respondents who passed 1, 459 who passed 2, and 1,604 who passed all 3. Total = 2,095 respondents retained.

To check specific attention check pass/fail counts: attn_q9_pass, attn_check_q17, attn_q114_pass attn_pass_count

### b. survey duration

```{r}
#| eval: false
#| include: false

raw_data_cleaned <- raw_data_cleaned %>%
 mutate(duration_in_seconds = as.numeric(duration_in_seconds))

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(duration_minutes = duration_in_seconds / 60)

#get pre-cleaned survey duration info: will want this to justify cleaning cutoffs

#summary(raw_data_cleaned$duration_minutes)

# visual of response times

uncleaned_response_times <-
ggplot(raw_data_cleaned,
       aes(x = duration_minutes, y = 0)) +
  geom_quasirandom(
    groupOnX = FALSE,
    width = .35,
    alpha = .35,
    size = 1.2
  ) +
  coord_cartesian(xlim = c(0, 120)) +
  labs(
    x = "Survey duration (minutes)",
    y = NULL,
    title = "Survey Duration"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )


uncleaned_response_times # save uncleaned_response_times as image


ggsave(
  "uncleaned_response_times.png",
  plot = uncleaned_response_times,
  width = 14,
  height = 9,
  dpi = 300,
)

```

Grouping ppl into speed

```{r}

raw_data_cleaned <- raw_data_cleaned %>%
  filter(duration_minutes >= 8)

raw_data_cleaned$speeder <- ifelse(raw_data_cleaned$duration_minutes < 12, 1, 0) 

```

```{r}
speed_review <- raw_data_cleaned %>%
  filter(duration_minutes >= 8) %>%
  mutate(
    speed_group = case_when(
      duration_minutes < 12 ~ "8–<12 min",
      duration_minutes < 20 ~ "12–<20 min",
      TRUE ~ "20+ min"
    ),
    
    attn_fail_count = 3 - attn_pass_count,
    
    any_attn_fail = attn_fail_count >= 1,
    two_attn_fails = attn_fail_count >= 2
  )

speed_review %>%
  count(speed_group)
```

```{r}
speed_review %>%
  count(speed_group, attn_fail_count)
```

I should probably remove people who were too fast two slow and only passed one attention check. or flag.

remember: attn_q9_pass, attn_check_q17, attn_q114_pass attn_pass_count

```{r}
speed_review %>%
  group_by(speed_group) %>%
  summarise(
    n = n(),
    
    q9_fail_n = sum(attn_q9_pass == FALSE, na.rm = TRUE),
    q9_fail_pct = mean(attn_q9_pass == FALSE, na.rm = TRUE) * 100,
    
    q17_fail_n = sum(attn_check_q17 == FALSE, na.rm = TRUE),
    q17_fail_pct = mean(attn_check_q17 == FALSE, na.rm = TRUE) * 100,
    
    q1114_fail_n = sum(attn_q114_pass == FALSE, na.rm = TRUE),
    q1114_fail_pct = mean(attn_q114_pass == FALSE, na.rm = TRUE) * 100
  )
```

fast respondents who had another issue

```{r}
speed_review %>%
  filter(
    duration_minutes < 12,
    attn_fail_count >= 1
  ) %>%
  select(
    response_id,
    duration_minutes,
    attn_pass_count,
    attn_fail_count,
    attn_q9_pass,
    attn_check_q17,
    attn_q114_pass
  ) %>%
  arrange(duration_minutes)
```

```{r}
speed_review %>%
  filter(
    duration_minutes < 20,
    attn_fail_count >= 1
  ) %>%
  select(
    response_id,
    duration_minutes,
    attn_pass_count,
    attn_fail_count,
    attn_q9_pass,
    attn_check_q17,
    attn_q114_pass
  ) %>%
  arrange(duration_minutes) %>%
  View()
```

```{r}
speed_review %>%
  summarise(
    # 8 to <12 minutes
    n_under_12 = sum(duration_minutes < 12),
    n_under_12_attn_fail =
      sum(duration_minutes < 12 & any_attn_fail),
    pct_under_12_attn_fail =
      mean(any_attn_fail[duration_minutes < 12]) * 100,
    
    # 8 to <20 minutes
    n_under_20 = sum(duration_minutes < 20),
    n_under_20_attn_fail =
      sum(duration_minutes < 20 & any_attn_fail),
    pct_under_20_attn_fail =
      mean(any_attn_fail[duration_minutes < 20]) * 100
  )
```

### c. completion of vars -->

# 3. Deservingness

### A. Deservingness and Power consolidation

Objective: consolidate blocks (i.e. so only 1 Incarcerated People/Prisoners)

ABC loop to consolidate groups. Verified. After below, dataset in memory has unified variables.

```{r consolidation, eval=TRUE}

# Load the mapping
mapping <- read_excel("scratch for consolidation.xlsx")

# Goal: standardize variable names in the mapping sheet so they match the names in raw_data_cleaned
# Clean column names and values in mapping to match raw_data_cleaned
mapping <- mapping %>%
  rename( #rename simplifies the column name for use in this loop
    var_a = `Deserving Group A`,
    var_b = `Deserving Group B`,
    var_c = `Deserving Group C`
  ) %>%
  mutate(
    var_a = str_replace_all(tolower(trimws(var_a)), " ", "_"),# trim removes whitespace
    var_b = str_replace_all(tolower(trimws(var_b)), " ", "_"),#lowercases
    var_c = str_replace_all(tolower(trimws(var_c)), " ", "_"),
    target_slug = trimws(target_slug)
  )

#undiagnostic

for (i in 1:nrow(mapping)) {

  new_var <- mapping$target_slug[i]
  var_a   <- mapping$var_a[i]
  var_b   <- mapping$var_b[i]
  var_c   <- mapping$var_c[i]

  raw_data_cleaned[[new_var]] <- NA_real_

  if (!is.na(var_a) && var_a %in% names(raw_data_cleaned)) {
    x <- as.numeric(raw_data_cleaned[[var_a]])
    x[x == -99] <- NA_real_

    raw_data_cleaned[[new_var]][raw_data_cleaned$block == "A"] <-
      x[raw_data_cleaned$block == "A"]
  }

  if (!is.na(var_b) && var_b %in% names(raw_data_cleaned)) {
    x <- as.numeric(raw_data_cleaned[[var_b]])
    x[x == -99] <- NA_real_

    raw_data_cleaned[[new_var]][raw_data_cleaned$block == "B"] <-
      x[raw_data_cleaned$block == "B"]
  }

  if (!is.na(var_c) && var_c %in% names(raw_data_cleaned)) {
    x <- as.numeric(raw_data_cleaned[[var_c]])
    x[x == -99] <- NA_real_

    raw_data_cleaned[[new_var]][raw_data_cleaned$block == "C"] <-
      x[raw_data_cleaned$block == "C"]
  }
}

mapping$target_slug %in% names(raw_data_cleaned)
summary(raw_data_cleaned[mapping$target_slug])


```


LIMITED Deserving/power for 1 group only: recode groups directly

```{r}
#| eval: false
#| include: false

# Load the split mapping
split_map <- read.csv("./data/deserving_split.csv")

split_map <- split_map %>%
  rename(
    old_var = var.name,
    new_var = target_slug
  ) %>%
  mutate(
    old_var = str_replace_all(tolower(trimws(old_var)), " ", "_"),
    new_var = trimws(new_var)
  )

skipped <- list()

for (i in 1:nrow(split_map)) {

  old_var <- split_map$old_var[i]
  new_var <- split_map$new_var[i]

  if (!(old_var %in% names(raw_data_cleaned))) {
    skipped[[new_var]] <- paste("Missing column:", old_var)
    next
  }

  x <- as.numeric(raw_data_cleaned[[old_var]])
  x[x == -99] <- NA_real_

  if (all(is.na(x))) {
    skipped[[new_var]] <- paste("No valid data in", old_var)
    next
  }

  raw_data_cleaned[[new_var]] <- x
}

if (length(skipped) > 0) {
  print("Skipped variables:")
  print(skipped)
} else {
  print("All group-specific variables created successfully.")
}

names(raw_data_cleaned)

# OLD
# Loop through and rename each variable
#for (i in 1:nrow(split_map)) {
  
#  old_var <- split_map$var.name[i]     # e.g., "Deserving Group A_43"
#  new_var <- split_map$target_slug[i]  # e.g., "d_noncit_parent_a"
  
  # Just copy the values over
#  raw_data_cleaned[[new_var]] <- raw_data_cleaned[[old_var]]
#}

```

Checking this worked before removing the old variables
```{r}
#CHECKS
# 1. Did every mapped ABC target get created?
setdiff(mapping$target_slug, names(raw_data_cleaned))

# 2. Are there any -99s left in consolidated ABC variables?

sapply(
  raw_data_cleaned[mapping$target_slug],
  function(x) sum(x == -99, na.rm = TRUE)
)

# 3. Check one important target directly
raw_data_cleaned %>%
  select(
    block,
    deserving_group_a_26,
    d_incppl_pris_abc
  ) %>%
  filter(block == "A") %>%
  slice_head(n = 20)

```

Variables I can now prepare to drop
```{r}
block_vars_to_drop <- grep(
  "^(deserving_group|power_group)_[abc]_",
  names(raw_data_cleaned),
  value = TRUE
)
```

Check/inspect to make sure right ones to drop
```{r}
length(block_vars_to_drop)
head(block_vars_to_drop, 20)
tail(block_vars_to_drop, 20)

```

Save pre-dropped version

```{r}
clean_with_block_spec_vars <- raw_data_cleaned
write_csv(clean_with_block_spec_vars, "data/cleaned/clean_with_block_spec_vars.csv")
```

Save analysis dataset without block-spec/deserv vars (cleaned_data_light)
```{r}
cleaned_data_light <- raw_data_cleaned %>%
  dplyr::select(-all_of(block_vars_to_drop))

```

# 4. SAVE: cleaned_light.csv

```{r}
#light csv
write_csv(cleaned_data_light, "data/cleaned/cleaned_light.csv") # main file to use going forward

testing <- readr::read_csv("data/cleaned/cleaned_light.csv")

```

