---
title: "Data_Junction"
output: html_notebook
---

In this script, I just break into those who finished the survey (ie demographic questions) and those who did not.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(tidyr)
library(readr)
library(janitor)   # for clean_names()
library(tibble)
library(stringr)
library(haven)
library(dplyr)
library(tidyr)
library(readr)
library(janitor)   # for clean_names()
library(tibble)
library(stringr)
library(haven)
library(tibble)
library(gt)
library(psych)
library(ggplot2)
library(kableExtra)


```

# Setup

```{r}

raw_data_cleaned <-readr::read_csv("data/cleaned/all_vars_over_8.csv")

```

sept 15 N is 2327

```{r}

used <- c("d_crim_abc", "d_incppl_pris_abc","d_incwomen_abc",
          "impact", "incarc", "knowincarc", "vict",
          "local_police", "no_strangle", "police_tracker", "police_training", "charge_police","dpen", "min_wage", "free_calls", "ed_funds", "alt_sentencing", "lwop",
"gov_attn", "vote2020", "willvote", "auto_reg", "post_felony_vote", "early_vote", "election_holiday", "voter_id", "noncit_local", "jail_access_rev", "White", "Black", "Hisp_Latino", "Asian_Am", "Native_Am",  "Pac_Island", "MENA", "race_notlisted",
  "self_ideo", "fire_rare", "fire_privilege", "fire_angry", "fire_fear", "fiscal_ideology", "social_ideology")

used_missing <- raw_data_cleaned %>% 
  summarise(
    across(
      all_of(used),
      ~sum(is.na(.))
    )
  ) %>%
  pivot_longer(
    everything(),
    names_to = "variable",
    values_to = "n_missing"
  ) %>%
  mutate(
    n_answered = nrow(raw_data_cleaned)- n_missing,
    pct_missing = n_missing/nrow(raw_data_cleaned) * 100,
    pct_answered = n_answered/nrow(raw_data_cleaned) *100
  ) %>%
  arrange(desc(pct_missing))

View(used_missing)

```

```{r}

demos <- c(
  "White",
  "Black",
  "Hisp_Latino",
  "Asian_Am",
  "Native_Am",
  "Pac_Island",
  "MENA",
  "race_notlisted",
   "self_ideo", "fire_rare", "fire_privilege", "fire_angry", "fire_fear", "fiscal_ideology", "social_ideology"
)


raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    no_controls = if_all(
      all_of(demos),
      is.na
    )
  )

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    missing_controls = if_any(
      all_of(demos),
      is.na
    )
  )

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    no_controls = if_all(
      all_of(demos),
      is.na
    ),
    missing_controls = if_any(
      all_of(demos),
      is.na
    ),
    
    control_status = case_when(
      no_controls ~ "No controls",
      missing_controls ~ "Some controls missing",
      TRUE ~ "All controls present"
    )
  )

#how many
raw_data_cleaned %>%
  count(control_status)

#version of just ppl w no controls for comparison
df_no_control <- raw_data_cleaned %>%
  filter(missing_controls)

#remove ppl without controls from missing controls
data_full <- raw_data_cleaned %>%
    filter(!missing_controls)

nrow(data_full)

```

## impact x two dfs

```{r}
table(data_full$incarc)
table(data_full$knowincarc)
table(data_full$vict)
table(data_full$impact)

table(df_no_control$impact)
```

## ration x two dfs

```{r}

table(data_full$ration_scripts)
table(df_no_control$ration_scripts)

table(data_full$ration_food)
table(df_no_control$ration_food)

```

## Drop NON USED variables now

This could change, but these are variables that you are really probably not using

```{r}
deserving_block_drop <- c(
  # Block A
  "d_hsteen_a",
  "d_athiest_a",
  "d_cath_a",
  "d_evan_a",
  "d_jews_a",
  "d_pal_a",
  "d_muslim_a",
  "d_blackwom_a",
  "d_blackmen_a",
  "d_lds_a",
  "d_longcov_a",
  "d_whitecol_a",
  "d_tech_a",
  "d_smokers_a",
  "d_genx_a",
  "d_boomer_a",
  "d_mill_a",
  "d_genz_a",
  
  # Block B
  "d_metoo_b",
  "d_christnat_b",
  "d_ruralam_b",
  "d_urbanam_b",
  "d_suburbam_b",
  "d_billion_b",
  "d_blm_b",
  "d_maga_b",
  "d_progs_b",
  "d_soc_b",
  "d_proudboys_b",
  "d_attnys_b",
  "d_military_b",
  "d_vets_b",
  "d_solds_b",
  "d_essentials_b",
  "d_nurses_b",
  "d_drs_b",
  "d_dems_b",
  "d_reps_b",
  "d_media_b",
  
  # Block C
  "d_bluecol_c",
  "d_working_c",
  "d_midclass_c",
  "d_migrantwork_c",
  "d_smallbiz_c",
  "d_taxpayers_c",
  "d_elderly_c",
  "d_moms_c",
  "d_retires_c",
  "d_homeown_c",
  "d_regufee_c",
  "d_unins_c",
  "d_mentill_c",
  "d_aids_c",
  "d_gay_c",
  "d_lesbians_c"
)
```

```{r}
# ------------------------------------------------------------
# 3. Abortion
# ------------------------------------------------------------

abortion_drop <- c(
  "abortion_op",
  "abortion_death",
  "abortion_health",
  "abortion_rape",
  "abortion_defect",
  "abortion_miscar",
  "abortion_debt",
  "abortion_violence",
  "abortion_incarc",
  "abortion_teen",
  "abortion_want",
  "abortion_ability"
)


# ------------------------------------------------------------
# 4. Inequality policy items
# ------------------------------------------------------------

inequality_drop <- c(
  "reduce_ineq_minwage",
  "reduce_ineq_studdebt",
  "reduce_ineq_taxrich",
  "reduce_ineq_moremedcare",
  "reduce_ineq_lessmig",
  "reduce_ineq_poorben",
  "reduce_ineq_morecoll"
)


# ------------------------------------------------------------
# 5. Guns
# ------------------------------------------------------------

guns_drop <- c(
  "gun_sales",
  "guns_1",
  "guns_2",
  "guns_3",
  "guns_4",
  "guns_5",
  "guns_6"
)


# ------------------------------------------------------------
# 6. LGBT policy
# ------------------------------------------------------------

lgbt_policy_drop <- c(
  "lgbt_employers",
  "lgbt_medical",
  "lgbt_business"
)


# ------------------------------------------------------------
# 7. Immigration policy
# ------------------------------------------------------------

immigration_drop <- c(
  "increase_deportations",
  "increase_border_security",
  "accept_refugees",
  "easier_family_sponsorship",
  "allow_daca_stay",
  "legal_status_undocumented"
)


# ------------------------------------------------------------
# 8. Regulation items
# ------------------------------------------------------------

regulation_drop <- c(
  "reg_banks",
  "reg_socmedia",
  "reg_six_fig",
  "reg_biz"
)


# ------------------------------------------------------------
# 9. Who respondent knows
# ------------------------------------------------------------

known_people_drop <- unique(c(
  "no_known_gay",
  "gay_friend",
  "gay_fam",
  "no_known_trans",
  "trans_friend",
  "trans_fam"
))
```

```{r}
vars_to_drop <- c(
  deserving_block_drop,
  #deserving_other_drop,
  abortion_drop,
  inequality_drop,
  guns_drop,
  lgbt_policy_drop,
  immigration_drop,
  regulation_drop,
  known_people_drop
)


```

```{r}
data_full <- data_full %>%
  select(-any_of(vars_to_drop))

df_no_control <- df_no_control %>%
  select(-any_of(vars_to_drop))
```


## SAVE

#### WID: cleaned_control_data.csv

save the csv of df of only non missing ("full")

AKA data_full

```{r}

write_csv(data_full, "data/cleaned_control_data.csv")

nrow(data_full)
```

#### WIDOUT: cleaned_control_data.csv

save the csv of those not in the above

AKA df_no_control

```{r}
write_csv(df_no_control, "data/cleaned_sans_control_data.csv")
```

## states

```{r}

data_full %>%
  group_by(state) %>%
  summarise(
    n = n(),
    incarc = sum(incarc, na.rm = TRUE),
    knowincarc = sum(knowincarc, na.rm = TRUE),
    vict = sum(vict, na.rm = TRUE)
  )

data_full %>%
  group_by(state) %>%
  summarise(
    n = n(),
    impact = sum(incarc, na.rm = TRUE)
  )

```
