---
title: "Describe_Sample"
output: html_notebook
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(tidyr)
library(readr)
library(janitor)   # for clean_names()
library(tibble)
library(stringr)
library(haven)
library(dplyr)
library(tidyr)
library(readr)
library(janitor)   # for clean_names()
library(tibble)
library(stringr)
library(haven)
library(tibble)
library(gt)
library(psych)
library(ggplot2)
library(kableExtra)
library(usmap)
library(scales)

source("variable_labels.R")
```

Read in data
```{r}
data_full <- read_csv("data/cleaned_control_data.csv")

not_full <- read_csv("data/cleaned_sans_control_data.csv")


```



# quality of samples

```{r}

summary(data_full$duration_minutes)

```

# who is my sample
## gender, sexual ID

```{r}
table(data_full$trans)

view(data_full %>%
       group_by(trans)%>%
  summarise(
    n = n(),
    man = sum(man, na.rm = TRUE), 
    woman = sum(woman, na.rm = TRUE),
    gqueer       = sum(gqueer, na.rm = TRUE),
    g_other = sum(g_other, na.rm = TRUE)
  ))

view(data_full %>%
  group_by(man) %>%
  summarise(
    n = n(),
    White = sum(White, na.rm = TRUE), 
    Black = sum(Black, na.rm = TRUE),
    Hisp_Latino       = sum(Hisp_Latino, na.rm = TRUE),
    Asian_Am     = sum(Asian_Am, na.rm = TRUE),
    Native_am = sum(Native_Am, na.rm = TRUE),
    OtherRace = sum(OtherRace, na.rm = TRUE)
  ))

view(data_full %>%
  group_by(woman) %>%
  summarise(
    n = n(),
    White = sum(White, na.rm = TRUE), 
    Black = sum(Black, na.rm = TRUE),
    Hisp_Latino       = sum(Hisp_Latino, na.rm = TRUE),
    Asian_Am     = sum(Asian_Am, na.rm = TRUE),
    Native_am = sum(Native_Am, na.rm = TRUE),
    OtherRace = sum(OtherRace, na.rm = TRUE)
  ))

view(data_full %>%
  group_by(gqueer) %>%
  summarise(
    n = n(),
    White = sum(White, na.rm = TRUE), 
    Black = sum(Black, na.rm = TRUE),
    Hisp_Latino       = sum(Hisp_Latino, na.rm = TRUE),
    Asian_Am     = sum(Asian_Am, na.rm = TRUE), 
    Native_am = sum(Native_Am, na.rm = TRUE),
    OtherRace = sum(OtherRace, na.rm = TRUE)
  ))


view(data_full %>%
  group_by(g_other) %>%
  summarise(
    n = n(),
    White = sum(White, na.rm = TRUE), 
    Black = sum(Black, na.rm = TRUE),
    Hisp_Latino       = sum(Hisp_Latino, na.rm = TRUE),
    Asian_Am     = sum(Asian_Am, na.rm = TRUE),
    Native_am = sum(Native_Am, na.rm = TRUE),
    OtherRace = sum(OtherRace, na.rm = TRUE)
  ))
```

*don't really have gender data for those excluded

## state, county

```{r}
state_percent_2020 <- read_csv("state_percent_2020.csv")

state_percent_2020 <- state_percent_2020 %>%
  mutate(
    percent_2020 = readr::parse_number(oct)
  )

#view(state_percent_2020)

where <- data_full %>%
  filter(!is.na(state)) %>%
  group_by(state) %>%
  summarise(
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(
    percent = 100 * n / nrow(data_full)
  ) %>%
  left_join(
    state_percent_2020,
    by = "state"
  ) %>%
    mutate(
    delta = (percent - percent_2020)
  ) %>%
  arrange(desc(n))

#View(where)

where_appendix <- where %>%
  select(state, n, percent, percent_2020, delta) %>%
  mutate(
    percent = round(percent, 2),
    percent_2020 = round(percent_2020, 2),
    delta = round(delta, 2)
)
    

#View(where_appendix)   

state_table <- where_appendix %>%
  kable(
    format = "latex",
    booktabs = TRUE,
    longtable = TRUE,
    caption = "Geographic Distribution of Survey Respondents Compared with the 2020 U.S. Population",
    label = "tab:state_distribution",
    col.names = c(
      "State",
      "N",
      "Sample (\\%)",
      "2020 U.S. Population (\\%)",
      "$\\Delta$ (p.p.)"
    ),
    align = c("l", "r", "r", "r", "r"),
    escape = FALSE
  ) %>%
  kable_styling(
    latex_options = c("repeat_header"),
    font_size = 9
  )

#state_table

writeLines(
  as.character(state_table),
  "state_distribution_table.tex"
)
```

Map?

```{r}
state_map_data <- where %>%
  select(state, percent)

sample_map<-plot_usmap(data = state_map_data, values = "percent", color = "white", linewidth = 0.2) +
  scale_fill_continuous(
    name = "Sample (%)",
    labels = label_number(accuracy = 0.01, suffix = "%")
  ) +
  labs(
    title = "Geographic Distribution of Respondents",
    subtitle = "Share of the full sample residing in each state"
  ) +
  theme(
    legend.position = "right"
  )
sample_map

```

```{r}
ggsave("sample_state_map.pdf", sample_map, width = 9, height = 6)
```


Map of difs

```{r}

delta_map_data <- where %>%
  select(state, delta)

delta_map <- plot_usmap(data = delta_map_data, values = "delta", color = "white", linewidth = 0.2) +
  scale_fill_gradient2(
    name = expression(Delta ~ "(p.p.)"),
    low = "steelblue",
    mid = "white",
    high = "firebrick",
    midpoint = 0,
    labels = label_number(accuracy = 0.01)
  ) +
  labs(
    title = "State-Level Over- and Underrepresentation in the Sample",
    subtitle = "Positive values indicate a larger sample share than the state's 2020 share of the U.S. population"
  ) +
  theme(
    legend.position = "right"
  )
delta_map
```

```{r}
ggsave("delta_map.pdf", delta_map, width = 9, height = 6)
```

## party ID, pol ideology

```{r}

table(data_full$pid_3)

table(data_full$pid_other)

parties <- data_full %>%
  filter(!is.na(pid_3)) %>%
  group_by(pid_3) %>%
  summarise(
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(
    percent = 100 * n / nrow(data_full)
  )%>%
  arrange(desc(n))


parties <- data_full %>%
  mutate(
    pid_table = case_when(
      pid_other == 1 ~ "Other",
      pid_3 == 1 ~ "Democrat",
      pid_3 == 2 ~ "Independent",
      pid_3 == 3 ~ "Republican",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(pid_table)) %>%
  count(pid_table, name = "n") %>%
  mutate(
    percent = 100 * n / sum(n)
  ) %>%
  arrange(desc(n))

parties



```
Are difs from 'actual' or other estimates of actual meaningful? 
```{r}
party_counts <- c(
  Democrat = 620,
  Independent = 510,
  Republican = 395
)

gallup_2024 <- c(
  Democrat = .28,
  Independent = .43,
  Republican = .28
)

# Because the published rounded percentages sum to .99
gallup_2024 <- gallup_2024 / sum(gallup_2024)

chisq.test(
  x = party_counts,
  p = gallup_2024
)
```



```{r}
pid_table <- parties %>%
  kable(
    format = "latex",
    booktabs = TRUE,
    longtable = TRUE,
    caption = "Party Identification Across Sample",
    label = "tab:pid_distribution",
    col.names = c(
      "Party",
      "N",
      "Sample (\\%)"
    ),
    align = c("l", "r", "r"),
    escape = FALSE
  ) %>%
  kable_styling(
    latex_options = c("repeat_header"),
    font_size = 9
  )

View(pid_table)

writeLines(
  as.character(pid_table),
  "pid_table_1.tex"
)

```



```{r}

table(data_full$pid_3)

```

```{r}
table(data_full$self_ideo)

self_ideoS <- data_full %>%
  filter(!is.na(self_ideo)) %>%
  group_by(self_ideo) %>%
  summarise(
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(
    percent = 100 * n / nrow(data_full)
  )%>%
  arrange(desc(n))

self_ideoS



self_ideoS <- data_full %>%
  mutate(
    self_ideo_table = case_when(
      self_ideo == 1 ~ "Very Liberal",
      self_ideo == 2 ~ "Somewhat Liberal",
      self_ideo == 3 ~ "Moderate",
      self_ideo == 4 ~ "Somewhat Conservative",
      self_ideo == 5 ~ "Very Conservative",
      TRUE ~ NA_character_
    )
  ) %>% filter(!is.na(self_ideo_table)) %>%
  count(self_ideo_table, name = "n") %>% 
  mutate(
    percent = 100 * n / sum(n)
    )
                


self_ideoS


ideo_table <- self_ideoS %>%
  kable(
    format = "latex",
    booktabs = TRUE,
    longtable = TRUE,
    caption = "Self-Identified Pol. Ideology",
    label = "tab:self_ideo_distribution",
    col.names = c(
      "Ideo.",
      "N",
      "Sample (\\%)"
    ),
    align = c("l", "r", "r"),
    escape = FALSE
  ) %>%
  kable_styling(
    latex_options = c("repeat_header"),
    font_size = 9
  )

View(pid_table)

writeLines(
  as.character(ideo_table),
  "self_ideo_table.tex"
)
```
Compared to Gallup 2024
```{r}
sum(data_full$self_ideo == 1)

Liberal = (sum(data_full$self_ideo == 1)) + (sum(data_full$self_ideo == 2))
  
Moderate = sum(data_full$self_ideo == 3)
  
Conservative = (sum(data_full$self_ideo == 4)) + (sum(data_full$self_ideo == 5))

Liberal
Conservative


l_r_c <- data_full %>%
  mutate(
    l_r_c_table = case_when(
      self_ideo == 1 | self_ideo == 2 ~ "Liberal",
      self_ideo == 3 ~ "Moderate",
      self_ideo == 4 | self_ideo == 5 ~ "Conservative",
      TRUE ~ NA_character_
    )
  ) %>% filter(!is.na(l_r_c_table)) %>%
  count(l_r_c_table, name = "n") %>% 
  mutate(
    percent = 100 * n / sum(n)
    )

l_r_c
```
gallup liberal = 9% plus 17 % is 26 %

gallup conservative = 10% plus 27% is 37%

moderate 34%

```{r}
ideo_counts <- c(
  Very_Liberal = sum(data_full$self_ideo == 1),
  Somewhat_Liberal = sum(data_full$self_ideo == 2),
  Moderate = sum(data_full$self_ideo == 3),
  Somewhat_Conservative = sum(data_full$self_ideo == 4),
  Very_Conservative = sum(data_full$self_ideo == 5)
)



gallup_2024_ideo <- c(
  Very_Liberal = .09,
  Somewhat_Liberal = .17,
  Moderate = .34,
  Somewhat_Conservative = .27,
  Very_Conservative = .01
)

# Because the published rounded percentages sum to .99
gallup_2024_ideo <- gallup_2024_ideo / sum(gallup_2024_ideo)



chisq.test(
  x = ideo_counts,
  p = gallup_2024_ideo
)

```


```{r}

pol_tude_vars <- c(
  "fiscal_ideology",
  "social_ideology",
  "self_ideo"
)

pol_tude <- data_full %>%
  dplyr::select(dplyr::all_of(pol_tude_vars))

cor(
  pol_tude,
  use = "pairwise.complete.obs"
)


print(psych::corr.test(pol_tude))
```

Ideo Bandits

How many respondents identify as conservative but then do not identify as fiscally conservative or socially conservative
```{r}
library(dplyr)
library(tidyr)

ideo_by_party <- data_full %>%
  
  # ------------------------------------------
  # 1. Create four-category party ID
  # ------------------------------------------
  mutate(
    party4 = case_when(
      pid_other == 1 ~ "Other",
      pid_3 == 1 ~ "Democrat",
      pid_3 == 2 ~ "Independent",
      pid_3 == 3 ~ "Republican",
      TRUE ~ NA_character_
    ),
    
    party4 = factor(
      party4,
      levels = c(
        "Democrat",
        "Independent",
        "Republican",
        "Other"
      )
    )
  ) %>%
  
  # ------------------------------------------
  # 2. Keep ideology variables
  # ------------------------------------------
  dplyr::select(
    party4,
    self_ideo,
    fiscal_ideology,
    social_ideology
  ) %>%
  
  # ------------------------------------------
  # 3. Put ideology measures into long form
  # ------------------------------------------
  pivot_longer(
    cols = c(
      self_ideo,
      fiscal_ideology,
      social_ideology
    ),
    names_to = "ideology_type",
    values_to = "ideology_value"
  )
```



labels
```{r}
ideo_by_party <- ideo_by_party %>%
  mutate(
    
    ideology_label = case_when(
      
      # General ideology: 1-5 scale
      ideology_type == "self_ideo" & ideology_value == 1 ~ "Very Liberal",
      ideology_type == "self_ideo" & ideology_value == 2 ~ "Somewhat Liberal",
      ideology_type == "self_ideo" & ideology_value == 3 ~ "Moderate",
      ideology_type == "self_ideo" & ideology_value == 4 ~ "Somewhat Conservative",
      ideology_type == "self_ideo" & ideology_value == 5 ~ "Very Conservative",
      
      # Fiscal/social ideology: 0-5 scale
      ideology_type %in% c("fiscal_ideology", "social_ideology") &
        ideology_value == 0 ~ "Very Liberal",
      
      ideology_type %in% c("fiscal_ideology", "social_ideology") &
        ideology_value == 1 ~ "Liberal",
      
      ideology_type %in% c("fiscal_ideology", "social_ideology") &
        ideology_value == 2 ~ "Somewhat Liberal",
      
      ideology_type %in% c("fiscal_ideology", "social_ideology") &
        ideology_value == 3 ~ "Somewhat Conservative",
      
      ideology_type %in% c("fiscal_ideology", "social_ideology") &
        ideology_value == 4 ~ "Conservative",
      
      ideology_type %in% c("fiscal_ideology", "social_ideology") &
        ideology_value == 5 ~ "Very Conservative",
      
      TRUE ~ NA_character_
    ),
    
    ideology_type = recode(
      ideology_type,
      self_ideo = "General Ideology",
      fiscal_ideology = "Fiscal Ideology",
      social_ideology = "Social Ideology"
    )
  )
```

```{r}
ideo_party_table <- ideo_by_party %>%
  filter(
    !is.na(party4),
    !is.na(ideology_label)
  ) %>%
  count(
    party4,
    ideology_type,
    ideology_label,
    name = "n"
  ) %>%
  group_by(
    party4,
    ideology_type
  ) %>%
  mutate(
    percent = 100 * n / sum(n)
  ) %>%
  ungroup()
```

```{r}
View(ideo_party_table)

```

Turning above into 4 separate tables (more legible for appendix)

```{r}
ideo_party_table <- ideo_party_table %>%
  mutate(
    ideology_label = factor(
      ideology_label,
      levels = c(
        "Very Liberal",
        "Liberal",
        "Somewhat Liberal",
        "Moderate",
        "Somewhat Conservative",
        "Conservative",
        "Very Conservative"
      )
    )
  )

make_ideo_table <- function(data, party_name) {
  
  data %>%
    dplyr::filter(party4 == party_name) %>%
    dplyr::arrange(ideology_type, ideology_label) %>%
    dplyr::mutate(
      `N (%)` = sprintf("%d (%.1f\\%%)", n, percent)
    ) %>%
    dplyr::select(
      `Ideology Measure` = ideology_type,
      `Ideological Placement` = ideology_label,
      `N (%)`
    ) %>%
    kable(
      format = "latex",
      booktabs = TRUE,
      longtable = TRUE,
      caption = paste(
        "Ideological Self-Placement Among",
        party_name,
        "Respondents"
      ),
      escape = FALSE,
      align = c("l", "l", "r")
    ) %>%
    kable_styling(
      latex_options = "repeat_header",
      font_size = 9
    )
}

dem_ideo_table <- make_ideo_table(
  ideo_party_table,
  "Democrat"
)

rep_ideo_table <- make_ideo_table(
  ideo_party_table,
  "Republican"
)

ind_ideo_table <- make_ideo_table(
  ideo_party_table,
  "Independent"
)

other_ideo_table <- make_ideo_table(
  ideo_party_table,
  "Other"
)

writeLines(
  as.character(dem_ideo_table),
  "democrat_ideology_table.tex"
)

writeLines(
  as.character(rep_ideo_table),
  "republican_ideology_table.tex"
)

writeLines(
  as.character(ind_ideo_table),
  "independent_ideology_table.tex"
)

writeLines(
  as.character(other_ideo_table),
  "other_ideology_table.tex"
)
```


## race

```{r}
all_race_vars <- c(
  "White",
  "Black",
  "Hisp_Latino",
  "Asian_Am",
  "Native_Am",
  "Pac_Island",
  "MENA",
  "race_notlisted",
  "multiracial"
)

data_full %>%
  dplyr::select(all_of(all_race_vars)) %>%
  pivot_longer(everything(), names_to = "var", values_to = "val") %>%
  count(var, val) %>%
  pivot_wider(names_from = val, values_from = n, values_fill = 0)
```
```{r}

rid_2020 <- read_csv("race_vs_2020.csv")

view(rid_2020)

rid_2020_table <- rid_2020 %>%
  dplyr::mutate(
    race_ethnicity = dplyr::recode(
      rid,
      "Asian_Am"    = "Asian",
      "Black"       = "Black or African American",
      "Hisp_Latino" = "Hispanic or Latino",
      "MENA"        = "Middle Eastern or North African",
      "Native_Am"   = "American Indian or Alaska Native",
      "Pac_Island"  = "Native Hawaiian or Other Pacific Islander",
      "White"       = "White"
    ),

    # pct_sample is currently stored as a proportion
    sample_pct = sprintf("%.1f\\%%", pct_sample * 100),

    # 2020_pct is already stored as a percent, e.g. 7.2
    census_pct = sprintf("%.1f\\%%", `2020_pct`),

    # Format Census population in millions
    census_n = sprintf("%.1f", `2020_n_million`)
  ) %>%
  dplyr::select(
    race_ethnicity,
    n,
    sample_pct,
    census_n,
    census_pct
  )


rid_table_appendix <- rid_2020_table %>%
  kable(
    format = "latex",
    booktabs = TRUE,
    escape = FALSE,
    align = c("l", "r", "r", "r", "r"),
    col.names = c(
      "Race/Ethnicity",
      "Sample $N$",
      "Sample \\%",
      "2020 U.S. Population (millions)",
      "2020 Census \\%"
    ),
    caption = "Racial and Ethnic Composition of the Survey Sample and 2020 U.S. Population",
    label = "tab:race_census_comparison"
  ) %>%
  kable_styling(
    latex_options = "hold_position"
  ) %>%
  footnote(
    general = paste(
      "Sample racial categories are non-mutually exclusive; respondents could identify with more than one race.",
      "Accordingly, percentages need not sum to 100 percent.",
      "Census race estimates report each racial group alone or in combination with one or more other races.",
      "Hispanic or Latino origin is measured by the Census as an ethnicity and includes persons of any race.",
      "In the 2020 Census, Middle Eastern or North African (MENA) was not a separate race category and was formally classified within White;",
      "the MENA estimate shown here uses detailed Census responses identifying MENA origin and therefore overlaps with the Census White category.",
      "Census benchmarks refer to the total 2020 U.S. population."
    ),
    general_title = "Note:",
    threeparttable = TRUE,
    escape = FALSE
  )


writeLines(
  as.character(rid_table_appendix),
  "rid_2020_table_2.tex"
)

```





```{r}
# combinations
combo_counts <- data_full %>%
  dplyr::mutate(.id = row_number()) %>%
  dplyr::select(.id, all_of(all_race_vars)) %>%
  pivot_longer(-.id, names_to = "race", values_to = "selected") %>%
  filter(selected == 1) %>%                      # drops both 0s and NAs
  mutate(race = factor(race, levels = all_race_vars)) %>%
  arrange(.id, race) %>%
  group_by(.id) %>%
  summarise(
    n_races    = n(),
    race_combo = paste(race, collapse = " + "),
    .groups = "drop"
  ) %>%
  count(n_races, race_combo, name = "n") %>%
  arrange(desc(n))

View(combo_counts)                                # everything

filter(combo_counts, n_races >= 2)          # multiracial combos only


```


```{r}
view(data_full %>%
  group_by(state) %>%
  summarise(
    n = n(),
    White = sum(White, na.rm = TRUE), 
    Black = sum(Black, na.rm = TRUE),
    Hisp_Latino       = sum(Hisp_Latino, na.rm = TRUE),
    Asian_Am     = sum(Asian_Am, na.rm = TRUE),
    Native_am = sum(Native_Am, na.rm = TRUE),
    OtherRace = sum(OtherRace, na.rm = TRUE)
  ))

```

### Multi-Racial Whites? 
```{r}
table(data_full$AllOthers)

view(data_full) %>%
  group_by(White) %>% 
  summarise (
    n = n(), 
    Black = sum(Black, na.rm = TRUE),
    Hisp_Latino       = sum(Hisp_Latino, na.rm = TRUE),
    AllOthers = sum(AllOthers, na.rm = TRUE)
  )
```
```{r}

view(data_full) %>%
  group_by(Black) %>% 
  summarise (
    n = n(), 
    White = sum(White, na.rm = TRUE),
    Hisp_Latino       = sum(Hisp_Latino, na.rm = TRUE),
    AllOthers = sum(AllOthers, na.rm = TRUE)
  )
```
```{r}

view(data_full) %>%
  group_by(Hisp_Latino) %>% 
  summarise (
    n = n(), 
    White = sum(White, na.rm = TRUE),
    Black       = sum(Black, na.rm = TRUE),
    AllOthers = sum(AllOthers, na.rm = TRUE)
  )

```

```{r}
view(data_full) %>%
  group_by(AllOthers) %>% 
  summarise (
    n = n(), 
    White = sum(White, na.rm = TRUE),
    Black       = sum(Black, na.rm = TRUE),
    Hisp_Latino = sum(Hisp_Latino, na.rm = TRUE)
  )
```



# Experiences

## CJ

```{r}

table(data_full$vict_bin, useNA = "ifany")

table(data_full$vict, useNA = "ifany")



data_full %>%
  summarise(
    all_three_n = sum(
      incarc == 1 & knowincarc == 1 & vict == 1,
      na.rm = TRUE
    )
  )


```

```{r}
data_full %>%
  summarise(
    incarc_ki_n = sum(
      incarc == 1 & knowincarc == 1,
      na.rm = TRUE
    )
  )
```
```{r}
data_full %>%
  summarise(
    incarc_vict_n = sum(
      incarc == 1 & vict == 1,
      na.rm = TRUE
    )
  )
```



```{r}
data_full %>%
  group_by(knowincarc) %>%
  summarise(
    n = n(),
    incarc = sum(incarc, na.rm = TRUE),
    vict = sum(vict, na.rm = TRUE)
  )
```



```{r}
table(not_full$vict, useNA = "ifany")
```




```{r}
table(data_full$incarc, useNA = "ifany")
table(data_full$incarc_bin, useNA = "ifany")

table(not_full$incarc, useNA = "ifany")
```


```{r}
table(data_full$knowincarc, useNA = "ifany")
table(data_full$knowincarc_bin, useNA = "ifany")

table(not_full$knowincarc, useNA = "ifany")
```
```{r}

table(data_full$impact, useNA = "ifany")

table(not_full$impact, useNA = "ifany")
```

```{r}
data_full %>%
  group_by(impact)%>%
  summarise(
    n=n(),
    total = 1600,
    incarc = sum(incarc, na.rm = TRUE),
    incarc_pct = ((incarc / total) *100),
    knowincarc = sum(knowincarc, na.rm = TRUE),
    knowincarc_pct = ((knowincarc / total) *100),
    vict = sum(vict, na.rm = TRUE),
    vict_pct = ((vict / total) *100),
    any_impact = sum(impact, na.rm = TRUE),
    any_impact_pct = ((any_impact / total) *100)
  )
```



```{r}

data_full %>%
  group_by(impact)%>%
  summarise(
    n=n(),
    incarc = sum(incarc_bin, na.rm = TRUE),
    knowincarc = sum(knowincarc_bin, na.rm = TRUE),
    vict = sum(vict_bin, na.rm = TRUE)
  )

```


## df full race impact
```{r}
data_full %>%
  group_by(incarc) %>%
  summarise(
    n = n(),
    incarc_total = sum(incarc, na.rm = TRUE),
    White = sum(White, na.rm = TRUE), 
    pct_white = ((White / incarc_total) *100),
    Black = sum(Black, na.rm = TRUE),
    pct_black = ((Black / incarc_total) *100),
    Hisp_Latino       = sum(Hisp_Latino, na.rm = TRUE),
    pct_hi = ((Hisp_Latino / incarc_total) *100),
    Asian_Am     = sum(Asian_Am, na.rm = TRUE),
    pct_aa = ((Asian_Am / incarc_total) *100),
    Native_am = sum(Native_Am, na.rm = TRUE),
    pct_nat = ((Native_am / incarc_total) *100),
    OtherRace = sum(OtherRace, na.rm = TRUE),
    pct_other = ((OtherRace / incarc_total) *100)
  )

```


```{r}
data_full %>%
  group_by(impact) %>%
  summarise(
    n = n(),
    White = sum(White, na.rm = TRUE), 
    Black = sum(Black, na.rm = TRUE),
    Hisp_Latino       = sum(Hisp_Latino, na.rm = TRUE),
    Asian_Am     = sum(Asian_Am, na.rm = TRUE),
    Native_am = sum(Native_Am, na.rm = TRUE),
    OtherRace = sum(OtherRace, na.rm = TRUE)
  )
```


```{r}
data_full %>%
  group_by(vict) %>%
  summarise(
    n = n(),
    White = sum(White, na.rm = TRUE), 
    Black = sum(Black, na.rm = TRUE),
    Hisp_Latino       = sum(Hisp_Latino, na.rm = TRUE),
    Asian_Am     = sum(Asian_Am, na.rm = TRUE),
    Native_am = sum(Native_Am, na.rm = TRUE),
    OtherRace = sum(OtherRace, na.rm = TRUE)
  )
```

### Impact by Race
of being incarcerated, knowincarc, vict, impact by RID

```{r}

#-----------------------------------
# Incarceration by race
#-----------------------------------

race_incarc <- data_full %>%
  summarise(
    # Total respondents in each race category
    aa_total    = sum(Asian_Am == 1, na.rm = TRUE),
    black_total = sum(Black == 1, na.rm = TRUE),
    hisp_total  = sum(Hisp_Latino == 1, na.rm = TRUE),
    mena_total  = sum(MENA == 1, na.rm = TRUE),
    na_total    = sum(Native_Am == 1, na.rm = TRUE),
    pi_total    = sum(Pac_Island == 1, na.rm = TRUE),
    white_total = sum(White == 1, na.rm = TRUE),
    ro_total    = sum(OtherRace == 1, na.rm = TRUE),

    # Number within each race category who have been incarcerated
    n_inc_aa    = sum(Asian_Am == 1 & incarc == 1, na.rm = TRUE),
    n_inc_black = sum(Black == 1 & incarc == 1, na.rm = TRUE),
    n_inc_hisp  = sum(Hisp_Latino == 1 & incarc == 1, na.rm = TRUE),
    n_inc_mena  = sum(MENA == 1 & incarc == 1, na.rm = TRUE),
    n_inc_na    = sum(Native_Am == 1 & incarc == 1, na.rm = TRUE),
    n_inc_pi    = sum(Pac_Island == 1 & incarc == 1, na.rm = TRUE),
    n_inc_white = sum(White == 1 & incarc == 1, na.rm = TRUE),
    n_inc_ro    = sum(OtherRace == 1 & incarc == 1, na.rm = TRUE)
  ) %>%
  mutate(
    pct_inc_aa    = 100 * n_inc_aa / aa_total,
    pct_inc_black = 100 * n_inc_black / black_total,
    pct_inc_hisp  = 100 * n_inc_hisp / hisp_total,
    pct_inc_mena  = 100 * n_inc_mena / mena_total,
    pct_inc_na    = 100 * n_inc_na / na_total,
    pct_inc_pi    = 100 * n_inc_pi / pi_total,
    pct_inc_white = 100 * n_inc_white / white_total,
    pct_inc_ro    = 100 * n_inc_ro / ro_total
  )

View(race_incarc)

incarc_race_table <- tibble(
  Race = c(
    "Asian American",
    "Black",
    "Hispanic/Latino",
    "Middle Eastern/North African",
    "Native American",
    "Pacific Islander",
    "White",
    "Other Race"
  ),
  
  Total_N = c(
    race_incarc$aa_total,
    race_incarc$black_total,
    race_incarc$hisp_total,
    race_incarc$mena_total,
    race_incarc$na_total,
    race_incarc$pi_total,
    race_incarc$white_total,
    race_incarc$ro_total
  ),
  
  Incarcerated_N = c(
    race_incarc$n_inc_aa,
    race_incarc$n_inc_black,
    race_incarc$n_inc_hisp,
    race_incarc$n_inc_mena,
    race_incarc$n_inc_na,
    race_incarc$n_inc_pi,
    race_incarc$n_inc_white,
    race_incarc$n_inc_ro
  ),
  
  Percent_Incarcerated = c(
    race_incarc$pct_inc_aa,
    race_incarc$pct_inc_black,
    race_incarc$pct_inc_hisp,
    race_incarc$pct_inc_mena,
    race_incarc$pct_inc_na,
    race_incarc$pct_inc_pi,
    race_incarc$pct_inc_white,
    race_incarc$pct_inc_ro
  )
) %>%
  mutate(
    Percent_Incarcerated = round(Percent_Incarcerated, 2)
  )

print(incarc_race_table)

```
for overleaf
```{r}

inc_race_pcts_table <- incarc_race_table %>%
  knitr::kable(
    format = "latex",
    booktabs = TRUE,
    align = c("l", "r", "r", "r"),
    col.names = c(
      "Race",
      "Total $N$",
      "Incarcerated $N$",
      "\\% Incarcerated"
    ),
    caption = "Incarceration Experience by Race"
  )


writeLines(
  as.character(inc_race_pcts_table),
  "inc_race_pcts_table.tex"
)

```



### (e) Table of Impact x Race

```{r}
library(dplyr)
library(tidyr)
library(purrr)
library(stringr)
source("variable_labels_sept1.R")

total_n <- nrow(data_full)

#----------------------------
# Impact x race table
#----------------------------

impact_race_table <- map_dfr(impact_vars, function(iv) {
  
  row_dat <- data_full %>%
    filter(.data[[iv]] == 1)
  
  tibble(
    Impact = impact_labels[[iv]],
    Total_n = nrow(row_dat),
    Total_pct = 100 * nrow(row_dat) / total_n
  ) %>%
    bind_cols(
      map_dfc(race_vars, function(rv) {
        
        n_cell <- sum(row_dat[[rv]] == 1, na.rm = TRUE)
        pct_cell <- 100 * n_cell / total_n
        
        tibble(
          !!race_labels[[rv]] :=
            sprintf("%d (%.1f%%)", n_cell, pct_cell)
        )
      })
    ) %>%
    mutate(
      Total = sprintf("%d (%.1f%%)", Total_n, Total_pct)
    ) %>%
    dplyr::select(
      Impact,
      all_of(unname(race_labels)),
      Total
    )
})


#----------------------------
# Overall sample row
#----------------------------

overall_row <- tibble(
  Impact = "Overall sample"
) %>%
  bind_cols(
    map_dfc(race_vars, function(rv) {
      
      n_cell <- sum(data_full[[rv]] == 1, na.rm = TRUE)
      pct_cell <- 100 * n_cell / total_n
      
      tibble(
        !!race_labels[[rv]] :=
          sprintf("%d (%.1f%%)", n_cell, pct_cell)
      )
    })
  ) %>%
  mutate(
    Total = sprintf("%d (100.0%%)", total_n)
  )


# Add overall row
impact_race_table <- bind_rows(
  impact_race_table,
  overall_row
)

print(impact_race_table)
```

Bar plots of impact x race

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(forcats)

#----------------------------
# 1. Reshape race indicators long
#----------------------------
race_impact_long <- data_full %>%
  dplyr::select(all_of(race_vars), all_of(impact_vars)) %>%
  pivot_longer(
    cols = all_of(race_vars),
    names_to = "race_var",
    values_to = "in_race"
  ) %>%
  filter(in_race == 1) %>%
  mutate(
    race = recode(race_var, !!!race_labels)
  )

#----------------------------
# 2. Plot ANY impact by race
#----------------------------
impact_by_race <- race_impact_long %>%
  group_by(race) %>%
  summarise(
    pct_impacted = mean(impact, na.rm = TRUE),
    n_race = n(),
    .groups = "drop"
  )

ggplot(impact_by_race, aes(x = fct_reorder(race, pct_impacted), y = pct_impacted)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = NULL,
    y = "Share with any impact",
    title = "Any carceral contact"
  ) +
  coord_flip() +
  theme_minimal()
```

```{r}
impact_by_race_type <- race_impact_long %>%
  pivot_longer(
    cols = all_of(impact_vars),
    names_to = "impact_var",
    values_to = "has_impact"
  ) %>%
  mutate(
    impact_type = recode(impact_var, !!!impact_labels)
  ) %>%
  group_by(race, impact_type) %>%
  summarise(
    pct = mean(has_impact, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(impact_by_race_type, aes(x = race, y = pct, fill = impact_type)) +
  geom_col(position = position_dodge(width = 0.8)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = NULL,
    y = "Share within racial group",
    fill = NULL,
    title = "Sample Carceral Contact by Race"
  ) +
  theme_minimal()

```

Turning above into 4 tables

```{r}
library(dplyr)
library(purrr)
library(tibble)

# race vars and labels
race_vars <- c("White", "Black", "Hisp_Latino", "OtherRace")

race_labels <- c(
  White = "White",
  Black = "Black",
  Hisp_Latino = "Hisp_Latino",
  OtherRace = "Other Race"
)

impact_vars <- c("impact", "incarc", "knowincarc", "vict")

impact_labels <- c(
  impact = "Any Impact",
  incarc = "Direct",
  knowincarc = "Proximal",
  vict = "Victim"
)

# function: one table for one racial group
make_race_impact_table <- function(race_var) {
  
  dat_race <- data_full %>%
    dplyr::filter(.data[[race_var]] == 1)
  
  race_n <- nrow(dat_race)
  
  tibble(
    Impact = unname(impact_labels),
    n = purrr::map_int(impact_vars, ~ sum(dat_race[[.x]] == 1, na.rm = TRUE))
  ) %>%
    dplyr::mutate(
      pct = round(100 * n / race_n, 1),
      `n (%)` = sprintf("%d (%.1f\\%%)", n, pct)
    ) %>%
    dplyr::select(Impact, `n (%)`)
}

# create all four tables as a named list
race_impact_tables <- setNames(
  lapply(race_vars, make_race_impact_table),
  unname(race_labels)
)

# view one example
race_impact_tables[["White"]]
race_impact_tables[["Black"]]
race_impact_tables[["Hispanic"]]
race_impact_tables[["Other Race"]]



latex_tables <- imap(race_impact_tables, ~ knitr::kable(
  .x,
  format = "latex",
  booktabs = TRUE,
  escape = FALSE,
  caption = paste("Carceral contact among", .y, "respondents")
))

cat(latex_tables[["White"]])
cat(latex_tables[["Black"]])
cat(latex_tables[["Hispanic"]])
cat(latex_tables[["Other Race"]])

```

### Proportion Race Reporting Impact

```{r}

experiences <- c(
  "incarc",
  "knowincarc",
  "vict",
  "impact"
)

binary_versions <- c(
  "incarc_bin",
  "knowincarc_bin",
  "vict_bin"
)

# Race vars (in case r script not working)
race_vars <- c(
  "White",
  "Black",
  "Hisp_Latino",
  "Asian_Am",
  "Native_Am",
  "OtherRace"
)

race_labels <- c(
  White       = "White",
  Black       = "Black",
  Hisp_Latino = "Hispanic/Latino",
  Asian_Am    = "Asian American",
  Native_Am   = "Native American",
  OtherRace   = "Other Race"
)


props_data <- c(
  experiences,
  binary_versions,
  race_vars,
  "impact"
)

props_sample <- data_full %>%
  dplyr::select(props_data)

  
names(props_sample)

# rename experience

experience_vars <- c(
  incarc = "Incarcerated",
  knowincarc = "Know Incarcerated",
  vict = "Victimized",
  impact = "Any Impact"
)

# pull

race_experience_plot <- purrr::map_dfr(race_vars, function(rv) {
  
  race_name <- race_labels[[rv]]
  
  # Same racial-group denominator for every experience
  race_dat <- props_sample %>%
    filter(
      .data[[rv]] == 1,
      !is.na(impact)
    )
  
  bind_rows(
    
    # Incarcerated
    race_dat %>%
      summarise(
        Race = race_name,
        Experience = "Incarcerated",
        denominator = n(),
        numerator = sum(incarc == 1, na.rm = TRUE),
        proportion = numerator / denominator
      ),
    
    # Victimized
    race_dat %>%
      summarise(
        Race = race_name,
        Experience = "Victimized",
        denominator = n(),
        numerator = sum(vict == 1, na.rm = TRUE),
        proportion = numerator / denominator
      ),
    
    # Know incarcerated
    race_dat %>%
      summarise(
        Race = race_name,
        Experience = "Know Incarcerated",
        denominator = n(),
        numerator = sum(knowincarc == 1, na.rm = TRUE),
        proportion = numerator / denominator
      ),
    
    # Impact
    race_dat %>%
      summarise(
        Race = race_name,
        Experience = "Any Impact",
        denominator = n(),
        numerator = sum(impact == 1, na.rm = TRUE),
        proportion = numerator / denominator
      )
  )
})

print(n=24, race_experience_plot)

```

```{r}

race_experience_plot <- race_experience_plot %>%
  mutate(
    Race = factor(
      Race,
      levels = c(
        "Black",
        "Hispanic/Latino",
        "Native American",
        "Other Race",
        "Asian American",
        "White"
      )
    ),
    Experience = factor(
      Experience,
      levels = c(
        "Incarcerated",
        "Victimized",
        "Know Incarcerated",
        "Any Impact"
      )
    )
  )


print(n = 25,
race_experience_plot %>%
  dplyr::select(Race, Experience, denominator)
)

```
#### 2 x 2

```{r}

race_experience_figure <- ggplot(
  race_experience_plot,
  aes(
    x = Race,
    y = proportion
  )
) +
  geom_col(width = 0.75) +
  
  # Four panels, arranged 2 x 2
  facet_wrap(
    ~ Experience,
    ncol = 2,
    scales = "free_y"
  ) +
  
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    expand = expansion(mult = c(0, 0.05))
  ) +
  
  labs(
    x = NULL,
    y = "Proportion of Group Reporting Yes"
  ) +
  
  theme_minimal() +
  theme(
    strip.text = element_text(
      face = "bold",
      size = 11
    ),
    axis.text.x = element_text(
      angle = 30,
      hjust = 1
    ),
    panel.grid.major.x = element_blank()
  )

race_experience_figure

```

save for overleaf
```{r}
ggsave(
  filename = "figures/race_experience_figure.pdf",
  plot = race_experience_figure,
  width = 8,
  height = 6,
  units = "in"
)
```





## Rationing
```{r}

cor(data_full$ration_scripts, data_full$ration_food, use = "pairwise.complete.obs")


data_full %>%
  group_by(ration_food) %>% 
  summarise (
    n = n(), 
    White = sum(White, na.rm = TRUE),
    Hisp_Latino       = sum(Hisp_Latino, na.rm = TRUE),
    Black = sum(Black, na.rm = TRUE),
    AllOthers = sum(AllOthers, na.rm = TRUE)
  )

ration_model <- lm(ration_food ~ ration_scripts, data = data_full)

summary(ration_model)
```

## Econ Experiences

```{r}
"uninsured" %in% names(data_full)

names(data_full)
table(data_full$tanf)

table(data_full$snap)

table(data_full$uninsured)

# table(data_full$unionmember)

table(data_full$on_medicaid)


data_full%>%
  summarise(
    n = n(),
    total = 1600,
    tanf = sum(tanf, na.rm = TRUE),
    tanf_pct = ((tanf/total)*100),
    snap = sum(snap, na.rm = TRUE),
    snap_pct = ((snap/total)*100),
    uninsured = sum(uninsured, na.rm = TRUE),
    uninsured_pct = ((uninsured/total)*100),
    unionmember = sum(unionmember, na.rm = TRUE),
    union_pct = ((unionmember/total)*100),
    on_medicaid = sum(on_medicaid, na.rm = TRUE),
    medicaid_pct = ((on_medicaid/total)*100)
  )

View(data_full%>%
       group_by(incarc)%>%
  summarise(
    n = n(),
    total = 1600,
    tanf = sum(tanf, na.rm = TRUE),
    
    snap = sum(snap, na.rm = TRUE),
   
    uninsured = sum(uninsured, na.rm = TRUE),
   
    unionmember = sum(unionmember, na.rm = TRUE),
    
    on_medicaid = sum(on_medicaid, na.rm = TRUE),
    
  ))

```


Econ Experiences Across Impact Groups

```{r}
economic_vars <- c(
  "snap",
  "tanf",
  "medicaid",
  "unemployed",
  "uninsured"
)
# 
# economic_impact_summary <- purrr::map_dfr(
#   economic_vars,
#   function(var) {
#     
#     data_full %>%
#       filter(.data[[var]] == 1) %>%
#       summarise(
#         variable = var,
#         n_group = n(),
#         
#         incarc_n = sum(incarc == 1, na.rm = TRUE),
#         incarc_prop = mean(incarc == 1, na.rm = TRUE),
#         
#         knowincarc_n = sum(knowincarc == 1, na.rm = TRUE),
#         knowincarc_prop = mean(knowincarc == 1, na.rm = TRUE),
#         
#         vict_n = sum(vict == 1, na.rm = TRUE),
#         vict_prop = mean(vict == 1, na.rm = TRUE),
#         
#         impact_n = sum(impact == 1, na.rm = TRUE),
#         impact_prop = mean(impact == 1, na.rm = TRUE),
#         
#         no_impact_n = sum(impact == 0, na.rm = TRUE),
#         no_impact_prop = mean(impact == 0, na.rm = TRUE)
#       )
#   }
# )
# 
# economic_impact_summary

economic_by_impact <- purrr::map_dfr(
  economic_vars,
  function(var) {
    
    data_full %>%
      filter(impact %in% c(0, 1)) %>%
      group_by(impact) %>%
      summarise(
        variable = var,
        
        n_group = n(),
        n_nonmissing = sum(!is.na(.data[[var]])),
        n_yes = sum(.data[[var]] == 1, na.rm = TRUE),
        
        prop_yes = mean(.data[[var]] == 1, na.rm = TRUE),
        
        .groups = "drop"
      )
  }
) %>%
  mutate(
    impact_group = case_when(
      impact == 1 ~ "Any Impact",
      impact == 0 ~ "No Impact"
    )
  )

```


```{r}


economic_impact_plot <- economic_impact_summary %>%
  dplyr::select(
    variable,
    incarc_prop,
    knowincarc_prop,
    vict_prop,
    impact_prop,
    no_impact_prop
  ) %>%
  pivot_longer(
    cols = -variable,
    names_to = "impact_type",
    values_to = "proportion"
  ) %>%
  mutate(
    impact_type = recode(
      impact_type,
      incarc_prop = "Direct",
      knowincarc_prop = "Proximal",
      vict_prop = "Victim",
      impact_prop = "Any Impact",
      no_impact_prop = "No Impact"
    ),
    impact_type = factor(
      impact_type,
      levels = c(
        "Direct",
        "Proximal",
        "Victim",
        "Any Impact",
        "No Impact"
      )
    )
  )

economic_impact_plot 

```

```{r}

make_economic_plot <- function(var_name, plot_title) {
  
  economic_impact_plot %>%
    filter(variable == var_name) %>%
    ggplot(
      aes(
        x = impact_type,
        y = proportion
      )
    ) +
    geom_col() +
    geom_text(
      aes(label = scales::percent(proportion, accuracy = 0.1)),
      vjust = -0.4
    ) +
    scale_y_continuous(
      labels = scales::percent,
      limits = c(0, 0.75)
    ) +
    labs(
      title = plot_title,
      x = NULL,
      y = "Proportion of Group Reporting Yes"
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(
        angle = 30,
        hjust = 1
      )
    )
}

plot_snap <- make_economic_plot(
  "snap",
  "SNAP Recipients"
)

plot_tanf <- make_economic_plot(
  "tanf",
  "TANF Recipients"
)

plot_medicaid <- make_economic_plot(
  "medicaid",
  "Medicaid Recipients"
)

plot_unemployed <- make_economic_plot(
  "unemployed",
  "Unemployed Respondents"
)

plot_uninsured <- make_economic_plot(
  "uninsured",
  "Uninsured Respondents"
)

```

```{r}

plot_snap
plot_uninsured
plot_medicaid
plot_uninsured

plot_tanf

```

```{r}


# Save each figure as a PDF
ggsave(
  filename = "figures/snap_impact_groups_10_5.pdf",
  plot = plot_snap,
  width = 7,
  height = 5,
  units = "in"
)

ggsave(
  filename = "figures/tanf_impact_groups_10_5.pdf",
  plot = plot_snap,
  width = 7,
  height = 5,
  units = "in"
)

ggsave(
  filename = "figures/medicaid_impact_groups_10_5.pdf",
  plot = plot_medicaid,
  width = 7,
  height = 5,
  units = "in"
)

ggsave(
  filename = "figures/unemployed_impact_groups_10_5.pdf",
  plot = plot_unemployed,
  width = 7,
  height = 5,
  units = "in"
)

ggsave(
  filename = "figures/uninsured_impact_groups_10_5.pdf",
  plot = plot_uninsured,
  width = 7,
  height = 5,
  units = "in"
)
```




# FIRE

```{r}
data_full %>%
  group_by(fire_rare) %>% 
  summarise (
    n = n(), 
    White = sum(White, na.rm = TRUE),
    Hisp_Latino       = sum(Hisp_Latino, na.rm = TRUE),
    Black = sum(Black, na.rm = TRUE),
    AllOthers = sum(AllOthers, na.rm = TRUE)
  )

# data_full %>%
#   group_by(fire_rare) %>% 
#   summarise (
#     n = n(), 
#     White = sum(White, na.rm = TRUE),
#     White_pct = ((White / total) *100),
#     Hisp_Latino       = sum(Hisp_Latino, na.rm = TRUE),
#     HL_pct = ((Hisp_Latino / total) *100),
#     Black = sum(Black, na.rm = TRUE),
#     Black_pct = ((Black / total) *100),
#     AllOthers = sum(AllOthers, na.rm = TRUE)
#   )

```

# Trust Gov?

```{r}

m_trust <- lm(gov_attn ~ incarc, data = data_full)

summary(m_trust)

```


# ------------------------------------------------------------
# D Scores
# -----------------------------------------------------------

```{r}



```



## desor core vars 
### -- confirm correlation
 
```{r}

desor_core_vars <- c(
  "d_welfare_tanf_abc","d_fstamps_snap_abc","d_medicaid_abc","d_medicare_abc",
  "d_unemployed_abc","d_poorfam_abc","d_homeless_abc","d_immigrants_abc","d_unauth_undoc_abc"
)


desor_corr <- psych::corr.test(
  core_mat,
  use = "pairwise",
  method = "pearson",
  adjust = "none"
)

# Pairwise correlations
view(round(desor_corr$r, 2))


dcore_data <- sample %>%
  select(all_of(desor_core_vars))


cor(dcore_data,
    use = "pairwise.complete.obs")


print(psych::corr.test(dcore_data))

print(psych::alpha(dcore_data))

```

```{r}
# Eligibility rule for composite
# at least 5 of 9 items observed
core_keep <- core_n >= 5

table(core_keep, useNA = "ifany")


core_alpha <- psych::alpha(
  core_mat[core_keep, ],
  check.keys = TRUE
)

# Key reliability outputs
core_alpha$total$raw_alpha
core_alpha$total$std.alpha
core_alpha$total$average_r
core_alpha$n.obs

# Item diagnostics
core_alpha$alpha.drop
core_alpha$item.stats

# Check whether psych automatically reversed any items
core_alpha$keys
```

