---
title: "Likely Speeders"
output: html_notebook
---


```{r setup, include=TRUE}
knitr::opts_chunk$set(
  echo = FALSE,       # Hide code by default
  warning = FALSE,
  message = FALSE
)

# Load packages
library(dplyr)
library(tidyr)
library(readr)
library(janitor)   # for clean_names()
library(tibble)
library(stringr)
library(haven)
library(ggplot2)
library(ggbeeswarm)
library(readxl)
library(stringr)
library(dplyr)
library(kableExtra)


```

```{r}
df <- read_csv("data/cleaned/likely_speeders.csv")
```

# 5. Description of Consented, pre-screening sample

```{r}
ncol(df)
nrow(df)
count(df, block)


```

## attention checks

```{r}
count(df, attn_pass_count)
count(df, attn_answered_count)
```


table:

```{r}
df %>%
  summarise(
    n = n(),

    q9_pass_n = sum(attn_q9_pass),
    q9_pass_pct = mean(attn_q9_pass) * 100,
    
    attn_q9_typo_n = sum(attn_q9_typo),
    attn_q9_typo_pct = mean(attn_q9_typo) * 100,

    q17_pass_n = sum(attn_check_q17),
    q17_pass_pct = mean(attn_check_q17) * 100,

    q114_pass_n = sum(attn_q114_pass),
    q114_pass_pct = mean(attn_q114_pass) * 100
  )
```



```{r}
attn_check_summary <- tibble(
  check = c("Q9", "Q17", "Q114"),
  
  total_n = nrow(df),
  
  answered_n = c(
    sum(df$attn_q9_answered),
    sum(df$attn_q17_answered),
    sum(df$attn_q114_answered)
  ),
  
  nonanswer_n = c(
    sum(!df$attn_q9_answered),
    sum(!df$attn_q17_answered),
    sum(!df$attn_q114_answered)
  ),
  
  pass_n = c(
    sum(df$attn_q9_pass),
    sum(df$attn_check_q17),
    sum(df$attn_q114_pass)
  )
) %>%
  mutate(
    # Answered, but did not pass
    wrong_n = answered_n - pass_n,
    
    # Everyone who did not pass: wrong answer + nonanswer
    nonpass_n = total_n - pass_n,
    
    answered_pct = answered_n / total_n * 100,
    
    # Your current percentages
    pass_pct_all = pass_n / total_n * 100,
    
    # More useful for actual attention performance:
    # Of people who ANSWERED, how many passed?
    pass_pct_answered = pass_n / answered_n * 100,
    
    # Of all apparent non-passes, what % are just nonanswers?
    nonanswer_pct_of_nonpass =
      nonanswer_n / nonpass_n * 100,
    
    # Of people who answered, what % actually got it wrong?
    wrong_pct_answered =
      wrong_n / answered_n * 100
  )

attn_check_summary
```



\*\*save that table for overleaf.

```{r}
df %>%
  count(attn_pass_count, attn_answered_count)
```

## duration

what does survey duration actually look like?

```{r}
# visual of response times

uncleaned_response_times <-
ggplot(df,
       aes(x = duration_minutes, y = 0)) +
  geom_quasirandom(
    groupOnX = FALSE,
    width = .35,
    alpha = .35,
    size = 1.2
  ) +
  coord_cartesian(xlim = c(0, 120)) +
  labs(
    x = "Survey duration (minutes)",
    y = NULL,
    title = "Survey Duration"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )


uncleaned_response_times # save uncleaned_response_times as image


# ggsave(
#   "uncleaned_response_times.png",
#   plot = uncleaned_response_times,
#   width = 14,
#   height = 9,
#   dpi = 300,
# )

```

```{r}
summary(df$duration_minutes)
```

Took respondents from less than a second to days to "complete" the survey. flags: -sub_8 -sub_12 -sub_20 -sub_30 -over_52 -over_180

```{r}
df %>%
  summarise(
    n = n(),

    sub_8_n = sum(sub_8.45),
    sub_8_pct = mean(sub_8.45) * 100,
    
    sub_12_n = sum(sub_12.6),
    sub_12_pct = mean(sub_12.6) * 100,

    sub_20_n = sum(sub_20),
    sub_20_pct = mean(sub_20) * 100,

    sub_30_n = sum(sub_30),
    sub_30_pct = mean(sub_30) * 100,
    
    over_52_n = sum(over_52),
    over_52_pct = mean(over_52) * 100,
    
    over_180_n = sum(over_180),
    over_180_pct = mean(over_180) * 100
  )
```

Of the 2,694 consenting respondents, 13% (n = 352) reached the end of the survey in less or equal to 8 minutes; 20% (n = 539) in less than or equal to 12 minutes; 38% (n = 1021) in less than or equal to 20 minutes, 59% (n = 1592) in or less than 30, 21% (n = 563) in 52 minutes or more, and 7% or 200 respondents took two hours or more.

## finishers

nonfinisher

```{r}

df %>%
  summarise(
    n = n(),
    
    finished_n = sum(finished == 1),
    finished_pct = finished_n / n,
    
    nonfinished_n = sum(nonfinisher),
    nonfinished_pct = nonfinished_n / n
  )


table(df$finished, useNA = "ifany")
```

Of the 2,694 consenting respondents, 83% finished (as in, reached the end of the survey, irregardless of skipped questions) (n = 2,231) and 17% did not (n = 463).

## missingness?

```{r}
summary(df$na_count)
summary(df$common_na_count)
summary(df$block_na_count)

```

```{r}
# visual of nas/missingness

uncleaned_missingness <-
ggplot(df,
       aes(x = na_count, y = 0)) +
  geom_quasirandom(
    groupOnX = FALSE,
    width = .35,
    alpha = .35,
    size = 1.2
  ) +
  coord_cartesian(xlim = c(0, 120)) +
  labs(
    x = "na count",
    y = NULL,
    title = "na count"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )


uncleaned_missingness 

# 
# ggsave(
#   "uncleaned_missingness.png",
#   plot = uncleaned_missingness,
#   width = 14,
#   height = 9,
#   dpi = 300,
# )
```

```{r}
df %>%
  group_by(finished) %>%
  summarise(
    n = n(),
    min_na = min(na_count, na.rm = TRUE),
    q1_na = quantile(na_count, .25, na.rm = TRUE),
    median_na = median(na_count, na.rm = TRUE),
    mean_na = mean(na_count, na.rm = TRUE),
    q3_na = quantile(na_count, .75, na.rm = TRUE),
    max_na = max(na_count, na.rm = TRUE)
  )
```

```{r}
df %>%
  group_by(block) %>%
  summarise(
    n = n(),
    median_na = median(na_count, na.rm = TRUE),
    mean_na = mean(na_count, na.rm = TRUE),
    q75_na = quantile(na_count, .75, na.rm = TRUE),
    q90_na = quantile(na_count, .90, na.rm = TRUE),
    max_na = max(na_count, na.rm = TRUE)
  )
```

```{r}
df %>%
  mutate(
    progress_group = case_when(
      progress < 25  ~ "<25%",
      progress < 50  ~ "25–<50%",
      progress < 75  ~ "50–<75%",
      progress < 100 ~ "75–<100%",
      progress == 100 ~ "100%",
      TRUE ~ NA_character_
    )
  ) %>%
  group_by(progress_group) %>%
  summarise(
    n = n(),
    median_na = median(na_count, na.rm = TRUE),
    mean_na = mean(na_count, na.rm = TRUE),
    min_na = min(na_count, na.rm = TRUE),
    max_na = max(na_count, na.rm = TRUE)
  )
```

```{r}
df %>%
  group_by(attn_answered_count) %>%
  summarise(
    n = n(),
    median_na = median(na_count, na.rm = TRUE),
    mean_na = mean(na_count, na.rm = TRUE),
    median_progress = median(progress, na.rm = TRUE),
    pct_finished = mean(finished == 1, na.rm = TRUE) * 100
  )
```

```{r}
df %>%
  group_by(finished) %>%
  summarise(
    n = n(),
    
    median_progress =
      median(progress, na.rm = TRUE),
    
    median_na =
      median(na_count, na.rm = TRUE),
    
    median_common_na =
      median(common_na_count, na.rm = TRUE),
    
    median_block_na =
      median(block_na_count, na.rm = TRUE),
    
    median_attn_answered =
      median(attn_answered_count, na.rm = TRUE),
    
    median_attn_passed =
      median(attn_pass_count, na.rm = TRUE)
  )
```

## clustering of indicators?

overlap summary

```{r}
overlap_summary <- tibble(
  check = c("Q9", "Q17", "Q114"),
  
  total_n = nrow(df),
  
  answered_n = c(
    sum(df$attn_q9_answered),
    sum(df$attn_q17_answered),
    sum(df$attn_q114_answered)
  ),
  
  nonanswer_n = c(
    sum(!df$attn_q9_answered),
    sum(!df$attn_q17_answered),
    sum(!df$attn_q114_answered)
  ),
  
  pass_n = c(
    sum(df$attn_q9_pass),
    sum(df$attn_check_q17),
    sum(df$attn_q114_pass)
  )
) %>%
  mutate(
    # Answered, but did not pass
    wrong_n = answered_n - pass_n,
    
    # Everyone who did not pass: wrong answer + nonanswer
    nonpass_n = total_n - pass_n,
    
    answered_pct = answered_n / total_n * 100,
    
    # Your current percentages
    pass_pct_all = pass_n / total_n * 100,
    
    # More useful for actual attention performance:
    # Of people who either didn't answer or got all 3 , how many passed?
    pass_pct_answered = pass_n / answered_n * 100,
    
    # Of all apparent non-passes, what % are just nonanswers?
    nonanswer_pct_of_nonpass =
      nonanswer_n / nonpass_n * 100,
    
    # Of people who answered, what % actually got it wrong?
    wrong_pct_answered =
      wrong_n / answered_n * 100
  )

overlap_summary

```

### enough usable measurements?

Ie, did respondents supply enough of the questionnaire to furnish usable measurements?

Among people who broke off—or among finishers with substantial missingness—what analytically important information is actually present?

q15_1 q15_2 q15_3 q15_4 q15_5 q15 q25_1 q25_2 q25_3 q25_4 q25_1 q4_1_1 q4_1_2 q4_1_3 q4_2_1 q4_2_2 q4_2_3

q15_1 q15_2 q15_3 q15_4 q15_5 q15 q25_1 q25_2 q25_3 q25_4 q25_1 q4_1_1 q4_1_2 q4_1_3 q4_2_1 q4_2_2 q4_2_3, q62_1,q51_1, q51_2, q51_3, q51_4, q51_5, q51_6, q51_7, q51_8, dem101,

q32, q33, q34, q35, q36, q37, q38, q111,

used variables

```{r}

used_vars <- c("d_crim_abc", "d_incppl_pris_abc","d_incwomen_abc","q15_1", "q15_2", "q15_3", "q15_4", "q15_5",
"q25_1", "q25_2", "q25_3", "q25_4", "q4_1_1", "q4_1_2", "q4_1_3", "q4_2_1", "q4_2_2", "q4_2_3","q15_1", "q15_2", "q15_3", "q15_4", "q15_5", "q25_1", "q25_2", "q25_3", "q25_4", "q25_1", "q62_1", "q23_2","q23_7",
"q4_1_1", "q4_1_2", "q4_1_3", "q4_2_1", "q4_2_2", "q4_2_3",
"q62_1","q51_1", "q51_2", "q51_3", "q51_4", "q51_5", "q51_6", "q51_7", "q51_8","dem101","q32", "q33", "q34", "q35", "q36", "q37", "q38", "q111"
)

used_var_missing <- df %>% 
  summarise(
    across(
      all_of(used_vars),
      ~sum(is.na(.))
    )
  ) %>%
  pivot_longer(
    everything(),
    names_to = "variable",
    values_to = "n_missing"
  ) %>%
  mutate(
    n_answered = nrow(df)- n_missing,
    pct_missing = n_missing/nrow(df) * 100,
    pct_answered = n_answered/nrow(df) *100
  ) %>%
  arrange(desc(pct_missing))

View(used_var_missing)

```

for overleaf

```{r}
used_var_missing_table <- used_var_missing %>%
    mutate(
    pct_missing = round(pct_missing, 1),
    pct_answered = round(pct_answered, 1)
  ) %>%
  rename(
    Variable = variable,
    `N Missing` = n_missing,
    `N Answered` = n_answered,
    `% Missing` = pct_missing,
    `% Answered` = pct_answered
  )


used_var_missing_latex <- used_var_missing_table  %>%
  kable(
    format = "latex",
    booktabs = TRUE,
    longtable = TRUE,
    caption = "Missingness and Response Rates for Employed Survey Items",
    label = "tab:used_var_missingness",
    align = c("l", "r", "r", "r", "r"),
    digits = 1
  ) %>%
  kable_styling(
    latex_options = c("repeat_header"),
    font_size = 9
  )
# 
# writeLines(
#   used_var_missing_latex,
#   "used_var_missing.tex"
)
```

```{r}
all_item_missingness <- df %>%
  summarise(
    across(
      all_of(common_vars),
      ~ sum(is.na(.))
    )
  ) %>%
  pivot_longer(
    everything(),
    names_to = "variable",
    values_to = "n_missing"
  ) %>%
  mutate(
    pct_missing = n_missing / nrow(df) * 100
  ) %>%
  arrange(desc(pct_missing))

all_item_missingness
```

Branch questions: 
109
112
dem
q53

Demographics
q51_n (racial dem)
q56 (gender) (_1 is women, _2 man, _3 queer, _4 other)


Info I will need about respondents who answer DVs
- race
- gender
- political identity

Race vars uncleaned are 0, 1, NA, -99 

```{r}
#white respondents
class(df$q51_1)
table(df$q51_1, useNA = "ifany")
```
425 not-white respondents who made it to this question
1184 white respondents who made it to this question
1082 started but didn't make it to this question 

```{r}
#Black respondents
unique(df$q51_2)
table(df$q51_2, useNA = "ifany")
```
268 black respondents
1341 not black respondents
1082 NA

```{r}
#Hispanic respondents
table(df$q51_7, useNA = "ifany")
```
237 Hispanic respondents
1372 not hispanic respondents
1082 NA



# 6. Flag and or drop

### attn checks
```{r}

df <- df %>%
  mutate(
    attn_wrong_count = attn_answered_count - attn_pass_count
  )
```

### duration

-already did sub 8 through 30, over 52 over 180
Hillygus: standard is to drop 1/3 median, flag 1/2 median

median: 25.35
1/3 is 8.43

### progress/finished
```{r}
df <- df %>%
  mutate(
    breakoff = finished == 0,
    early_breakoff = progress < 25,
    sub_mean_breakoff = progress < 84.68
  )


```
