---
title: "outcome_var_prep"
output: html_notebook
date: "2026-08-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(tidyr)
library(readr)
library(janitor)   # for clean_names()
library(tibble)
library(stringr)
library(haven)
library(dplyr)
library(tidyr)
library(readr)
library(janitor)   # for clean_names()
library(tibble)
library(stringr)
library(haven)
library(tibble)
library(gt)
library(psych)
library(ggplot2)

```

# Setup

```{r}

raw_data_cleaned <- readr::read_csv("data/cleaned/cleaned_light.csv") 

```

# Policy Preferences


# Deservingness Scores

## Sympathy "desor_core" control

```{r, desor1}
# ------------------------------------------------------------
# Deservingness / "sympathy" orientation: CORE items
# ------------------------------------------------------------

desor_core_vars <- c(
  "d_welfare_tanf_abc","d_fstamps_snap_abc","d_medicaid_abc","d_medicare_abc",
  "d_unemployed_abc","d_poorfam_abc","d_homeless_abc","d_immigrants_abc","d_unauth_undoc_abc"
)
#no -99s, are all numeric

# ------------------------------------------------------------
# 1. Item-only data and missingness
# ------------------------------------------------------------

# Item-only data
core_mat <- raw_data_cleaned %>%
  select(all_of(desor_core_vars))

# Number of observed items per respondent
core_n <- rowSums(!is.na(core_mat))

capture.output(table(core_n, useNA = "ifany"))

# ------------------------------------------------------------
# 2. Should they be index? Correlations among component items
# ------------------------------------------------------------

desor_corr <- psych::corr.test(
  core_mat,
  use = "pairwise",
  method = "pearson",
  adjust = "none"
)

# Pairwise correlations
view(round(desor_corr$r, 2))

# N used for each pairwise correlation
print(desor_corr$n)

```



```{r, desor2}
# ------------------------------------------------------------
# 3. Reliability / internal consistency
# ------------------------------------------------------------

# Eligibility rule for composite
# at least 5 of 9 items observed
core_keep <- core_n >= 5

print(table(core_keep, useNA = "ifany"))


core_alpha <- psych::alpha(
  core_mat[core_keep, ],
  check.keys = TRUE
)

# Key reliability outputs
print(core_alpha$total$raw_alpha)
print(core_alpha$total$std.alpha)
print(core_alpha$total$average_r)
print(core_alpha$n.obs)

# Item diagnostics
print(core_alpha$alpha.drop)
print(core_alpha$item.stats)

# Check whether psych automatically reversed any items
print(core_alpha$keys)
```


```{r, desor3}
# ------------------------------------------------------------
# 4. Construct respondent-level composite
# ------------------------------------------------------------

raw_data_cleaned$desor_core <- ifelse(
  core_keep,
  rowMeans(core_mat, na.rm = TRUE),
  NA_real_
)

print(summary(raw_data_cleaned$desor_core))


# ------------------------------------------------------------
# 5. Standardized version in case useful later
# ------------------------------------------------------------

raw_data_cleaned$desor_core_z <-
  as.numeric(scale(raw_data_cleaned$desor_core))

print(summary(raw_data_cleaned$desor_core_z))

```

raw_alpha: 0.9078 and 0.9142 internal consistency -- suggests items hang together well. average_r = 0.5422 -- high inter-item correlation. cor: fairly strong correlation to the DV (0.547) R\^2 = 0.30 bivariately. Soaks up general generosity toward assistance-framed groups. Should justify use of control for testing whether contact predicts deservingness of incarcerated people, above and beyond that general orientation.

