---
title: "outcome_var_prep"
output: html_notebook
date: "2026-08-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(tidyr)
library(readr)
library(janitor)   # for clean_names()
library(tibble)
library(stringr)
library(haven)
library(dplyr)
library(tidyr)
library(readr)
library(janitor)   # for clean_names()
library(tibble)
library(stringr)
library(haven)
library(tibble)
library(gt)
library(psych)
library(ggplot2)

```

# Setup

```{r}

raw_data_cleaned <- readr::read_csv("data/cleaned/cleaned_light.csv") 

```

# Deservingness Scores

## Sympathy "desor_core" control

```{r, desor1}
# ------------------------------------------------------------
# Deservingness / "sympathy" orientation: CORE items
# ------------------------------------------------------------

desor_core_vars <- c(
  "d_welfare_tanf_abc","d_fstamps_snap_abc","d_medicaid_abc","d_medicare_abc",
  "d_unemployed_abc","d_poorfam_abc","d_homeless_abc","d_immigrants_abc","d_unauth_undoc_abc"
)
#no -99s, are all numeric

# ------------------------------------------------------------
# 1. Item-only data and missingness
# ------------------------------------------------------------

# Item-only data
core_mat <- raw_data_cleaned %>%
  select(all_of(desor_core_vars))

# Number of observed items per respondent
core_n <- rowSums(!is.na(core_mat))

capture.output(table(core_n, useNA = "ifany"))

# ------------------------------------------------------------
# 2. Should they be index? Correlations among component items
# ------------------------------------------------------------

desor_corr <- psych::corr.test(
  core_mat,
  use = "pairwise",
  method = "pearson",
  adjust = "none"
)

# Pairwise correlations
view(round(desor_corr$r, 2))

# N used for each pairwise correlation
print(desor_corr$n)

```

```{r, desor2}
# ------------------------------------------------------------
# 3. Reliability / internal consistency
# ------------------------------------------------------------

# Eligibility rule for composite
# at least 5 of 9 items observed
core_keep <- core_n >= 5

print(table(core_keep, useNA = "ifany"))


core_alpha <- psych::alpha(
  core_mat[core_keep, ],
  check.keys = TRUE
)

# Key reliability outputs
print(core_alpha$total$raw_alpha)
print(core_alpha$total$std.alpha)
print(core_alpha$total$average_r)
print(core_alpha$n.obs)

# Item diagnostics
print(core_alpha$alpha.drop)
print(core_alpha$item.stats)

# Check whether psych automatically reversed any items
print(core_alpha$keys)
```

```{r, desor3}
# ------------------------------------------------------------
# 4. Construct respondent-level composite
# ------------------------------------------------------------

raw_data_cleaned$desor_core <- ifelse(
  core_keep,
  rowMeans(core_mat, na.rm = TRUE),
  NA_real_
)

print(summary(raw_data_cleaned$desor_core))


# ------------------------------------------------------------
# 5. Standardized version in case useful later
# ------------------------------------------------------------

raw_data_cleaned$desor_core_z <-
  as.numeric(scale(raw_data_cleaned$desor_core))

print(summary(raw_data_cleaned$desor_core_z))

```

raw_alpha: 0.9078 and 0.9142 internal consistency -- suggests items hang together well. average_r = 0.5422 -- high inter-item correlation. cor: fairly strong correlation to the DV (0.547) R\^2 = 0.30 bivariately. Soaks up general generosity toward assistance-framed groups. Should justify use of control for testing whether contact predicts deservingness of incarcerated people, above and beyond that general orientation.

# Policy Preferences

## Purpose of Incarceration

```{r, punishorrehab}

#Originally 62_1 with 1 = purpose = punish and 7 = purpose = rehabilitate people

summary(raw_data_cleaned$q62_1)

# Rename 
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(q62_1 = na_if(as.numeric(q62_1), -99))

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(penalpoint = 8 - q62_1) #reverse coded so 7 = most punitive

# summary(raw_data_cleaned$penalpoint) #confirmed I flipped

# ordered factor version of the reverse-coded scale (7 = most punitive)
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(penalpoint_ord = factor(
    penalpoint,
    levels = 1:7,
    labels = c(
      "Rehab (1)", "Rehab (2)", "Rehab (3)", "Neutral (4)",
      "Punish (5)", "Punish (6)", "Punish (7)"
    ),
    ordered = TRUE
  ))

# NOTE: Factor version of PenalPunitiveness (reverse-coded already: 1 = least punitive, 7 = most punitive)
```

## Dpen & LWOP

```{r, prisonpen}
#ORIGINALLY Q25_1 and Q25_5, death penalty and LWOP (1 = strongly favor 4 = strongly oppose, 5 = don't know)

         
#rename, -99, ensure numeric
raw_data_cleaned <- raw_data_cleaned %>% 
  mutate(
    dpen         = na_if(as.numeric(q25_1), -99),
    lwop = na_if(as.numeric(q25_5), -99)
  )

pen_raw_vars <- c("dpen", "lwop") #character vector 

# Reverse code (higher num should mean more support, lower less support, and  "Don't know" becomes midpoint)

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(
    all_of(pen_raw_vars),
    ~ case_when(
    . == 1 ~ 4,           # strongly favor → strongly support help
    . == 2 ~ 3,
    . == 3 ~ 2,
    . == 4 ~ 1,
    . == 5 ~ 2.5,         # DK = neutral midpoint
    TRUE ~ NA_real_
  ),  # note: we skip 5 = DK
    .names = "{.col}_rev"
  ))

pen_vars <-c("dpen_rev", "lwop_rev") #pen_vars saved here

# step 3: save just the means of pen questions as prisonpen

raw_data_cleaned <- raw_data_cleaned %>%
  rowwise() %>%
  mutate(prisonpen = mean(c_across(c(dpen_rev, lwop_rev)), na.rm = TRUE)) %>%
  ungroup()

```

Takeaway: prison pen, prison pen 1 = dpen, prison pen 5 = lwop now -confirmed that prisonpen works -- constructed with average of death penalty and lwop. USE dpen_rev and lwop_rev and prisonpen.

### CBA & Corr

```{r}

pen_data <- raw_data_cleaned %>%
  select(all_of(pen_vars))


cor(pen_data,
    use = "pairwise.complete.obs")


print(psych::corr.test(pen_data))

print(psych::alpha(pen_data))

# Quick look
psych::describe(pen_data)

# Correlation (more interpretable than alpha for 2 items)
cor(raw_data_cleaned[pen_vars], use = "pairwise.complete.obs")

# Cronbach's alpha (fine to report, but remember it's 2 items)
psych::alpha(pen_data, check.keys = FALSE)

summary(raw_data_cleaned$dpen_rev)

summary(raw_data_cleaned$lwop_rev)

```

results - The **correlation** between the two items is **r = 0.238**.

- With only two items, Cronbach’s alpha is essentially a rescaled version of that correlation. (So don’t oversell alpha here; the correlation is the clean summary.)

Interpretation

- **These two punitive items are positively related, but weakly so.** People who favor the death penalty are *somewhat* more likely to favor LWOP, but plenty of respondents don’t treat them as interchangeable.

- That’s not necessarily a “problem.” Death penalty and LWOP are conceptually related (punitive severity), but they can tap **different moral/legal intuitions** (irreversibility, innocence risk, retribution vs incapacitation), so moderate association is plausible.

- The response distribution shows a **nontrivial chunk of “Don’t know”** recoded to 2.5 (15–17%) and about **8% missing** for each item. The DK-to-midpoint choice can slightly inflate apparent consistency by shrinking variance, so reliability is a conservative-ish check but not definitive.

### Visual

```{r}
# hist(raw_data_cleaned$dpen_rev)
# print(hist(raw_data_cleaned$lwop_rev))


# ------------------------------------------------------------
# bar plot for overleaf
# ------------------------------------------------------------

pen_data <- raw_data_cleaned %>%
  select(all_of(pen_vars)) %>%
  pivot_longer(
    cols = everything(),
    names_to = "item",
    values_to = "position"
  ) %>%
  filter(!is.na(position)) %>%
  mutate(
    item = factor(
      item,
      levels = c(
        "dpen_rev",
        "lwop_rev"
      ),
      labels = c(
        "Death Penalty",
        "Life Without Parole"
      )
    ),
    
    # Treat the response categories as discrete/ordinal categories
    position = factor(
      position,
      levels = c(1, 2, 2.5, 3, 4)
    )
  ) %>%
  count(item, position, .drop = FALSE)
# ------------------------------------------------------------
# make figure
# ------------------------------------------------------------
pen_barplot <- ggplot(
  pen_data,
  aes(x = position, y = n)
) +
  geom_col(
    width = 0.70,
    fill = "grey35"
  ) +
  facet_wrap(
    ~ item,
    ncol = 2
  ) +
  scale_x_discrete(drop = FALSE) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05))
  ) +
  labs(
    x = "Position",
    y = "Number of Respondents",
    caption = paste0(
      "Position: 1 = Strongly Oppose; 2 = Somewhat Oppose; ",
      "2.5 = Unsure; 3 = Somewhat Support; 4 = Strongly Support."
    )
  ) +
  theme_classic(base_size = 11) +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(
      face = "bold",
      size = 11
    ),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 9.5),
    panel.spacing = grid::unit(1, "lines"),
    plot.caption = element_text(
      size = 9,
      hjust = 0,
      margin = margin(t = 8)
    )
  )

# ------------------------------------------------------------
# save/print
# ------------------------------------------------------------

ggsave(
  filename = "figures/pen_barplot.pdf",
  plot = pen_barplot,
  width = 6.5,
  height = 5,
  units = "in",
)
```

## Prison Help

```{r, prisonhelp}
# Originally 1 = strong favor, 4 = strong oppose, 5 = don't know
# Define Q25 help-oriented items
q25_help_items <- c("q25_2", "q25_3", "q25_4", "q25_6")

# Step 1: Recode -99 to NA and ensure numeric
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(all_of(q25_help_items), ~ na_if(as.numeric(.), -99)))

# Step 2: Create reverse-coded versions (1 = strong favor → 4 = strong oppose becomes 4 → 1 so higher means more support for help oriented policyaQ                                                 
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(
    all_of(q25_help_items),
    ~ case_when(
    . == 1 ~ 4,           # strongly favor → strongly support help
    . == 2 ~ 3,
    . == 3 ~ 2,
    . == 4 ~ 1,
    . == 5 ~ 2.5,         # DK = neutral midpoint
    TRUE ~ NA_real_
  ),  # note: we skip 5 = DK
    .names = "{.col}_rev"
  ))


# now, raw dta cleaned has q25_n_rev (each reverse coded so that higher number means more support)

# step 3
raw_data_cleaned <- raw_data_cleaned %>%
  rename(
    min_wage       = q25_2_rev,
    free_calls     = q25_3_rev,
    ed_funds       = q25_4_rev,
    alt_sentencing = q25_6_rev
  )

# step 4: save just the means of q25 help questions as prisonhelp

raw_data_cleaned <- raw_data_cleaned %>% 
  rowwise() %>% #dplyr should treat each row as own group so c_across() can return row level vector of cols for that row
  mutate(prisonhelp = mean(c_across(c(min_wage, free_calls, ed_funds, alt_sentencing)), na.rm = TRUE)) %>%
  ungroup()

```

### Descriptive

```{r}

#checking variance + SD here 08-2026)
summary(raw_data_cleaned$prisonhelp)

summary(raw_data_cleaned$alt_sentencing)


q25_help_vars <- c("min_wage", "free_calls", "ed_funds", "alt_sentencing")

# 
# prisonhelp_data <- raw_data_cleaned %>%
#   select(all_of(q25_help_vars))


psych::describe(raw_data_cleaned[q25_help_vars])


#I get min 1 1st qu 2.5 median 2.875 mean 2.88 3q 3.250 max 4 nas = 186

#histograms
print(hist(raw_data_cleaned$min_wage))
print(hist(raw_data_cleaned$free_calls))
print(hist(raw_data_cleaned$ed_funds))
print(hist(raw_data_cleaned$alt_sentencing))

library(dplyr)
library(tidyr)
library(ggplot2)


q25_long <- raw_data_cleaned %>%
  select(all_of(q25_help_vars)) %>%
  pivot_longer(
    cols = everything(),
    names_to = "item",
    values_to = "response"
  )

ggplot(q25_long, aes(x = response)) +
  geom_bar() +
  facet_wrap(~ item) +
  labs(
    x = "Response",
    y = "Count"
  )

#barplots

# par(mfrow = c(2, 2))
# 
# barplot(table(raw_data_cleaned$min_wage),
#         main = "min_wage",
#         xlab = "Response")
# 
# barplot(table(raw_data_cleaned$free_calls),
#         main = "free_calls",
#         xlab = "Response")
# 
# barplot(table(raw_data_cleaned$ed_funds),
#         main = "ed_funds",
#         xlab = "Response")
# 
# barplot(table(raw_data_cleaned$alt_sentencing),
#         main = "alt_sentencing",
#         xlab = "Response")
# 
# par(mfrow = c(1, 1))


#bar plot for overleaf

# Create plotting data
q25_plot_data <- raw_data_cleaned %>%
  select(
    min_wage,
    free_calls,
    ed_funds,
    alt_sentencing
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "item",
    values_to = "position"
  ) %>%
  filter(!is.na(position)) %>%
  mutate(
    item = factor(
      item,
      levels = c(
        "min_wage",
        "free_calls",
        "ed_funds",
        "alt_sentencing"
      ),
      labels = c(
        "Minimum Wage",
        "No Cost Calls",
        "Fund Education",
        "Parent Alts"
      )
    ),
    
    # Treat the response categories as discrete/ordinal categories
    position = factor(
      position,
      levels = c(1, 2, 2.5, 3, 4)
    )
  ) %>%
  count(item, position, .drop = FALSE)


# Make figure
q25_barplot <- ggplot(
  q25_plot_data,
  aes(x = position, y = n)
) +
  geom_col(
    width = 0.70,
    fill = "grey35"
  ) +
  facet_wrap(
    ~ item,
    ncol = 2
  ) +
  scale_x_discrete(drop = FALSE) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05))
  ) +
  labs(
    x = "Position",
    y = "Number of Respondents",
    caption = paste0(
      "Position: 1 = Strongly Oppose; 2 = Somewhat Oppose; ",
      "2.5 = Unsure; 3 = Somewhat Support; 4 = Strongly Support."
    )
  ) +
  theme_classic(base_size = 11) +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(
      face = "bold",
      size = 11
    ),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 9.5),
    panel.spacing = grid::unit(1, "lines"),
    plot.caption = element_text(
      size = 9,
      hjust = 0,
      margin = margin(t = 8)
    )
  )

ggsave(
  filename = "figures/q25_help_items_barplot.pdf",
  plot = q25_barplot,
  width = 6.5,
  height = 5,
  units = "in",
)
q25_barplot


```

### CBA & Corr
```{r}

q25_help_data <- raw_data_cleaned %>%
  select(all_of(q25_help_vars))

cor(q25_help_data,
    use = "pairwise.complete.obs")

print(psych::corr.test(q25_help_data))

print(psych::alpha(q25_help_data))

# Correlation (more interpretable than alpha for 2 items)
cor(q25_help_data, use = "pairwise.complete.obs")

# Cronbach's alpha (fine to report, but remember it's 2 items)
psych::alpha(q25_help_data, check.keys = FALSE)

q25_help_data <- raw_data_cleaned %>%
  select(all_of(q25_help_vars))

cor(q25_help_data,
    use = "pairwise.complete.obs")
```

## Election Items

```{r, electionhelpqs}
names(raw_data_cleaned) %>% stringr::str_subset("q25|q23")

# Define Q23 help-oriented items (election - q23_2 is elects_reenfranchise and q23_7 is voting in jail (Requiring that jail and election officers ensure that people who are eligible to vote while incarcerated are able to exercise that right.) OG 1 = favor, 2 = oppose, 3 = DK

#rename, -99, ensure numeric
raw_data_cleaned <- raw_data_cleaned %>% 
  mutate(
    auto_reg         = na_if(as.numeric(q23_1), -99),
    post_felony_vote = na_if(as.numeric(q23_2), -99),
    early_vote       = na_if(as.numeric(q23_3), -99),
    election_holiday = na_if(as.numeric(q23_4), -99),
    voter_id         = na_if(as.numeric(q23_5), -99),
    noncit_local     = na_if(as.numeric(q23_6), -99),
    jail_access      = na_if(as.numeric(q23_7), -99)
  )

all_electoral_vars <- c("auto_reg", "post_felony_vote", "early_vote", "election_holiday", "voter_id", "noncit_local", "jail_access") #character vector 

#recode 1-3 scale to match direction of prison help items (also intuition that higher number means more support)
# here, 1 = favor, 2 = oppose, 3 = DK
# we want higher = more support for help-oriented policy so 1-> 3, 2-> 1, 3 -> 2
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(
    all_of(all_electoral_vars),
    ~ case_when(
      . == 1 ~ 3,
      . == 2 ~ 1,
      . == 3 ~ 2,
      TRUE ~ NA_real_
    ),
    .names = "{.col}_rev"
  ))

# respondent level means: all electoral vars
# 
# raw_data_cleaned <- raw_data_cleaned %>% 
#   rowwise() %>% #dplyr should treat each row as own group so c_across() can return row level vector of cols for that row
#   mutate(all_electoral_means = mean(c_across(c(auto_reg_rev, post_felony_vote_rev, early_vote_rev, election_holiday_rev, voter_id_rev, noncit_local_rev, jail_access_rev)), na.rm = TRUE)) %>%
#   ungroup()

#summary(raw_data_cleaned$all_electoral_means)#all_election_items now exists as #mean across all 7 Q23 questions, reverse coded, for each observation. 

## Jail voting and rights restoration

incarc_electoral_vars <- c("post_felony_vote_rev", "jail_access_rev")

# respondent level means: just the two (incarc_electoral_means)

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    incarc_electoral_means = if_else(
      !is.na(post_felony_vote_rev) & !is.na(jail_access_rev),
      (post_felony_vote_rev + jail_access_rev) / 2,
      NA_real_
    )
  )

# old: q23_help_items <- c("q23_2", "q23_7")
# new: incarc_electoral_means, post_felony_vote_rev, jail_access_rev

# to compare to other prison policy questions on same scale, rescale these two  the 1–3 items to 1–4 range

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(
    all_of(incarc_electoral_vars),
    ~ case_when(
      . == 3 ~ 4,
      . == 1 ~ 1,
      . == 2 ~ 2.5,
      TRUE ~ NA_real_
    ),
    .names = "{.col}_1_to_4"
  ))

# produces: post_felony_vote_rev_1_to_4, jail_access_rev_1_to_4

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    incarc_electoral_mean_1_to_4 = if_else(
      !is.na(post_felony_vote_rev_1_to_4) &
        !is.na(jail_access_rev_1_to_4),
      (post_felony_vote_rev_1_to_4 +
         jail_access_rev_1_to_4) / 2,
      NA_real_
    )
  )
hist(raw_data_cleaned$post_felony_vote_rev)
hist(raw_data_cleaned$jail_access_rev)

print(summary(raw_data_cleaned$post_felony_vote_rev))
print(summary(raw_data_cleaned$jail_access_rev))
```

Come out of above block with: post_felony_vote_rev individual item, 1–3 jail_access_rev individual item, 1–3 incarc_electoral_mean combined measure, 1–3

post_felony_vote_rev_1_to_4 comparable item, 1–4 jail_access_rev_1_to_4 comparable item, 1–4 incarc_electoral_mean_1_to_4 comparable combined measure, 1–4

#### D.4_C Prison/Election Items Descriptive look

What's up

```{r}

#bar plot for overleaf

# Create plotting data
incarc_electoral_data <- raw_data_cleaned %>%
  select(
    post_felony_vote_rev,
    jail_access_rev
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "item",
    values_to = "position"
  ) %>%
  filter(!is.na(position)) %>%
  mutate(
    item = factor(
      item,
      levels = c(
        "post_felony_vote_rev",
        "jail_access_rev"
      ),
      labels = c(
        "Rights Restoration",
        "Jail Vote Access"
      )
    ),
    
    # Treat the response categories as discrete/ordinal categories
    position = factor(
      position,
      levels = c(1, 2, 3)
    )
  ) %>%
  count(item, position, .drop = FALSE)


# Make figure
incarc_electoral_barplot <- ggplot(
  incarc_electoral_data ,
  aes(x = position, y = n)
) +
  geom_col(
    width = 0.70,
    fill = "grey35"
  ) +
  facet_wrap(
    ~ item,
    ncol = 2
  ) +
  scale_x_discrete(drop = FALSE) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05))
  ) +
  labs(
    x = "Position",
    y = "Number of Respondents",
    caption = paste0(
      "Position: 1 = Oppose; 2 = Unsure; ",
      "3 = Somewhat Support."
    )
  ) +
  theme_classic(base_size = 11) +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(
      face = "bold",
      size = 11
    ),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 9.5),
    panel.spacing = grid::unit(1, "lines"),
    plot.caption = element_text(
      size = 9,
      hjust = 0,
      margin = margin(t = 8)
    )
  )


ggsave(
  filename = "figures/incarc_electoral_barplot.pdf",
  plot = q25_barplot,
  width = 6.5,
  height = 5,
  units = "in",
)
incarc_electoral_barplot

```

###CBA & Corr: JBV & RR
```{r}

incarc_electoral_data <- raw_data_cleaned %>%
  select(all_of(incarc_electoral_vars))

cor(incarc_electoral_data,
    use = "pairwise.complete.obs")

print(psych::corr.test(incarc_electoral_data))

print(psych::alpha(incarc_electoral_data))

```

###CBA & Corr: Help, JBV & RR

Need

post_felony_vote_rev jail_access_rev_1_to_4 q25_help_data or min_wage, free_calls, ed_funds, alt_sentencing

```{r, electionhelpqs_2}

# ALL HELP ITEMS

all_help_items <- c(
  "min_wage",
  "free_calls",
  "ed_funds",
  "alt_sentencing",
  "post_felony_vote_rev_1_to_4",
  "jail_access_rev_1_to_4"
)

#df for the six total vars
all_help_data <- raw_data_cleaned %>%
  select(all_of(all_help_items))

# now, pairwise corrs
View(round(
  cor(all_help_data, use = "pairwise.complete.obs"),
  2
))

# Correlations + N + significance information
all_help_corr <- psych::corr.test(
  all_help_data,
  use = "pairwise",
  method = "pearson",
  adjust = "none"
)

print(all_help_corr)

# Internal consistency of the six items
all_help_alpha <- psych::alpha(
  all_help_data,
  check.keys = FALSE
)

print(all_help_alpha)

```

```{r}
# ALL ITEMS except help as index

combined_items <- c(
  "post_felony_vote_rev_1_to_4",
  "jail_access_rev_1_to_4",
  "min_wage",
  "free_calls",
  "ed_funds",
  "alt_sentencing",
  "dpen_rev",
  "lwop_rev"
)


combined_data <- raw_data_cleaned %>%
  select(all_of(combined_items))

# Pairwise correlations
View(round(
  cor(combined_data, use = "pairwise.complete.obs"),
  2
))

combined_corr <- psych::corr.test(
  combined_data,
  use = "pairwise",
  method = "pearson",
  adjust = "none"
)

print(combined_corr)

combined_alpha <- psych::alpha(
  combined_data,
  check.keys = FALSE
)

print(combined_alpha)
```

#### D5. Variance and Stand Dev

```{r}
#don't need to redo this
#help_policy_items <- c("q25_2", "q25_3", "q25_4", "q25_6", "q23_2", "q23_7")

# Variance and SD across help policy items
raw_data_cleaned <- raw_data_cleaned %>%
  rowwise() %>%
  mutate(
    variancehelp = var(c_across(all_of(help_policy_items)), na.rm = TRUE),
    sdhelp = sd(c_across(all_of(help_policy_items)), na.rm = TRUE)
  ) %>%
  ungroup()


```


# Write and Save Data
```{r}

```

