---
title: "outcome_var_prep"
output: html_notebook
date: "2026-08-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(tidyr)
library(readr)
library(janitor)   # for clean_names()
library(tibble)
library(stringr)
library(haven)
library(dplyr)
library(tidyr)
library(readr)
library(janitor)   # for clean_names()
library(tibble)
library(stringr)
library(haven)
library(tibble)
library(gt)
library(psych)
library(ggplot2)

```

# Setup

```{r}

raw_data_cleaned <- readr::read_csv("data/cleaned/coded_data__09_18_26.csv") 

# raw_data_cleaned <- read_csv("data/cleaned_control_data.csv")

# raw_data_cleaned <- read_csv("data/cleaned_sans_control_data.csv")

```

# Deservingness Scores


Standardize? No -- in my prelim, I write: To ensure that I do not control for a variable that is conceptually or empirically too close to the deservingness of incarcerated people when it is the dependent variable, I exclude any deservingness ratings for groups that are substantively overlapping or highly correlated. Including those could absorb variance in the outcome or introduce post-treatment bias by controlling for a related evaluative construct. I also avoid using global composite or average deservingness scores across all groups – for example, a respondent's mean deservingnesss rating across all 65 groups they evaluated – because that composite would include the rating of incarcerated people itself, introducing potential collinearity and making it difficult to isolate the unique effect of contact. This modeling strategy attempts to preserve interpretability and ensuring that the effect of policy contact is estimated with respect to unstandardized, target-specific deservingness scores. 12

12 While constructing composite or control variables involving other deservingness targets, I assess whether their inclusion introduces collinearity or conceptual overlap with the dependent variable. In particular, I examine correlations between group-level deservingness ratings to ensure that any included groups are not too closely tied to perceptions of incarcerated people. If preliminary checks reveal that including a particular group (e.g., “criminals” or “formerly incarcerated individuals”) would obscure key variance or conflate target constructs, I exclude it from the model and treat it as conceptually non-distinct for the purposes of this analysis.

## B. Means?


## B. Sympathy "desor_core" control

```{r, desor1}
# ------------------------------------------------------------
# Deservingness / "sympathy" orientation: CORE items
# ------------------------------------------------------------

desor_core_vars <- c(
  "d_welfare_tanf_abc","d_fstamps_snap_abc","d_medicaid_abc","d_medicare_abc",
  "d_unemployed_abc","d_poorfam_abc","d_homeless_abc","d_immigrants_abc","d_unauth_undoc_abc"
)

#no -99s, are all numeric

# ------------------------------------------------------------
# 1. Item-only data and missingness
# ------------------------------------------------------------

# Item-only data
core_mat <- raw_data_cleaned %>%
  select(all_of(desor_core_vars))

#names(raw_data_cleaned)
# Number of observed items per respondent
core_n <- rowSums(!is.na(core_mat))

capture.output(table(core_n, useNA = "ifany"))

# ------------------------------------------------------------
# 2. Should they be index? Correlations among component items
# ------------------------------------------------------------

desor_corr <- psych::corr.test(
  core_mat,
  use = "pairwise",
  method = "pearson",
  adjust = "none"
)

# Pairwise correlations
#view(round(desor_corr$r, 2))

# N used for each pairwise correlation
desor_corr$n

```

```{r, desor2}
# ------------------------------------------------------------
# 3. Reliability / internal consistency
# ------------------------------------------------------------

# Eligibility rule for composite
# at least 5 of 9 items observed
core_keep <- core_n >= 5

table(core_keep, useNA = "ifany")


core_alpha <- psych::alpha(
  core_mat[core_keep, ],
  check.keys = TRUE
)

# Key reliability outputs
core_alpha$total$raw_alpha
core_alpha$total$std.alpha
core_alpha$total$average_r
core_alpha$n.obs

# Item diagnostics
core_alpha$alpha.drop
core_alpha$item.stats

# Check whether psych automatically reversed any items
core_alpha$keys
```

```{r, desor3}
# ------------------------------------------------------------
# 4. Construct respondent-level composite
# ------------------------------------------------------------

raw_data_cleaned$desor_core <- ifelse(
  core_keep,
  rowMeans(core_mat, na.rm = TRUE),
  NA_real_
)

summary(raw_data_cleaned$desor_core)


# ------------------------------------------------------------
# 5. Standardized version in case useful later
# ------------------------------------------------------------

raw_data_cleaned$desor_core_z <-
  as.numeric(scale(raw_data_cleaned$desor_core))

summary(raw_data_cleaned$desor_core_z)

```

raw_alpha: 0.9078 and 0.9142 internal consistency -- suggests items hang together well. average_r = 0.5422 -- high inter-item correlation. cor: fairly strong correlation to the DV (0.547) R\^2 = 0.30 bivariately. Soaks up general generosity toward assistance-framed groups. Should justify use of control for testing whether contact predicts deservingness of incarcerated people, above and beyond that general orientation.

# Policy Preferences

## Purpose of Incarceration

```{r, punishorrehab}

#Originally 62_1 with 1 = purpose = punish and 7 = purpose = rehabilitate people

summary(raw_data_cleaned$prison_purpose)

# Rename 
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(q62_1 = na_if(as.numeric(prison_purpose), -99))

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(penalpoint = 8 - prison_purpose) #reverse coded so 7 = most punitive

# summary(raw_data_cleaned$penalpoint) #confirmed I flipped

# ordered factor version of the reverse-coded scale (7 = most punitive)
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(penalpoint_ord = factor(
    penalpoint,
    levels = 1:7,
    labels = c(
      "Rehab (1)", "Rehab (2)", "Rehab (3)", "Neutral (4)",
      "Punish (5)", "Punish (6)", "Punish (7)"
    ),
    ordered = TRUE
  ))

# NOTE: Factor version of PenalPunitiveness (reverse-coded already: 1 = least punitive, 7 = most punitive)
```

## Dpen & LWOP

```{r, prisonpen}
#ORIGINALLY Q25_1 and Q25_5, death penalty and LWOP (1 = strongly favor 4 = strongly oppose, 5 = don't know)

         
#rename, -99, ensure numeric
raw_data_cleaned <- raw_data_cleaned %>% 
  mutate(
    dpen         = na_if(as.numeric(dpen), -99),
    lwop = na_if(as.numeric(lwop), -99)
  )

pen_raw_vars <- c("dpen", "lwop") #character vector 

# Reverse code (higher num should mean more support, lower less support, and  "Don't know" becomes midpoint)

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(
    all_of(pen_raw_vars),
    ~ case_when(
    . == 1 ~ 4,           # strongly favor → strongly support help
    . == 2 ~ 3,
    . == 3 ~ 2,
    . == 4 ~ 1,
    . == 5 ~ 2.5,         # DK = neutral midpoint
    TRUE ~ NA_real_
  ),  # note: we skip 5 = DK
    .names = "{.col}_rev"
  ))

pen_vars <-c("dpen_rev", "lwop_rev") #pen_vars saved here

# step 3: save just the means of pen questions as prisonpen

raw_data_cleaned <- raw_data_cleaned %>%
  rowwise() %>%
  mutate(prisonpen = mean(c_across(c(dpen_rev, lwop_rev)), na.rm = TRUE)) %>%
  ungroup()

```

Takeaway: prison pen, prison pen 1 = dpen, prison pen 5 = lwop now -confirmed that prisonpen works -- constructed with average of death penalty and lwop. USE dpen_rev and lwop_rev and prisonpen.

### CBA & Corr

```{r}

pen_data <- raw_data_cleaned %>%
  select(all_of(pen_vars))


cor(pen_data,
    use = "pairwise.complete.obs")


#print(psych::corr.test(pen_data))

#print(psych::alpha(pen_data))

# Quick look
#psych::describe(pen_data)

# Correlation (more interpretable than alpha for 2 items)
cor(raw_data_cleaned[pen_vars], use = "pairwise.complete.obs")

# Cronbach's alpha (fine to report, but remember it's 2 items)
psych::alpha(pen_data, check.keys = FALSE)

summary(raw_data_cleaned$dpen_rev)

summary(raw_data_cleaned$lwop_rev)

```

results - The **correlation** between the two items is **r = 0.238**.

- With only two items, Cronbach’s alpha is essentially a rescaled version of that correlation. (So don’t oversell alpha here; the correlation is the clean summary.)

Interpretation

- **These two punitive items are positively related, but weakly so.** People who favor the death penalty are *somewhat* more likely to favor LWOP, but plenty of respondents don’t treat them as interchangeable.

- That’s not necessarily a “problem.” Death penalty and LWOP are conceptually related (punitive severity), but they can tap **different moral/legal intuitions** (irreversibility, innocence risk, retribution vs incapacitation), so moderate association is plausible.

- The response distribution shows a **nontrivial chunk of “Don’t know”** recoded to 2.5 (15–17%) and about **8% missing** for each item. The DK-to-midpoint choice can slightly inflate apparent consistency by shrinking variance, so reliability is a conservative-ish check but not definitive.

### Visual

```{r}
# hist(raw_data_cleaned$dpen_rev)
# print(hist(raw_data_cleaned$lwop_rev))


# ------------------------------------------------------------
# bar plot for overleaf
# ------------------------------------------------------------

pen_data <- raw_data_cleaned %>%
  select(all_of(pen_vars)) %>%
  pivot_longer(
    cols = everything(),
    names_to = "item",
    values_to = "position"
  ) %>%
  filter(!is.na(position)) %>%
  mutate(
    item = factor(
      item,
      levels = c(
        "dpen_rev",
        "lwop_rev"
      ),
      labels = c(
        "Death Penalty",
        "Life Without Parole"
      )
    ),
    
    # Treat the response categories as discrete/ordinal categories
    position = factor(
      position,
      levels = c(1, 2, 2.5, 3, 4)
    )
  ) %>%
  count(item, position, .drop = FALSE)
# ------------------------------------------------------------
# make figure
# ------------------------------------------------------------
pen_barplot <- ggplot(
  pen_data,
  aes(x = position, y = n)
) +
  geom_col(
    width = 0.70,
    fill = "grey35"
  ) +
  facet_wrap(
    ~ item,
    ncol = 2
  ) +
  scale_x_discrete(drop = FALSE) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05))
  ) +
  labs(
    x = "Position",
    y = "Number of Respondents",
    caption = paste0(
      "Position: 1 = Strongly Oppose; 2 = Somewhat Oppose; ",
      "2.5 = Unsure; 3 = Somewhat Support; 4 = Strongly Support."
    )
  ) +
  theme_classic(base_size = 11) +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(
      face = "bold",
      size = 11
    ),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 9.5),
    panel.spacing = grid::unit(1, "lines"),
    plot.caption = element_text(
      size = 9,
      hjust = 0,
      margin = margin(t = 8)
    )
  )

# ------------------------------------------------------------
# save/print
# ------------------------------------------------------------

ggsave(
  filename = "figures/pen_barplot.pdf",
  plot = pen_barplot,
  width = 6.5,
  height = 5,
  units = "in",
)
```

## Prison Help

```{r, prisonhelp}
# Originally 1 = strong favor, 4 = strong oppose, 5 = don't know
# Define Q25 help-oriented items
q25_help_items <- c("min_wage", "free_calls", "ed_funds", "alt_sentencing")

# Step 1: Recode -99 to NA and ensure numeric
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(all_of(q25_help_items), ~ na_if(as.numeric(.), -99)))

# Step 2: Create reverse-coded versions (1 = strong favor → 4 = strong oppose becomes 4 → 1 so higher means more support for help oriented policyaQ                                                 
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(
    all_of(q25_help_items),
    ~ case_when(
    . == 1 ~ 4,           # strongly favor → strongly support help
    . == 2 ~ 3,
    . == 3 ~ 2,
    . == 4 ~ 1,
    . == 5 ~ 2.5,         # DK = neutral midpoint
    TRUE ~ NA_real_
  )))


# now, raw dta cleaned has q25_n_rev (each reverse coded so that higher number means more support)

# step 3 (renamed now in data prep)


# step 4: save just the means of q25 help questions as prisonhelp

raw_data_cleaned <- raw_data_cleaned %>% 
  rowwise() %>% #dplyr should treat each row as own group so c_across() can return row level vector of cols for that row
  mutate(prisonhelp = mean(c_across(c(min_wage, free_calls, ed_funds, alt_sentencing)), na.rm = TRUE)) %>%
  ungroup()

```

### Descriptive

```{r}

#checking variance + SD here 08-2026)
summary(raw_data_cleaned$prisonhelp)

summary(raw_data_cleaned$alt_sentencing)


q25_help_vars <- c("min_wage", "free_calls", "ed_funds", "alt_sentencing")

# 
# prisonhelp_data <- raw_data_cleaned %>%
#   select(all_of(q25_help_vars))


psych::describe(raw_data_cleaned[q25_help_vars])



#I get min 1 1st qu 2.5 median 2.875 mean 2.88 3q 3.250 max 4 nas = 186

#histograms
# print(hist(raw_data_cleaned$min_wage))
# print(hist(raw_data_cleaned$free_calls))
# print(hist(raw_data_cleaned$ed_funds))
# print(hist(raw_data_cleaned$alt_sentencing))

library(dplyr)
library(tidyr)
library(ggplot2)


q25_long <- raw_data_cleaned %>%
  select(all_of(q25_help_vars)) %>%
  pivot_longer(
    cols = everything(),
    names_to = "item",
    values_to = "response"
  )

ggplot(q25_long, aes(x = response)) +
  geom_bar() +
  facet_wrap(~ item) +
  labs(
    x = "Response",
    y = "Count"
  )

#barplots

# par(mfrow = c(2, 2))
# 
# barplot(table(raw_data_cleaned$min_wage),
#         main = "min_wage",
#         xlab = "Response")
# 
# barplot(table(raw_data_cleaned$free_calls),
#         main = "free_calls",
#         xlab = "Response")
# 
# barplot(table(raw_data_cleaned$ed_funds),
#         main = "ed_funds",
#         xlab = "Response")
# 
# barplot(table(raw_data_cleaned$alt_sentencing),
#         main = "alt_sentencing",
#         xlab = "Response")
# 
# par(mfrow = c(1, 1))


#bar plot for overleaf

# Create plotting data
q25_plot_data <- raw_data_cleaned %>%
  select(
    min_wage,
    free_calls,
    ed_funds,
    alt_sentencing
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "item",
    values_to = "position"
  ) %>%
  filter(!is.na(position)) %>%
  mutate(
    item = factor(
      item,
      levels = c(
        "min_wage",
        "free_calls",
        "ed_funds",
        "alt_sentencing"
      ),
      labels = c(
        "Minimum Wage",
        "No Cost Calls",
        "Fund Education",
        "Parent Alts"
      )
    ),
    
    # Treat the response categories as discrete/ordinal categories
    position = factor(
      position,
      levels = c(1, 2, 2.5, 3, 4)
    )
  ) %>%
  count(item, position, .drop = FALSE)


# Make figure
q25_barplot <- ggplot(
  q25_plot_data,
  aes(x = position, y = n)
) +
  geom_col(
    width = 0.70,
    fill = "grey35"
  ) +
  facet_wrap(
    ~ item,
    ncol = 2
  ) +
  scale_x_discrete(drop = FALSE) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05))
  ) +
  labs(
    x = "Position",
    y = "Number of Respondents",
    caption = paste0(
      "Position: 1 = Strongly Oppose; 2 = Somewhat Oppose; ",
      "2.5 = Unsure; 3 = Somewhat Support; 4 = Strongly Support."
    )
  ) +
  theme_classic(base_size = 11) +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(
      face = "bold",
      size = 11
    ),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 9.5),
    panel.spacing = grid::unit(1, "lines"),
    plot.caption = element_text(
      size = 9,
      hjust = 0,
      margin = margin(t = 8)
    )
  )

ggsave(
  filename = "figures/q25_help_items_barplot_09_26.pdf",
  plot = q25_barplot,
  width = 6.5,
  height = 5,
  units = "in",
)
q25_barplot


```

### CBA & Corr
```{r}

q25_help_data <- raw_data_cleaned %>%
  select(all_of(q25_help_vars))

cor(q25_help_data,
    use = "pairwise.complete.obs")

print(psych::corr.test(q25_help_data))

print(psych::alpha(q25_help_data))

# Correlation (more interpretable than alpha for 2 items)
cor(q25_help_data, use = "pairwise.complete.obs")

# Cronbach's alpha (fine to report, but remember it's 2 items)
psych::alpha(q25_help_data, check.keys = FALSE)

q25_help_data <- raw_data_cleaned %>%
  select(all_of(q25_help_vars))

cor(q25_help_data,
    use = "pairwise.complete.obs")
```

## Election Items

```{r, electionhelpqs}
names(raw_data_cleaned) %>% stringr::str_subset("q25|q23")

# Define Q23 help-oriented items (election - q23_2 is elects_reenfranchise and q23_7 is voting in jail (Requiring that jail and election officers ensure that people who are eligible to vote while incarcerated are able to exercise that right.) OG 1 = favor, 2 = oppose, 3 = DK

#rename, -99, ensure numeric
raw_data_cleaned <- raw_data_cleaned %>% 
  mutate(
    auto_reg         = na_if(as.numeric(auto_reg), -99),
    post_felony_vote = na_if(as.numeric(post_felony_vote), -99),
    early_vote       = na_if(as.numeric(early_vote), -99),
    election_holiday = na_if(as.numeric(election_holiday), -99),
    voter_id         = na_if(as.numeric(voter_id), -99),
    noncit_local     = na_if(as.numeric(noncit_local), -99),
    jail_access      = na_if(as.numeric(jail_access), -99)
  )

all_electoral_vars <- c("auto_reg", "post_felony_vote", "early_vote", "election_holiday", "voter_id", "noncit_local", "jail_access") #character vector 

#recode 1-3 scale to match direction of prison help items (also intuition that higher number means more support)
# here, 1 = favor, 2 = oppose, 3 = DK
# we want higher = more support for help-oriented policy so 1-> 3, 2-> 1, 3 -> 2
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(
    all_of(all_electoral_vars),
    ~ case_when(
      . == 1 ~ 3,
      . == 2 ~ 1,
      . == 3 ~ 2,
      TRUE ~ NA_real_
    ),
    .names = "{.col}_rev"
  ))

# respondent level means: all electoral vars
# 
# raw_data_cleaned <- raw_data_cleaned %>% 
#   rowwise() %>% #dplyr should treat each row as own group so c_across() can return row level vector of cols for that row
#   mutate(all_electoral_means = mean(c_across(c(auto_reg_rev, post_felony_vote_rev, early_vote_rev, election_holiday_rev, voter_id_rev, noncit_local_rev, jail_access_rev)), na.rm = TRUE)) %>%
#   ungroup()

#summary(raw_data_cleaned$all_electoral_means)#all_election_items now exists as #mean across all 7 Q23 questions, reverse coded, for each observation. 

## Jail voting and rights restoration

incarc_electoral_vars <- c("post_felony_vote_rev", "jail_access_rev")

# respondent level means: just the two (incarc_electoral_means)

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    incarc_electoral_means = if_else(
      !is.na(post_felony_vote_rev) & !is.na(jail_access_rev),
      (post_felony_vote_rev + jail_access_rev) / 2,
      NA_real_
    )
  )

# old: q23_help_items <- c("q23_2", "q23_7")
# new: incarc_electoral_means, post_felony_vote_rev, jail_access_rev

# to compare to other prison policy questions on same scale, rescale these two  the 1–3 items to 1–4 range

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(
    all_of(incarc_electoral_vars),
    ~ case_when(
      . == 3 ~ 4,
      . == 1 ~ 1,
      . == 2 ~ 2.5,
      TRUE ~ NA_real_
    ),
    .names = "{.col}_1_to_4"
  ))

# produces: post_felony_vote_rev_1_to_4, jail_access_rev_1_to_4

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    incarc_electoral_mean_1_to_4 = if_else(
      !is.na(post_felony_vote_rev_1_to_4) &
        !is.na(jail_access_rev_1_to_4),
      (post_felony_vote_rev_1_to_4 +
         jail_access_rev_1_to_4) / 2,
      NA_real_
    )
  )
hist(raw_data_cleaned$post_felony_vote_rev)
hist(raw_data_cleaned$jail_access_rev)

print(summary(raw_data_cleaned$post_felony_vote_rev))
print(summary(raw_data_cleaned$jail_access_rev))
```

Come out of above block with: post_felony_vote_rev individual item, 1–3 jail_access_rev individual item, 1–3 incarc_electoral_mean combined measure, 1–3

post_felony_vote_rev_1_to_4 comparable item, 1–4 jail_access_rev_1_to_4 comparable item, 1–4 incarc_electoral_mean_1_to_4 comparable combined measure, 1–4

#### D.4_C Prison/Election Items Descriptive look

What's up

```{r}

#bar plot for overleaf

# Create plotting data
incarc_electoral_data <- raw_data_cleaned %>%
  select(
    post_felony_vote_rev,
    jail_access_rev
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "item",
    values_to = "position"
  ) %>%
  filter(!is.na(position)) %>%
  mutate(
    item = factor(
      item,
      levels = c(
        "post_felony_vote_rev",
        "jail_access_rev"
      ),
      labels = c(
        "Rights Restoration",
        "Jail Vote Access"
      )
    ),
    
    # Treat the response categories as discrete/ordinal categories
    position = factor(
      position,
      levels = c(1, 2, 3)
    )
  ) %>%
  count(item, position, .drop = FALSE)


# Make figure
incarc_electoral_barplot <- ggplot(
  incarc_electoral_data ,
  aes(x = position, y = n)
) +
  geom_col(
    width = 0.70,
    fill = "grey35"
  ) +
  facet_wrap(
    ~ item,
    ncol = 2
  ) +
  scale_x_discrete(drop = FALSE) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05))
  ) +
  labs(
    x = "Position",
    y = "Number of Respondents",
    caption = paste0(
      "Position: 1 = Oppose; 2 = Unsure; ",
      "3 = Somewhat Support."
    )
  ) +
  theme_classic(base_size = 11) +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(
      face = "bold",
      size = 11
    ),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 9.5),
    panel.spacing = grid::unit(1, "lines"),
    plot.caption = element_text(
      size = 9,
      hjust = 0,
      margin = margin(t = 8)
    )
  )


ggsave(
  filename = "figures/incarc_electoral_barplot.pdf",
  plot = q25_barplot,
  width = 6.5,
  height = 5,
  units = "in",
)
incarc_electoral_barplot

```

###CBA & Corr: JBV & RR
```{r}

incarc_electoral_data <- raw_data_cleaned %>%
  select(all_of(incarc_electoral_vars))

cor(incarc_electoral_data,
    use = "pairwise.complete.obs")

print(psych::corr.test(incarc_electoral_data))

print(psych::alpha(incarc_electoral_data))

```

###CBA & Corr: Help, JBV & RR

Need

post_felony_vote_rev jail_access_rev_1_to_4 q25_help_data or min_wage, free_calls, ed_funds, alt_sentencing

```{r, electionhelpqs_2}

# ALL HELP ITEMS

all_help_items <- c(
  "min_wage",
  "free_calls",
  "ed_funds",
  "alt_sentencing",
  "post_felony_vote_rev_1_to_4",
  "jail_access_rev_1_to_4"
)

#df for the six total vars
all_help_data <- raw_data_cleaned %>%
  select(all_of(all_help_items))

# now, pairwise corrs
#View(round(
#  cor(all_help_data, use = "pairwise.complete.obs"),
#  2
#))

# Correlations + N + significance information
all_help_corr <- psych::corr.test(
  all_help_data,
  use = "pairwise",
  method = "pearson",
  adjust = "none"
)

print(all_help_corr)

# Internal consistency of the six items
all_help_alpha <- psych::alpha(
  all_help_data,
  check.keys = FALSE
)

print(all_help_alpha)

```

```{r}
# ALL ITEMS except help as index

combined_items <- c(
  "post_felony_vote_rev_1_to_4",
  "jail_access_rev_1_to_4",
  "min_wage",
  "free_calls",
  "ed_funds",
  "alt_sentencing",
  "dpen_rev",
  "lwop_rev"
)


combined_data <- raw_data_cleaned %>%
  select(all_of(combined_items))

# Pairwise correlations
# View(round(
#   cor(combined_data, use = "pairwise.complete.obs"),
#   2
# ))

combined_corr <- psych::corr.test(
  combined_data,
  use = "pairwise",
  method = "pearson",
  adjust = "none"
)

print(combined_corr)

combined_alpha <- psych::alpha(
  combined_data,
  check.keys = FALSE
)

#print(combined_alpha)
```

#### D5. Variance and Stand Dev

```{r}
#don't need to redo this
#help_policy_items <- c("q25_2", "q25_3", "q25_4", "q25_6", "q23_2", "q23_7")

# Variance and SD across help policy items
# raw_data_cleaned <- raw_data_cleaned %>%
#   rowwise() %>%
#   mutate(
#     variancehelp = var(c_across(all_of(help_policy_items)), na.rm = TRUE),
#     sdhelp = sd(c_across(all_of(help_policy_items)), na.rm = TRUE)
#   ) %>%
#   ungroup()


```


## Police

```{r}

# police policies

# Originally 6 = strong oppose, 7= somewhat oppose, 8 = neutral, 9 = somewhat support, 10 = strong support

# Define Q25 police items
police_items <- c("local_police", "no_strangle", "police_tracker", "police_training", "charge_police")

# Step 1: Recode -99 to NA and ensure numeric
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(all_of(police_items), ~ na_if(as.numeric(.), -99)))

# Step 2: Recode so 1-5 (6 = strong oppose → 1 = strong oppose, 
# 7 = somewhat oppose → 2 = somewhat oppose, 
# 8 = neutral → 3 = neutral, 
# 9 = somewhat support → 4 = somewhat support,
#10 = strong support → 5 = strong support)

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(
    all_of(police_items),
    ~ case_when(
    . == 6 ~ 1,          
    . == 7 ~ 2,
    . == 8 ~ 3,
    . == 9 ~ 4,
    . == 10 ~ 5,        
    TRUE ~ NA_real_
  ),  
    .names = "{.col}_1_to_5"
  ))

police_items_1_to_5 <- c(
  "local_police",
  "no_strangle",
  "police_tracker",
  "police_training",
  "charge_police"
)

```

# Civics

```{r}

## trust in government 

summary(raw_data_cleaned$gov_attn)
unique(raw_data_cleaned$gov_attn)

# Originally 1 = a lot, 2 = some, 3 = not muc

# rename
## trust in government attention
# q30 originally: 1 = a lot, 2 = some, 3 = not much
# recode so higher = more government attention/trust

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    gov_attn = case_when(
      as.numeric(gov_attn) == -99 ~ NA_real_,
      as.numeric(gov_attn) == 1 ~ 3,
      as.numeric(gov_attn) == 2 ~ 2,
      as.numeric(gov_attn) == 3 ~ 1,
      TRUE ~ NA_real_
    )
  )

table(raw_data_cleaned$gov_attn)

##add gets issues
##unique(raw_data$gets_issues)

## did vote in 2020
# q31: 1 = no, not eligible; 2 = no, eligible; 3 = yes

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    vote2020 = na_if(as.numeric(vote2020), -99)
  )

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    voteD2020 = case_when(
      as.numeric(vote2020) == 1 ~ 0, 
      as.numeric(vote2020) == 3 ~ 1
    ) 
  )

## plans to vote
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    willvote = na_if(as.numeric(willvote), -99)
  )

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    vote_plan = case_when(
      as.numeric(willvote) == 2 ~ 0, 
      as.numeric(willvote) == 3 ~ 1
    ) 
  )

## "gets issues"
summary(raw_data_cleaned$gets_issues)
unique(raw_data_cleaned$gets_issues)

# recode so higher = gets issues very well
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    gets_issues = case_when(
      as.numeric(gets_issues) == -99 ~ NA_real_,
      as.numeric(gets_issues) == 1 ~ 5,
      as.numeric(gets_issues) == 2 ~ 4,
      as.numeric(gets_issues) == 3 ~ 3,
      as.numeric(gets_issues) == 4 ~ 2,
      as.numeric(gets_issues) == 5 ~ 1,
      TRUE ~ NA_real_
    )
  )
```

Note:

Approximately 561 respondents did not provide responses to the political behavior items (government attention, past vote, intended vote). Missingness is perfectly overlapping across these items, indicating that these questions were likely administered to a subset of respondents rather than reflecting item-level nonresponse.

# Write and Save Data
```{r}
#sept 14 -- saving to now look at completeness problem (+ representativeness)
nrow(raw_data_cleaned)
write_csv(raw_data_cleaned, "data/cleaned/all_vars_over_8.csv")

#names(raw_data_cleaned)

```

