---
title: "Policy Analyses without IDKs"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE, echo=FALSE}

knitr::opts_chunk$set(
  echo = FALSE,       # Hide code by default
  warning = FALSE,
  message = FALSE
)

# Load packages
library(dplyr)
library(tidyr)
library(readr)
library(janitor)   # for clean_names()
library(tibble)
library(stringr)
library(haven)
library(gt)
library(marginaleffects)
library(mediation)
library(naniar)
library(rsq)
library(patchwork)
library(modelsummary)
library(rlang)
library(forcats)
library(ggplot2)
library(broom)
library(knitr)
library(dplyr)
library(purrr)
library(lmtest)
library(sandwich)

source("variable_labels_sept1.R")

```

Note on samples --

This uses sample_all-> sample -> sample_dk -> sample_no_dk

In Sample, as in prev analyses policy variables have "don't know/no answer" coded as a middle value (2.5 or 2).

Sample_dk has new ambivalence DK variables (ie dk_free_calls) indicate whether a respondent gave a policy "dunno."
- Can use this for analysis of ambivalence, or can use below

sample_no_dk changes the policy variables so that "dunno" is now an NA. It also has the IDk indicator variables. 
- This (use for analysis of preferences) asks preferences after DK responses have been treated as missing.

# I. Setup

```{r}
# load data
sample_all <- read_csv("data/cleaned_control_data.csv")

not_full <- read_csv("data/cleaned_sans_control_data.csv")

# drop variables not working with, keep variables working with 
## Some variables were dropped in Data_Junction.rmd
## See variable_labels_sept1


more_needed_vars <- unique(c( ##variables I need that are n
  "state",
  "county"
))

policy_analysis_vars <- c(
  impact_vars, 
  q25_help_vars,
  policy_vars,
  police_items,
  deserving_vars,
  political_behavior_vars,
  race_vars,
  demo_controls,
  party_ideo_controls,
  desor_core_vars,
  fire_controls,
  ration_vars,
  experience_buffet,
  
  more_needed_vars
)


sample <- sample_all %>%
  dplyr::select(any_of(policy_analysis_vars))


# Vars where DK is coded 2.5
q25_help_vars <- c("min_wage", "free_calls","ed_funds","alt_sentencing")

punish_vars <- c("dpen_rev", "lwop_rev")

# Combine
dk_2.5_policy_vars <- c(
  q25_help_vars,
  punish_vars
)


# Variables where DK is coded 2
incar_vote_vars <- c(
  "post_felony_vote_rev",
  "jail_access_rev"
)


# All policy variables we want to examine
all_policy_vars <- c(
  q25_help_vars,
  punish_vars,
  incar_vote_vars
)


```

Confirm Counts

```{r}
table(sample$impact)

sample %>%
  group_by(impact) %>%
  summarise(
    n = n(),
    incarc = sum(incarc, na.rm = TRUE),
    knowincarc = sum(knowincarc, na.rm = TRUE),
    vict = sum(vict, na.rm = TRUE),
  )

# Confirm the comparison-group coding
table(sample$impact, sample$incarc, useNA = "ifany")
table(sample$impact, sample$knowincarc, useNA = "ifany")
table(sample$impact, sample$vict, useNA = "ifany")
```

## A. New DK Variables

One binary DK variable for each policy item

```{r}
# ------------------------------------------------------------
# Item-level DK indicators
# ------------------------------------------------------------

sample_dk <- sample %>%
  mutate(
    
    # DK = 2 for incarceration/voting items
    across(
      all_of(incar_vote_vars),
      ~ case_when(
        .x == 2 ~ 1L,          # DK
        !is.na(.x) ~ 0L,       # answered, but not DK
        TRUE ~ NA_integer_     # did not answer
      ),
      .names = "dk_{.col}"
    ),
    
    # DK = 2.5 for help + punishment items
    across(
      all_of(dk_2.5_policy_vars),
      ~ case_when(
        .x == 2.5 ~ 1L,        # DK
        !is.na(.x) ~ 0L,       # answered, but not DK
        TRUE ~ NA_integer_     # did not answer
      ),
      .names = "dk_{.col}"
    )
  )


#look at min wage:

table(
  sample_dk$dk_min_wage,
  useNA = "ifany" #should use usena if any
)

```

Create group DKs to match how the vars are grouped

```{r}
# ------------------------------------------------------------
# Names the new DK indicator variables
# ------------------------------------------------------------

dk_incar_vote_vars <- paste0(
  "dk_",
  incar_vote_vars
)

dk_q25_help_vars <- paste0(
  "dk_",
  q25_help_vars
)

dk_punish_vars <- paste0(
  "dk_",
  punish_vars
)

dk_all_vars <- c(
  dk_incar_vote_vars,
  dk_q25_help_vars,
  dk_punish_vars
)
```

Create total DK counts for each respondent

```{r}
# ------------------------------------------------------------
# Respondent-level DK totals
# ------------------------------------------------------------

sample_dk <- sample_dk %>%
  mutate(
    
    # --------------------------------
    # Number of DK responses
    # --------------------------------
    
    dk_total_all =
      rowSums(
        pick(all_of(dk_all_vars)),
        na.rm = TRUE
      ),
    
    dk_total_incar_vote =
      rowSums(
        pick(all_of(dk_incar_vote_vars)),
        na.rm = TRUE
      ),
    
    dk_total_q25_help =
      rowSums(
        pick(all_of(dk_q25_help_vars)),
        na.rm = TRUE
      ),
    
    dk_total_punish =
      rowSums(
        pick(all_of(dk_punish_vars)),
        na.rm = TRUE
      ),
    
    
    # --------------------------------
    # Number of items actually answered
    # --------------------------------
    
    dk_answered_all =
      rowSums(
        !is.na(pick(all_of(dk_all_vars)))
      ),
    
    dk_answered_incar_vote =
      rowSums(
        !is.na(pick(all_of(dk_incar_vote_vars)))
      ),
    
    dk_answered_q25_help =
      rowSums(
        !is.na(pick(all_of(dk_q25_help_vars)))
      ),
    
    dk_answered_punish =
      rowSums(
        !is.na(pick(all_of(dk_punish_vars)))
      )
  )

```

Respondents might not have the same number of observed answers.

Proportions of answered items that were DK

```{r}

sample_dk <- sample_dk %>%
  mutate(
    
    dk_prop_all = if_else(
      dk_answered_all > 0,
      dk_total_all / dk_answered_all,
      NA_real_
    ),
    
    dk_prop_incar_vote = if_else(
      dk_answered_incar_vote > 0,
      dk_total_incar_vote / dk_answered_incar_vote,
      NA_real_
    ),
    
    dk_prop_q25_help = if_else(
      dk_answered_q25_help > 0,
      dk_total_q25_help / dk_answered_q25_help,
      NA_real_
    ),
    
    dk_prop_punish = if_else(
      dk_answered_punish > 0,
      dk_total_punish / dk_answered_punish,
      NA_real_
    )
  )

```

## B. Edit Sample 

### "sample_no_idk"
- has dk_policy indicator variables 
- IDK responses now remove from policy vars themselves by converting to NA

```{r}
sample_no_dk <- sample_dk %>%
  mutate(
    
    # Incarceration/voting variables: DK = 2
    across(
      all_of(incar_vote_vars),
      ~ na_if(.x, 2)
    ),
    
    # Help + punishment variables: DK = 2.5
    across(
      all_of(dk_2.5_policy_vars),
      ~ na_if(.x, 2.5)
    ),
    
    # Police variables
    across(
      all_of(police_items),
      ~ na_if(.x, 2.5)
    )
  )

View(sample_no_dk)

```

## C. Counts code (revise if needed)

```{r}
# ------------------------------------------------------------
# Count DK responses overall and by impact status
# ------------------------------------------------------------

dk_policy_summary <- sample %>%
  
  # Put policy items into long format
  pivot_longer(
    cols = all_of(all_policy_vars),
    names_to = "policy",
    values_to = "response"
  ) %>%
  
  # Identify whether each response is DK
  mutate(
    is_dk = case_when(
      policy %in% incar_vote_vars ~ response == 2,
      policy %in% dk_2.5_policy_vars ~ response == 2.5,
      TRUE ~ FALSE
    )
  ) %>%
  
  group_by(policy) %>%
  
  summarise(
    
    # Total number of DK responses
    total_dk = sum(is_dk, na.rm = TRUE),
    
    # Number of DK responses from each group
    incarc_dk = sum(is_dk & incarc == 1, na.rm = TRUE),
    knowincarc_dk = sum(is_dk & knowincarc == 1, na.rm = TRUE),
    vict_dk = sum(is_dk & vict == 1, na.rm = TRUE),
    impact_dk = sum(is_dk & impact == 1, na.rm = TRUE),
    
    .groups = "drop"
  ) %>%
  
  # Calculate each group's share of all DK responses
  mutate(
    incarc_pct = if_else(
      total_dk > 0,
      100 * incarc_dk / total_dk,
      NA_real_
    ),
    
    knowincarc_pct = if_else(
      total_dk > 0,
      100 * knowincarc_dk / total_dk,
      NA_real_
    ),
    
    vict_pct = if_else(
      total_dk > 0,
      100 * vict_dk / total_dk,
      NA_real_
    ),
    
    impact_pct = if_else(
      total_dk > 0,
      100 * impact_dk / total_dk,
      NA_real_
    )
  ) %>%
  
  # Optional: make percentages easier to read
  mutate(
    across(
      ends_with("_pct"),
      ~ round(.x, 2)
    )
  )

dk_policy_summary

```

```{r}

# Group sizes in the original sample
group_totals <- sample %>%
  summarise(
    incarc_total = sum(incarc == 1, na.rm = TRUE),
    knowincarc_total = sum(knowincarc == 1, na.rm = TRUE),
    vict_total = sum(vict == 1, na.rm = TRUE),
    impact_total = sum(impact == 1, na.rm = TRUE)
  )


dk_policy_summary <- dk_policy_summary %>%
  mutate(
    incarc_dk_rate =
      100 * incarc_dk / group_totals$incarc_total,
    
    knowincarc_dk_rate =
      100 * knowincarc_dk / group_totals$knowincarc_total,
    
    vict_dk_rate =
      100 * vict_dk / group_totals$vict_total,
    
    impact_dk_rate =
      100 * impact_dk / group_totals$impact_total
  ) %>%
  mutate(
    across(
      ends_with("_rate"),
      ~ round(.x, 2)
    )
  )

dk_policy_summary
```

# II. Analysis: Preferences



## A. Name waves

Each wave is controls, so first group controls (tho these should
already/eventually be clear in variable labels R file)

```{r}
demo_controls <- c(
  "Black", "Hisp_Latino", "AllOthers", "gender"
)

demo_controls_labels <- c(
  Black = "Black",
  Hisp_Latino    = "Hispanic/Latino", 
  AllOthers = "Other RID",
  gender = "Men"
)

party_ideo_controls <- c(
  "Republican",
  "Democrat",
  "fiscal_ideology",
  "social_ideology"
)

party_ideo_controls_labels <- c(
  # pid_3 = "Party ID",
  Republican = "Republican",
  Democrat = "Democrat",
  fiscal_ideology = "Fiscal ideology",
  social_ideology = "Social ideology"
)

fire_controls <- c(
  "fire_rare",
  "fire_privilege",
  "fire_angry",
  "fire_fear"
)

fire_controls_labels <- c(
  fire_rare = "Racism rare",
  fire_privilege = "White privilege",
  fire_angry = "Angry about racism",
  fire_fear = "Fear other races"
)

control_waves <- list(
  nocontrol = character(0),
  demo = demo_controls,
  party_ideo = c(demo_controls, party_ideo_controls),
  full_attitudes = c(demo_controls, party_ideo_controls,
                     fire_controls)
)

# , desor_core


```

```{r}

wave_order <- c(
  "nocontrol",
  "demo",
  "party_ideo",
  "full_attitudes"
)

wave_labels <- c(
  nocontrol = "No controls",
  demo = "Demographics",
  party_ideo = "+ Party/Ideology",
  full_attitudes = "+ Attitudes"
)

# DV labels already defined in variable_labels_sept1.R through policy_map
policy_labels <- setNames(
  policy_map$label,
  policy_map$var
)

```

## B. Define model

The waves are waves of a model. So now define that model:

"whenever (later) I give this function a dataset (df), a dependent
variable (dv), an independent variable (iv), and some controls
(controls), R should:

(a) start with a default of assuming there are no controls, if no
    controls are given (character (0));
(b) in the right-hand side (rhs) or the equation, use the supplied iv
    and supplied controls to make a vector with all vraibles that should
    appear to the right of \~
(c) (fml) construct an ordinary regression syntax with the supplied dv
    on the left of \~, then supplied rhs info and turn it into a signle
    string (collapse = "+"), and
(d) treat this as a statistical formula rather than a string
    (as.formula())
(e) run the regression

```{r}

make_policy_model <- function(df, dv, iv, controls = character(0)) { #character(0) just means empty character vector so this function works for uncontrolled model then for controlled model
  rhs <- c(iv, controls)
  fml <- as.formula(paste(dv, "~", paste(rhs, collapse = " + ")))
  lm(fml, data = df)
}


```

Create bucket to store things (policy_dummies_waves)

```{r}
# When I run the models it will produce stuff, so first I need to create buckets to put the models: 
policy_dummies_waves <- list() #this is actually now main model generating loop

```

Created bucket to store things (policy_dummies_waves), and created the
basic model structure (make_policy_model), now explain how to fill
bucket.

How to fill the things: with waves of models that loop through the DVs
and iteratively add controls for each DV, generating 4 substantive
comparisons right now: 1. impact versus non impact 2. incarc versus non
impact 3. knowincarc versus non impact 4. victimized vs non impact

```{r}

# Now I am saying how I will fill the buckets: 

for (wave_name in names(control_waves)) {#calls list of waves in control waves

  model_list <- list() #fresh model list within each wave

  for (dv in policy_vars) {
    for (iv in impact_vars) { #loop thru dvs, ivs, labels file gives the 4 ivs; fits 4 models for each outcome

      model_name <- paste0("m_", dv, "_", iv) #ie model using prisonhelp and incarc will be named m_prisonhelp_incarc

      # NOW, actually telling it to run and store the regressions
      model_list[[model_name]] <- make_policy_model(
        df = sample_no_dk,
        dv = dv, 
        iv = iv,
        controls = control_waves[[wave_name]]
      )
    }
  }

  policy_dummies_waves[[wave_name]] <- model_list
}

```

This should have made something like:

policy_dummies_waves \| \|-- nocontrol \| \|-- m_prisonhelp_impact \|
\|-- m_prisonhelp_incarc \| \|-- m_prisonhelp_knowincarc \| \|--
m_prisonhelp_vict \| \|-- ... \| \|-- demo \| \|-- m_prisonhelp_impact
\| \|-- m_prisonhelp_incarc \| \|-- ... \| \|-- party_ideo \| \|--
full_attitudes

It is helpful to not need to type out the full
(policy_dummies_waves[["nocontrol"]][["m_prisonhelp_incarc"]]).

Making shortcuts/names

```{r}

policy_dummies_models <- policy_dummies_waves[["nocontrol"]]

policy_dummies_demo_models <- policy_dummies_waves[["demo"]]

policy_dummies_partyideo_models <- policy_dummies_waves[["party_ideo"]]

policy_dummies_full_models <- policy_dummies_waves[["full_attitudes"]]

```

For example "policy_dummies_full_models" shows the list of models with
all controls included

```{r}

names(policy_dummies_full_models)


```

## C. Run the models?

```{r}

# ------------------------------------------------------------
# 2. Helpers for extracting and formatting one focal estimate
# ------------------------------------------------------------

p_to_stars <- function(p) {
  dplyr::case_when(
    is.na(p) ~ "",
    p < .001 ~ "***",
    p < .01  ~ "**",
    p < .05  ~ "*",
    TRUE     ~ ""
  )
}

fmt_num <- function(x, digits = 3) {
  ifelse(
    is.na(x),
    "",
    sprintf(paste0("%.", digits, "f"), x)
  )
}


extract_policy_wave_stats <- function(
    dv_name,
    iv_name,
    wave_name,
    model_waves = policy_dummies_waves,
    robust = TRUE
) {

  # Model names follow the naming system created in your main loop:
  # e.g. m_prisonhelp_incarc
  model_name <- paste0("m_", dv_name, "_", iv_name)

  # Check that requested wave exists
  if (!wave_name %in% names(model_waves)) {
    stop("Wave not found: ", wave_name)
  }

  # Pull that wave's model list
  model_list <- model_waves[[wave_name]]

  # Check that requested model exists
  if (!model_name %in% names(model_list)) {
    stop(
      "Model not found in wave '",
      wave_name,
      "': ",
      model_name
    )
  }

  # Pull the actual lm object
  mod <- model_list[[model_name]]

  # Extract coefficient information
  if (robust) {

    ct <- lmtest::coeftest(
      mod,
      vcov. = sandwich::vcovHC(mod, type = "HC1")
    )

    coef_table <- data.frame(
      term = rownames(ct),
      estimate = ct[, 1],
      std.error = ct[, 2],
      p.value = ct[, 4],
      row.names = NULL
    )

  } else {

    sm <- summary(mod)$coefficients

    coef_table <- data.frame(
      term = rownames(sm),
      estimate = sm[, 1],
      std.error = sm[, 2],
      p.value = sm[, 4],
      row.names = NULL
    )
  }

  # The focal coefficient should have the same name as the IV:
  # impact, incarc, knowincarc, or vict
  if (!iv_name %in% coef_table$term) {
    stop(
      "Term '",
      iv_name,
      "' not found in model ",
      model_name
    )
  }

  focal_row <- coef_table[
    coef_table$term == iv_name,
  ]

  # Return one tidy row
  tibble::tibble(
    dv = dv_name,
    dv_label = policy_labels[[dv_name]],
    iv = iv_name,
    iv_label = impact_labels[[iv_name]],
    wave = wave_name,
    wave_label = wave_labels[[wave_name]],
    estimate = focal_row$estimate,
    std.error = focal_row$std.error,
    p.value = focal_row$p.value,
    n = stats::nobs(mod),
    r2 = summary(mod)$r.squared
  )
}

```

```{r}
# ------------------------------------------------------------
# 3. Gather all four IVs x all four waves for one DV
# ------------------------------------------------------------

make_wave_results_long <- function(
    dv_name,
    model_waves = policy_dummies_waves,
    row_vars = impact_vars,
    waves = wave_order,
    robust = TRUE
) {

  # Confirm requested DV belongs to current policy outcome set
  if (!dv_name %in% policy_vars) {
    stop(
      "DV not found in policy_vars: ",
      dv_name
    )
  }

  rows <- lapply(row_vars, function(iv_name) {

    wave_rows <- lapply(waves, function(wave_name) {

      extract_policy_wave_stats(
        dv_name = dv_name,
        iv_name = iv_name,
        wave_name = wave_name,
        model_waves = model_waves,
        robust = robust
      )
    })

    dplyr::bind_rows(wave_rows)
  })

  dplyr::bind_rows(rows) %>%
    dplyr::mutate(
      iv = factor(
        iv,
        levels = row_vars
      ),
      wave = factor(
        wave,
        levels = waves
      )
    )
}
```

```{r}
# ------------------------------------------------------------
# 4. Build one wide inspection table for one DV
# ------------------------------------------------------------

make_wave_results_table <- function(
    dv_name,
    model_waves = policy_dummies_waves,
    row_vars = impact_vars,
    waves = wave_order,
    robust = TRUE
) {

  results_long <- make_wave_results_long(
    dv_name = dv_name,
    model_waves = model_waves,
    row_vars = row_vars,
    waves = waves,
    robust = robust
  ) %>%
    dplyr::mutate(

      # What will eventually appear in each LaTeX table cell
      result = paste0(
        fmt_num(estimate, 3),
        p_to_stars(p.value),
        " (",
        fmt_num(std.error, 3),
        ")"
      ),

      # Preserve desired display order
      iv_label = factor(
        iv_label,
        levels = unname(impact_labels[row_vars])
      ),

      wave_label = factor(
        wave_label,
        levels = unname(wave_labels[waves])
      )
    )

  results_wide <- results_long %>%
    dplyr::select(
      iv_label,
      wave_label,
      result
    ) %>%
    tidyr::pivot_wider(
      names_from = wave_label,
      values_from = result
    ) %>%
    dplyr::arrange(iv_label) %>%
    dplyr::rename(
      `Carceral contact` = iv_label
    )

  results_wide
}
```



# Results:

## Model outputs for 2.5s excluded

```{r}
# Show the four wave names
names(policy_dummies_waves)

# Show every model stored in a given wave name (for ex, full-attitudes wave)
names(policy_dummies_full_models)
# Demographics + party/ideology + FIRE + core deservingness


# No controls
summary(policy_dummies_models[["m_prisonhelp_incarc"]])

# Demographic controls
summary(policy_dummies_demo_models[["m_prisonhelp_incarc"]])

# Demographics + party/ideology
summary(policy_dummies_partyideo_models[["m_prisonhelp_incarc"]])

summary(policy_dummies_full_models[["m_prisonhelp_incarc"]])

```

```{r}
summary(policy_dummies_full_models[["m_min_wage_vict"]])
```


### Voting Variables

```{r}
summary(policy_dummies_full_models[["m_post_felony_vote_rev_impact"]])
```

```{r}
summary(policy_dummies_full_models[["m_jail_access_rev_impact"]])
```

```{r}
summary(policy_dummies_full_models[["m_post_felony_vote_rev_vict"]])
```

```{r}
summary(policy_dummies_full_models[["m_jail_access_rev_incarc"]])
```

# Latex 

This assumes that analses steps A, B, and C have already run. This "picks up" from the end of C.

## Per DV, 1 4W Table:


```{r}
# Assumes Steps 1-4 have already run, including:
#   make_wave_results_long()
#   fmt_num()
#   p_to_stars()
#   wave_order
#   wave_labels
#
# Outcome and comparison labels are already attached by
# make_wave_results_long(), using policy_labels and impact_labels.

# Step A. Escape plain-text labels for Latex

wave_tex_escape <- function(x) {

  replacements <- c(
    "\\" = "\\textbackslash{}",
    "&" = "\\&",
    "%" = "\\%",
    "$" = "\\$",
    "#" = "\\#",
    "_" = "\\_",
    "{" = "\\{",
    "}" = "\\}",
    "~" = "\\textasciitilde{}",
    "^" = "\\textasciicircum{}"
  )

  vapply(as.character(x), function(txt) {

    if (is.na(txt)) {
      stop("A table label is missing.", call. = FALSE)
    }

    chars <- strsplit(txt, "", fixed = TRUE)[[1]]
    positions <- match(chars, names(replacements))
    replace <- !is.na(positions)

    chars[replace] <- unname(
      replacements[positions[replace]]
    )

    paste0(chars, collapse = "")

  }, character(1), USE.NAMES = FALSE)
}

# Step B. What to print

## for each comparison, print: coefficient with SE underneath, N, R-squared

make_wave_latex_table <- function(
    dv_name,
    model_waves = policy_dummies_waves,
    row_vars = impact_vars,
    waves = wave_order,
    robust = TRUE,
    digits = 3
) {

  if (length(waves) != 4L || anyDuplicated(waves)) {
    stop(
      "Supply four distinct model waves.",
      call. = FALSE
    )
  }

  # Use the checked extraction function from Step 3.
  # This retains unrounded statistics, N, R2, and existing labels.
  results <- make_wave_results_long(
    dv_name = dv_name,
    model_waves = model_waves,
    row_vars = row_vars,
    waves = waves,
    robust = robust
  )

  results$iv <- as.character(results$iv)
  results$wave <- as.character(results$wave)

  # Do not silently export missing or duplicated results.
  if (
    nrow(results) != length(row_vars) * length(waves) ||
    anyDuplicated(results[c("iv", "wave")])
  ) {
    stop(
      "Missing or duplicated model results for: ",
      dv_name,
      call. = FALSE
    )
  }

  numeric_cols <- c(
    "estimate", "std.error", "p.value", "n", "r2"
  )

  if (any(!is.finite(as.matrix(results[numeric_cols])))) {
    stop(
      "Non-finite model statistics for: ",
      dv_name,
      call. = FALSE
    )
  }

  # Use the outcome label already obtained from policy_map.
  dv_label <- unique(as.character(results$dv_label))

  if (
    length(dv_label) != 1L ||
    anyNA(dv_label) ||
    !nzchar(dv_label)
  ) {
    stop(
      "Check the policy_map label for: ",
      dv_name,
      call. = FALSE
    )
  }

  # Join cells and add the LaTeX end-of-row marker.
  tex_row <- function(cells) {
    paste0(paste(cells, collapse = " & "), " \\\\")
  }

  body_lines <- character(0)

  # Build one comparison block at a time.
  for (iv_name in row_vars) {

    block <- results[results$iv == iv_name, ]

    # Put the four results in the specified wave order.
    block <- block[match(waves, block$wave), ]

    if (anyNA(block$wave)) {
      stop(
        "A wave is missing for ",
        dv_name, " / ", iv_name,
        call. = FALSE
      )
    }

    # Use the comparison label supplied through impact_labels.
    iv_label <- as.character(block$iv_label[1])

    # Round only for display.
    # Significance stars use the unrounded p-values.
    cells <- paste0(
      "\\makecell[c]{",
      fmt_num(block$estimate, digits),
      p_to_stars(block$p.value),
      " \\\\ (",
      fmt_num(block$std.error, digits),
      ")}"
    )

    body_lines <- c(
      body_lines,

      # Coefficient and SE row
      tex_row(c(
        wave_tex_escape(iv_label),
        cells
      )),

      # Sample size for each model in this comparison
      tex_row(c(
        "\\quad N",
        formatC(
          as.integer(block$n),
          format = "d",
          big.mark = ","
        )
      )),

      # R-squared for each model in this comparison
      tex_row(c(
        "\\quad $R^2$",
        fmt_num(block$r2, digits)
      )),

      "\\addlinespace"
    )
  }

  # Keep the note consistent with the requested SE method.
  se_note <- if (robust) {
    "Entries are OLS coefficients with HC1 robust standard errors in parentheses."
  } else {
    "Entries are OLS coefficients with conventional OLS standard errors in parentheses."
  }

  # Reference keys are identifiers, not printed labels:
  # do not escape underscores here.
  label_key <- paste0("tab:policy-waves-", dv_name)

  # Assemble the complete table.
  c(
    "\\begin{table}[!htbp]",
    "\\centering",
    "\\begin{threeparttable}",

    paste0(
      "\\caption{",
      wave_tex_escape(paste0("OLS models of ", dv_label)),
      "}"
    ),

    paste0("\\label{", label_key, "}"),

    "\\small",
    "\\setlength{\\tabcolsep}{4pt}",
    "\\renewcommand{\\arraystretch}{1.12}",

    "\\begin{tabularx}{\\textwidth}{@{}>{\\raggedright\\arraybackslash}p{0.22\\textwidth}*{4}{>{\\centering\\arraybackslash}X}@{}}",

    "\\toprule",

    tex_row(c(
      "Carceral contact",
      wave_tex_escape(unname(wave_labels[waves]))
    )),

    "\\midrule",
    body_lines,
    "\\bottomrule",
    "\\end{tabularx}",

    "\\begin{tablenotes}[flushleft]",
    "\\footnotesize",

    paste0("\\item ", se_note),

    "\\item Each focal coefficient comes from a separate regression. The reference group reports none of the three focal experiences: personal incarceration, knowing someone incarcerated, or personal victimization. Exposure groups may overlap.",

    "\\item Controls accumulate from left to right: none; demographics; demographics plus party/ideology; then additionally FIRE and core deservingness controls. Control coefficients are not displayed.",

    "\\item N and $R^2$ refer to the model immediately above them. Each model retains its own estimation sample; sample sizes may differ across comparisons and waves.",

    "\\item * $p<0.05$, ** $p<0.01$, *** $p<0.001$.",

    "\\end{tablenotes}",
    "\\end{threeparttable}",
    "\\end{table}"
  )
}


```

## Export the Per DV 1 4W Tables

```{r}

# ------------------------------------------------------------
# 6A. Generate one table for every DV in policy_vars
# Uses policy_vars, NOT the narrower results_dvs.
# ------------------------------------------------------------

# Keep filenames and LaTeX reference keys safe and unique.
if (
  length(policy_vars) == 0L ||
  anyNA(policy_vars) ||
  anyDuplicated(policy_vars) ||
  any(!grepl("^[A-Za-z0-9_][A-Za-z0-9_.-]*$", policy_vars))
) {
  stop(
    "policy_vars must contain unique, nonmissing, filename-safe variable names."
  )
}

wave_latex_tables <- setNames(
  lapply(policy_vars, function(dv_name) {

    make_wave_latex_table(
      dv_name = dv_name,
      model_waves = policy_dummies_waves,
      row_vars = impact_vars,
      waves = wave_order,
      robust = TRUE
    )

  }),
  policy_vars
)


# ------------------------------------------------------------
# 6B. Write one .tex file per DV
# ------------------------------------------------------------

wave_table_dir <- file.path("tables", "policy_waves")

dir.create(
  wave_table_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

if (!dir.exists(wave_table_dir)) {
  stop("Could not create output folder: ", wave_table_dir)
}

# Named vector linking each DV to its output filename.
wave_table_files <- setNames(
  paste0(policy_vars, "_wave_table.tex"),
  policy_vars
)

for (dv_name in policy_vars) {

  writeLines(
    wave_latex_tables[[dv_name]],
    con = file.path(
      wave_table_dir,
      wave_table_files[[dv_name]]
    )
  )

}


# ------------------------------------------------------------
# 6C. Create an Overleaf main document
#
# The main file and individual table files belong together
# in the same folder.
#
# Each DV starts on a fresh page, in policy_vars order.
# ------------------------------------------------------------

input_lines <- unlist(
  lapply(unname(wave_table_files), function(filename) {

    c(
      paste0("\\input{", filename, "}"),
      "\\clearpage"
    )

  }),
  use.names = FALSE
)

master_tex <- c(
  "\\documentclass[11pt,letterpaper]{article}",
  "\\usepackage[margin=0.75in]{geometry}",
  "\\usepackage[T1]{fontenc}",
  "\\usepackage[utf8]{inputenc}",
  "\\usepackage{booktabs,tabularx,array,threeparttable,makecell}",
  "\\begin{document}",
  input_lines,
  "\\end{document}"
)

writeLines(
  master_tex,
  con = file.path(
    wave_table_dir,
    "all_policy_wave_tables.tex"
  )
)


# ------------------------------------------------------------
# 6D. Print the completion status and exact output location
# ------------------------------------------------------------

cat(
  "Exported ",
  length(wave_latex_tables),
  " four-wave DV tables.\n",

  "Folder: ",
  normalizePath(wave_table_dir, winslash = "/"),
  "\n",

  "Overleaf main file: all_policy_wave_tables.tex\n",

  sep = ""
)
```


# III. Analysis: Ambivalence

Can use either sample_no_dk or sample_dk for this. 


## A. Name waves

DVs are different here.

```{r}

vars <- c(
  "dk_post_felony_vote_rev",
  "dk_jail_access_rev",
  "dk_min_wage",
  "dk_free_calls",
  "dk_ed_funds",
  "dk_alt_sentencing",
  "dk_dpen_rev",
  "dk_lwop_rev",
  "dk_total_all",
  "dk_total_incar_vote",
  "dk_total_q25_help",
  "dk_total_punish",
  "dk_prop_all",
  "dk_prop_incar_vote",
  "dk_prop_q25_help",
  "dk_prop_punish"
)

vars_labels <- c(
  "DK: Rest. voting rights",
  "DK: Jail voting access",
  "DK: Min wage",
  "DK: Free calls",
  "DK: Ed funding",
  "DK: Sentencing Alts",
  "DK: Death penalty",
  "DK: Life without parole",
  "Total DK responses: All policy items",
  "Total DK responses: Incarceration/voting policies",
  "Total DK responses: Help policies",
  "Total DK responses: Punishment policies",
  "Proportion DK: All policy items",
  "Proportion DK: Incarceration/voting policies",
  "Proportion DK: Help policies",
  "Proportion DK: Punishment policies"
)


# info about answered per 

vars_1 <- c(
  "dk_answered_all",
  "dk_answered_incar_vote",
  "dk_answered_q25_help",
  "dk_answered_punish"
)

vars_1_labels <- c(
  "Number of policy items answered: All",
  "Number of policy items answered: Incarceration/voting",
  "Number of policy items answered: Help policies",
  "Number of policy items answered: Punishment policies"
)


```


Controls will be applied in waves, so  group controls (tho these should already/eventually be clear in variable labels R file)

```{r}

demo_controls <- c(
  "Black", "Hisp_Latino", "AllOthers", "gender"
)

demo_controls_labels <- c(
  Black = "Black",
  Hisp_Latino    = "Hispanic/Latino", 
  AllOthers = "Other RID",
  gender = "Men"
)

party_ideo_controls <- c(
  "Republican",
  "Democrat",
  "fiscal_ideology",
  "social_ideology"
)

party_ideo_controls_labels <- c(
  # pid_3 = "Party ID",
  Republican = "Republican",
  Democrat = "Democrat",
  fiscal_ideology = "Fiscal ideology",
  social_ideology = "Social ideology"
)

fire_controls <- c(
  "fire_rare",
  "fire_privilege",
  "fire_angry",
  "fire_fear"
)

fire_controls_labels <- c(
  fire_rare = "Racism rare",
  fire_privilege = "White privilege",
  fire_angry = "Angry about racism",
  fire_fear = "Fear other races"
)

control_waves <- list(
  nocontrol = character(0),
  demo = demo_controls,
  party_ideo = c(demo_controls, party_ideo_controls),
  full_attitudes = c(demo_controls, party_ideo_controls,
                     fire_controls)
)

# , desor_core


```

```{r}

wave_order <- c(
  "nocontrol",
  "demo",
  "party_ideo",
  "full_attitudes"
)

wave_labels <- c(
  nocontrol = "No controls",
  demo = "Demographics",
  party_ideo = "+ Party/Ideology",
  full_attitudes = "+ Attitudes"
)

```

## B. Define model

### Logit or OLS? 

Basic idea is dk_policy ~ impact. The DV's are binary. 

lm() with a binary DV would give the wrong error structure and produce heteroskedastic errors.
-also assumes effects are constant in probability units

Logistic regression might be better. 
- it models a transformation of the probability (log odds)

### Logit Exploration

#### Free Calls

```{r}

m_freecalls_1 <- glm(
  dk_free_calls ~ impact,
  data = sample_no_dk,
  family = binomial(link = "logit")
)

summary(m_freecalls_1)

exp(coef(m_freecalls_1))


```
Interpretation - 

$$
log( \frac{P(DK=1)}{1-P(DK=1)} ) = B(intercept) +B(impact) = -1.785 - .968
$$
Intercept -- 

Coefficients are in log-odds units, so the -0.968 coefficient is not really substantive to interpret. Intercept is 0/1, so it describes non-impacted respondents. It is exponentiated (exp(-1.784)) giving the 0.167 Intercept. 

Convert the intercept to probability:

$$
P(DK = 1 | impact = 0) = \frac{e^-1.78}{1 + e^-1.78} = 0.144
$$
So the intercept tells that: 

Among respondents with impact = 0, the predictability of answering DK on the free calls item is about 14.4%. 

Impact --

Impact coefficient: -0.96786

The reported 0.3798937 is the odds ratio:
  - exp(-0.96786) = 0.3799 ~ 0.380
  
  - it means that the odds of answering DK among respondents with impact = 1 are about 0.38 times the odds among respondents with impact = 0. IE: impacted respondents have approximately a 62% lower odds of answering DK than do respondents with impact = 1.

In terms of probabilities, the linear predictor becomes 

$$
-1.78489-.96786=-2.75275
$$


Respondents with carceral/victimization impact were substantially less likely to select “don't know/no opinion” on the free-calls item. The predicted probability of a DK response was approximately 14.4% among respondents without impact and 6.0% among impacted respondents, a difference of roughly 8.4 percentage points.


```{r}

m_freecalls_2 <- glm(
  dk_free_calls ~ impact + Black + Hisp_Latino + AllOthers + gender,
  data = sample_no_dk,
  family = binomial(link = "logit")
)

summary(m_freecalls_2)

exp(coef(m_freecalls_2))

```

Interpretation via odds:


Model fit test with ANOVA asks:

"Does adding this set of four demographic controls improve model fit enough to provide evidence that Model 2 fits the outcome better than the simpler impact-only model?"

```{r}
anova(m_freecalls_1, m_freecalls_2, test = "Chisq")
```

Adding these four demographic variables does not significantly improve model fit over the impact-only model according to the likelihood-ratio test.

The two models have essentially indistinguishable AIC values; adding the demographic controls does not yield a clear improvement in the fit–complexity tradeoff.

Model 1: Impact is strongly associated with a lower likelihood of DK responding on free calls. The observed/model-implied probability is approximately 14.4% among non-impacted respondents and 6.0% among impacted respondents, an unadjusted difference of about 8.4 percentage points. The corresponding odds ratio is .38.

Model 2: That relationship remains essentially intact after introducing the four demographic controls. The adjusted odds ratio for impact is .39. None of the demographic variables individually provides strong evidence of an association with DK responding, and the demographic block does not significantly improve model fit over the impact-only model (\(\chi^2(4)=7.40,\ p=.116\)). Both models use the same 1,600 observations.

#### Education Funding

```{r}

m_ed_funds_1 <- glm(
  dk_ed_funds ~ impact,
  data = sample_no_dk,
  family = binomial(link = "logit")
)

summary(m_ed_funds_1)

exp(coef(m_ed_funds_1))


```
Interpretation - 

#### Alternative Sentencing

```{r}

m_alt_sentencing_1 <- glm(
  dk_alt_sentencing ~ impact,
  data = sample_no_dk,
  family = binomial(link = "logit")
)

summary(m_alt_sentencing_1)

exp(coef(m_alt_sentencing_1))


```
Interpretation - 

#### Min Wage

```{r}

m_min_wage_1 <- glm(
  dk_min_wage ~ impact,
  data = sample_no_dk,
  family = binomial(link = "logit")
)

summary(m_min_wage_1)

exp(coef(m_min_wage_1))


```
Interpretation - 

#### Rights Restoration

```{r}

m_rr_1 <- glm(
  dk_post_felony_vote_rev ~ impact,
  data = sample_no_dk,
  family = binomial(link = "logit")
)

summary(m_rr_1)

exp(coef(m_rr_1))


```
Interpretation - 

#### Jail Voting

```{r}

m_jailvote_1 <- glm(
  dk_jail_access_rev ~ impact,
  data = sample_no_dk,
  family = binomial(link = "logit")
)

summary(m_jailvote_1)

exp(coef(m_jailvote_1))

```

#### Death Penalty
```{r}

m_dpen_1 <- glm(
  dk_dpen_rev ~ impact,
  data = sample_no_dk,
  family = binomial(link = "logit")
)

summary(m_dpen_1)

exp(coef(m_dpen_1))

```


#### LWOP

```{r}

m_lwop_1 <- glm(
  dk_lwop_rev ~ impact,
  data = sample_no_dk,
  family = binomial(link = "logit")
)

summary(m_lwop_1)

exp(coef(m_lwop_1))

```
