---
title: "policy_analysis_sept25"
format: html
editor: visual
---

## Note to self

On Friday Sept 25, I am just using this doc to stay focused on fixing the section of my analysis script that iteratively loops through:

- models with no controls

- models with just demographic controls

- models with demographics and party/ideology

- models with all controls

I care about these models for:

1.  Any impact vs no impact
2.  Direct impact (incarc) vs no impact
3.  Proximal impact (knowincarc) vs no impact
4.  Victim impact (vict) vs no impact

This code also should make tables and figures. I'll come back to adjusting tables/figures later, for now I just want to make sure that I have fixed the code so it reflects a change: that the impact, incarc, knowincarc, and vict are all set so that when = 0, it means impact = 0. I want to report out, for example, how people who selected vict = 1 differ from those whose impact = 0, rather than just those whose vict = 0.

```{r setup, include=FALSE, echo=FALSE}

knitr::opts_chunk$set(
  echo = FALSE,       # Hide code by default
  warning = FALSE,
  message = FALSE
)

# Load packages
library(dplyr)
library(tidyr)
library(readr)
library(janitor)   # for clean_names()
library(tibble)
library(stringr)
library(haven)
library(gt)
library(marginaleffects)
library(mediation)
library(naniar)
library(rsq)
library(patchwork)
library(modelsummary)
library(rlang)
library(forcats)
library(ggplot2)
library(broom)
library(knitr)
library(dplyr)
library(purrr)
library(lmtest)
library(sandwich)

source("variable_labels_sept1.R")
```


```{r}

sample <- read_csv("data/cleaned_control_data.csv")

not_full <- read_csv("data/cleaned_sans_control_data.csv")
```

# ORIGINAL STARTS HERE

This section continues to use variable_labels.R. The goal of this section was to avoid and clean up extensive seperated regressions, and instead incorporate use loop that iteratively work through outcome variables and incorporate controls.

PRE 08-26, it had two sets of regression models.

Set 1 are asks "across the full sample, how is each form of impact associated with each policy outcome before adding controls (ie, prisonhelp \~ incarc) and then after/as we add controls (ie, prisonhelp \~incarc + Black + Hispanic + AllOthers + gender). (policy_dummies_waves)

Set 2 asks instead how each impacted subgroup compares specifically compares with respondents who are fully not-impacted. ie, I throw out respondents who are impacted in some other way, but not directly incarcerated; I compare impact == 0 to incarc == 1).

As of now, it (policy_vs_nonimp_waves) compares those groups (ie incarc == 1 and impact == 0) and then does so with controls.

Then this chunk builds tables.

\*\*\*08-26 I HAVE NOW RECODED incarc, knowincarc, and vict so that 1 = impacted in that particular way and 0 = NOT IMPACTED at all (ie, impact == 0).

To confirm:

```{r}
table(sample$impact)

sample %>%
  group_by(impact) %>%
  summarise(
    n = n(),
    incarc = sum(incarc, na.rm = TRUE),
    knowincarc = sum(knowincarc, na.rm = TRUE),
    vict = sum(vict, na.rm = TRUE),
  )

# Confirm the comparison-group coding
table(sample$impact, sample$incarc, useNA = "ifany")
table(sample$impact, sample$knowincarc, useNA = "ifany")
table(sample$impact, sample$vict, useNA = "ifany")
```



This means that I don't need the same 2 goal situation. I am now cleaning this up so that the main models will say:

For every policy DV:

1.  Any impact vs no impact
2.  Direct impact vs no impact
3.  Proximal impact vs no impact
4.  Victim impact vs no impact

Run each through: Model 1: no controls Model 2: demographics Model 3: demographics + party/ideology Model 4: full attitudes

### 3.1A: Framework for model sequence

control_waves \| \|-- nocontrol \|-- demo \|-- party_ideo \|-- full_attitudes

```{r}

names(sample)
# ============================================================
# 3. TO START - big picture here is defining the model "waves." The first is bivariate o no controls ("nocontrol"), then demo is models with just demographic controls (8-26 this is just race I think, then demographic controls and part ID and fiscal ideology and social ideology (party_ideo) then with all controls, so aforementioned plus FIRE and *eventually, need to fix* deservingness orientation. 
# The following controls are grouped in vriable_labels, but prepped in variable_prep_script:
#   demo_controls
#   party_ideo_controls
#   desor_core***
#   fire_controls
# ============================================================
demo_controls <- c(
  "Black", "Hisp_Latino", "AllOthers", "gender"
)

demo_controls_labels <- c(
  Black = "Black",
  Hisp_Latino    = "Hispanic/Latino", 
  AllOthers = "Other RID",
  gender = "Men"
)

# problematic_race_vars <- c(
#   "White",
#   "Black",
#   "Hisp_Latino",
#   "AllOthers"
# )
# problematic_race_labels <- c(
#   White = "White",
#   Black = "Black",
#   Hisp_Latino = "Hisp_Latino",
#   AllOtherss = "AllOthers"
# )

party_ideo_controls <- c(
  "pid_3",
  "fiscal_ideology",
  "social_ideology"
)

party_ideo_controls_labels <- c(
  pid_3 = "Party ID",
  # Republican = "Republican",
  # Democrat = "Democrat",
  fiscal_ideology = "Fiscal ideology",
  social_ideology = "Social ideology"
)

fire_controls <- c(
  "fire_rare",
  "fire_privilege",
  "fire_angry",
  "fire_fear"
)

fire_controls_labels <- c(
  fire_rare = "Racism rare",
  fire_privilege = "White privilege",
  fire_angry = "Angry about racism",
  fire_fear = "Fear other races"
)

control_waves <- list(
  nocontrol = character(0),
  demo = demo_controls,
  party_ideo = c(demo_controls, party_ideo_controls),
  full_attitudes = c(demo_controls, party_ideo_controls,
                     fire_controls, desor_core)
)
```

### 3.2 Define lm function that will waves will operate thru
# ============================================================
# This starts by defining my reusable function. That function says:

#"whenever (later) I give this function a dataset (df), a dependent variable (dv), an independent variable (iv), and some controls (controls), R should:

##(a) start with a default of assuming there are no controls, if no controls are given (character (0)); 
## (b) in the right-hand side (rhs) or the equation, use the supplied iv and supplied controls to make a vector with all vraibles that should appear to the right of ~
## (c) (fml) construct an ordinary regression syntax with the supplied dv on the left of ~, then supplied rhs info and turn it into a signle string (collapse = "+"), and 
## (d) treat this as a statistical formula rather than a string (as.formula())
## (e) run the regression
# ============================================================

```{r}
make_policy_model <- function(df, dv, iv, controls = character(0)) { #character(0) just means empty character vector so this function works for uncontrolled model then for controlled model
  rhs <- c(iv, controls)
  fml <- as.formula(paste(dv, "~", paste(rhs, collapse = " + ")))
  lm(fml, data = df)
}

```

# ============================================================
# MAIN MODEL GENERATING LOOOP Now, I need to line up buckets to put the models 
## that place is "policy_dummies_waves"
#
# Then, I need to say how I will fill it. I will fill it with waves of models that loop through dvs and iteratively add controls 
#for each dv, generates 4 substantive comparisons:
### impact versus non impact
### incarc versus non impact
### knowincarc versus non impact
### victimized vs non impact
# ============================================================

```{r}
policy_dummies_waves <- list() #this is actually now main model generating loop

for (wave_name in names(control_waves)) {#calls list of waves in control waves

  model_list <- list() #fresh model list within each wave

  for (dv in policy_vars) {
    for (iv in impact_vars) { #loop thru dvs, ivs, labels file gives the 4 ivs; fits 4 models for each outcome

      model_name <- paste0("m_", dv, "_", iv) #ie model using prisonhelp and incarc will be named m_prisonhelp_incarc

      # NOW, actually telling it to run and store the regressions
      model_list[[model_name]] <- make_policy_model(
        df = sample,
        dv = dv, 
        iv = iv,
        controls = control_waves[[wave_name]]
      )
    }
  }

  policy_dummies_waves[[wave_name]] <- model_list
}
```
##after the two inner loops finish organization for storing them. 
## object is roughly: policy_dummies_waves : A no control -> A1 no control m_prisonhelp_impact, A2 no control m_prisonhelp_incarc. . ..etc; B1 demographic controls m_prisonhelp_impact....(see below)
#policy_dummies_waves
# |
# |-- nocontrol
# |     |-- m_prisonhelp_impact
# |     |-- m_prisonhelp_incarc
# |     |-- m_prisonhelp_knowincarc
# |     |-- m_prisonhelp_vict
# |     |-- ...
# |
# |-- demo
# |     |-- m_prisonhelp_impact
# |     |-- m_prisonhelp_incarc
# |     |-- ...
# |
# |-- party_ideo
# |
# |-- full_attitudes

# ============================================================
# NEXT -- making hopefully more convenient shortcuts/references so that  instead of typing out the full "policy_dummies_waves[["nocontrol"]][["m_prisonhelp_incarc"]] I can type "policy_dummies_models[{"m_prisonhelp_incarc"}] (really tho, the table making code came first and this lets me adopt above outputs to )
# ============================================================

```{r}

policy_dummies_models <- policy_dummies_waves[["nocontrol"]]
policy_dummies_demo_models <- policy_dummies_waves[["demo"]]
policy_dummies_partyideo_models <- policy_dummies_waves[["party_ideo"]]
policy_dummies_full_models <- policy_dummies_waves[["full_attitudes"]]

```

### 3.3 Appendix Tables - Each IV Model Grouped around DV

```{r}
# ============================================================
# Build one LaTeX regression-results table per DV
# Assumes  variable-label script has already been sourced,
# so the following objects already exist:
#   - impact_vars
#   - impact_labels
#   - policy_vars
#   - policy_map
# Also assumes policy_dummies_models already exists.
# ============================================================


# ------------------------------------------------------------
# sanity checks: make sure the variables referenced in 
# helper objects actually exist in the data
# -----------------------------------------------------------

missing_policy_vars <- setdiff(policy_vars, names(sample))
missing_impact_vars <- setdiff(impact_vars, names(sample))

if (length(missing_policy_vars) > 0) {
  stop("These policy_vars are missing from sample: ",
       paste(missing_policy_vars, collapse = ", "))
}

if (length(missing_impact_vars) > 0) {
  stop("These impact_vars are missing from sample: ",
       paste(missing_impact_vars, collapse = ", "))
}

# ------------------------------------------------------------
# create a named vector of DV labels from policy_map
# ------------------------------------------------------------

policy_labels <- setNames(policy_map$label, policy_map$var)

# ------------------------------------------------------------
# helper function: convert p-values to significance stars
# ------------------------------------------------------------

p_to_stars <- function(p) {
  dplyr::case_when(
    is.na(p)  ~ "",
    p < .001  ~ "***",
    p < .01   ~ "**",
    p < .05   ~ "*",
    TRUE      ~ ""
  )
}

# ------------------------------------------------------------
# function: build a clean one-DV results table
# ------------------------------------------------------------

make_dv_results_table <- function(dv_name, model_list) {

  rows <- lapply(impact_vars, function(iv_name) {

    model_name <- paste0("m_", dv_name, "_", iv_name)

    if (!model_name %in% names(model_list)) {
      stop("Model not found in policy_dummies_models: ", model_name)
    }

    mod <- model_list[[model_name]]

    coef_row <- broom::tidy(mod) %>%
      dplyr::filter(term != "(Intercept)") %>%
      dplyr::mutate(
        iv_name = iv_name,
        iv_label = impact_labels[[iv_name]],
        coef = round(estimate, 3),
        se = round(std.error, 3),
        p = round(p.value, 3),
        stars = p_to_stars(p.value),
        coef_with_stars = paste0(format(coef, nsmall = 3), stars),
        n = broom::glance(mod)$nobs,
        r2 = round(broom::glance(mod)$r.squared, 3)
      ) %>%
      dplyr::select(iv_label, coef_with_stars, se, p, n, r2)

    coef_row
  })

  dplyr::bind_rows(rows) %>%
    dplyr::rename(
      `Independent Variable` = iv_label,
      `Coefficient` = coef_with_stars,
      `SE` = se,
      `p-value` = p,
      `N` = n,
      `R^2` = r2
    )
}

# ------------------------------------------------------------
# function: convert one DV results table into LaTeX code
# ------------------------------------------------------------

make_dv_latex_table <- function(dv_name, model_list) {

  results_df <- make_dv_results_table(dv_name, model_list)

  caption_text <- paste0("Bivariate Models Predicting ", policy_labels[[dv_name]])
  label_text <- paste0("tab:", dv_name, "_bivariate")

  knitr::kable(
    results_df,
    format = "latex",
    booktabs = TRUE,
    caption = caption_text,
    label = label_text,
    align = c("l", "c", "c", "c", "c", "c"),
    escape = TRUE
  )
}

# ------------------------------------------------------------
# create one LaTeX table per DV and store them in a named list
# ------------------------------------------------------------

dv_latex_tables <- lapply(policy_vars, function(dv_name) {
  make_dv_latex_table(dv_name, policy_dummies_models)
})

names(dv_latex_tables) <- policy_vars

# ------------------------------------------------------------
# example: print one table to the console so you can copy/paste
# into Overleaf
# ------------------------------------------------------------

#cat(dv_latex_tables[["prisonhelp"]])

# ------------------------------------------------------------
# write all tables to separate .tex files
# ------------------------------------------------------------

for (dv_name in policy_vars) {
  writeLines(
    dv_latex_tables[[dv_name]],
    con = paste0(dv_name, "_bivariate_table.tex")
  )
}
```



### 3.4 Compare Each Impacted Subgroup to Non-Impacted

### 3.5 Tables Loop
# ============================================================
# Tables
# Goal: compare each impacted subgroup to non-impacted only
# Model 2 = no-controls models
# Model 3 = placeholder for later controlled models
#
# Packages used:
# dplyr, stringr, purrr, lmtest, sandwich
# ============================================================

```{r}
# ------------------------------------------------------------
# Labels from variable script
# ------------------------------------------------------------
policy_labels <- setNames(policy_map$label, policy_map$var)

# ------------------------------------------------------------
# Formatting helpers (regression output to latex)
# ------------------------------------------------------------
p_to_stars <- function(p) {
  dplyr::case_when(
    is.na(p) ~ "",
    p < .001 ~ "***",
    p < .01  ~ "**",
    p < .05  ~ "*",
    TRUE     ~ ""
  )
}

fmt_num <- function(x, digits = 3) {
  ifelse(is.na(x), "", sprintf(paste0("%.", digits, "f"), x))
}

latex_escape <- function(x) {
  x %>%
    str_replace_all("\\\\", "\\\\\\\\") %>%
    str_replace_all("([#$%&_{}])", "\\\\\\1") %>%
    str_replace_all("~", "\\\\textasciitilde{}") %>%
    str_replace_all("\\^", "\\\\textasciicircum{}")
}

# ------------------------------------------------------------
# Split a vector into chunks for appendix tables
# ------------------------------------------------------------
chunk_vector <- function(x, chunk_size = 2) {
  split(x, ceiling(seq_along(x) / chunk_size))
}
```

# ------------------------------------------------------------
# MAIN TABLE BUILDER HERE
# -----------------------------------------------------------

```{r}
make_policy_panel_latex_table <- function(
    dv_vec,
    model2_list,
    model3_list = NULL,
    caption_text,
    label_text,
    model2_title = "Model 2",
    model3_title = "Model 3",
    row_vars = c("impact", "incarc", "knowincarc", "vict"),
    control_vars = c("Black", "Hisp_Latino", "AllOthers", "gender"),
    control_labels = c(
      Black = "Black",
      Hisp_Latino = "Hispanic/Latino",
      AllOthers = "Other Race",
      gender = "Male"
    ),
    robust = TRUE
) {

  # ----- column setup -----
  n_dv <- length(dv_vec)
  n_model_blocks <- 2
  total_data_cols <- n_dv * n_model_blocks
  align_string <- paste0("l", paste(rep("c", total_data_cols), collapse = ""))

  # ----- header rows -----
  header_top <- c(
    "",
    paste0("\\multicolumn{", n_dv, "}{c}{", latex_escape(model2_title), "}"),
    paste0("\\multicolumn{", n_dv, "}{c}{", latex_escape(model3_title), "}")
  )

  header_second <- c(
    "",
    latex_escape(policy_labels[dv_vec]),
    latex_escape(policy_labels[dv_vec])
  )

  cmid_model2 <- paste0("\\cmidrule(lr){2-", 1 + n_dv, "}")
  cmid_model3 <- paste0("\\cmidrule(lr){", 2 + n_dv, "-", 1 + (2 * n_dv), "}")
  cmid_line <- paste(cmid_model2, cmid_model3, collapse = " ")

  # ----- helper: extract a specific term from a model -----#
  
  get_term_cell <- function(model_list, dv, iv, term_name, robust = TRUE) {

    if (is.null(model_list)) return("")

    model_name <- paste0("m_", dv, "_", iv)
    if (!model_name %in% names(model_list)) return("")

    mod <- model_list[[model_name]]

    if (robust) {
      ct <- lmtest::coeftest(mod, vcov. = sandwich::vcovHC(mod, type = "HC1"))
      out <- data.frame(
        term      = rownames(ct),
        estimate  = ct[, 1],
        std.error = ct[, 2],
        p.value   = ct[, 4],
        stringsAsFactors = FALSE
      )
    } else {
      sm <- summary(mod)$coefficients
      out <- data.frame(
        term      = rownames(sm),
        estimate  = sm[, 1],
        std.error = sm[, 2],
        p.value   = sm[, 4],
        stringsAsFactors = FALSE
      )
    }

    if (!term_name %in% out$term) return("")

    row <- out[out$term == term_name, ]

    coef_txt <- paste0(
      fmt_num(row$estimate, 3),
      p_to_stars(row$p.value)
    )

    se_txt <- paste0("(", fmt_num(row$std.error, 3), ")")

    paste0("\\makecell[c]{", coef_txt, " \\\\ ", se_txt, "}")
  }

  # ----- body rows -----
  body_lines <- c()

  # Section header
  body_lines <- c(
    body_lines,
    paste(
      c(
        "\\textit{Carceral Contact (ref. Not impacted)}",
        rep("", total_data_cols)
      ),
      collapse = " & "
    ),
    "\\\\"
  )

  # For each subgroup, print coefficient row + repeated control rows
  for (iv in row_vars) {

    row_label <- impact_labels[[iv]]
    main_term <- iv

    # main subgroup row
    model2_cells <- sapply(dv_vec, function(dv) {
      get_term_cell(model2_list, dv, iv, main_term, robust = robust)
    })

    model3_cells <- sapply(dv_vec, function(dv) {
      get_term_cell(model3_list, dv, iv, main_term, robust = robust)
    })

    subgroup_line <- paste(
      c(latex_escape(row_label), model2_cells, model3_cells),
      collapse = " & "
    )

    body_lines <- c(body_lines, paste0(subgroup_line, " \\\\"))

    # repeated control rows under this subgroup
    for (ctrl in control_vars) {

      model2_ctrl_cells <- sapply(dv_vec, function(dv) {
        get_term_cell(model2_list, dv, iv, ctrl, robust = robust)
      })

      model3_ctrl_cells <- sapply(dv_vec, function(dv) {
        get_term_cell(model3_list, dv, iv, ctrl, robust = robust)
      })

      ctrl_line <- paste(
        c(latex_escape(control_labels[[ctrl]]), model2_ctrl_cells, model3_ctrl_cells),
        collapse = " & "
      )

      body_lines <- c(body_lines, paste0(ctrl_line, " \\\\"))
    }

    body_lines <- c(body_lines, "\\addlinespace")
  }

  # ----- assemble LaTeX -----
  out <- c(
    "\\begin{table}[!htbp]",
    "\\centering",
    paste0("\\caption{", latex_escape(caption_text), "}"),
    paste0("\\label{", latex_escape(label_text), "}"),
    "\\begin{threeparttable}",
    paste0("\\begin{tabular}{", align_string, "}"),
    "\\toprule",
    paste(header_top, collapse = " & "),
    " \\\\",
    cmid_line,
    paste(header_second, collapse = " & "),
    " \\\\",
    "\\midrule",
    body_lines,
    "\\bottomrule",
    "\\end{tabular}",
    "\\begin{tablenotes}[flushleft]",
    "\\footnotesize",
    "\\item Entries are OLS coefficients with HC1 robust standard errors in parentheses.",
    "\\item * p$<$0.05, ** p$<$0.01, *** p$<$0.001.",
    "\\end{tablenotes}",
    "\\end{threeparttable}",
    "\\end{table}"
  )

  out
}
```

TEll the function which DVs and model waves to display 

```{r}
results_dvs <- c("prisonhelp", "dpen_rev", "lwop_rev")

results_table_tex <- make_policy_panel_latex_table(
  dv_vec = results_dvs,
  model2_list = policy_dummies_models,
  model3_list = policy_dummies_demo_models,
  caption_text = "OLS Models of Policy Ors - None to Demographic Controls",
  label_text = "tab:carceral_policy_ors_nocontrol_demo",
  model2_title = "Model 2",
  model3_title = "Model 3",
  row_vars = c("impact", "incarc", "knowincarc", "vict"),
  control_vars = c("Black", "Hisp_Latino", "AllOthers", "gender"),
  control_labels = c(
    Black = "Black",
    Hisp_Latino = "Hispanic/Latino",
    AllOthers = "Other_Race",
    gender = "Male"
  ),
  robust = TRUE
)

writeLines(results_table_tex, "results_policy_orientations_nocontrol_demo_table.tex")
```


```{r}

##model for just desor

dep_variable <-(sample$knowincarc)
desor_model <- lm(desor_core ~ dep_variable + Black + Hisp_Latino + AllOthers + fiscal_ideology + social_ideology, data = sample)

writeLines(results_table_tex, "desor_controls_table.tex")
```



```{r}
# ------------------------------------------------------------
# Results Table
# ------------------------------------------------------------

results_dvs <- c("prisonhelp", "dpen_rev", "lwop_rev")

results_table_tex <- make_policy_panel_latex_table(
  dv_vec = results_dvs,
  model2_list = policy_vs_nonimp_nocontrol_models,
  model3_list = policy_vs_nonimp_demo_models,
  caption_text = "OLS Models of Carceral Policy Orientations - No controls to Demographic Controls",
  label_text = "tab:carceral_policy_orientations_nocontrol_demo",
  model2_title = "Model 2",
  model3_title = "Model 3",
  row_vars = c("impact", "incarc", "knowincarc", "vict"),
  control_vars = c("Black", "Hisp_Latino", "AllOthers", "gender"),
  control_labels = c(
    Black = "Black",
    Hisp_Latino = "Hispanic/Latino",
    AllOthers = "Other race",
    gender = "Male"
  ),
  robust = TRUE
)

writeLines(results_table_tex, "results_policy_orientations_nocontrol_demo_table.tex")

# Optional preview in console
# cat(results_table_tex, sep = "\n")


# ------------------------------------------------------------
# 2. APPENDIX TABLES FOR REMAINING POLICY DVs
#    Same architecture, chunked two DVs at a time
# ------------------------------------------------------------

appendix_dvs <- setdiff(policy_vars, results_dvs)

# You can keep help/punish separated if you want cleaner appendix organization
appendix_help_dvs <- appendix_dvs[policy_map$section[match(appendix_dvs, policy_map$var)] == "Help"]
appendix_punish_dvs <- appendix_dvs[policy_map$section[match(appendix_dvs, policy_map$var)] == "Punish"]

appendix_help_chunks <- chunk_vector(appendix_help_dvs, chunk_size = 2)
appendix_punish_chunks <- chunk_vector(appendix_punish_dvs, chunk_size = 2)

appendix_tables <- list()

# Help tables
for (i in seq_along(appendix_help_chunks)) {
  dv_chunk <- appendix_help_chunks[[i]]

  tex <- make_policy_panel_latex_table(
    dv_vec = dv_chunk,
    model2_list = policy_vs_nonimp_nocontrol_models,
    model3_list = NULL,
    caption_text = paste0("Restricted-Sample OLS Models of Help-Oriented Policy Outcomes, Appendix Table ", i),
    label_text = paste0("tab:appendix_help_restricted_", i),
    model2_title = "Model 2",
    model3_title = "Model 3",
    row_vars = c("impact", "incarc", "knowincarc", "vict"),
    include_placeholder_control_rows = TRUE,
    robust = TRUE
  )

  appendix_tables[[paste0("help_", i)]] <- tex
  writeLines(tex, paste0("appendix_help_restricted_table_", i, ".tex"))
}

# Punish tables
for (i in seq_along(appendix_punish_chunks)) {
  dv_chunk <- appendix_punish_chunks[[i]]

  tex <- make_policy_panel_latex_table(
    dv_vec = dv_chunk,
    model2_list = policy_vs_nonimp_nocontrol_models,
    model3_list = NULL,
    caption_text = paste0("Restricted-Sample OLS Models of Punitive Policy Outcomes, Appendix Table ", i),
    label_text = paste0("tab:appendix_punish_restricted_", i),
    model2_title = "Model 2",
    model3_title = "Model 3",
    row_vars = c("impact", "incarc", "knowincarc", "vict"),
    include_placeholder_control_rows = TRUE,
    robust = TRUE
  )

  appendix_tables[[paste0("punish_", i)]] <- tex
  writeLines(tex, paste0("appendix_punish_restricted_table_", i, ".tex"))
}
```

Looking just at prison help and prison pun after adding in demographic factors

```{r}
# summary(policy_vs_nonimp_demo_models$comp_prisonhelp_impact)
# summary(policy_vs_nonimp_demo_models$comp_prisonpen_impact)
# 
# summary(policy_vs_nonimp_demo_models$comp_prisonhelp_incarc)
# summary(policy_vs_nonimp_demo_models$comp_prisonpen_incarc)
# 
# summary(policy_vs_nonimp_demo_models$comp_prisonhelp_knowincarc)
# summary(policy_vs_nonimp_demo_models$comp_prisonpen_knowincarc)
# 
# summary(policy_vs_nonimp_demo_models$comp_prisonhelp_vict)
# summary(policy_vs_nonimp_demo_models$comp_prisonpen_vict)
# 
# 
# comp_prisonhelp_incarc     
# comp_prisonhelp_knowincarc  
# comp_prisonhelp_vict       
# comp_prisonpen_impact       
# comp_prisonpen_incarc      
# comp_prisonpen_knowincarc  
# comp_prisonpen_vict        


results_dvs <- c("prisonhelp", "dpen_rev", "lwop_rev")

results_table_tex <- make_policy_panel_latex_table(
  dv_vec = results_dvs,
  model2_list = policy_vs_nonimp_nocontrol_models,
  model3_list = policy_vs_nonimp_demo_models,
  caption_text = "OLS Models of Carceral Policy Orientations - No controls to Demographic Controls",
  label_text = "tab:carceral_policy_orientations_nocontrol_demo",
  model2_title = "Model 2",
  model3_title = "Model 3",
  row_vars = c("impact", "incarc", "knowincarc", "vict"),
  include_placeholder_control_rows = TRUE,
  robust = TRUE
)

writeLines(results_table_tex, "results_policy_orientations_nocontrol_demo_table.tex")
```
