---
title: "policy_analysis_sept25"
format: html
editor: visual
---

## Note to self

On Friday Sept 25, I am just using this doc to stay focused on fixing the section of my analysis script that iteratively loops through:

- models with no controls

- models with just demographic controls

- models with demographics and party/ideology

- models with all controls

I care about these models for:

1.  Any impact vs no impact
2.  Direct impact (incarc) vs no impact
3.  Proximal impact (knowincarc) vs no impact
4.  Victim impact (vict) vs no impact

This code also should make tables and figures. I'll come back to adjusting tables/figures later, for now I just want to make sure that I have fixed the code so it reflects a change: that the impact, incarc, knowincarc, and vict are all set so that when = 0, it means impact = 0. I want to report out, for example, how people who selected vict = 1 differ from those whose impact = 0, rather than just those whose vict = 0.

```{r setup, include=FALSE, echo=FALSE}

knitr::opts_chunk$set(
  echo = FALSE,       # Hide code by default
  warning = FALSE,
  message = FALSE
)

# Load packages
library(dplyr)
library(tidyr)
library(readr)
library(janitor)   # for clean_names()
library(tibble)
library(stringr)
library(haven)
library(gt)
library(marginaleffects)
library(mediation)
library(naniar)
library(rsq)
library(patchwork)
library(modelsummary)
library(rlang)
library(forcats)
library(ggplot2)
library(broom)
library(knitr)
library(dplyr)
library(purrr)
library(lmtest)
library(sandwich)

source("variable_labels_sept1.R")
```


```{r}

sample <- read_csv("data/cleaned_control_data.csv")

not_full <- read_csv("data/cleaned_sans_control_data.csv")

names(sample)
```



# Policy Certainty

Difs in 2.5 by impact groups
```{r}
sample <- sample %>%
  mutate(
    noimpact = ifelse(impact == 0, 1, 0)
    )

table(sample$knowincarc)

help_nums_by_impact <- sample %>%
  group_by(prisonhelp) %>%
  summarise(
    n = n(),
    incarc = sum(incarc, na.rm = TRUE),
    knowincarc = sum(knowincarc, na.rm = TRUE),
    vict = sum(vict, na.rm = TRUE),
    impact = sum(impact, na.rm = TRUE),
    noimpact = sum(noimpact, na.rm = TRUE)
  )
#print(help_nums_by_impact, n = 23)
```

Uncertainty across help vars
```{r}
m <- m_prisonhelp_impact_verify
mf <- model.frame(m)

prison_items <- c(
  policy_vars,
  police_items
)

# 
# prison_items <- c(
#   "min_wage",
#   "free_calls",
#   "ed_funds",
#   "alt_sentencing",
#   "lwop_rev",
#   "dpen_rev",
#   
# )

# Separate audit object
audit <- sample

# Identify complete cases for the current model's variables
model_vars <- all.vars(formula(m))

audit$in_model <- complete.cases(
  audit[, model_vars, drop = FALSE]
)

# Check that these reproduce the fitted model's observations
audit_mf <- audit[audit$in_model, names(mf), drop = FALSE]

stopifnot(
  nrow(audit_mf) == nrow(mf),
  all(vapply(
    names(mf),
    function(v) {
      isTRUE(all.equal(
        audit_mf[[v]],
        mf[[v]],
        check.attributes = FALSE
      ))
    },
    logical(1)
  ))
)

# Individual policy responses, not the prisonhelp average
help_mat <- as.matrix(audit[, help_items, drop = FALSE])

# Check the expected coding
stopifnot(
  is.numeric(help_mat),
  all(is.na(help_mat) | help_mat %in% c(1, 2, 2.5, 3, 4))
)

audit$n_help_answered <- rowSums(!is.na(help_mat))
audit$n_help_dk <- rowSums(help_mat == 2.5, na.rm = TRUE)

# Compare actual DK use within the regression sample
audit %>%
  dplyr::filter(in_model) %>%
  dplyr::group_by(impact) %>%
  dplyr::summarise(
    n = dplyr::n(),
    fewer_than_four_answers = sum(n_help_answered < 4),
    any_dk_n = sum(n_help_dk > 0),
    any_dk_pct = 100 * mean(n_help_dk > 0),
    all_four_dk_n = sum(n_help_dk == 4),
    midpoint_with_four_answers_and_no_dk = sum(
      prisonhelp == 2.5 &
        n_help_answered == 4 &
        n_help_dk == 0
    ),
    .groups = "drop"
  )

# Decompose the midpoint respondents by their actual response pattern
audit %>%
  dplyr::filter(in_model, prisonhelp == 2.5) %>%
  dplyr::count(impact, n_help_answered, n_help_dk)
```




# Policy Prefs Loop

This section continues to use variable_labels.R. The goal of this section was to avoid and clean up extensive seperated regressions, and instead incorporate use loop that iteratively work through outcome variables and incorporate controls.

PRE 08-26, it had two sets of regression models.

Set 1 are asks "across the full sample, how is each form of impact associated with each policy outcome before adding controls (ie, prisonhelp \~ incarc) and then after/as we add controls (ie, prisonhelp \~incarc + Black + Hispanic + AllOthers + gender). (policy_dummies_waves)

Set 2 asks instead how each impacted subgroup compares specifically compares with respondents who are fully not-impacted. ie, I throw out respondents who are impacted in some other way, but not directly incarcerated; I compare impact == 0 to incarc == 1).

As of now, it (policy_vs_nonimp_waves) compares those groups (ie incarc == 1 and impact == 0) and then does so with controls.

Then this chunk builds tables.

\*\*\*08-26 I HAVE NOW RECODED incarc, knowincarc, and vict so that 1 = impacted in that particular way and 0 = NOT IMPACTED at all (ie, impact == 0).

To confirm:

```{r}
table(sample$impact)

sample %>%
  group_by(impact) %>%
  summarise(
    n = n(),
    incarc = sum(incarc, na.rm = TRUE),
    knowincarc = sum(knowincarc, na.rm = TRUE),
    vict = sum(vict, na.rm = TRUE),
  )

# Confirm the comparison-group coding
table(sample$impact, sample$incarc, useNA = "ifany")
table(sample$impact, sample$knowincarc, useNA = "ifany")
table(sample$impact, sample$vict, useNA = "ifany")
```



This means that I don't need the same 2 goal situation. I am now cleaning this up so that the main models will say:

For every policy DV:

1.  Any impact vs no impact
2.  Direct impact vs no impact
3.  Proximal impact vs no impact
4.  Victim impact vs no impact

Run each through: Model 1: no controls Model 2: demographics Model 3: demographics + party/ideology Model 4: full attitudes

### 3.1A: Framework for model sequence

control_waves \| \|-- nocontrol \|-- demo \|-- party_ideo \|-- full_attitudes

```{r}

names(sample)
# ============================================================
# 3. TO START - big picture here is defining the model "waves." The first is bivariate o no controls ("nocontrol"), then demo is models with just demographic controls (8-26 this is just race I think, then demographic controls and part ID and fiscal ideology and social ideology (party_ideo) then with all controls, so aforementioned plus FIRE and *eventually, need to fix* deservingness orientation. 
# The following controls are grouped in vriable_labels, but prepped in variable_prep_script:
#   demo_controls
#   party_ideo_controls
#   desor_core***
#   fire_controls
# ============================================================
demo_controls <- c(
  "Black", "Hisp_Latino", "AllOthers", "gender"
)

demo_controls_labels <- c(
  Black = "Black",
  Hisp_Latino    = "Hispanic/Latino", 
  AllOthers = "Other RID",
  gender = "Men"
)

# problematic_race_vars <- c(
#   "White",
#   "Black",
#   "Hisp_Latino",
#   "AllOthers"
# )
# problematic_race_labels <- c(
#   White = "White",
#   Black = "Black",
#   Hisp_Latino = "Hisp_Latino",
#   AllOtherss = "AllOthers"
# )

party_ideo_controls <- c(
  "Republican",
  "Democrat",
  "fiscal_ideology",
  "social_ideology"
)

party_ideo_controls_labels <- c(
  # pid_3 = "Party ID",
  Republican = "Republican",
  Democrat = "Democrat",
  fiscal_ideology = "Fiscal ideology",
  social_ideology = "Social ideology"
)

fire_controls <- c(
  "fire_rare",
  "fire_privilege",
  "fire_angry",
  "fire_fear"
)

fire_controls_labels <- c(
  fire_rare = "Racism rare",
  fire_privilege = "White privilege",
  fire_angry = "Angry about racism",
  fire_fear = "Fear other races"
)

control_waves <- list(
  nocontrol = character(0),
  demo = demo_controls,
  party_ideo = c(demo_controls, party_ideo_controls),
  full_attitudes = c(demo_controls, party_ideo_controls,
                     fire_controls, desor_core)
)
```

### 3.2 Define lm function that will waves will operate thru
# ============================================================
# Define Resusable function
# ============================================================
This starts by defining my reusable function. That function says:

"whenever (later) I give this function a dataset (df), a dependent variable (dv), an independent variable (iv), and some controls (controls), R should:

(a) start with a default of assuming there are no controls, if no controls are given (character (0)); 
(b) in the right-hand side (rhs) or the equation, use the supplied iv and supplied controls to make a vector with all vraibles that should appear to the right of ~
(c) (fml) construct an ordinary regression syntax with the supplied dv on the left of ~, then supplied rhs info and turn it into a signle string (collapse = "+"), and 
 (d) treat this as a statistical formula rather than a string (as.formula())
(e) run the regression
```{r}
make_policy_model <- function(df, dv, iv, controls = character(0)) { #character(0) just means empty character vector so this function works for uncontrolled model then for controlled model
  rhs <- c(iv, controls)
  fml <- as.formula(paste(dv, "~", paste(rhs, collapse = " + ")))
  lm(fml, data = df)
}

```

# ============================================================
# MAIN MODEL GENERATING LOOOP 
# ============================================================
Now, I need to line up buckets to put the models 
 that place is "policy_dummies_waves"

 Then, I need to say how I will fill it. I will fill it with waves of models that loop through dvs and iteratively add controls 
for each dv, generates 4 substantive comparisons:
 impact versus non impact incarc versus non impact
 knowincarc versus non impact
victimized vs non impact
```{r}
policy_dummies_waves <- list() #this is actually now main model generating loop

for (wave_name in names(control_waves)) {#calls list of waves in control waves

  model_list <- list() #fresh model list within each wave

  for (dv in policy_vars) {
    for (iv in impact_vars) { #loop thru dvs, ivs, labels file gives the 4 ivs; fits 4 models for each outcome

      model_name <- paste0("m_", dv, "_", iv) #ie model using prisonhelp and incarc will be named m_prisonhelp_incarc

      # NOW, actually telling it to run and store the regressions
      model_list[[model_name]] <- make_policy_model(
        df = sample,
        dv = dv, 
        iv = iv,
        controls = control_waves[[wave_name]]
      )
    }
  }

  policy_dummies_waves[[wave_name]] <- model_list
}
```
##after the two inner loops finish organization for storing them. 
## object is roughly: policy_dummies_waves : A no control -> A1 no control m_prisonhelp_impact, A2 no control m_prisonhelp_incarc. . ..etc; B1 demographic controls m_prisonhelp_impact....(see below)
#policy_dummies_waves
# |
# |-- nocontrol
# |     |-- m_prisonhelp_impact
# |     |-- m_prisonhelp_incarc
# |     |-- m_prisonhelp_knowincarc
# |     |-- m_prisonhelp_vict
# |     |-- ...
# |
# |-- demo
# |     |-- m_prisonhelp_impact
# |     |-- m_prisonhelp_incarc
# |     |-- ...
# |
# |-- party_ideo
# |
# |-- full_attitudes

# ============================================================
# NEXT -- making hopefully more convenient shortcuts/references so that  instead of typing out the full "policy_dummies_waves[["nocontrol"]][["m_prisonhelp_incarc"]] I can type "policy_dummies_models[{"m_prisonhelp_incarc"}] (really tho, the table making code came first and this lets me adopt above outputs to )
# ============================================================

```{r}

policy_dummies_models <- policy_dummies_waves[["nocontrol"]]
policy_dummies_demo_models <- policy_dummies_waves[["demo"]]
policy_dummies_partyideo_models <- policy_dummies_waves[["party_ideo"]]
policy_dummies_full_models <- policy_dummies_waves[["full_attitudes"]]

#names(policy_dummies_full_models)
```

# ------------------------------------------------------------
# START 4-Wave Results Table SEPT 27 -- CREATING LATEX FOR ALL TABLES FOR INSPECTION
# ------------------------------------------------------------

First, define order and printed names of the 4 model waves 

Second, create one helper whose only job is to extract the focal IV result from one model

Third, create a function that builds one table for one DV

Fourth, arrange extracted results into the 4 column table

Fifth, create second function that turns that one DV results object into LaTex

Sixth, loop across every DV and write one .tex file per DV. 

# ------------------------------------------------------------
# 1. Define model-wave order and display labels
# ------------------------------------------------------------


```{r}

wave_order <- c(
  "nocontrol",
  "demo",
  "party_ideo",
  "full_attitudes"
)

wave_labels <- c(
  nocontrol = "No controls",
  demo = "Demographics",
  party_ideo = "+ Party/Ideology",
  full_attitudes = "+ Attitudes"
)

# DV labels already defined in variable_labels_sept1.R through policy_map
policy_labels <- setNames(
  policy_map$label,
  policy_map$var
)

# Sanity check: all four desired waves should exist
missing_waves <- setdiff(wave_order, names(policy_dummies_waves))

if (length(missing_waves) > 0) {
  stop(
    "These model waves are missing from policy_dummies_waves: ",
    paste(missing_waves, collapse = ", ")
  )
}

```

```{r}
# ------------------------------------------------------------
# 2. Helpers for extracting and formatting one focal estimate
# ------------------------------------------------------------

p_to_stars <- function(p) {
  dplyr::case_when(
    is.na(p) ~ "",
    p < .001 ~ "***",
    p < .01  ~ "**",
    p < .05  ~ "*",
    TRUE     ~ ""
  )
}

fmt_num <- function(x, digits = 3) {
  ifelse(
    is.na(x),
    "",
    sprintf(paste0("%.", digits, "f"), x)
  )
}


extract_policy_wave_stats <- function(
    dv_name,
    iv_name,
    wave_name,
    model_waves = policy_dummies_waves,
    robust = TRUE
) {

  # Model names follow the naming system created in your main loop:
  # e.g. m_prisonhelp_incarc
  model_name <- paste0("m_", dv_name, "_", iv_name)

  # Check that requested wave exists
  if (!wave_name %in% names(model_waves)) {
    stop("Wave not found: ", wave_name)
  }

  # Pull that wave's model list
  model_list <- model_waves[[wave_name]]

  # Check that requested model exists
  if (!model_name %in% names(model_list)) {
    stop(
      "Model not found in wave '",
      wave_name,
      "': ",
      model_name
    )
  }

  # Pull the actual lm object
  mod <- model_list[[model_name]]

  # Extract coefficient information
  if (robust) {

    ct <- lmtest::coeftest(
      mod,
      vcov. = sandwich::vcovHC(mod, type = "HC1")
    )

    coef_table <- data.frame(
      term = rownames(ct),
      estimate = ct[, 1],
      std.error = ct[, 2],
      p.value = ct[, 4],
      row.names = NULL
    )

  } else {

    sm <- summary(mod)$coefficients

    coef_table <- data.frame(
      term = rownames(sm),
      estimate = sm[, 1],
      std.error = sm[, 2],
      p.value = sm[, 4],
      row.names = NULL
    )
  }

  # The focal coefficient should have the same name as the IV:
  # impact, incarc, knowincarc, or vict
  if (!iv_name %in% coef_table$term) {
    stop(
      "Term '",
      iv_name,
      "' not found in model ",
      model_name
    )
  }

  focal_row <- coef_table[
    coef_table$term == iv_name,
  ]

  # Return one tidy row
  tibble::tibble(
    dv = dv_name,
    dv_label = policy_labels[[dv_name]],
    iv = iv_name,
    iv_label = impact_labels[[iv_name]],
    wave = wave_name,
    wave_label = wave_labels[[wave_name]],
    estimate = focal_row$estimate,
    std.error = focal_row$std.error,
    p.value = focal_row$p.value,
    n = stats::nobs(mod),
    r2 = summary(mod)$r.squared
  )
}
```

```{r}
# ------------------------------------------------------------
# 3. Gather all four IVs x all four waves for one DV
# ------------------------------------------------------------

make_wave_results_long <- function(
    dv_name,
    model_waves = policy_dummies_waves,
    row_vars = impact_vars,
    waves = wave_order,
    robust = TRUE
) {

  # Confirm requested DV belongs to current policy outcome set
  if (!dv_name %in% policy_vars) {
    stop(
      "DV not found in policy_vars: ",
      dv_name
    )
  }

  rows <- lapply(row_vars, function(iv_name) {

    wave_rows <- lapply(waves, function(wave_name) {

      extract_policy_wave_stats(
        dv_name = dv_name,
        iv_name = iv_name,
        wave_name = wave_name,
        model_waves = model_waves,
        robust = robust
      )
    })

    dplyr::bind_rows(wave_rows)
  })

  dplyr::bind_rows(rows) %>%
    dplyr::mutate(
      iv = factor(
        iv,
        levels = row_vars
      ),
      wave = factor(
        wave,
        levels = waves
      )
    )
}
```


```{r}
# ------------------------------------------------------------
# 4. Build one wide inspection table for one DV
# ------------------------------------------------------------

make_wave_results_table <- function(
    dv_name,
    model_waves = policy_dummies_waves,
    row_vars = impact_vars,
    waves = wave_order,
    robust = TRUE
) {

  results_long <- make_wave_results_long(
    dv_name = dv_name,
    model_waves = model_waves,
    row_vars = row_vars,
    waves = waves,
    robust = robust
  ) %>%
    dplyr::mutate(

      # What will eventually appear in each LaTeX table cell
      result = paste0(
        fmt_num(estimate, 3),
        p_to_stars(p.value),
        " (",
        fmt_num(std.error, 3),
        ")"
      ),

      # Preserve desired display order
      iv_label = factor(
        iv_label,
        levels = unname(impact_labels[row_vars])
      ),

      wave_label = factor(
        wave_label,
        levels = unname(wave_labels[waves])
      )
    )

  results_wide <- results_long %>%
    dplyr::select(
      iv_label,
      wave_label,
      result
    ) %>%
    tidyr::pivot_wider(
      names_from = wave_label,
      values_from = result
    ) %>%
    dplyr::arrange(iv_label) %>%
    dplyr::rename(
      `Carceral contact` = iv_label
    )

  results_wide
}
```

# =================================================
# VALIDATE THE FOUR-WAVE TABLE BEFORE ADDING LATEX
# =================================================

```{r}

# 
# # 1. Recreate both outputs using the functions from Steps 3 and 4
# prisonhelp_wave_long <- make_wave_results_long(
#   dv_name = "prisonhelp",
#   robust = TRUE
# )
# 
# prisonhelp_wave_table <- make_wave_results_table(
#   dv_name = "prisonhelp",
#   robust = TRUE
# )
# 
# # ------------------------------------------------------------
# # 2. Check the structure and labels
# # ------------------------------------------------------------
# 
# stopifnot(
#   # Four comparisons x four waves = 16 model results
#   nrow(prisonhelp_wave_long) == 16L,
# 
#   # No duplicated comparison-wave combinations
#   anyDuplicated(prisonhelp_wave_long[c("iv", "wave")]) == 0L,
# 
#   # Display table: four rows and five columns
#   # (comparison label + four model-wave columns)
#   nrow(prisonhelp_wave_table) == 4L,
#   ncol(prisonhelp_wave_table) == 5L,
# 
#   # Columns appear in the intended wave order
#   identical(
#     names(prisonhelp_wave_table),
#     c("Carceral contact", unname(wave_labels[wave_order]))
#   ),
# 
#   # Rows use the existing impact labels in the intended order
#   identical(
#     as.character(prisonhelp_wave_table[["Carceral contact"]]),
#     as.character(impact_labels[impact_vars])
#   )
# )
# 
# # ------------------------------------------------------------
# # 3. Check all 16 extracted results against their stored models
# # ------------------------------------------------------------
# 
# for (i in seq_len(nrow(prisonhelp_wave_long))) {
# 
#   check_row <- prisonhelp_wave_long[i, ]
# 
#   # These columns were made into factors; use their text names
#   iv_name <- as.character(check_row$iv)
#   wave_name <- as.character(check_row$wave)
# 
#   model_name <- paste0("m_prisonhelp_", iv_name)
# 
#   # Retrieve the original stored model directly
#   mod <- policy_dummies_waves[[wave_name]][[model_name]]
# 
#   # Calculate HC1 inference directly from that model
#   ct <- lmtest::coeftest(
#     mod,
#     vcov. = sandwich::vcovHC(mod, type = "HC1")
#   )
# 
#   # Values obtained directly from the stored model:
#   # coefficient, HC1 SE, p-value, N, R-squared
#   expected <- c(
#     unname(ct[iv_name, c(1, 2, 4)]),
#     stats::nobs(mod),
#     summary(mod)$r.squared
#   )
# 
#   # Corresponding values returned by our extraction function
#   extracted <- c(
#     check_row$estimate,
#     check_row$std.error,
#     check_row$p.value,
#     check_row$n,
#     check_row$r2
#   )
# 
#   # Stop and identify the affected model if anything is missing
#   # or fails to match within a very small numerical tolerance
#   if (
#     any(!is.finite(extracted)) ||
#     !isTRUE(all.equal(
#       unname(extracted),
#       unname(expected),
#       tolerance = 1e-10
#     ))
#   ) {
#     stop(
#       "Validation failed for ", model_name,
#       " in wave '", wave_name, "'.",
#       call. = FALSE
#     )
#   }
# }
# 
# message(
#   "PASS: table structure and labels are correct; ",
#   "all 16 extracted results match their stored models."
# )

# ------------------------------------------------------------
# 4. Display the four-wave table
# ------------------------------------------------------------

# print(prisonhelp_wave_table, width = Inf)

```



```{r}
# ============================================================
# STEP 5: ONE FOUR-WAVE LATEX TABLE PER DV
#
# Assumes Steps 1-4 have already run, including:
#   make_wave_results_long()
#   fmt_num()
#   p_to_stars()
#   wave_order
#   wave_labels
#
# Outcome and comparison labels are already attached by
# make_wave_results_long(), using policy_labels and impact_labels.
# ============================================================

# ------------------------------------------------------------
# 5A. Escape plain-text labels for LaTeX
# This leaves your older latex_escape() function untouched.
# ------------------------------------------------------------

wave_tex_escape <- function(x) {

  replacements <- c(
    "\\" = "\\textbackslash{}",
    "&" = "\\&",
    "%" = "\\%",
    "$" = "\\$",
    "#" = "\\#",
    "_" = "\\_",
    "{" = "\\{",
    "}" = "\\}",
    "~" = "\\textasciitilde{}",
    "^" = "\\textasciicircum{}"
  )

  vapply(as.character(x), function(txt) {

    if (is.na(txt)) {
      stop("A table label is missing.", call. = FALSE)
    }

    chars <- strsplit(txt, "", fixed = TRUE)[[1]]
    positions <- match(chars, names(replacements))
    replace <- !is.na(positions)

    chars[replace] <- unname(
      replacements[positions[replace]]
    )

    paste0(chars, collapse = "")

  }, character(1), USE.NAMES = FALSE)
}


# ------------------------------------------------------------
# 5B. Build one four-wave LaTeX table for one DV
#
# For each comparison, print:
#   coefficient with SE underneath
#   N
#   R-squared
# ------------------------------------------------------------

make_wave_latex_table <- function(
    dv_name,
    model_waves = policy_dummies_waves,
    row_vars = impact_vars,
    waves = wave_order,
    robust = TRUE,
    digits = 3
) {

  if (length(waves) != 4L || anyDuplicated(waves)) {
    stop(
      "Supply four distinct model waves.",
      call. = FALSE
    )
  }

  # Use the checked extraction function from Step 3.
  # This retains unrounded statistics, N, R2, and existing labels.
  results <- make_wave_results_long(
    dv_name = dv_name,
    model_waves = model_waves,
    row_vars = row_vars,
    waves = waves,
    robust = robust
  )

  results$iv <- as.character(results$iv)
  results$wave <- as.character(results$wave)

  # Do not silently export missing or duplicated results.
  if (
    nrow(results) != length(row_vars) * length(waves) ||
    anyDuplicated(results[c("iv", "wave")])
  ) {
    stop(
      "Missing or duplicated model results for: ",
      dv_name,
      call. = FALSE
    )
  }

  numeric_cols <- c(
    "estimate", "std.error", "p.value", "n", "r2"
  )

  if (any(!is.finite(as.matrix(results[numeric_cols])))) {
    stop(
      "Non-finite model statistics for: ",
      dv_name,
      call. = FALSE
    )
  }

  # Use the outcome label already obtained from policy_map.
  dv_label <- unique(as.character(results$dv_label))

  if (
    length(dv_label) != 1L ||
    anyNA(dv_label) ||
    !nzchar(dv_label)
  ) {
    stop(
      "Check the policy_map label for: ",
      dv_name,
      call. = FALSE
    )
  }

  # Join cells and add the LaTeX end-of-row marker.
  tex_row <- function(cells) {
    paste0(paste(cells, collapse = " & "), " \\\\")
  }

  body_lines <- character(0)

  # Build one comparison block at a time.
  for (iv_name in row_vars) {

    block <- results[results$iv == iv_name, ]

    # Put the four results in the specified wave order.
    block <- block[match(waves, block$wave), ]

    if (anyNA(block$wave)) {
      stop(
        "A wave is missing for ",
        dv_name, " / ", iv_name,
        call. = FALSE
      )
    }

    # Use the comparison label supplied through impact_labels.
    iv_label <- as.character(block$iv_label[1])

    # Round only for display.
    # Significance stars use the unrounded p-values.
    cells <- paste0(
      "\\makecell[c]{",
      fmt_num(block$estimate, digits),
      p_to_stars(block$p.value),
      " \\\\ (",
      fmt_num(block$std.error, digits),
      ")}"
    )

    body_lines <- c(
      body_lines,

      # Coefficient and SE row
      tex_row(c(
        wave_tex_escape(iv_label),
        cells
      )),

      # Sample size for each model in this comparison
      tex_row(c(
        "\\quad N",
        formatC(
          as.integer(block$n),
          format = "d",
          big.mark = ","
        )
      )),

      # R-squared for each model in this comparison
      tex_row(c(
        "\\quad $R^2$",
        fmt_num(block$r2, digits)
      )),

      "\\addlinespace"
    )
  }

  # Keep the note consistent with the requested SE method.
  se_note <- if (robust) {
    "Entries are OLS coefficients with HC1 robust standard errors in parentheses."
  } else {
    "Entries are OLS coefficients with conventional OLS standard errors in parentheses."
  }

  # Reference keys are identifiers, not printed labels:
  # do not escape underscores here.
  label_key <- paste0("tab:policy-waves-", dv_name)

  # Assemble the complete table.
  c(
    "\\begin{table}[!htbp]",
    "\\centering",
    "\\begin{threeparttable}",

    paste0(
      "\\caption{",
      wave_tex_escape(paste0("OLS models of ", dv_label)),
      "}"
    ),

    paste0("\\label{", label_key, "}"),

    "\\small",
    "\\setlength{\\tabcolsep}{4pt}",
    "\\renewcommand{\\arraystretch}{1.12}",

    "\\begin{tabularx}{\\textwidth}{@{}>{\\raggedright\\arraybackslash}p{0.22\\textwidth}*{4}{>{\\centering\\arraybackslash}X}@{}}",

    "\\toprule",

    tex_row(c(
      "Carceral contact",
      wave_tex_escape(unname(wave_labels[waves]))
    )),

    "\\midrule",
    body_lines,
    "\\bottomrule",
    "\\end{tabularx}",

    "\\begin{tablenotes}[flushleft]",
    "\\footnotesize",

    paste0("\\item ", se_note),

    "\\item Each focal coefficient comes from a separate regression. The reference group reports none of the three focal experiences: personal incarceration, knowing someone incarcerated, or personal victimization. Exposure groups may overlap.",

    "\\item Controls accumulate from left to right: none; demographics; demographics plus party/ideology; then additionally FIRE and core deservingness controls. Control coefficients are not displayed.",

    "\\item N and $R^2$ refer to the model immediately above them. Each model retains its own estimation sample; sample sizes may differ across comparisons and waves.",

    "\\item * $p<0.05$, ** $p<0.01$, *** $p<0.001$.",

    "\\end{tablenotes}",
    "\\end{threeparttable}",
    "\\end{table}"
  )
}
```

```{r}
# ============================================================
# STEP 6: EXPORT ALL FOUR-WAVE DV TABLES
# ============================================================

# ------------------------------------------------------------
# 6A. Generate one table for every DV in policy_vars
# Uses policy_vars, NOT the narrower results_dvs.
# ------------------------------------------------------------

# Keep filenames and LaTeX reference keys safe and unique.
if (
  length(policy_vars) == 0L ||
  anyNA(policy_vars) ||
  anyDuplicated(policy_vars) ||
  any(!grepl("^[A-Za-z0-9_][A-Za-z0-9_.-]*$", policy_vars))
) {
  stop(
    "policy_vars must contain unique, nonmissing, filename-safe variable names."
  )
}

wave_latex_tables <- setNames(
  lapply(policy_vars, function(dv_name) {

    make_wave_latex_table(
      dv_name = dv_name,
      model_waves = policy_dummies_waves,
      row_vars = impact_vars,
      waves = wave_order,
      robust = TRUE
    )

  }),
  policy_vars
)


# ------------------------------------------------------------
# 6B. Write one .tex file per DV
# ------------------------------------------------------------

wave_table_dir <- file.path("tables", "policy_waves")

dir.create(
  wave_table_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

if (!dir.exists(wave_table_dir)) {
  stop("Could not create output folder: ", wave_table_dir)
}

# Named vector linking each DV to its output filename.
wave_table_files <- setNames(
  paste0(policy_vars, "_wave_table.tex"),
  policy_vars
)

for (dv_name in policy_vars) {

  writeLines(
    wave_latex_tables[[dv_name]],
    con = file.path(
      wave_table_dir,
      wave_table_files[[dv_name]]
    )
  )

}


# ------------------------------------------------------------
# 6C. Create an Overleaf main document
#
# The main file and individual table files belong together
# in the same folder.
#
# Each DV starts on a fresh page, in policy_vars order.
# ------------------------------------------------------------

input_lines <- unlist(
  lapply(unname(wave_table_files), function(filename) {

    c(
      paste0("\\input{", filename, "}"),
      "\\clearpage"
    )

  }),
  use.names = FALSE
)

master_tex <- c(
  "\\documentclass[11pt,letterpaper]{article}",
  "\\usepackage[margin=0.75in]{geometry}",
  "\\usepackage[T1]{fontenc}",
  "\\usepackage[utf8]{inputenc}",
  "\\usepackage{booktabs,tabularx,array,threeparttable,makecell}",
  "\\begin{document}",
  input_lines,
  "\\end{document}"
)

writeLines(
  master_tex,
  con = file.path(
    wave_table_dir,
    "all_policy_wave_tables.tex"
  )
)


# ------------------------------------------------------------
# 6D. Print the completion status and exact output location
# ------------------------------------------------------------

cat(
  "Exported ",
  length(wave_latex_tables),
  " four-wave DV tables.\n",

  "Folder: ",
  normalizePath(wave_table_dir, winslash = "/"),
  "\n",

  "Overleaf main file: all_policy_wave_tables.tex\n",

  sep = ""
)
```


# ------------------------------------------------------------
# END TEMPORARY SEPT 27 -- CREATING LATEX FOR ALL TABLES FOR INSPECTION
# ------------------------------------------------------------


# ============================================================
# JUST WANT TO LOOK AT MODELS? 
# ============================================================

NAMES
```{r}
# Show the four wave names
names(policy_dummies_waves)

# Show every model stored in a given wave name (for ex, full-attitudes wave)
names(policy_dummies_full_models)
```

Remember standard errors though: summary () reports conventional OLS standard errors, my tables in Latex report HC1 Robust

```{r}

# No controls
summary(policy_dummies_models[["m_prisonhelp_incarc"]])

# Demographic controls
summary(policy_dummies_demo_models[["m_prisonhelp_incarc"]])

# Demographics + party/ideology
summary(policy_dummies_partyideo_models[["m_prisonhelp_incarc"]])

# Demographics + party/ideology + FIRE + core deservingness
summary(policy_dummies_full_models[["m_prisonhelp_incarc"]])

```

To look at w/HC1 Robust SEs:

```{r}
model_to_check <- policy_dummies_full_models[["m_prisonhelp_incarc"]]

lmtest::coeftest(
  model_to_check,
  vcov. = sandwich::vcovHC(model_to_check, type = "HC1")
)
```


```{r}
model_to_check <- policy_dummies_full_models[["m_prisonhelp_impact"]]

lmtest::coeftest(
  model_to_check,
  vcov. = sandwich::vcovHC(model_to_check, type = "HC1")
)
```

Variables have 0 value?
fiscal and social at 0 are most liberal
```{r}
table(sample$fiscal_ideology)

table(sample$social_ideology)
```

pid_3 doesn't have coded 0 value, gender at 0 is women, fire vars don't have a 0, desorcore has a respondent who gave 0 

desorcore
```{r}
table(sample$pid_3)
table(sample$gender)
table(sample$fire_angry)
```


```{r}
sample %>%
  group_by(pid_3) %>%
  summarise(
    n=n(),
    Democrat = sum(Democrat, na.rm = TRUE),
    Republican = sum(Republican, na.rm = TRUE),

  )
```




```{r}
m_prisonhelp_impact_verify <- lm(prisonhelp ~ impact + Black + Hisp_Latino + AllOthers + gender + Republican + Democrat + fiscal_ideology + social_ideology + fire_rare + fire_privilege + fire_angry + fire_fear + desor_core, data = sample)

lmtest::coeftest(
  m_prisonhelp_impact_verify,
  vcov. = sandwich::vcovHC(m_prisonhelp_impact_verify, type = "HC1")
)


```



# ============================================================
# Questions raised Sept 28
# ============================================================

```{r}
m <- m_prisonhelp_impact_verify

# The actual estimation sample retained by this fitted model
mf <- model.frame(m)

# Model size and composition
nobs(m)
df.residual(m)

table(mf$impact, useNA = "ifany")
table(mf$Republican, useNA = "ifany")
table(mf$Democrat, useNA = "ifany")

# Verify data types and actual ranges/codes
str(mf)
summary(mf)

# Missingness in the source data for variables used in this model
# These counts overlap; do not sum them to obtain total exclusions.
model_vars <- all.vars(formula(m))

sort(
  colSums(is.na(sample[, model_vars, drop = FALSE])),
  decreasing = TRUE
)

# Outcome scale and fitted-value checks
range(mf$prisonhelp)
sd(mf$prisonhelp)
range(fitted(m))

# Impact contrast expressed in observed outcome-SD units
unname(coef(m)["impact"] / sd(mf$prisonhelp))

# HC1 coefficient tests and corresponding confidence intervals
V_HC1 <- sandwich::vcovHC(m, type = "HC1")

lmtest::coeftest(m, vcov. = V_HC1)
lmtest::coefci(m, vcov. = V_HC1)

# Inference sensitivity check: coefficients will not change
V_HC3 <- sandwich::vcovHC(m, type = "HC3")

lmtest::coeftest(m, vcov. = V_HC3)
```


# ------------------------------------------------------------
# What about when we make controls our outcome variables
# ------------------------------------------------------------

```{r}
# 
# hist(sample$fire_rare)
# 
# m_fire_fear <- lm(fire_rare ~ incarc + Black + Hisp_Latino + pid_3 + AllOthers, data = sample)
# 
# summary(m_fire_fear)
# 
# sample %>%
#   group_by(fire_rare) %>%
#   summarise(
#     n = n(),
#     White = sum(White)
#   )

```




### 3.3 Appendix Tables - Each IV Model Grouped around DV

```{r}
# ============================================================
# Build one LaTeX regression-results table per DV
# Assumes  variable-label script has already been sourced,
# so the following objects already exist:
#   - impact_vars
#   - impact_labels
#   - policy_vars
#   - policy_map
# Also assumes policy_dummies_models already exists.
# ============================================================


# ------------------------------------------------------------
# sanity checks: make sure the variables referenced in 
# helper objects actually exist in the data
# -----------------------------------------------------------

missing_policy_vars <- setdiff(policy_vars, names(sample))
missing_impact_vars <- setdiff(impact_vars, names(sample))

if (length(missing_policy_vars) > 0) {
  stop("These policy_vars are missing from sample: ",
       paste(missing_policy_vars, collapse = ", "))
}

if (length(missing_impact_vars) > 0) {
  stop("These impact_vars are missing from sample: ",
       paste(missing_impact_vars, collapse = ", "))
}

# ------------------------------------------------------------
# create a named vector of DV labels from policy_map
# ------------------------------------------------------------

policy_labels <- setNames(policy_map$label, policy_map$var)

# ------------------------------------------------------------
# helper function: convert p-values to significance stars
# ------------------------------------------------------------

p_to_stars <- function(p) {
  dplyr::case_when(
    is.na(p)  ~ "",
    p < .001  ~ "***",
    p < .01   ~ "**",
    p < .05   ~ "*",
    TRUE      ~ ""
  )
}

# ------------------------------------------------------------
# function: build a clean one-DV results table
# ------------------------------------------------------------

make_dv_results_table <- function(dv_name, model_list) {

  rows <- lapply(impact_vars, function(iv_name) {

    model_name <- paste0("m_", dv_name, "_", iv_name)

    if (!model_name %in% names(model_list)) {
      stop("Model not found in policy_dummies_models: ", model_name)
    }

    mod <- model_list[[model_name]]

    coef_row <- broom::tidy(mod) %>%
      dplyr::filter(term != "(Intercept)") %>%
      dplyr::mutate(
        iv_name = iv_name,
        iv_label = impact_labels[[iv_name]],
        coef = round(estimate, 3),
        se = round(std.error, 3),
        p = round(p.value, 3),
        stars = p_to_stars(p.value),
        coef_with_stars = paste0(format(coef, nsmall = 3), stars),
        n = broom::glance(mod)$nobs,
        r2 = round(broom::glance(mod)$r.squared, 3)
      ) %>%
      dplyr::select(iv_label, coef_with_stars, se, p, n, r2)

    coef_row
  })

  dplyr::bind_rows(rows) %>%
    dplyr::rename(
      `Independent Variable` = iv_label,
      `Coefficient` = coef_with_stars,
      `SE` = se,
      `p-value` = p,
      `N` = n,
      `R^2` = r2
    )
}

# ------------------------------------------------------------
# function: convert one DV results table into LaTeX code
# ------------------------------------------------------------

make_dv_latex_table <- function(dv_name, model_list) {

  results_df <- make_dv_results_table(dv_name, model_list)

  caption_text <- paste0("Bivariate Models Predicting ", policy_labels[[dv_name]])
  label_text <- paste0("tab:", dv_name, "_bivariate")

  knitr::kable(
    results_df,
    format = "latex",
    booktabs = TRUE,
    caption = caption_text,
    label = label_text,
    align = c("l", "c", "c", "c", "c", "c"),
    escape = TRUE
  )
}

# ------------------------------------------------------------
# create one LaTeX table per DV and store them in a named list
# ------------------------------------------------------------

dv_latex_tables <- lapply(policy_vars, function(dv_name) {
  make_dv_latex_table(dv_name, policy_dummies_models)
})

names(dv_latex_tables) <- policy_vars

# ------------------------------------------------------------
# example: print one table to the console so you can copy/paste
# into Overleaf
# ------------------------------------------------------------

#cat(dv_latex_tables[["prisonhelp"]])

# ------------------------------------------------------------
# write all tables to separate .tex files
# ------------------------------------------------------------

for (dv_name in policy_vars) {
  writeLines(
    dv_latex_tables[[dv_name]],
    con = paste0(dv_name, "_bivariate_table.tex")
  )
}
```


### 3.5 Tables Loop
# ============================================================
# Tables
# Goal: compare each impacted subgroup to non-impacted only
# Model 2 = no-controls models
# Model 3 = placeholder for later controlled models
#
# Packages used:
# dplyr, stringr, purrr, lmtest, sandwich
# ============================================================

```{r}
# ------------------------------------------------------------
# Labels from variable script
# ------------------------------------------------------------
policy_labels <- setNames(policy_map$label, policy_map$var)

# ------------------------------------------------------------
# Formatting helpers (regression output to latex)
# ------------------------------------------------------------
p_to_stars <- function(p) {
  dplyr::case_when(
    is.na(p) ~ "",
    p < .001 ~ "***",
    p < .01  ~ "**",
    p < .05  ~ "*",
    TRUE     ~ ""
  )
}

fmt_num <- function(x, digits = 3) {
  ifelse(is.na(x), "", sprintf(paste0("%.", digits, "f"), x))
}

latex_escape <- function(x) {
  x %>%
    str_replace_all("\\\\", "\\\\\\\\") %>%
    str_replace_all("([#$%&_{}])", "\\\\\\1") %>%
    str_replace_all("~", "\\\\textasciitilde{}") %>%
    str_replace_all("\\^", "\\\\textasciicircum{}")
}

# ------------------------------------------------------------
# Split a vector into chunks for appendix tables
# ------------------------------------------------------------
chunk_vector <- function(x, chunk_size = 2) {
  split(x, ceiling(seq_along(x) / chunk_size))
}
```

# ------------------------------------------------------------
# MAIN TABLE BUILDER HERE
# -----------------------------------------------------------

```{r}
make_policy_panel_latex_table <- function(
    dv_vec,
    model2_list,
    model3_list = NULL,
    caption_text,
    label_text,
    model2_title = "Model 2",
    model3_title = "Model 3",
    row_vars = c("impact", "incarc", "knowincarc", "vict"),
    control_vars = c("Black", "Hisp_Latino", "AllOthers", "gender"),
    control_labels = c(
      Black = "Black",
      Hisp_Latino = "Hispanic/Latino",
      AllOthers = "Other Race",
      gender = "Male"
    ),
    robust = TRUE
) {

  # ----- column setup -----
  n_dv <- length(dv_vec)
  n_model_blocks <- 2
  total_data_cols <- n_dv * n_model_blocks
  align_string <- paste0("l", paste(rep("c", total_data_cols), collapse = ""))

  # ----- header rows -----
  header_top <- c(
    "",
    paste0("\\multicolumn{", n_dv, "}{c}{", latex_escape(model2_title), "}"),
    paste0("\\multicolumn{", n_dv, "}{c}{", latex_escape(model3_title), "}")
  )

  header_second <- c(
    "",
    latex_escape(policy_labels[dv_vec]),
    latex_escape(policy_labels[dv_vec])
  )

  cmid_model2 <- paste0("\\cmidrule(lr){2-", 1 + n_dv, "}")
  cmid_model3 <- paste0("\\cmidrule(lr){", 2 + n_dv, "-", 1 + (2 * n_dv), "}")
  cmid_line <- paste(cmid_model2, cmid_model3, collapse = " ")

  # ----- helper: extract a specific term from a model -----#
  
  get_term_cell <- function(model_list, dv, iv, term_name, robust = TRUE) {

    if (is.null(model_list)) return("")

    model_name <- paste0("m_", dv, "_", iv)
    if (!model_name %in% names(model_list)) return("")

    mod <- model_list[[model_name]]

    if (robust) {
      ct <- lmtest::coeftest(mod, vcov. = sandwich::vcovHC(mod, type = "HC1"))
      out <- data.frame(
        term      = rownames(ct),
        estimate  = ct[, 1],
        std.error = ct[, 2],
        p.value   = ct[, 4],
        stringsAsFactors = FALSE
      )
    } else {
      sm <- summary(mod)$coefficients
      out <- data.frame(
        term      = rownames(sm),
        estimate  = sm[, 1],
        std.error = sm[, 2],
        p.value   = sm[, 4],
        stringsAsFactors = FALSE
      )
    }

    if (!term_name %in% out$term) return("")

    row <- out[out$term == term_name, ]

    coef_txt <- paste0(
      fmt_num(row$estimate, 3),
      p_to_stars(row$p.value)
    )

    se_txt <- paste0("(", fmt_num(row$std.error, 3), ")")

    paste0("\\makecell[c]{", coef_txt, " \\\\ ", se_txt, "}")
  }

  # ----- body rows -----
  body_lines <- c()

  # Section header
  body_lines <- c(
    body_lines,
    paste(
      c(
        "\\textit{Carceral Contact (ref. Not impacted)}",
        rep("", total_data_cols)
      ),
      collapse = " & "
    ),
    "\\\\"
  )

  # For each subgroup, print coefficient row + repeated control rows
  for (iv in row_vars) {

    row_label <- impact_labels[[iv]]
    main_term <- iv

    # main subgroup row
    model2_cells <- sapply(dv_vec, function(dv) {
      get_term_cell(model2_list, dv, iv, main_term, robust = robust)
    })

    model3_cells <- sapply(dv_vec, function(dv) {
      get_term_cell(model3_list, dv, iv, main_term, robust = robust)
    })

    subgroup_line <- paste(
      c(latex_escape(row_label), model2_cells, model3_cells),
      collapse = " & "
    )

    body_lines <- c(body_lines, paste0(subgroup_line, " \\\\"))

    # repeated control rows under this subgroup
    for (ctrl in control_vars) {

      model2_ctrl_cells <- sapply(dv_vec, function(dv) {
        get_term_cell(model2_list, dv, iv, ctrl, robust = robust)
      })

      model3_ctrl_cells <- sapply(dv_vec, function(dv) {
        get_term_cell(model3_list, dv, iv, ctrl, robust = robust)
      })

      ctrl_line <- paste(
        c(latex_escape(control_labels[[ctrl]]), model2_ctrl_cells, model3_ctrl_cells),
        collapse = " & "
      )

      body_lines <- c(body_lines, paste0(ctrl_line, " \\\\"))
    }

    body_lines <- c(body_lines, "\\addlinespace")
  }

  # ----- assemble LaTeX -----
  out <- c(
    "\\begin{table}[!htbp]",
    "\\centering",
    paste0("\\caption{", latex_escape(caption_text), "}"),
    paste0("\\label{", latex_escape(label_text), "}"),
    "\\begin{threeparttable}",
    paste0("\\begin{tabular}{", align_string, "}"),
    "\\toprule",
    paste(header_top, collapse = " & "),
    " \\\\",
    cmid_line,
    paste(header_second, collapse = " & "),
    " \\\\",
    "\\midrule",
    body_lines,
    "\\bottomrule",
    "\\end{tabular}",
    "\\begin{tablenotes}[flushleft]",
    "\\footnotesize",
    "\\item Entries are OLS coefficients with HC1 robust standard errors in parentheses.",
    "\\item * p$<$0.05, ** p$<$0.01, *** p$<$0.001.",
    "\\end{tablenotes}",
    "\\end{threeparttable}",
    "\\end{table}"
  )

  out
}
```

TEll the function which DVs and model waves to display 

```{r}
results_dvs <- c("prisonhelp", "dpen_rev", "lwop_rev")

results_table_tex <- make_policy_panel_latex_table(
  dv_vec = results_dvs,
  model2_list = policy_dummies_models,
  model3_list = policy_dummies_demo_models,
  caption_text = "OLS Models of Policy Ors - None to Demographic Controls",
  label_text = "tab:carceral_policy_ors_nocontrol_demo",
  model2_title = "Model 2",
  model3_title = "Model 3",
  row_vars = c("impact", "incarc", "knowincarc", "vict"),
  control_vars = c("Black", "Hisp_Latino", "AllOthers", "gender"),
  control_labels = c(
    Black = "Black",
    Hisp_Latino = "Hispanic/Latino",
    AllOthers = "Other_Race",
    gender = "Male"
  ),
  robust = TRUE
)

writeLines(results_table_tex, "results_policy_orientations_nocontrol_demo_table.tex")
```


```{r}

##model for just desor

dep_variable <-(sample$knowincarc)
desor_model <- lm(desor_core ~ dep_variable + Black + Hisp_Latino + AllOthers + fiscal_ideology + social_ideology, data = sample)

writeLines(results_table_tex, "desor_controls_table.tex")
```

# ------------------------------------------------------------
# following has not yet been updated
# ------------------------------------------------------------





```{r}
# ------------------------------------------------------------
# Results Table
# ------------------------------------------------------------

results_dvs <- c("prisonhelp", "dpen_rev", "lwop_rev")

results_table_tex <- make_policy_panel_latex_table(
  dv_vec = results_dvs,
  model2_list = policy_vs_nonimp_nocontrol_models,
  model3_list = policy_vs_nonimp_demo_models,
  caption_text = "OLS Models of Carceral Policy Orientations - No controls to Demographic Controls",
  label_text = "tab:carceral_policy_orientations_nocontrol_demo",
  model2_title = "Model 2",
  model3_title = "Model 3",
  row_vars = c("impact", "incarc", "knowincarc", "vict"),
  control_vars = c("Black", "Hisp_Latino", "AllOthers", "gender"),
  control_labels = c(
    Black = "Black",
    Hisp_Latino = "Hispanic/Latino",
    AllOthers = "Other race",
    gender = "Male"
  ),
  robust = TRUE
)

writeLines(results_table_tex, "results_policy_orientations_nocontrol_demo_table.tex")

# Optional preview in console
# cat(results_table_tex, sep = "\n")


# ------------------------------------------------------------
# 2. APPENDIX TABLES FOR REMAINING POLICY DVs
#    Same architecture, chunked two DVs at a time
# ------------------------------------------------------------

appendix_dvs <- setdiff(policy_vars, results_dvs)

# You can keep help/punish separated if you want cleaner appendix organization
appendix_help_dvs <- appendix_dvs[policy_map$section[match(appendix_dvs, policy_map$var)] == "Help"]
appendix_punish_dvs <- appendix_dvs[policy_map$section[match(appendix_dvs, policy_map$var)] == "Punish"]

appendix_help_chunks <- chunk_vector(appendix_help_dvs, chunk_size = 2)
appendix_punish_chunks <- chunk_vector(appendix_punish_dvs, chunk_size = 2)

appendix_tables <- list()

# Help tables
for (i in seq_along(appendix_help_chunks)) {
  dv_chunk <- appendix_help_chunks[[i]]

  tex <- make_policy_panel_latex_table(
    dv_vec = dv_chunk,
    model2_list = policy_vs_nonimp_nocontrol_models,
    model3_list = NULL,
    caption_text = paste0("Restricted-Sample OLS Models of Help-Oriented Policy Outcomes, Appendix Table ", i),
    label_text = paste0("tab:appendix_help_restricted_", i),
    model2_title = "Model 2",
    model3_title = "Model 3",
    row_vars = c("impact", "incarc", "knowincarc", "vict"),
    include_placeholder_control_rows = TRUE,
    robust = TRUE
  )

  appendix_tables[[paste0("help_", i)]] <- tex
  writeLines(tex, paste0("appendix_help_restricted_table_", i, ".tex"))
}

# Punish tables
for (i in seq_along(appendix_punish_chunks)) {
  dv_chunk <- appendix_punish_chunks[[i]]

  tex <- make_policy_panel_latex_table(
    dv_vec = dv_chunk,
    model2_list = policy_vs_nonimp_nocontrol_models,
    model3_list = NULL,
    caption_text = paste0("Restricted-Sample OLS Models of Punitive Policy Outcomes, Appendix Table ", i),
    label_text = paste0("tab:appendix_punish_restricted_", i),
    model2_title = "Model 2",
    model3_title = "Model 3",
    row_vars = c("impact", "incarc", "knowincarc", "vict"),
    include_placeholder_control_rows = TRUE,
    robust = TRUE
  )

  appendix_tables[[paste0("punish_", i)]] <- tex
  writeLines(tex, paste0("appendix_punish_restricted_table_", i, ".tex"))
}
```

# ------------------------------------------------------------
# ------------------------------------------------------------
# Analysis with IDKs Excluded
# ------------------------------------------------------------
# ------------------------------------------------------------


```{r}

sample <- sample %>%
  mutate(
    dpen_no_idk
  )


```




