---
title: "scratch"
output: html_document
date: "2026-09-08"
---




you are moving flags to a first part of data_prep_script so that you can find out the counts across the whole sample before applying exclusions. 

Leave in data_prep_sept1 those lines of code that create flags.Here, I'm just paste the parts that will ultimately be bumped to the end of the flag-creating section

vars:
raw_data$attn_q9_pass
raw_data$attn_q9_typo
attn_check_q17
accepted_varients
attn_q114_pass
attn_q9_answered
attn_q17_answered
attn_q114_answered
attn_fail_count
attn_pass_count
attn_answered_count
duration_in_minutes

speed categories
sub_8
sub_12
sub_20
sub_30
over_52
over_180

progress
nonfinished
finished

common_vars
all_block_vars
quality_derived_vars
non_question_vars

common_na_count
block_na_count
na_count

Goes to application of exclusion:

From attention check section
```{r}

count(raw_data, attn_pass_count)


write_csv(raw_data, "data/cleaned/raw_data_consent_pre_attn_rest.csv")

data_pre_attn_exclusion <- raw_data # saving pre exclusion, or just use raw_data

raw_data_cleaned <- raw_data %>%
  filter(attn_pass_count > 0)#switched to raw_data_cleaned

count(raw_data_cleaned, attn_pass_count)

```

With dropped respondents (failed all 3), I have 32 respondents who passed 1, 459 who passed 2, and 1,604 who passed all 3. Total = 2,095 respondents retained.

To check specific attention check pass/fail counts: attn_q9_pass, attn_check_q17, attn_q114_pass attn_pass_count

Speeds
-have created sub_8 variable and sub_12 variable


```{r}
speed_review <- raw_data_cleaned %>%
  filter(duration_minutes >= 8) %>%
  mutate(
    speed_group = case_when(
      duration_minutes < 12 ~ "8–<12 min",
      duration_minutes < 20 ~ "12–<20 min",
      TRUE ~ "20+ min"
    ),
    
    attn_fail_count = 3 - attn_pass_count,
    
    any_attn_fail = attn_fail_count >= 1,
    two_attn_fails = attn_fail_count >= 2
  )

speed_review %>%
  count(speed_group)
```





```{r}
speed_review %>%
  count(speed_group, attn_fail_count)
```

I should probably remove people who were too fast two slow and only passed one attention check. or flag.

remember: attn_q9_pass, attn_check_q17, attn_q114_pass attn_pass_count

fast respondents who had another issue

```{r}
speed_review %>%
  filter(
    duration_minutes < 12,
    attn_fail_count >= 1
  ) %>%
  select(
    response_id,
    duration_minutes,
    attn_pass_count,
    attn_fail_count,
    attn_q9_pass,
    attn_check_q17,
    attn_q114_pass
  ) %>%
  arrange(duration_minutes)
```

```{r}
speed_review %>%
  filter(
    duration_minutes < 20,
    attn_fail_count >= 1
  ) %>%
  select(
    response_id,
    duration_minutes,
    attn_pass_count,
    attn_fail_count,
    attn_q9_pass,
    attn_check_q17,
    attn_q114_pass
  ) %>%
  arrange(duration_minutes) %>%
  View()
```

```{r}
speed_review %>%
  summarise(
    # 8 to <12 minutes
    n_under_12 = sum(duration_minutes < 12),
    n_under_12_attn_fail =
      sum(duration_minutes < 12 & any_attn_fail),
    pct_under_12_attn_fail =
      mean(any_attn_fail[duration_minutes < 12]) * 100,
    
    # 8 to <20 minutes
    n_under_20 = sum(duration_minutes < 20),
    n_under_20_attn_fail =
      sum(duration_minutes < 20 & any_attn_fail),
    pct_under_20_attn_fail =
      mean(any_attn_fail[duration_minutes < 20]) * 100
  )
```

