---
title: "variable_prep_script"
format: html
editor: visual
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,       # Hide code by default
  warning = FALSE,
  message = FALSE
)

# Load packages
library(dplyr)
library(tidyr)
library(readr)
library(janitor)   # for clean_names()
library(tibble)
library(stringr)
library(haven)
library(tibble)
library(gt)
library(psych)
library(ggplot2)
library(tidyverse)

```

## 1. Directory of Most Recent Data

Load cleaned survey data

File created in data_prep.qmd

```{r}
#view(raw_data_cleaned)

raw_data_cleaned <- readr::read_csv("data/cleaned/over_8_cleaned.csv") 


#names(raw_data_cleaned)

nrow(raw_data_cleaned)

```

## 2. Variables

### A. Deservingness and Power consolidation

```{r, prisonerdeserves}

# pretty much done, maybe standardize later?

```

Standardize? No -- in my prelim, I write: To ensure that I do not control for a variable that is conceptually or empirically too close to the deservingness of incarcerated people when it is the dependent variable, I exclude any deservingness ratings for groups that are substantively overlapping or highly correlated. Including those could absorb variance in the outcome or introduce post-treatment bias by controlling for a related evaluative construct. I also avoid using global composite or average deservingness scores across all groups – for example, a respondent's mean deservingnesss rating across all 65 groups they evaluated – because that composite would include the rating of incarcerated people itself, introducing potential collinearity and making it difficult to isolate the unique effect of contact. This modeling strategy attempts to preserve interpretability and ensuring that the effect of policy contact is estimated with respect to unstandardized, target-specific deservingness scores. 12

12 While constructing composite or control variables involving other deservingness targets, I assess whether their inclusion introduces collinearity or conceptual overlap with the dependent variable. In particular, I examine correlations between group-level deservingness ratings to ensure that any included groups are not too closely tied to perceptions of incarcerated people. If preliminary checks reveal that including a particular group (e.g., “criminals” or “formerly incarcerated individuals”) would obscure key variance or conflate target constructs, I exclude it from the model and treat it as conceptually non-distinct for the purposes of this analysis.

### B. CJ Experiences - Basic Coding

#### Impact Plus (Constructing For-Analysis)

```{r, impactvar}

#counting the *total* of those who said they were incarcerated, know incarcerated, or victims)

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    impact = if_else(
      incarc == 1 | knowincarc == 1 | vict == 1,
      1, 0,
      missing = NA_real_
    )
  )

# post Aug 2026 version with casewhen,where  incarc = 0 to mean impact = 0 (no impact, not just no incarceration):

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    incarc = case_when(
      incarc == 1 ~ 1, 
      impact == 0 ~ 0, 
      TRUE ~ NA_real_ #"otherwise, assign NA"
    )
  )

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    knowincarc = case_when(
      knowincarc == 1 ~ 1, 
      impact == 0 ~ 0, 
      TRUE ~ NA_real_ #"otherwise, assign NA"
    )
  )

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    vict = case_when(
      vict == 1 ~ 1, 
      impact == 0 ~ 0, 
      TRUE ~ NA_real_ #"otherwise, assign NA"
    )
  )

```

#### Extra binaries (incarc, KI, vict)

```{r}
#extra binary
 raw_data_cleaned <- raw_data_cleaned %>%
   mutate(
     incarc_bin = if_else(
       incarc == 1,
       1,
       0,
       missing = NA_real_
     )
   )
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    noincarc_bin = if_else(
  (incarc == 0 & knowincarc == 0) | noincarc == 1,
  1, 0,
  missing = NA_real_
))


 raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    knowincarc_bin = if_else(
      knowincarc == 1,
      1,
      0,
      missing = NA_real_
    )
  )

 raw_data_cleaned <- raw_data_cleaned %>%
   mutate(
     vict_bin = if_else(
       vict == 1,
       1,
       0,
       missing = NA_real_
     )
   )


```

### C. Other Experiences

#### Impact Plus (Constructing For-Analysis)

Coding in same way as CJ experiences but in giant loop

```{r}

# Function to code one experience family ---------------------------------

code_experience <- function(data, self_var, know_var, no_var, prefix) {
  
  impact_var   <- paste0(prefix, "_impact")
  self_bin     <- paste0(self_var, "_bin")
  know_bin     <- paste0(know_var, "_bin")
  no_bin       <- paste0(no_var, "_bin")
  
  data %>%
    # 1. Overall exposure:
    #    1 = self OR know someone
    #    0 = neither
    mutate(
      !!impact_var := if_else(
        .data[[self_var]] == 1 | .data[[know_var]] == 1,
        1, 0,
        missing = NA_real_
      )
    ) %>%
    
    # 2. Recode self/know variables:
    #    1 = this specific kind of exposure
    #    0 = completely unexposed
    #    NA = exposed, but not in this particular way
    mutate(
      !!self_var := case_when(
        .data[[self_var]] == 1 ~ 1,
        .data[[impact_var]] == 0 ~ 0,
        TRUE ~ NA_real_
      ),
      
      !!know_var := case_when(
        .data[[know_var]] == 1 ~ 1,
        .data[[impact_var]] == 0 ~ 0,
        TRUE ~ NA_real_
      )
    ) %>%
    
    # 3. Binary versions
    mutate(
      !!self_bin := if_else(
        .data[[self_var]] == 1,
        1, 0,
        missing = NA_real_
      ),
      
      !!know_bin := if_else(
        .data[[know_var]] == 1,
        1, 0,
        missing = NA_real_
      ),
      
      # 1 = explicitly in neither category
      #     OR explicitly selected the "no" option
      !!no_bin := if_else(
        (.data[[self_var]] == 0 & .data[[know_var]] == 0) |
          .data[[no_var]] == 1,
        1, 0,
        missing = NA_real_
      )
    )
}

```

9 experience groups

```{r}
experience_sets <- list(
  
  medicaid = c(
    self = "on_medicaid",
    know = "knowmedicaid",
    no   = "no_medicaid"
  ),
  
  tanf = c(
    self = "tanf",
    know = "knowtanf",
    no   = "notanf"
  ),
  
  snap = c(
    self = "snap",
    know = "knowsnap",
    no   = "nosnap"
  ),
  
  immigrated = c(
    self = "immigrated",
    know = "knowimmigrated",
    no   = "noimmigrated"
  ),
  
  union = c(
    self = "unionmember",
    know = "knowunion",
    no   = "nounion"
  ),
  
  abortion = c(
    self = "aborted",
    know = "knowaborted",
    no   = "noaborted"
  ),
  
  unemployment = c(
    self = "unemployed",
    know = "knowunemployed",
    no   = "nounemployed"
  ),
  
  uninsured = c(
    self = "uninsured",
    know = "knowuninsured",
    no   = "nouninsured"
  ),
  
  unable_vote = c(
    self = "unable_vote",
    know = "know_unable_vote",
    no   = "no_unable_vote"
  )
)
```

apply experience groups

```{r}
for (experience in names(experience_sets)) {
  
  vars <- experience_sets[[experience]]
  
  raw_data_cleaned <- code_experience(
    data     = raw_data_cleaned,
    self_var = vars["self"],
    know_var = vars["know"],
    no_var   = vars["no"],
    prefix   = experience
  )
}


```

### D. Q6, 7, 8 - Ration Block

### 

```{r}

# clean and transform
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(
    c(ration_scripts, ration_period, ration_food), 
    ~ na_if(as.numeric(.x), -99)
    )) %>% mutate(across(
    c(ration_scripts, ration_period, ration_food),
    ~ factor(.x, levels = 1:4,
             labels = c("Frequently", "Sometimes", "Rarely", "Never"),
             ordered = TRUE)
  ))

```

### F. FIRE controls

```{r}


fire_vars <- c("fire_rare", "fire_privilege", "fire_angry", "fire_fear")


raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(
    all_of(fire_vars),
    ~ as.numeric(na_if(as.character(.x), "-99"))
  ))


# summary(raw_data_cleaned$fire_angry)
# summary(raw_data_cleaned$fire_privilege)
# summary(raw_data_cleaned$fire_fear)
# summary(raw_data_cleaned$fire_rare)


```

### G. Race

q51\_ ▢ White (1) ▢ Black or African American (2) ▢ Native American / Alaska Native (3) ▢ Asian / Asian American (4) ▢ Native Hawaiian / Other Pacific Islander (5) ▢ Middle Eastern/ North African (6) ▢ Hispanic/Latino/a/e (7) ▢ My preferred response not listed (8)

-measured as set of binary indicator variables in the survey, rather than as one categorical race variable. - 1 = selected that category - 0 = ? did not select? - NA = c

```{r}

all_race_vars <- c("White", "Black", "Hisp_Latino", "Asian_Am", "Native_Am", "Pac_Island", "MENA", "race_notlisted")

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(all_of(all_race_vars), ~ na_if(as.numeric(.x), -99)))

```

Responses -- noting consistent NAs so NAs are missing

```{r}

raw_data_cleaned %>%
  select(all_of(all_race_vars)) %>%
  pivot_longer(everything(), names_to = "var", values_to = "val") %>%
  count(var, val) %>%
  pivot_wider(names_from = val, values_from = n, values_fill = 0)

```

How many multiracial respondents?

```{r}

#row level counts
race_df <- raw_data_cleaned %>%
  mutate(
    n_races   = rowSums(across(all_of(all_race_vars), ~ .x == 1), na.rm = TRUE),
    n_missing = rowSums(across(all_of(all_race_vars), is.na)),
    race_item_answered = n_races > 0,
    multiracial = case_when(
      !race_item_answered ~ NA,
      n_races >= 2        ~ TRUE,
      n_races == 1        ~ FALSE
    )
  )

count(race_df, n_races)
count(race_df, multiracial)

summary(race_df$n_races)

```

I have 218 multiracial respondents.

#### multiracial var

```{r}

raw_data_cleaned$multiracial <- race_df$multiracial

all_race_vars <- c(
  "White",
  "Black",
  "Hisp_Latino",
  "Asian_Am",
  "Native_Am",
  "Pac_Island",
  "MENA",
  "race_notlisted",
  "multiracial"
)

```

```{r}
# combinations
combo_counts <- raw_data_cleaned %>%
  mutate(.id = row_number()) %>%
  select(.id, all_of(all_race_vars)) %>%
  pivot_longer(-.id, names_to = "race", values_to = "selected") %>%
  filter(selected == 1) %>%                      # drops both 0s and NAs
  mutate(race = factor(race, levels = all_race_vars)) %>%
  arrange(.id, race) %>%
  group_by(.id) %>%
  summarise(
    n_races    = n(),
    race_combo = paste(race, collapse = " + "),
    .groups = "drop"
  ) %>%
  count(n_races, race_combo, name = "n") %>%
  arrange(desc(n))

View(combo_counts)                                # everything

filter(combo_counts, n_races >= 2)          # multiracial combos only
```

```{r}
# pairs
race_mat <- raw_data_cleaned %>%
  select(all_of(all_race_vars)) %>%
  mutate(across(everything(), ~ replace_na(as.integer(.x == 1), 0L))) %>%
  as.matrix()

co_occur <- crossprod(race_mat)   # diagonal = marginal totals
co_occur
```

#### condensed race var

```{r}
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    OtherRace = case_when(
      MENA == 1 | Pac_Island == 1 | race_notlisted == 1 ~ 1,
      MENA == 0 & Pac_Island == 0 & race_notlisted == 0 ~ 0,
      TRUE ~ NA_real_
    )
  )

#table(raw_data_cleaned$OtherRace, useNA = "ifany")

race_vars <- c(
  "White",
  "Black",
  "Hisp_Latino",
  "Asian_Am",
  "Native_Am",
  "OtherRace"
)

```

### H. Party + Ideology controls

```{r}

#raw_data_cleaned %>% count(q47)

#clean dem 101
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(dem101 = na_if(as.numeric(dem101), -99))

#party ID dummies
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    Republican = case_when(
      dem101 == 1 ~ 1,
      dem101 %in% c(2, 3, 4) ~ 0,
      TRUE ~ NA_real_
    ),
    Democrat = case_when(
      dem101 == 2 ~ 1,
      dem101 %in% c(1, 3, 4) ~ 0,
      TRUE ~ NA_real_
    )
  )

#clean ideology
#summary(raw_data_cleaned$q46)
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    fiscal_ideology = na_if(as.numeric(fiscal_ideology), -99),
    social_ideology = na_if(as.numeric(social_ideology), -99)
  )

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    fiscal_ideology = ifelse(!is.na(fiscal_ideology), 6 - fiscal_ideology, NA),#now 0 = very lib, 5 = very conservative (left to right)
    social_ideology = ifelse(!is.na(social_ideology), 6 - social_ideology, NA)
  )

#summary(raw_data_cleaned$fiscal_ideology)

#self-reported ideo ident
table(raw_data_cleaned$politcs) # already coded so that left to right
# 
# raw_data_cleaned <- raw_data_cleaned %>%
#    rename(
#      self_ideo = politcs
#      )
```

```{r}
pol_tude_vars <- c("fiscal_ideology", "social_ideology", "self_ideo")

pol_tude <- raw_data_cleaned %>%
  select(all_of(pol_tude_vars))

cor(pol_tude,
    use = "pairwise.complete.obs")

print(psych::corr.test(pol_tude))
```

### I. Gender

```{r, gender}
#raw_data_cleaned %>% count(q55)

#trans clean

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(trans = na_if(as.numeric(trans), -99))


#trans indicator 
raw_data_cleaned <- raw_data_cleaned %>% mutate(trans_id = case_when( trans == 1 ~ 1, trans %in% c(2, 3) ~ 0, # Conservative approach: counts only self-identified trans as 1 TRUE ~ NA_real_ ))

#gender clean
gender_vars <- c("woman", "man", "gqueer", "g_other")

table(raw_data_cleaned$trans)

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(all_of(gender_vars), ~ na_if(as.numeric(.x), -99)))


#constructing the gender binary
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(gender = case_when(
    man == 1 ~ 1,  # Selected man
    woman == 1 & (man != 1 | is.na(man)) ~ 0,  # Woman only or with genderqueer/other
    gqueer == 1 | g_other == 1 ~ 0,  # Genderqueer or other, no man selected
    TRUE ~ NA_real_
  ))


```

### J. State/County

-numeric variables that are (I think) Qualtrics drill down internal answer codes

```{r}
#unique(raw_data_cleaned$state)

```

### K. LAST OF CLEANING - WRITE CSV

```{r}

#nrow(raw_data_cleaned)
write_csv(raw_data_cleaned, "data/cleaned/coded_data__09_18_26.csv")
#write_csv(raw_data_cleaned, "data/cleaned/coded_data_may9.csv")

```

# ARCHIVED

```{r}
#| eval: false
#| include: false

getwd()
impacted <- read.csv("/Users/kristinamensik/Library/Mobile Documents/com~apple~CloudDocs/Documents/Duke/2024-2025/902 Legislators Pre-Ananlysis Plan/Prelim Folder/Prelim 3-23/data/impacted.csv")

view(impacted)

# Count unique combinations
impacted_summary <- impacted %>%
  select(been_inc, know_inc, none_inc, been_vict, know_vict, no_vict) %>%
  group_by(across(everything())) %>%
  summarise(count = n(), .groups = "drop") %>%
  arrange(desc(count))

# View the result
view(impacted_summary)

#only incarcerated
sum(
  impacted$been_inc == 1 &
  impacted$no_vict == 1 &
  impacted$know_inc == 0 &
  impacted$none_inc == 0 &
  impacted$been_vict == 0 &
  impacted$know_vict == 0
)

#incarcerated & Victimized
sum(
  impacted$been_inc == 1 &
  impacted$no_vict == 0 &
  impacted$know_inc == 0 &
  impacted$none_inc == 0 &
  impacted$been_vict == 1 &
  impacted$know_vict == 0
)

#incarcerated & Victimized & Know Inc
sum(
  impacted$been_inc == 1 &
  impacted$no_vict == 0 &
  impacted$know_inc == 1 &
  impacted$none_inc == 0 &
  impacted$been_vict == 1 &
  impacted$know_vict == 0
)

#only know incarcerated
sum(
  impacted$been_inc == 0 &
  impacted$no_vict == 1 &
  impacted$know_inc == 1 &
  impacted$none_inc == 0 &
  impacted$been_vict == 0 &
  impacted$know_vict == 0
)

#only vict
sum(
  impacted$been_inc == 0 &
  impacted$no_vict == 0 &
  impacted$know_inc == 0 &
  impacted$none_inc == 1 &
  impacted$been_vict == 1 &
  impacted$know_vict == 0
)

#only know vict
sum(
  impacted$been_inc == 0 &
  impacted$no_vict == 0 &
  impacted$know_inc == 0 &
  impacted$none_inc == 1 &
  impacted$been_vict == 0 &
  impacted$know_vict == 1
)

#know vict, know incarc, been vict
sum(
  impacted$been_inc == 0 &
  impacted$no_vict == 0 &
  impacted$know_inc == 1 &
  impacted$none_inc == 0 &
  impacted$been_vict == 1 &
  impacted$know_vict == 1
)


```
