---
title: "variable_prep_script"
format: html
editor: visual
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,       # Hide code by default
  warning = FALSE,
  message = FALSE
)

# Load packages
library(dplyr)
library(tidyr)
library(readr)
library(janitor)   # for clean_names()
library(tibble)
library(stringr)
library(haven)
library(tibble)
library(gt)
library(psych)
library(ggplot2)

```

## 1. Directory of Most Recent Data

Load cleaned survey data

File created in data_prep.qmd

```{r}
#view(raw_data_cleaned)

raw_data_cleaned <- readr::read_csv("data/cleaned/cleaned_light.csv") 
#CONTINGENT note that alhough this is called raw_data_cleaned, it is the equivalent of "cleaned_data_light" from data_prep.qmd. See that file to pull the version with old var names for D and P. Note: no power here.

names(raw_data_cleaned)

```

## 2. Variables

### A. Deservingness and Power consolidation

```{r, prisonerdeserves}

# pretty much done, maybe standardize later?

```

Standardize? No -- in my prelim, I write: To ensure that I do not control for a variable that is conceptually or empirically too close to the deservingness of incarcerated people when it is the dependent variable, I exclude any deservingness ratings for groups that are substantively overlapping or highly correlated. Including those could absorb variance in the outcome or introduce post-treatment bias by controlling for a related evaluative construct. I also avoid using global composite or average deservingness scores across all groups – for example, a respondent's mean deservingnesss rating across all 65 groups they evaluated – because that composite would include the rating of incarcerated people itself, introducing potential collinearity and making it difficult to isolate the unique effect of contact. This modeling strategy attempts to preserve interpretability and ensuring that the effect of policy contact is estimated with respect to unstandardized, target-specific deservingness scores. 12

12 While constructing composite or control variables involving other deservingness targets, I assess whether their inclusion introduces collinearity or conceptual overlap with the dependent variable. In particular, I examine correlations between group-level deservingness ratings to ensure that any included groups are not too closely tied to perceptions of incarcerated people. If preliminary checks reveal that including a particular group (e.g., “criminals” or “formerly incarcerated individuals”) would obscure key variance or conflate target constructs, I exclude it from the model and treat it as conceptually non-distinct for the purposes of this analysis.


### B. Experiences - Basic Coding

#### Impact Plus (Constructing For-Analysis)

```{r, impactvar}

#counting the *total* of those who said they were incarcerated, know incarcerated, or victims)

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    impact = if_else(
      q4_1_1 == 1 | q4_1_2 == 1 | q4_2_1 == 1,
      1, 0,
      missing = NA_real_
    )
  )

# post Aug 2026 version with casewhen,where  incarc = 0 to mean impact = 0 (no impact, not just no incarceration):

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    incarc = case_when(
      q4_1_1 == 1 ~ 1, 
      impact == 0 ~ 0, 
      TRUE ~ NA_real_ #"otherwise, assign NA"
    )
  )

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    knowincarc = case_when(
      q4_1_2 == 1 ~ 1, 
      impact == 0 ~ 0, 
      TRUE ~ NA_real_ #"otherwise, assign NA"
    )
  )

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    vict = case_when(
      q4_2_1 == 1 ~ 1, 
      impact == 0 ~ 0, 
      TRUE ~ NA_real_ #"otherwise, assign NA"
    )
  )

```

#### Extra binaries (incarc, KI, vict)

```{r}

 raw_data_cleaned <- raw_data_cleaned %>%
   mutate(
     incarc_bin = if_else(
       q4_1_1 == 1,
       1,
       0,
       missing = NA_real_
     )
   )

 raw_data_cleaned <- raw_data_cleaned %>%
   mutate(
     noincarc_bin = if_else(
       q4_1_1 == 0 | q4_1_2 == 0 | q4_1_3 == 1,
       1, 0,
       missing = NA_real_
     )
   )


 raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    knowincarc_bin = if_else(
      q4_1_2 == 1,
      1,
      0,
      missing = NA_real_
    )
  )

 raw_data_cleaned <- raw_data_cleaned %>%
   mutate(
     vict_bin = if_else(
       q4_2_1 == 1,
       1,
       0,
       missing = NA_real_
     )
   )


```

### C. Q6, 7, 8 - Ration Block

```{r}

#inspect

#raw_data %>% select(q6, q7, q8) %>% summarise_all(~ list(unique(.)))

# clean and transform
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(
    c(q6, q7, q8), 
    ~ na_if(as.numeric(.x), -99)
    )) %>% mutate(across(
    c(q6, q7, q8),
    ~ factor(.x, levels = 1:4,
             labels = c("Frequently", "Sometimes", "Rarely", "Never"),
             ordered = TRUE)
  ))

 raw_data_cleaned %>% count(q8)

```

### E.Policy Orientation (polor) control

Decided to skip because of how questions are worded/coded. Will need to return to this later

10-27 Diagnostics

```{r}
# names(raw_data_cleaned) %>% stringr::str_subset("q25|q23|q62|prison|help|punish|rehab")
# 
# glimpse(raw_data_cleaned %>% select(starts_with("q25"), starts_with("q23"), starts_with("q62"), starts_with("prison"), starts_with("penal")))
# 
# 
# glimpse(dplyr::select(raw_data_cleaned,
#                       starts_with("q25_"),
#                       starts_with("q23_"),
#                       starts_with("q62"),
#                       starts_with("prison"),
#                       starts_with("penal")))

```

### F. FIRE controls

```{r}

#raw_data_cleaned %>% count(q32)

#raw_data_cleaned %>% count(Q33)

#raw_data_cleaned %>% count(Q34)

#raw_data_cleaned %>% count(Q111)

fire_vars <- c("q32", "q33", "q34", "q111")
names(raw_data_cleaned)
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(
    all_of(fire_vars),
    ~ as.numeric(na_if(as.character(.x), "-99"))
  ))

raw_data_cleaned <- raw_data_cleaned %>%
  rename(
    fire_rare = q32,
    fire_privilege = q33,
    fire_angry = q34,
    fire_fear = q111
  )

summary(raw_data_cleaned$fire_angry)
summary(raw_data_cleaned$fire_privilege)
summary(raw_data_cleaned$fire_fear)
summary(raw_data_cleaned$fire_rare)


```

### G. Race

```{r}

#convert + clean
race_vars <- paste0("q51_", 1:8)

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(all_of(race_vars), ~ na_if(as.numeric(.x), -99)))

# REPLACE your current race dummies with this block
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    # Ethnicity first: Hispanic gets priority
    Hispanic = ifelse(q51_7 == 1, 1, 0),

    # Among non-Hispanics, assign Black and White
    Black    = ifelse(q51_2 == 1 & q51_7 != 1, 1, 0),
    White    = ifelse(q51_1 == 1 & q51_7 != 1, 1, 0)
  ) %>%
  mutate(
    # Everyone else with any race selection becomes OtherRace (mutually exclusive)
    OtherRace = ifelse(
      (Black + Hispanic + White) == 0 &
        rowSums(dplyr::select(., dplyr::all_of(race_vars)), na.rm = TRUE) > 0,
      1, 0
    )
  )

```

DOUBLE CHECK

```{r}

# no overlaps among the 4 dummies
sum((raw_data_cleaned$Black + raw_data_cleaned$White +
     raw_data_cleaned$Hispanic + raw_data_cleaned$OtherRace) > 1, na.rm = TRUE)

# counts by category (mutually exclusive)
table(
  with(raw_data_cleaned,
       dplyr::case_when(
         Hispanic == 1 ~ "Hispanic",
         Black    == 1 ~ "Black",
         White    == 1 ~ "White",
         OtherRace== 1 ~ "OtherRace",
         TRUE          ~ NA_character_
       )),
  useNA = "ifany"
)


```

Diagnosed problem in model where Black + Other dropped - Oct 27 Suggests all Black respondents had missing data so were dropped

### H. Party + Ideology controls

```{r}

#raw_data_cleaned %>% count(q47)

#clean dem 101
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(dem101 = na_if(as.numeric(dem101), -99))

#party ID dummies
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    Republican = case_when(
      dem101 == 1 ~ 1,
      dem101 %in% c(2, 3, 4) ~ 0,
      TRUE ~ NA_real_
    ),
    Democrat = case_when(
      dem101 == 2 ~ 1,
      dem101 %in% c(1, 3, 4) ~ 0,
      TRUE ~ NA_real_
    )
  )

#clean ideology
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    q46 = na_if(as.numeric(q46), -99),
    q47 = na_if(as.numeric(q47), -99)
  )

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    fiscal_ideology = ifelse(!is.na(q46), 6 - q46, NA),
    social_ideology = ifelse(!is.na(q47), 6 - q47, NA)
  )


```

### I. Gender

```{r, gender}
#raw_data_cleaned %>% count(q55)

#trans clean

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(q55 = na_if(as.numeric(q55), -99))

#trans indicator
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(trans_id = case_when(
    q55 == 1 ~ 1,
    q55 %in% c(2, 3) ~ 0,  # Conservative approach: counts only self-identified trans as 1
    TRUE ~ NA_real_
  ))


#gender clean
gender_vars <- paste0("q56_", 1:4)

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(all_of(gender_vars), ~ na_if(as.numeric(.x), -99)))


#constructing the gender binary
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(gender = case_when(
    q56_2 == 1 ~ 1,  # Selected man
    q56_1 == 1 & (q56_2 != 1 | is.na(q56_2)) ~ 0,  # Woman only or with genderqueer/other
    q56_3 == 1 | q56_4 == 1 ~ 0,  # Genderqueer or other, no man selected
    TRUE ~ NA_real_
  ))


```

### J. Victim Analysis Vars

```{r}

# police policies

# Originally 6 = strong oppose, 7= somewhat oppose, 8 = neutral, 9 = somewhat support, 10 = strong support

# Define Q25 police items
police_items <- c("q15_1", "q15_2", "q15_3", "q15_4", "q15_5")

# Step 1: Recode -99 to NA and ensure numeric
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(all_of(police_items), ~ na_if(as.numeric(.), -99)))

# Step 2: Recode so 1-5 (6 = strong oppose → 1 = strong oppose, 
# 7 = somewhat oppose → 2 = somewhat oppose, 
# 8 = neutral → 3 = neutral, 
# 9 = somewhat support → 4 = somewhat support,
#10 = strong support → 5 = strong support)

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(
    all_of(police_items),
    ~ case_when(
    . == 6 ~ 1,          
    . == 7 ~ 2,
    . == 8 ~ 3,
    . == 9 ~ 4,
    . == 10 ~ 5,        
    TRUE ~ NA_real_
  ),  
    .names = "{.col}_1_to_5"
  ))

police_items_1_to_5 <- c(
  "q15_1_1_to_5",
  "q15_2_1_to_5",
  "q15_3_1_to_5",
  "q15_4_1_to_5",
  "q15_5_1_to_5"
)

```

```{r}

## trust in government 

summary(raw_data_cleaned$q30)
unique(raw_data_cleaned$gov_attn)

# Originally 1 = a lot, 2 = some, 3 = not muc

# rename
## trust in government attention
# q30 originally: 1 = a lot, 2 = some, 3 = not much
# recode so higher = more government attention/trust

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    gov_attn = case_when(
      as.numeric(q30) == -99 ~ NA_real_,
      as.numeric(q30) == 1 ~ 3,
      as.numeric(q30) == 2 ~ 2,
      as.numeric(q30) == 3 ~ 1,
      TRUE ~ NA_real_
    )
  )

hist(raw_data_cleaned$gov_attn)

## did vote in 2020
# q31: 1 = no, not eligible; 2 = no, eligible; 3 = yes

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    vote2020 = na_if(as.numeric(q31), -99)
  )


## plans to vote
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    willvote = na_if(as.numeric(q110), -99)
  )


```

Note:

Approximately 561 respondents did not provide responses to the political behavior items (government attention, past vote, intended vote). Missingness is perfectly overlapping across these items, indicating that these questions were likely administered to a subset of respondents rather than reflecting item-level nonresponse.

### K. LAST OF CLEANING - WRITE CSV

```{r}

names(raw_data_cleaned)
write_csv(raw_data_cleaned, "data/cleaned/coded_data__08_2026_3.csv")
#write_csv(raw_data_cleaned, "data/cleaned/coded_data_may9.csv")

```

# ARCHIVED

```{r}
#| eval: false
#| include: false

getwd()
impacted <- read.csv("/Users/kristinamensik/Library/Mobile Documents/com~apple~CloudDocs/Documents/Duke/2024-2025/902 Legislators Pre-Ananlysis Plan/Prelim Folder/Prelim 3-23/data/impacted.csv")

view(impacted)

# Count unique combinations
impacted_summary <- impacted %>%
  select(been_inc, know_inc, none_inc, been_vict, know_vict, no_vict) %>%
  group_by(across(everything())) %>%
  summarise(count = n(), .groups = "drop") %>%
  arrange(desc(count))

# View the result
view(impacted_summary)

#only incarcerated
sum(
  impacted$been_inc == 1 &
  impacted$no_vict == 1 &
  impacted$know_inc == 0 &
  impacted$none_inc == 0 &
  impacted$been_vict == 0 &
  impacted$know_vict == 0
)

#incarcerated & Victimized
sum(
  impacted$been_inc == 1 &
  impacted$no_vict == 0 &
  impacted$know_inc == 0 &
  impacted$none_inc == 0 &
  impacted$been_vict == 1 &
  impacted$know_vict == 0
)

#incarcerated & Victimized & Know Inc
sum(
  impacted$been_inc == 1 &
  impacted$no_vict == 0 &
  impacted$know_inc == 1 &
  impacted$none_inc == 0 &
  impacted$been_vict == 1 &
  impacted$know_vict == 0
)

#only know incarcerated
sum(
  impacted$been_inc == 0 &
  impacted$no_vict == 1 &
  impacted$know_inc == 1 &
  impacted$none_inc == 0 &
  impacted$been_vict == 0 &
  impacted$know_vict == 0
)

#only vict
sum(
  impacted$been_inc == 0 &
  impacted$no_vict == 0 &
  impacted$know_inc == 0 &
  impacted$none_inc == 1 &
  impacted$been_vict == 1 &
  impacted$know_vict == 0
)

#only know vict
sum(
  impacted$been_inc == 0 &
  impacted$no_vict == 0 &
  impacted$know_inc == 0 &
  impacted$none_inc == 1 &
  impacted$been_vict == 0 &
  impacted$know_vict == 1
)

#know vict, know incarc, been vict
sum(
  impacted$been_inc == 0 &
  impacted$no_vict == 0 &
  impacted$know_inc == 1 &
  impacted$none_inc == 0 &
  impacted$been_vict == 1 &
  impacted$know_vict == 1
)


```
