---
title: "variable_prep_script"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,       # Hide code by default
  warning = FALSE,
  message = FALSE
)

# Load packages
library(dplyr)
library(tidyr)
library(readr)
library(janitor)   # for clean_names()
library(tibble)
library(stringr)
library(haven)
library(tibble)
library(gt)

```

## 1. Directory of Most Recent Data

Load cleaned survey data

File created in data_prep.qmd

```{r}
#view(raw_data_cleaned)

raw_data_cleaned <- readr::read_csv("data/cleaned/cleaned_light.csv") 
#CONTINGENT note that alhough this is called raw_data_cleaned, it is the equivalent of "cleaned_data_light" from data_prep.qmd. See that file to pull the version with old var names for D and P. Note: no power here.

#names(raw_data_cleaned)

```

## 2. Variables

### A. Deservingness and Power consolidation

```{r, prisonerdeserves}

# pretty much done, maybe standardize later?

```

Standardize? No -- in my prelim, I write: To ensure that I do not control for a variable that is conceptually or empirically too close to the deservingness of incarcerated people when it is the dependent variable, I exclude any deservingness ratings for groups that are substantively overlapping or highly correlated. Including those could absorb variance in the outcome or introduce post-treatment bias by controlling for a related evaluative construct. I also avoid using global composite or average deservingness scores across all groups – for example, a respondent's mean deservingnesss rating across all 65 groups they evaluated – because that composite would include the rating of incarcerated people itself, introducing potential collinearity and making it difficult to isolate the unique effect of contact. This modeling strategy attempts to preserve interpretability and ensuring that the effect of policy contact is estimated with respect to unstandardized, target-specific deservingness scores. 12

12 While constructing composite or control variables involving other deservingness targets, I assess whether their inclusion introduces collinearity or conceptual overlap with the dependent variable. In particular, I examine correlations between group-level deservingness ratings to ensure that any included groups are not too closely tied to perceptions of incarcerated people. If preliminary checks reveal that including a particular group (e.g., “criminals” or “formerly incarcerated individuals”) would obscure key variance or conflate target constructs, I exclude it from the model and treat it as conceptually non-distinct for the purposes of this analysis.

#### Deservingness Orientation Control

Creating a core composite score for groups evaluated by all 3 ABC first. Flagging will want to use extended version with targets evaluated by just one of the 3 ABC: "d_unins_c", "d_noncit_parent_a", "d_mentill_c"

```{r desor}

#names(raw_data_cleaned)

#summary(raw_data_cleaned$d_mentill_c)

#NEW (OCT 27) DESOR

desor_core_vars <- c(
  "d_welfare_tanf_abc","d_fstamps_snap_abc","d_medicaid_abc","d_medicare_abc",
  "d_unemployed_abc","d_poorfam_abc","d_homeless_abc","d_immigrants_abc","d_unauth_undoc_abc"
)

# below, try to keep only things in current dataset / prevent old code constructs I don't want to include
core_vars <- intersect(desor_core_vars, names(raw_data_cleaned))
# Takes the vector  defined earlier (desor_core_vars) and keeps only those names that actually exist in the current data. Prevents errors if a column is missing or was renamed

core_mat  <- dplyr::select(raw_data_cleaned, dplyr::all_of(core_vars))
#Builds a mini data frame (matrix) containing only the core variables (columns) from dataset.

core_n    <- rowSums(!is.na(core_mat))
#For each respondent (row), counts how many core items are non-missing. Will be used to enforce minimum answered-items rule

raw_data_cleaned$desor_core   <- ifelse(core_n >= 5, rowMeans(core_mat, na.rm = TRUE), NA_real_)
# Creates the core composite. With 9 core items, “≥5” ≈ a 50% completeness rule—balances coverage and reliability.

raw_data_cleaned$desor_core_z <- as.numeric(scale(raw_data_cleaned$desor_core))
# Makes a standardized version (z-score) of the core composite (mean 0, sd 1).
# Convenient for regression comparability and coefficient interpretation.


# EXTENDED VERSION -- groups only evaluated by 1 of ABC. For robustness. 
desor_ext_vars <- c(desor_core_vars, "d_unins_c", "d_mentill_c", "d_noncit_parent_a")
ext_vars <- intersect(desor_ext_vars, names(raw_data_cleaned))
ext_mat  <- dplyr::select(raw_data_cleaned, dplyr::all_of(ext_vars))
ext_n    <- rowSums(!is.na(ext_mat))
raw_data_cleaned$desor_ext    <- ifelse(ext_n >= 6, rowMeans(ext_mat, na.rm = TRUE), NA_real_)
raw_data_cleaned$desor_ext_z  <- as.numeric(scale(raw_data_cleaned$desor_ext))

# Check
summary(raw_data_cleaned$desor_core); table(core_n)
summary(raw_data_cleaned$desor_ext);  table(ext_n)

summary(raw_data_cleaned$desor_core); table(core_n, useNA="ifany")
summary(raw_data_cleaned$desor_ext);  table(ext_n,  useNA="ifany")


```

Section Conclusion: - creates composite desor Main specs: use desor_core_z.

Robustness: re-estimate with desor_ext_z and add + factor(block) (or + block) to your regressions.

Keep justice-adjacent targets out of both composites.

#### Deservingness Orientation Control -- Checks

```{r}
library(psych)


# keep only vars that exist
core_vars <- intersect(desor_core_vars, names(raw_data_cleaned))

# build item matrix/dataframe
core_mat <- dplyr::select(raw_data_cleaned, dplyr::all_of(core_vars))

# apply your completeness rule (>= 5 answered out of 9)
core_keep <- rowSums(!is.na(core_mat)) >= 5

# Cronbach's alpha on respondents who meet the rule
core_alpha <- psych::alpha(core_mat[core_keep, ], check.keys = TRUE)

# key outputs
print(core_alpha$total$raw_alpha)
print(core_alpha$total$std.alpha)
print(core_alpha$total$average_r)
print(core_alpha$n.obs)
print(core_alpha$alpha.drop)
print(core_alpha$item.stats)


```

Interpretation: raw_alpha: 0.9078 and 0.9142 internal consistency -- suggests items hang together well. average_r = 0.5422 -- high inter-item correlation. cor: fairly strong correlation to the DV (0.547) R\^2 = 0.30 bivariately. Soaks up general generosity toward assistance-framed groups. Should justify use of control for testing whether contact predicts deservingness of incarcerated people, above and beyond that general orientation.

```{r}
# reliability -- EXT

ext_mat  <- dplyr::select(raw_data_cleaned, dplyr::all_of(ext_vars))
ext_keep <- rowSums(!is.na(ext_mat)) >= 6
# alpha on complete cases after the ≥6 rule
psych::alpha(na.omit(ext_mat[ext_keep, ]), check.keys = TRUE)

#  per-block alphas
psych::alpha(na.omit(ext_mat[ext_keep & raw_data_cleaned$block=="A", ]), check.keys = TRUE)
psych::alpha(na.omit(ext_mat[ext_keep & raw_data_cleaned$block=="B", ]), check.keys = TRUE)
psych::alpha(na.omit(ext_mat[ext_keep & raw_data_cleaned$block=="C", ]), check.keys = TRUE)

```

```{r}
#standardize
raw_data_cleaned$desor_core_z <- as.numeric(scale(raw_data_cleaned$desor_core))
raw_data_cleaned$desor_ext_z  <- as.numeric(scale(raw_data_cleaned$desor_ext))

```

#### A1. Deserving Validation

```{r}
#CHECK INDIVIDUALS
#raw_data_cleaned %>%
#  filter(ResponseId == "R_5DugmlzJSCFvja1") %>%
 # select(block, `Deserving Group A_1`, d_black_afam_abc)

```

### B. Q4 - Experiences Block -- ie Incarcerated Combos

```{r, cleaningexperiences1}

#view(raw_data_cleaned$Q4_1_1)

# Basic Totals
#I’m counting how many people said they themselves have been incarcerated by checking where Q4_1_1 equals 1. I’m ignoring missing values
total_incarcerated <- sum(raw_data_cleaned$q4_1_1 == 1, na.rm = TRUE)

#"I’m counting how many people said they know someone who’s been incarcerated (Q4_1_2 == 1). Again, I’m skipping over any missing answers."
total_knows_incarcerated <- sum(raw_data_cleaned$q4_1_2 == 1, na.rm = TRUE)

#I’m counting how many people selected ‘no / unsure’ when asked about incarceration experience (Q4_1_3 == 1)
total_no_incarcerated <- sum(raw_data_cleaned$q4_1_3 == 1, na.rm = TRUE)

#I’m counting how many respondents said they themselves were a victim of violent crime.
total_victim <- sum(raw_data_cleaned$q4_2_1 == 1, na.rm = TRUE)

#I’m counting how many said they know someone who was a victim of violent crime
total_knows_victim<- sum(raw_data_cleaned$q4_2_2 == 1, na.rm = TRUE)

#I’m counting those who selected ‘no / unsure’ for victimization experience
total_no_victim <- sum(raw_data_cleaned$q4_2_3 == 1, na.rm = TRUE)
```

#### Impact Variable (Constructing For-Analysis)

```{r, impactvar}

#counting the *total* of those who said they were incarcerated, know incarcerated, or victims)

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    impact = if_else(
      q4_1_1 == 1 | q4_1_2 == 1 | q4_2_1 == 1,
      1, 0,
      missing = NA_real_
    )
  )


```

#### Incarc Variable (Constructing For-Analysis)

```{r}

# pre August 2026 version:
# raw_data_cleaned <- raw_data_cleaned %>%
#   mutate(
#     incarc = if_else(
#       q4_1_1 == 1,
#       1, 
#       0,
#       missing = NA_real_
#     )
#   )

#for now - aug 24 2026 - I don't need to change noincarc
# raw_data_cleaned <- raw_data_cleaned %>%
#   mutate(
#     noincarc = if_else(
#       q4_1_1 == 0 | q4_1_2 == 0 | q4_1_3 == 1,
#       1, 0,
#       missing = NA_real_
#     )
#   )

# post Aug 2026 version with casewhen, becuase if_else runs as a true false and I want incarc = 0 to mean impact = 0 (no impact, not just no incarceration):

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    incarc = case_when(
      q4_1_1 == 1 ~ 1, 
      impact == 0 ~ 0, 
      TRUE ~ NA_real_ #"otherwise, assign NA"
    )
  )

```

#### Knowincarc Variable (Constructing For-Analysis)

```{r}
# pre August 2026 version:
# raw_data_cleaned <- raw_data_cleaned %>%
#  mutate(
#    knowincarc = if_else(
#      q4_1_2 == 1,
#      1, 
#      0,
#      missing = NA_real_
#    )
#  )

# post Aug 2026 version with casewhen


raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    knowincarc = case_when(
      q4_1_2 == 1 ~ 1, 
      impact == 0 ~ 0, 
      TRUE ~ NA_real_ #"otherwise, assign NA"
    )
  )

```

#### Vict Variable (Constructing For-Analysis)
```{r}

# raw_data_cleaned <- raw_data_cleaned %>%
#   mutate(
#     vict = if_else(
#       q4_2_1 == 1,
#       1, 
#       0,
#       missing = NA_real_
#     )
#   )


raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    vict = case_when(
      q4_2_1 == 1 ~ 1, 
      impact == 0 ~ 0, 
      TRUE ~ NA_real_ #"otherwise, assign NA"
    )
  )

```

Create fingerprint

- create a new variable named contact_pattern

- for each respondent, create a 6 digit code that summarizes how they answered each Q4_1 and Q4_2 question

  - if the person said yes (1), use that 1. If they didn't answer, treat it as 0

- combine all 6 digits starting with incarceration responses then victimization to create pattern to put in var

```{r, experiencesfingerprints1}
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    contact_pattern = paste0(
      coalesce(q4_1_1, 0),
      coalesce(q4_1_2, 0),
      coalesce(q4_1_3, 0),
      coalesce(q4_2_1, 0),
      coalesce(q4_2_2, 0),
      coalesce(q4_2_3, 0)
    )
  )

count(raw_data_cleaned, contact_pattern)

```

#### Patterns and Coding Note –

Some respondents selected both a contact experience (e.g., "been incarcerated") and also marked “none/unsure”, which is logically inconsistent. These may reflect confusion, misclicks, or low-quality responses.

Conflict Type Count Q4_1_1 (Been incarcerated) & Q4_1_3 (None/Unsure) 1 Q4_1_2 (Knows incarcerated) & Q4_1_3 (None/Unsure) 1 Q4_2_1 (Been victimized) & Q4_2_3 (None/Unsure) 10 Q4_2_2 (Knows victim) & Q4_2_3 (None/Unsure) 10

```{r, experiencesfingerprints}
#table(raw_data_cleaned$Q4_1_3, useNA = "always")
#table(raw_data_cleaned$Q4_2_3, useNA = "always")


strange1 <- raw_data_cleaned %>%
  filter(q4_1_1 == 1 & q4_1_3 == 1)  

strange2 <- raw_data_cleaned %>%
  filter(q4_1_2 == 1 & q4_1_3 == 1) 

strange3 <- raw_data_cleaned %>%
  filter(q4_2_1 == 1 & q4_1_3 == 1) 

strange4 <- raw_data_cleaned %>%
  filter(q4_2_2 == 1 & q4_1_3 == 1) 

print(strange1)
print(strange2)
print(strange3)
print(strange4)


#
count(raw_data_cleaned, contact_pattern)


```

Tabulate unique patterns

```{r, allexperienceuniquecombos}

combo_summary <- raw_data_cleaned %>%
  group_by(contact_pattern) %>%
  summarise(
    N = n(),
    pct_total = 100 * n() / nrow(raw_data_cleaned),
    pct_of_been_incarcerated = 100 * sum(q4_1_1 == 1, na.rm = TRUE) / total_incarcerated,
    pct_of_knows_incarcerated = 100 * sum(q4_1_2 == 1, na.rm = TRUE) / total_knows_incarcerated,
    pct_of_been_victim = 100 * sum(q4_2_1 == 1, na.rm = TRUE) / total_victim
  ) %>%
  arrange(desc(N))
```

```{r}

gracious_recode <- tibble::tibble(
 contact_pattern = c(
  "100000", "100001", "101001", "101000", #Incarcerated Only
  "010000", "010001", "011001", "011000",  # Knows Inc Only
  "001100", "001101", "000100", "000101",  # Victim Only
  "110000", "110001", "111001", "111000", # Inc, Knows Inc
  "010100", "010101", "011101", "011100",  
  "110100", "110101", "111101", # Inc, Knows Inc, Vict
  "010110", "011111", "011110", "010111",  # Knows Inc, Vict, Knows Vict
  "110110",  "111110", "110111", "111111,"  # All
 ),
  label = c(
    "Incarcerated Only",
    "Knows Inc Only",
    "Victim Only",
    "Inc, Knows Inc",
    "Knows Inc, Vict",
    "Inc, Knows Inc, Vict",
    "Knows Inc, Vict, Knows Vict",
    "Also Vict",
    "Also and only Vict",
    "Victim Only",
    "Also and only KI"
  )
)


```

```{r, exploratoryimpactdataframe}

# Step 1: Create the pattern-to-label lookup
pattern_lookup <- tibble::tibble(
  contact_pattern = c(
    "100000", "100001", "101001", "101000",
    "010000", "010001", "011001", "011000",
    "001100", "001101", "000100", "000101",
    "110000", "110001", "111001", "111000",
    "100100", "100101", "101100", "101101",
    "100010", "101011", "100011", "101010",
    "010100", "010101", "011101", "011100",
    "110100", "110101", "111101",
    "010110", "011111", "011110", "010111",
    "110110", "111110", "110111", "111111"
  ),
  label = c(
    rep("Incarcerated Only", 4),
    rep("Knows Incarcerated Only", 4),
    rep("Victim Only", 4),
    rep("Inc & Knows Inc", 4),
    rep("Inc & Vict", 4),
    rep("Inc & Knows Vict", 4),
    rep("Knows Inc & Vict", 4),
    rep("Inc, Knows Inc, Vict", 3),
    rep("Knows Inc, Vict, Knows Vict", 4),
    rep("All", 4)
  )
)

# Step 2: Join this to dataset and count
pattern_counts <- raw_data_cleaned %>%
  inner_join(pattern_lookup, by = "contact_pattern") %>%
  count(label, contact_pattern, name = "count") %>%
  arrange(label, desc(count))

# View the result
print(pattern_counts)


```

### charts

```{r}

# Calculate basic totals
table_a <- tibble::tibble(
  Type = c("Been Incarcerated", "Knows Incarcerated", "Been Victimized"),
  Count = c(total_incarcerated, total_knows_incarcerated, total_victim),
  Percent_of_Sample = round(100 * Count / nrow(raw_data_cleaned), 1)
)

table_a %>%
  gt() %>%
  cols_label(
    Type = "Type",
    Count = "Total Count",
    Percent_of_Sample = "Percent of Sample"
  ) %>%
  tab_header(title = "Appendix Table A. Basic Contact Types") %>%
  tab_options(table.font.size = "small")

```

```{r}

combo_appendix_data %>%
  filter(label %in% c(
    "Incarcerated Only",
    "Also KI",
    "Also and only KI",
    "Also Vict",
    "Also and only Vict",
    "Also KI and Vict",
    "Also and only KI and Vict"
  )) %>%
  select(label, N, pct_total, pct_of_incarcerated) %>%
  arrange(desc(N)) %>%
  gt() %>%
  cols_label(
    label = "Contact Combo",
    N = "Count",
    pct_total = "Percent of Sample",
    pct_of_incarcerated = "Percent of Total Incarcerated Count"
  ) %>%
  fmt_number(columns = where(is.numeric), decimals = 1) %>%
  tab_header(title = "Appendix Table B. Incarcerated (Inc) Combos") %>%
  tab_options(table.font.size = "small")

```

```{r}

combo_appendix_data %>%
  filter(label %in% c(
    "KI Only",
    "Also Vict",
    "Also and only Vict"
  )) %>%
  select(label, N, pct_total, pct_of_ki) %>%
  arrange(desc(N)) %>%
  gt() %>%
  cols_label(
    label = "Contact Combo",
    N = "Count",
    pct_total = "Percent of Sample",
    pct_of_ki = "Percent of Total KI Count"
  ) %>%
  fmt_number(columns = where(is.numeric), decimals = 1) %>%
  tab_header(title = "Appendix Table C. Knows Incarcerated (KI) Combos") %>%
  tab_options(table.font.size = "small")
```

```{r}

combo_appendix_data %>%
  filter(label %in% c(
    "Victim Only",
    "Also KI",
    "Also and only KI"
  )) %>%
  select(label, N, pct_total, pct_of_victim) %>%
  arrange(desc(N)) %>%
  gt() %>%
  cols_label(
    label = "Contact Combo",
    N = "Count",
    pct_total = "Percent of Sample",
    pct_of_victim = "Percent of Total Victim Count"
  ) %>%
  fmt_number(columns = where(is.numeric), decimals = 1) %>%
  tab_header(title = "Appendix Table D. Victim (VI) Combos") %>%
  tab_options(table.font.size = "small")

```

```{r Incarcerated Combos}
#| eval: false
#| include: false

incarcerated_only <- sum(
  raw_data_cleaned$Q4_1_1 == 1 &
  raw_data_cleaned$Q4_1_2 == 0 &
  raw_data_cleaned$Q4_1_3 == 0 &
  raw_data_cleaned$Q4_2_1 == 0 &
  raw_data_cleaned$Q4_2_2 == 0 &
  raw_data_cleaned$Q4_2_3 == 0 &
  !is.na(raw_data_cleaned$Q4_1_1) &
  !is.na(raw_data_cleaned$Q4_1_2) &
  !is.na(raw_data_cleaned$Q4_1_3) &
  !is.na(raw_data_cleaned$Q4_2_1) &
  !is.na(raw_data_cleaned$Q4_2_2) &
  !is.na(raw_data_cleaned$Q4_2_3)
)

incarcerated_knowsinc <- sum(
  raw_data_cleaned$Q4_1_1 == 1 &
  raw_data_cleaned$Q4_1_2 == 1 &
  raw_data_cleaned$Q4_1_3 == 0 &
  raw_data_cleaned$Q4_2_1 == 0 &
  raw_data_cleaned$Q4_2_2 == 0 &
  raw_data_cleaned$Q4_2_3 == 0 &
  !is.na(raw_data_cleaned$Q4_1_1) &
  !is.na(raw_data_cleaned$Q4_1_2) &
  !is.na(raw_data_cleaned$Q4_1_3) &
  !is.na(raw_data_cleaned$Q4_2_1) &
  !is.na(raw_data_cleaned$Q4_2_2) &
  !is.na(raw_data_cleaned$Q4_2_3)
)

inc_knowsinc_victim <- sum(
  raw_data_cleaned$Q4_1_1 == 1 &
  raw_data_cleaned$Q4_1_2 == 1 &
  raw_data_cleaned$Q4_1_3 == 0 &
  raw_data_cleaned$Q4_2_1 == 1 &
  raw_data_cleaned$Q4_2_2 == 0 &
  raw_data_cleaned$Q4_2_3 == 0 &
  !is.na(raw_data_cleaned$Q4_1_1) &
  !is.na(raw_data_cleaned$Q4_1_2) &
  !is.na(raw_data_cleaned$Q4_1_3) &
  !is.na(raw_data_cleaned$Q4_2_1) &
  !is.na(raw_data_cleaned$Q4_2_2) &
  !is.na(raw_data_cleaned$Q4_2_3)
)

inc_knowsinc_knowsvict <- sum(
  raw_data_cleaned$Q4_1_1 == 1 &
  raw_data_cleaned$Q4_1_2 == 1 &
  raw_data_cleaned$Q4_1_3 == 0 &
  raw_data_cleaned$Q4_2_1 == 0 &
  raw_data_cleaned$Q4_2_2 == 1 &
  raw_data_cleaned$Q4_2_3 == 0 &
  !is.na(raw_data_cleaned$Q4_1_1) &
  !is.na(raw_data_cleaned$Q4_1_2) &
  !is.na(raw_data_cleaned$Q4_1_3) &
  !is.na(raw_data_cleaned$Q4_2_1) &
  !is.na(raw_data_cleaned$Q4_2_2) &
  !is.na(raw_data_cleaned$Q4_2_3)
)


inc_knowsinc_victim_knowsvict <- sum(
  raw_data_cleaned$Q4_1_1 == 1 &
  raw_data_cleaned$Q4_1_2 == 1 &
  raw_data_cleaned$Q4_1_3 == 0 &
  raw_data_cleaned$Q4_2_1 == 1 &
  raw_data_cleaned$Q4_2_2 == 1 &
  raw_data_cleaned$Q4_2_3 == 0 &
  !is.na(raw_data_cleaned$Q4_1_1) &
  !is.na(raw_data_cleaned$Q4_1_2) &
  !is.na(raw_data_cleaned$Q4_1_3) &
  !is.na(raw_data_cleaned$Q4_2_1) &
  !is.na(raw_data_cleaned$Q4_2_2) &
  !is.na(raw_data_cleaned$Q4_2_3)
)

incarcerated_victim <- sum(
  raw_data_cleaned$Q4_1_1 == 1 &
  raw_data_cleaned$Q4_1_2 == 0 &
  raw_data_cleaned$Q4_1_3 == 0 &
  raw_data_cleaned$Q4_2_1 == 1 &
  raw_data_cleaned$Q4_2_2 == 0 &
  raw_data_cleaned$Q4_2_3 == 0 &
  !is.na(raw_data_cleaned$Q4_1_1) &
  !is.na(raw_data_cleaned$Q4_1_2) &
  !is.na(raw_data_cleaned$Q4_1_3) &
  !is.na(raw_data_cleaned$Q4_2_1) &
  !is.na(raw_data_cleaned$Q4_2_2) &
  !is.na(raw_data_cleaned$Q4_2_3)
)

inc_victim_knowsvict <- sum(
  raw_data_cleaned$Q4_1_1 == 1 &
  raw_data_cleaned$Q4_1_2 == 0 &
  raw_data_cleaned$Q4_1_3 == 0 &
  raw_data_cleaned$Q4_2_1 == 1 &
  raw_data_cleaned$Q4_2_2 == 1 &
  raw_data_cleaned$Q4_2_3 == 0 &
  !is.na(raw_data_cleaned$Q4_1_1) &
  !is.na(raw_data_cleaned$Q4_1_2) &
  !is.na(raw_data_cleaned$Q4_1_3) &
  !is.na(raw_data_cleaned$Q4_2_1) &
  !is.na(raw_data_cleaned$Q4_2_2) &
  !is.na(raw_data_cleaned$Q4_2_3)
)

incarcerated_knowsvict <- sum(
  raw_data_cleaned$Q4_1_1 == 1 &
  raw_data_cleaned$Q4_1_2 == 0 &
  raw_data_cleaned$Q4_1_3 == 0 &
  raw_data_cleaned$Q4_2_1 == 0 &
  raw_data_cleaned$Q4_2_2 == 1 &
  raw_data_cleaned$Q4_2_3 == 0 &
  !is.na(raw_data_cleaned$Q4_1_1) &
  !is.na(raw_data_cleaned$Q4_1_2) &
  !is.na(raw_data_cleaned$Q4_1_3) &
  !is.na(raw_data_cleaned$Q4_2_1) &
  !is.na(raw_data_cleaned$Q4_2_2) &
  !is.na(raw_data_cleaned$Q4_2_3)
)


knowsinc_only <- sum(
  raw_data_cleaned$Q4_1_1 == 0 &
  raw_data_cleaned$Q4_1_2 == 1 &
  raw_data_cleaned$Q4_1_3 == 0 &
  raw_data_cleaned$Q4_2_1 == 0 &
  raw_data_cleaned$Q4_2_2 == 0 &
  raw_data_cleaned$Q4_2_3 == 0 &
  !is.na(raw_data_cleaned$Q4_1_1) &
  !is.na(raw_data_cleaned$Q4_1_2) &
  !is.na(raw_data_cleaned$Q4_1_3) &
  !is.na(raw_data_cleaned$Q4_2_1) &
  !is.na(raw_data_cleaned$Q4_2_2) &
  !is.na(raw_data_cleaned$Q4_2_3)
)

knowsinc_victim <- sum(
  raw_data_cleaned$Q4_1_1 == 0 &
  raw_data_cleaned$Q4_1_2 == 1 &
  raw_data_cleaned$Q4_1_3 == 0 &
  raw_data_cleaned$Q4_2_1 == 1 &
  raw_data_cleaned$Q4_2_2 == 0 &
  raw_data_cleaned$Q4_2_3 == 0 &
  !is.na(raw_data_cleaned$Q4_1_1) &
  !is.na(raw_data_cleaned$Q4_1_2) &
  !is.na(raw_data_cleaned$Q4_1_3) &
  !is.na(raw_data_cleaned$Q4_2_1) &
  !is.na(raw_data_cleaned$Q4_2_2) &
  !is.na(raw_data_cleaned$Q4_2_3)
)

knowsinc_knowsvict <- sum(
  raw_data_cleaned$Q4_1_1 == 0 &
  raw_data_cleaned$Q4_1_2 == 1 &
  raw_data_cleaned$Q4_1_3 == 0 &
  raw_data_cleaned$Q4_2_1 == 0 &
  raw_data_cleaned$Q4_2_2 == 1 &
  raw_data_cleaned$Q4_2_3 == 0 &
  !is.na(raw_data_cleaned$Q4_1_1) &
  !is.na(raw_data_cleaned$Q4_1_2) &
  !is.na(raw_data_cleaned$Q4_1_3) &
  !is.na(raw_data_cleaned$Q4_2_1) &
  !is.na(raw_data_cleaned$Q4_2_2) &
  !is.na(raw_data_cleaned$Q4_2_3)
)

knowsinc_victim_knowsvict <- sum(
  raw_data_cleaned$Q4_1_1 == 0 &
  raw_data_cleaned$Q4_1_2 == 1 &
  raw_data_cleaned$Q4_1_3 == 0 &
  raw_data_cleaned$Q4_2_1 == 1 &
  raw_data_cleaned$Q4_2_2 == 1 &
  raw_data_cleaned$Q4_2_3 == 0 &
  !is.na(raw_data_cleaned$Q4_1_1) &
  !is.na(raw_data_cleaned$Q4_1_2) &
  !is.na(raw_data_cleaned$Q4_1_3) &
  !is.na(raw_data_cleaned$Q4_2_1) &
  !is.na(raw_data_cleaned$Q4_2_2) &
  !is.na(raw_data_cleaned$Q4_2_3)
)

vict_only <- sum(
  raw_data_cleaned$Q4_1_1 == 0 &
  raw_data_cleaned$Q4_1_2 == 0 &
  raw_data_cleaned$Q4_1_3 == 0 &
  raw_data_cleaned$Q4_2_1 == 1 &
  raw_data_cleaned$Q4_2_2 == 0 &
  raw_data_cleaned$Q4_2_3 == 0 &
  !is.na(raw_data_cleaned$Q4_1_1) &
  !is.na(raw_data_cleaned$Q4_1_2) &
  !is.na(raw_data_cleaned$Q4_1_3) &
  !is.na(raw_data_cleaned$Q4_2_1) &
  !is.na(raw_data_cleaned$Q4_2_2) &
  !is.na(raw_data_cleaned$Q4_2_3)
)

victim_knowsvict <- sum(
  raw_data_cleaned$Q4_1_1 == 0 &
  raw_data_cleaned$Q4_1_2 == 0 &
  raw_data_cleaned$Q4_1_3 == 0 &
  raw_data_cleaned$Q4_2_1 == 1 &
  raw_data_cleaned$Q4_2_2 == 1 &
  raw_data_cleaned$Q4_2_3 == 0 &
  !is.na(raw_data_cleaned$Q4_1_1) &
  !is.na(raw_data_cleaned$Q4_1_2) &
  !is.na(raw_data_cleaned$Q4_1_3) &
  !is.na(raw_data_cleaned$Q4_2_1) &
  !is.na(raw_data_cleaned$Q4_2_2) &
  !is.na(raw_data_cleaned$Q4_2_3)
)

knowsvict_only <- sum(
  raw_data_cleaned$Q4_1_1 == 0 &
  raw_data_cleaned$Q4_1_2 == 0 &
  raw_data_cleaned$Q4_1_3 == 0 &
  raw_data_cleaned$Q4_2_1 == 1 &
  raw_data_cleaned$Q4_2_2 == 0 &
  raw_data_cleaned$Q4_2_3 == 0 &
  !is.na(raw_data_cleaned$Q4_1_1) &
  !is.na(raw_data_cleaned$Q4_1_2) &
  !is.na(raw_data_cleaned$Q4_1_3) &
  !is.na(raw_data_cleaned$Q4_2_1) &
  !is.na(raw_data_cleaned$Q4_2_2) &
  !is.na(raw_data_cleaned$Q4_2_3)
)

#data frame
contact_counts <- tibble::tibble(
  total_incarcerated = total_incarcerated,
  total_knows_incarcerated = total_knows_incarcerated,
  total_no_incarcerated = total_no_incarcerated,
  total_victim = total_victim,
  total_knows_victim = total_knows_victim,
  total_no_victim = total_no_victim,
  incarcerated_only = incarcerated_only,
  incarcerated_knowsinc = incarcerated_knowsinc,
  inc_knowsinc_victim = inc_knowsinc_victim,
  inc_knowsinc_knowsvict = inc_knowsinc_knowsvict,
  inc_knowsinc_victim_knowsvict = inc_knowsinc_victim_knowsvict,
  incarcerated_victim = incarcerated_victim,
  inc_victim_knowsvict = inc_victim_knowsvict,
  incarcerated_knowsvict = incarcerated_knowsvict,
  knowsinc_only = knowsinc_only,
  knowsinc_victim = knowsinc_victim,
  knowsinc_knowsvict = knowsinc_knowsvict,
  knowsinc_victim_knowsvict = knowsinc_victim_knowsvict,
  vict_only = vict_only,
  victim_knowsvict = victim_knowsvict,
  knowsvict_only = knowsvict_only
)

write_csv(contact_counts, "filled_contact_combo_counts.csv")


#view(contact_counts)

```

#### Tables

```{r}
combo_summary_labeled %>%
  select(label, contact_pattern, N, pct_total) %>%
  arrange(desc(N)) %>%
  gt() %>%
  cols_label(
    contact_pattern = "Q4 Pattern",
    label = "Description",
    N = "N",
    pct_total = "% of Sample"
  ) %>%
  fmt_number(columns = where(is.numeric), decimals = 1) %>%
  tab_header(title = "Appendix Table A. Full Distribution of Q4 Contact Combinations") %>%
  tab_options(table.font.size = "small")


print(combo_summary_labeled)
```

### C. Q6, 7, 8 - Ration Block

```{r}

#inspect

#raw_data %>% select(q6, q7, q8) %>% summarise_all(~ list(unique(.)))

# clean and transform
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(
    c(q6, q7, q8), 
    ~ na_if(as.numeric(.x), -99)
    )) %>% mutate(across(
    c(q6, q7, q8),
    ~ factor(.x, levels = 1:4,
             labels = c("Frequently", "Sometimes", "Rarely", "Never"),
             ordered = TRUE)
  ))

 raw_data_cleaned %>% count(q8)

```

### D. Prison & Election Policy Questions

#### D.1 "Purpose of Incarceration" - punishment or rehab?

```{r, punishorrehab}
#Originally 62_1 with 1  punish and 7  rehabilitate people

names(raw_data_cleaned)
# Renamed raw column to a more intuitive name
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(q62_1 = na_if(as.numeric(q62_1), -99))

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(penalpoint = 8 - q62_1) #reverse coded so 7 = most punitive

# Create ordered factor version of the reverse-coded scale (7 = most punitive)
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(penalpoint_ord = factor(
    penalpoint,
    levels = 1:7,
    labels = c(
      "Rehab (1)", "Rehab (2)", "Rehab (3)", "Neutral (4)",
      "Punish (5)", "Punish (6)", "Punish (7)"
    ),
    ordered = TRUE
  ))

# NOTE: Factor version of PenalPunitiveness (reverse-coded already: 1 = least punitive, 7 = most punitive)

# Tabulate raw values in Q62_1

# Count how many times -99 appears in the whole dataset
#sum(raw_data_cleaned == -99, na.rm = TRUE)



```

#### D.2 Punitive Items: (death penalty and LWOP) - prisonpen

```{r, prisonpen}
#ORIGINALLY Q25_1 and Q25_5, death penalty and LWOP (1 = strongly favor 4 = strongly oppose, 5 = don't know)

         
#rename, -99, ensure numeric
raw_data_cleaned <- raw_data_cleaned %>% 
  mutate(
    dpen         = na_if(as.numeric(q25_1), -99),
    lwop = na_if(as.numeric(q25_5), -99)
  )

pen_vars <- c("dpen", "lwop") #character vector 

# Reverse code (higher num should mean more support, lower less support, and  "Don't know" becomes midpoint)

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(
    all_of(pen_vars),
    ~ case_when(
    . == 1 ~ 4,           # strongly favor → strongly support help
    . == 2 ~ 3,
    . == 3 ~ 2,
    . == 4 ~ 1,
    . == 5 ~ 2.5,         # DK = neutral midpoint
    TRUE ~ NA_real_
  ),  # note: we skip 5 = DK
    .names = "{.col}_rev"
  ))


# step 3: save just the means of pen questions as prisonpen

raw_data_cleaned <- raw_data_cleaned %>%
  rowwise() %>%
  mutate(prisonpen = mean(c_across(c(dpen_rev, lwop_rev)), na.rm = TRUE)) %>%
  ungroup()

summary(raw_data_cleaned$dpen_rev)
summary(raw_data_cleaned$lwop_rev)
```

Takeaway: prison pen, prison pen 1 = dpen, prison pen 5 = lwop now -confirmed that prisonpen works -- constructed with average of death penalty and lwop. USE dpen_rev and lwop_rev ! and prisonpen

OPTIONAL/potentially helpful for tables: factor version

```{r, q25_1_ord}
# #OPTIONAL/potentially helpful for tables: factor version
# raw_data_cleaned <- raw_data_cleaned %>%
#   mutate(q25_1_ord = factor(
#     q25_1,
#     levels = 1:5,
#     labels = c("Strongly favor", "Favor", "Oppose", "Strongly oppose", "Don't know"),
#     ordered = TRUE
#   ))
```

#### D.3 Punitive Items: Doing Cronback's Alpha

```{r}
library(psych)
library(dplyr)

#prisonpen, prisonpen_1, prisonpen_5
pen_vars <- c("dpen_rev", "lwop_rev")

pen_data <- raw_data_cleaned %>%
  select(all_of(pen_vars))

cor(pen_data,
    use = "pairwise.complete.obs")

print(psych::corr.test(pen_data))

print(psych::alpha(pen_data))

# Quick look
psych::describe(pen_data)

# Correlation (more interpretable than alpha for 2 items)
cor(raw_data_cleaned[pen_vars], use = "pairwise.complete.obs")

# Cronbach's alpha (fine to report, but remember it's 2 items)
psych::alpha(pen_data, check.keys = FALSE)

```

results - The **correlation** between the two items is **r = 0.238**.

- With only two items, Cronbach’s alpha is essentially a rescaled version of that correlation. (So don’t oversell alpha here; the correlation is the clean summary.)

Interpretation

- **These two punitive items are positively related, but  weakly so.** People who favor the death penalty are *somewhat* more likely to favor LWOP, but plenty of respondents don’t treat them as interchangeable.

- That’s not necessarily a “problem.” Death penalty and LWOP are conceptually related (punitive severity), but they can tap **different moral/legal intuitions** (irreversibility, innocence risk, retribution vs incapacitation), so moderate association is plausible.

- The response distribution shows a **nontrivial chunk of “Don’t know”** recoded to 2.5 (15–17%) and about **8% missing** for each item. The DK-to-midpoint choice can slightly inflate apparent consistency by shrinking variance, so reliability is a conservative-ish check but not definitive.

Visual





#### D.4_A Q25 'Help-Oriented' Prison Policy.

These will include prison and election policies. Here, I first set up the prison policy Q's, then prison/voting Q's.

```{r, prisonhelp}
# Originally 1 = strong favor, 4 = strong oppose, 5 = don't know
# Define Q25 help-oriented items
q25_help_items <- c("q25_2", "q25_3", "q25_4", "q25_6")

# Step 1: Recode -99 to NA and ensure numeric
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(all_of(q25_help_items), ~ na_if(as.numeric(.), -99)))

# Step 2: Create reverse-coded versions (1 = strong favor → 4 = strong oppose becomes 4 → 1 so higher means more support for help oriented policyaQ                                                 )
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(
    all_of(q25_help_items),
    ~ case_when(
    . == 1 ~ 4,           # strongly favor → strongly support help
    . == 2 ~ 3,
    . == 3 ~ 2,
    . == 4 ~ 1,
    . == 5 ~ 2.5,         # DK = neutral midpoint
    TRUE ~ NA_real_
  ),  # note: we skip 5 = DK
    .names = "{.col}_rev"
  ))

# now, raw dta cleaned has q25_n_rev (each reverse coded so that higher number means more support)

# step 3: save just the means of q25 help questions as q25_help_items

raw_data_cleaned <- raw_data_cleaned %>% 
  rowwise() %>% #dplyr should treat each row as own group so c_across() can return row level vector of cols for that row
  mutate(q25_help_items = mean(c_across(c(q25_2_rev, q25_3_rev, q25_4_rev, q25_6_rev)), na.rm = TRUE)) %>%
  ungroup()

names(raw_data_cleaned)

```

##### D.4_A Descriptive
q25_help_items is the average of means of the help items (just prison policy "help" excluding the prison and voting ones); q25_help_items is to look at them separately
```{r}

#checking variance + SD here 08-2026)
summary(raw_data_cleaned$q25_help_items)

summary(raw_data_cleaned$q25_6_rev)


q25_help_vars <- c("q25_2_rev", "q25_3_rev", "q25_4_rev", "q25_6_rev")

psych::describe(raw_data_cleaned[q25_help_vars])


#I get min 1 1st qu 2.5 median 2.875 mean 2.88 3q 3.250 max 4 nas = 186

#histograms
print(hist(raw_data_cleaned$q25_2_rev))
print(hist(raw_data_cleaned$q25_3_rev))
print(hist(raw_data_cleaned$q25_4_rev))
print(hist(raw_data_cleaned$q25_6_rev))

library(dplyr)
library(tidyr)
library(ggplot2)

q25_long <- raw_data_cleaned %>%
  select(all_of(q25_help_vars)) %>%
  pivot_longer(
    cols = everything(),
    names_to = "item",
    values_to = "response"
  )

ggplot(q25_long, aes(x = response)) +
  geom_bar() +
  facet_wrap(~ item) +
  labs(
    x = "Response",
    y = "Count"
  )

#barplots

par(mfrow = c(2, 2))

barplot(table(raw_data_cleaned$q25_2_rev),
        main = "q25_2_rev",
        xlab = "Response")

barplot(table(raw_data_cleaned$q25_3_rev),
        main = "q25_3_rev",
        xlab = "Response")

barplot(table(raw_data_cleaned$q25_4_rev),
        main = "q25_4_rev",
        xlab = "Response")

barplot(table(raw_data_cleaned$q25_6_rev),
        main = "q25_6_rev",
        xlab = "Response")

par(mfrow = c(1, 1))


#bar plot for overleaf

# Create plotting data
q25_plot_data <- raw_data_cleaned %>%
  select(
    q25_2_rev,
    q25_3_rev,
    q25_4_rev,
    q25_6_rev
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "item",
    values_to = "position"
  ) %>%
  filter(!is.na(position)) %>%
  mutate(
    item = factor(
      item,
      levels = c(
        "q25_2_rev",
        "q25_3_rev",
        "q25_4_rev",
        "q25_6_rev"
      ),
      labels = c(
        "Minimum Wage",
        "No Cost Calls",
        "Fund Education",
        "Parental Alts"
      )
    ),
    
    # Treat the response categories as discrete/ordinal categories
    position = factor(
      position,
      levels = c(1, 2, 2.5, 3, 4)
    )
  ) %>%
  count(item, position, .drop = FALSE)


# Make figure
q25_barplot <- ggplot(
  q25_plot_data,
  aes(x = position, y = n)
) +
  geom_col(
    width = 0.70,
    fill = "grey35"
  ) +
  facet_wrap(
    ~ item,
    ncol = 2
  ) +
  scale_x_discrete(drop = FALSE) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05))
  ) +
  labs(
    x = "Position",
    y = "Number of Respondents",
    caption = paste0(
      "Position: 1 = Strongly Oppose; 2 = Somewhat Oppose; ",
      "2.5 = Unsure; 3 = Somewhat Support; 4 = Strongly Support."
    )
  ) +
  theme_classic(base_size = 11) +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(
      face = "bold",
      size = 11
    ),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 9.5),
    panel.spacing = grid::unit(1, "lines"),
    plot.caption = element_text(
      size = 9,
      hjust = 0,
      margin = margin(t = 8)
    )
  )

ggsave(
  filename = "figures/q25_help_items_barplot.pdf",
  plot = q25_barplot,
  width = 6.5,
  height = 5,
  units = "in",
  device = cairo_pdf
)
q25_barplot
# box plot 
boxplot(
  raw_data_cleaned[q25_help_vars],
  main = "Q25 Help Items",
  ylab = "Response"


```
Correlations, Crohnbach's alpha

```{r}

q25_help_data <- raw_data_cleaned %>%
  select(all_of(q25_help_vars))

cor(q25_help_data,
    use = "pairwise.complete.obs")

print(psych::corr.test(q25_help_data))

print(psych::alpha(q25_help_data))

# Correlation (more interpretable than alpha for 2 items)
cor(prisonpen, use = "pairwise.complete.obs")

# Cronbach's alpha (fine to report, but remember it's 2 items)
psych::alpha(prisonpen, check.keys = FALSE)

q25_help_data <- raw_data_cleaned %>%
  select(all_of(q25_help_vars))

cor(q25_help_data,
    use = "pairwise.complete.obs")
```




Ordered factor versions- could be useful for plotting/tables/labeling) but as of 08-2026 don't use

```{r, prisonhelp_factor}
# raw_data_cleaned <- raw_data_cleaned %>%
#   mutate(across(
#     all_of(q25_help_items),
#     ~ factor(.,
#              levels = 1:5,
#              labels = c("Strongly favor", "Favor", "Oppose", "Strongly oppose", "Don't know"),
#              ordered = TRUE),
#     .names = "{.col}_ord"
#   ))

```

#### D.4_B Q35 'Help-Oriented' Prison/Election Items (Adding to Above)

```{r, electionhelpqs}
names(raw_data_cleaned) %>% stringr::str_subset("q25|q23")

# Define Q23 help-oriented items (election - q23_2 is elects_reenfranchise and q23_7 is voting in jail (Requiring that jail and election officers ensure that people who are eligible to vote while incarcerated are able to exercise that right.) OG 1 = favor, 2 = oppose, 3 = DK

#rename, -99, ensure numeric
raw_data_cleaned <- raw_data_cleaned %>% 
  mutate(
    auto_reg         = na_if(as.numeric(q23_1), -99),
    post_felony_vote = na_if(as.numeric(q23_2), -99),
    early_vote       = na_if(as.numeric(q23_3), -99),
    election_holiday = na_if(as.numeric(q23_4), -99),
    voter_id         = na_if(as.numeric(q23_5), -99),
    noncit_local     = na_if(as.numeric(q23_6), -99),
    jail_access      = na_if(as.numeric(q23_7), -99)
  )

all_electoral_vars <- c("auto_reg", "post_felony_vote", "early_vote", "election_holiday", "voter_id", "noncit_local", "jail_access") #character vector 

#recode 1-3 scale to match direction of prison help items (also intuition that higher number means more support)
# here, 1 = favor, 2 = oppose, 3 = DK
# we want higher = more support for help-oriented policy so 1-> 3, 2-> 1, 3 -> 2
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(
    all_of(all_electoral_vars),
    ~ case_when(
      . == 1 ~ 3,
      . == 2 ~ 1,
      . == 3 ~ 2,
      TRUE ~ NA_real_
    ),
    .names = "{.col}_rev"
  ))

# respondent level means: all electoral vars
# 
# raw_data_cleaned <- raw_data_cleaned %>% 
#   rowwise() %>% #dplyr should treat each row as own group so c_across() can return row level vector of cols for that row
#   mutate(all_electoral_means = mean(c_across(c(auto_reg_rev, post_felony_vote_rev, early_vote_rev, election_holiday_rev, voter_id_rev, noncit_local_rev, jail_access_rev)), na.rm = TRUE)) %>%
#   ungroup()

summary(raw_data_cleaned$all_electoral_means)#all_election_items now exists as mean across all 7 Q23 questions, reverse coded, for each observation. 

## Jail voting and rights restoration

incarc_electoral_vars <- c("post_felony_vote_rev", "jail_access_rev")

# respondent level means: just the two (incarc_electoral_means)

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    incarc_electoral_means = if_else(
      !is.na(post_felony_vote_rev) & !is.na(jail_access_rev),
      (post_felony_vote_rev + jail_access_rev) / 2,
      NA_real_
    )
  )

# old: q23_help_items <- c("q23_2", "q23_7")
# new: incarc_electoral_means, post_felony_vote_rev, jail_access_rev

# to compare to other prison policy questions on same scale, rescale these two  the 1–3 items to 1–4 range

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(
    all_of(incarc_electoral_vars),
    ~ case_when(
      . == 3 ~ 4,
      . == 1 ~ 1,
      . == 2 ~ 2.5,
      TRUE ~ NA_real_
    ),
    .names = "{.col}_1_to_4"
  ))

# produces: post_felony_vote_rev_1_to_4, jail_access_rev_1_to_4

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    incarc_electoral_mean_1_to_4 = if_else(
      !is.na(post_felony_vote_rev_1_to_4) &
        !is.na(jail_access_rev_1_to_4),
      (post_felony_vote_rev_1_to_4 +
         jail_access_rev_1_to_4) / 2,
      NA_real_
    )
  )
hist(raw_data_cleaned$post_felony_vote_rev)
hist(raw_data_cleaned$jail_access_rev)

print(summary(raw_data_cleaned$post_felony_vote_rev))
print(summary(raw_data_cleaned$jail_access_rev))
``` 
Come out of above block with: 
post_felony_vote_rev          individual item, 1–3
jail_access_rev               individual item, 1–3
incarc_electoral_mean         combined measure, 1–3

post_felony_vote_rev_1_to_4   comparable item, 1–4
jail_access_rev_1_to_4        comparable item, 1–4
incarc_electoral_mean_1_to_4  comparable combined measure, 1–4

#### D.4_C Prison/Election Items Descriptive look 
What's up
```{r}

#bar plot for overleaf

# Create plotting data
incarc_electoral_data <- raw_data_cleaned %>%
  select(
    post_felony_vote_rev,
    jail_access_rev
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "item",
    values_to = "position"
  ) %>%
  filter(!is.na(position)) %>%
  mutate(
    item = factor(
      item,
      levels = c(
        "post_felony_vote_rev",
        "jail_access_rev"
      ),
      labels = c(
        "Rights Restoration",
        "Jail Vote Access"
      )
    ),
    
    # Treat the response categories as discrete/ordinal categories
    position = factor(
      position,
      levels = c(1, 2, 3)
    )
  ) %>%
  count(item, position, .drop = FALSE)


# Make figure
incarc_electoral_barplot <- ggplot(
  incarc_electoral_data ,
  aes(x = position, y = n)
) +
  geom_col(
    width = 0.70,
    fill = "grey35"
  ) +
  facet_wrap(
    ~ item,
    ncol = 2
  ) +
  scale_x_discrete(drop = FALSE) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05))
  ) +
  labs(
    x = "Position",
    y = "Number of Respondents",
    caption = paste0(
      "Position: 1 = Oppose; 2 = Unsure; ",
      "3 = Somewhat Support."
    )
  ) +
  theme_classic(base_size = 11) +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(
      face = "bold",
      size = 11
    ),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 9.5),
    panel.spacing = grid::unit(1, "lines"),
    plot.caption = element_text(
      size = 9,
      hjust = 0,
      margin = margin(t = 8)
    )
  )


ggsave(
  filename = "figures/incarc_electoral_barplot.pdf",
  plot = q25_barplot,
  width = 6.5,
  height = 5,
  units = "in",
  device = cairo_pdf
)
incarc_electoral_barplot

```


Correlation between jail voting and rights restoration 
```{r}

incarc_electoral_data <- raw_data_cleaned %>%
  select(all_of(incarc_electoral_vars))

cor(incarc_electoral_data,
    use = "pairwise.complete.obs")

print(psych::corr.test(incarc_electoral_data))

print(psych::alpha(incarc_electoral_data))
```


What about relationship between all the help variables (2 election, 4 others?)

Need

post_felony_vote_rev
jail_access_rev_1_to_4
q25_help_data or q25_2_rev,
    q25_3_rev,
    q25_4_rev,
    q25_6_rev


```{r, electionhelpqs}

# ALL HELP ITEMS

all_help_items <- c(
  "q25_2_rev",
  "q25_3_rev",
  "q25_4_rev",
  "q25_6_rev",
  "post_felony_vote_rev_1_to_4",
  "jail_access_rev_1_to_4"
)

#df for the six total vars
all_help_data <- raw_data_cleaned %>%
  select(all_of(all_help_items))

# now, pairwise corrs
view(cor(all_help_data,
    use = "pairwise.complete.obs"))

```

#### D5. Variance and Stand Dev

```{r}
#don't need to redo this
#help_policy_items <- c("q25_2", "q25_3", "q25_4", "q25_6", "q23_2", "q23_7")

# Variance and SD across help policy items
raw_data_cleaned <- raw_data_cleaned %>%
  rowwise() %>%
  mutate(
    variancehelp = var(c_across(all_of(help_policy_items)), na.rm = TRUE),
    sdhelp = sd(c_across(all_of(help_policy_items)), na.rm = TRUE)
  ) %>%
  ungroup()


```

### E.Policy Orientation (polor) control

Decided to skip because of how questions are worded/coded. Will need to return to this later

10-27 Diagnostics

```{r}
names(raw_data_cleaned) %>% stringr::str_subset("q25|q23|q62|prison|help|punish|rehab")

glimpse(raw_data_cleaned %>% select(starts_with("q25"), starts_with("q23"), starts_with("q62"), starts_with("prison"), starts_with("penal")))


glimpse(dplyr::select(raw_data_cleaned,
                      starts_with("q25_"),
                      starts_with("q23_"),
                      starts_with("q62"),
                      starts_with("prison"),
                      starts_with("penal")))

```

### F. FIRE controls

```{r}

#raw_data_cleaned %>% count(q32)

#raw_data_cleaned %>% count(Q33)

#raw_data_cleaned %>% count(Q34)

#raw_data_cleaned %>% count(Q111)

fire_vars <- c("q32", "q33", "q34", "q111")

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(
    all_of(fire_vars),
    ~ as.numeric(na_if(as.character(.x), "-99"))
  ))

raw_data_cleaned <- raw_data_cleaned %>%
  rename(
    fire_rare = q32,
    fire_privilege = q33,
    fire_angry = q34,
    fire_fear = q111
  )

summary(raw_data_cleaned$fire_angry)
summary(raw_data_cleaned$fire_privilege)
summary(raw_data_cleaned$fire_fear)
summary(raw_data_cleaned$fire_rare)


```

### G. Race

```{r}

#convert + clean
race_vars <- paste0("q51_", 1:8)

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(all_of(race_vars), ~ na_if(as.numeric(.x), -99)))

# REPLACE your current race dummies with this block
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    # Ethnicity first: Hispanic gets priority
    Hispanic = ifelse(q51_7 == 1, 1, 0),

    # Among non-Hispanics, assign Black and White
    Black    = ifelse(q51_2 == 1 & q51_7 != 1, 1, 0),
    White    = ifelse(q51_1 == 1 & q51_7 != 1, 1, 0)
  ) %>%
  mutate(
    # Everyone else with any race selection becomes OtherRace (mutually exclusive)
    OtherRace = ifelse(
      (Black + Hispanic + White) == 0 &
        rowSums(dplyr::select(., dplyr::all_of(race_vars)), na.rm = TRUE) > 0,
      1, 0
    )
  )

```

DOUBLE CHECK

```{r}

# no overlaps among the 4 dummies
sum((raw_data_cleaned$Black + raw_data_cleaned$White +
     raw_data_cleaned$Hispanic + raw_data_cleaned$OtherRace) > 1, na.rm = TRUE)

# counts by category (mutually exclusive)
table(
  with(raw_data_cleaned,
       dplyr::case_when(
         Hispanic == 1 ~ "Hispanic",
         Black    == 1 ~ "Black",
         White    == 1 ~ "White",
         OtherRace== 1 ~ "OtherRace",
         TRUE          ~ NA_character_
       )),
  useNA = "ifany"
)


```

Diagnosed problem in model where Black + Other dropped - Oct 27 Suggests all Black respondents had missing data so were dropped

### H. Party + Ideology controls

```{r}

#raw_data_cleaned %>% count(q47)

#clean dem 101
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(dem101 = na_if(as.numeric(dem101), -99))

#party ID dummies
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    Republican = case_when(
      dem101 == 1 ~ 1,
      dem101 %in% c(2, 3, 4) ~ 0,
      TRUE ~ NA_real_
    ),
    Democrat = case_when(
      dem101 == 2 ~ 1,
      dem101 %in% c(1, 3, 4) ~ 0,
      TRUE ~ NA_real_
    )
  )

#clean ideology
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    q46 = na_if(as.numeric(q46), -99),
    q47 = na_if(as.numeric(q47), -99)
  )

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    fiscal_ideology = ifelse(!is.na(q46), 6 - q46, NA),
    social_ideology = ifelse(!is.na(q47), 6 - q47, NA)
  )


```

### I. Gender

```{r, gender}
#raw_data_cleaned %>% count(q55)

#trans clean

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(q55 = na_if(as.numeric(q55), -99))

#trans indicator
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(trans_id = case_when(
    q55 == 1 ~ 1,
    q55 %in% c(2, 3) ~ 0,  # Conservative approach: counts only self-identified trans as 1
    TRUE ~ NA_real_
  ))


#gender clean
gender_vars <- paste0("q56_", 1:4)

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(all_of(gender_vars), ~ na_if(as.numeric(.x), -99)))


#constructing the gender binary
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(gender = case_when(
    q56_2 == 1 ~ 1,  # Selected man
    q56_1 == 1 & (q56_2 != 1 | is.na(q56_2)) ~ 0,  # Woman only or with genderqueer/other
    q56_3 == 1 | q56_4 == 1 ~ 0,  # Genderqueer or other, no man selected
    TRUE ~ NA_real_
  ))


```

### J. Victim Analysis Vars

```{r}

# police policies

# Originally 6 = strong oppose, 7= somewhat oppose, 8 = neutral, 9 = somewhat support, 10 = strong support

# Define Q25 police items
police_items <- c("q15_1", "q15_2", "q15_3", "q15_4", "q15_5")

# Step 1: Recode -99 to NA and ensure numeric
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(all_of(police_items), ~ na_if(as.numeric(.), -99)))

# Step 2: Recode so 1-5 (6 = strong oppose → 1 = strong oppose, 
# 7 = somewhat oppose → 2 = somewhat oppose, 
# 8 = neutral → 3 = neutral, 
# 9 = somewhat support → 4 = somewhat support,
#10 = strong support → 5 = strong support)

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(
    all_of(police_items),
    ~ case_when(
    . == 6 ~ 1,          
    . == 7 ~ 2,
    . == 8 ~ 3,
    . == 9 ~ 4,
    . == 10 ~ 5,        
    TRUE ~ NA_real_
  ),  
    .names = "{.col}_1_to_5"
  ))

police_items_1_to_5 <- c(
  "q15_1_1_to_5",
  "q15_2_1_to_5",
  "q15_3_1_to_5",
  "q15_4_1_to_5",
  "q15_5_1_to_5"
)

```

```{r}

## trust in government 

summary(raw_data_cleaned$q30)
unique(raw_data_cleaned$gov_attn)

# Originally 1 = a lot, 2 = some, 3 = not muc

# rename
## trust in government attention
# q30 originally: 1 = a lot, 2 = some, 3 = not much
# recode so higher = more government attention/trust

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    gov_attn = case_when(
      as.numeric(q30) == -99 ~ NA_real_,
      as.numeric(q30) == 1 ~ 3,
      as.numeric(q30) == 2 ~ 2,
      as.numeric(q30) == 3 ~ 1,
      TRUE ~ NA_real_
    )
  )

hist(raw_data_cleaned$gov_attn)

## did vote in 2020
# q31: 1 = no, not eligible; 2 = no, eligible; 3 = yes

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    vote2020 = na_if(as.numeric(q31), -99)
  )


## plans to vote
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    willvote = na_if(as.numeric(q110), -99)
  )


```

Note:

Approximately 561 respondents did not provide responses to the political behavior items (government attention, past vote, intended vote). Missingness is perfectly overlapping across these items, indicating that these questions were likely administered to a subset of respondents rather than reflecting item-level nonresponse.

### K. LAST OF CLEANING - WRITE CSV

```{r}

names(raw_data_cleaned)
write_csv(raw_data_cleaned, "data/cleaned/coded_data__08_2026.csv")
#write_csv(raw_data_cleaned, "data/cleaned/coded_data_may9.csv")

```

# ARCHIVED

```{r}
#| eval: false
#| include: false

getwd()
impacted <- read.csv("/Users/kristinamensik/Library/Mobile Documents/com~apple~CloudDocs/Documents/Duke/2024-2025/902 Legislators Pre-Ananlysis Plan/Prelim Folder/Prelim 3-23/data/impacted.csv")

view(impacted)

# Count unique combinations
impacted_summary <- impacted %>%
  select(been_inc, know_inc, none_inc, been_vict, know_vict, no_vict) %>%
  group_by(across(everything())) %>%
  summarise(count = n(), .groups = "drop") %>%
  arrange(desc(count))

# View the result
view(impacted_summary)

#only incarcerated
sum(
  impacted$been_inc == 1 &
  impacted$no_vict == 1 &
  impacted$know_inc == 0 &
  impacted$none_inc == 0 &
  impacted$been_vict == 0 &
  impacted$know_vict == 0
)

#incarcerated & Victimized
sum(
  impacted$been_inc == 1 &
  impacted$no_vict == 0 &
  impacted$know_inc == 0 &
  impacted$none_inc == 0 &
  impacted$been_vict == 1 &
  impacted$know_vict == 0
)

#incarcerated & Victimized & Know Inc
sum(
  impacted$been_inc == 1 &
  impacted$no_vict == 0 &
  impacted$know_inc == 1 &
  impacted$none_inc == 0 &
  impacted$been_vict == 1 &
  impacted$know_vict == 0
)

#only know incarcerated
sum(
  impacted$been_inc == 0 &
  impacted$no_vict == 1 &
  impacted$know_inc == 1 &
  impacted$none_inc == 0 &
  impacted$been_vict == 0 &
  impacted$know_vict == 0
)

#only vict
sum(
  impacted$been_inc == 0 &
  impacted$no_vict == 0 &
  impacted$know_inc == 0 &
  impacted$none_inc == 1 &
  impacted$been_vict == 1 &
  impacted$know_vict == 0
)

#only know vict
sum(
  impacted$been_inc == 0 &
  impacted$no_vict == 0 &
  impacted$know_inc == 0 &
  impacted$none_inc == 1 &
  impacted$been_vict == 0 &
  impacted$know_vict == 1
)

#know vict, know incarc, been vict
sum(
  impacted$been_inc == 0 &
  impacted$no_vict == 0 &
  impacted$know_inc == 1 &
  impacted$none_inc == 0 &
  impacted$been_vict == 1 &
  impacted$know_vict == 1
)


```
