---
title: "variable_prep_script"
format: html
editor: visual
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,       # Hide code by default
  warning = FALSE,
  message = FALSE
)

# Load packages
library(dplyr)
library(tidyr)
library(readr)
library(janitor)   # for clean_names()
library(tibble)
library(stringr)
library(haven)
library(tibble)
library(gt)
library(psych)
library(ggplot2)
library(tidyverse)

```

## 1. Directory of Most Recent Data

Load cleaned survey data

File created in data_prep.qmd

```{r}
#view(raw_data_cleaned)

raw_data_cleaned <- readr::read_csv("data/cleaned/over_8_cleaned.csv") 


#names(raw_data_cleaned)

nrow(raw_data_cleaned)

```

## 2. Variables

### A. CJ Experiences - Basic Coding

#### Extra binaries (incarc, KI, vict)

```{r}
#extra binary
 raw_data_cleaned <- raw_data_cleaned %>%
   mutate(
     incarc_bin = if_else(
       incarc_raw == 1,
       1,
       0,
       missing = NA_real_
     )
   )
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    noincarc_bin = if_else(
  (incarc_raw == 0 & knowincarc_raw == 0) | noincarc_raw == 1,
  1, 0,
  missing = NA_real_
))


 raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    knowincarc_bin = if_else(
      knowincarc_raw == 1,
      1,
      0,
      missing = NA_real_
    )
  )

 raw_data_cleaned <- raw_data_cleaned %>%
   mutate(
     vict_bin = if_else(
       vict_raw == 1,
       1,
       0,
       missing = NA_real_
     )
   )


```

#### Impact Plus (Constructing For-Analysis)

```{r, impactvar}

#counting the *total* of those who said they were incarcerated, know incarcerated, or victims)

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    impact = if_else(
      incarc_raw == 1 | knowincarc_raw == 1 | vict_raw == 1,
      1, 0,
      missing = NA_real_
    )
  )

# post Aug 2026 version with casewhen,where  incarc = 0 to mean impact = 0 (no impact, not just no incarceration):

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    incarc = case_when(
      incarc_raw == 1 ~ 1, 
      impact == 0 ~ 0, 
      TRUE ~ NA_real_ #"otherwise, assign NA"
    )
  )

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    knowincarc = case_when(
      knowincarc_raw == 1 ~ 1, 
      impact == 0 ~ 0, 
      TRUE ~ NA_real_ #"otherwise, assign NA"
    )
  )

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    vict = case_when(
      vict_raw == 1 ~ 1, 
      impact == 0 ~ 0, 
      TRUE ~ NA_real_ #"otherwise, assign NA"
    )
  )

```

### B. Other Experiences

#### Impact Plus (Constructing For-Analysis)

Coding in same way as CJ experiences but in giant loop

```{r}

# Function to code one experience family ---------------------------------

code_experience <- function(data, self_var, know_var, no_var, prefix) {
  
  impact_var   <- paste0(prefix, "_impact")
  self_bin     <- paste0(self_var, "_bin")
  know_bin     <- paste0(know_var, "_bin")
  no_bin       <- paste0(no_var, "_bin")
  
  data %>%
    # 1. Overall exposure:
    #    1 = self OR know someone
    #    0 = neither
    mutate(
      !!impact_var := if_else(
        .data[[self_var]] == 1 | .data[[know_var]] == 1,
        1, 0,
        missing = NA_real_
      )
    ) %>%
    
    # 2. Recode self/know variables:
    #    1 = this specific kind of exposure
    #    0 = completely unexposed
    #    NA = exposed, but not in this particular way
    mutate(
      !!self_var := case_when(
        .data[[self_var]] == 1 ~ 1,
        .data[[impact_var]] == 0 ~ 0,
        TRUE ~ NA_real_
      ),
      
      !!know_var := case_when(
        .data[[know_var]] == 1 ~ 1,
        .data[[impact_var]] == 0 ~ 0,
        TRUE ~ NA_real_
      )
    ) %>%
    
    # 3. Binary versions
    mutate(
      !!self_bin := if_else(
        .data[[self_var]] == 1,
        1, 0,
        missing = NA_real_
      ),
      
      !!know_bin := if_else(
        .data[[know_var]] == 1,
        1, 0,
        missing = NA_real_
      ),
      
      # 1 = explicitly in neither category
      #     OR explicitly selected the "no" option
      !!no_bin := if_else(
        (.data[[self_var]] == 0 & .data[[know_var]] == 0) |
          .data[[no_var]] == 1,
        1, 0,
        missing = NA_real_
      )
    )
}

```

9 experience groups

```{r}
experience_sets <- list(
  
  medicaid = c(
    self = "on_medicaid",
    know = "knowmedicaid",
    no   = "no_medicaid"
  ),
  
  tanf = c(
    self = "tanf",
    know = "knowtanf",
    no   = "notanf"
  ),
  
  snap = c(
    self = "snap",
    know = "knowsnap",
    no   = "nosnap"
  ),
  
  immigrated = c(
    self = "immigrated",
    know = "knowimmigrated",
    no   = "noimmigrated"
  ),
  
  union = c(
    self = "unionmember",
    know = "knowunion",
    no   = "nounion"
  ),
  
  abortion = c(
    self = "aborted",
    know = "knowaborted",
    no   = "noaborted"
  ),
  
  unemployment = c(
    self = "unemployed",
    know = "knowunemployed",
    no   = "nounemployed"
  ),
  
  uninsured = c(
    self = "uninsured",
    know = "knowuninsured",
    no   = "nouninsured"
  ),
  
  unable_vote = c(
    self = "unable_vote",
    know = "know_unable_vote",
    no   = "no_unable_vote"
  )
)
```

apply experience groups

```{r}
for (experience in names(experience_sets)) {
  
  vars <- experience_sets[[experience]]
  
  raw_data_cleaned <- code_experience(
    data     = raw_data_cleaned,
    self_var = vars["self"],
    know_var = vars["know"],
    no_var   = vars["no"],
    prefix   = experience
  )
}


```

### C. Ration, Hardship Block

### 

```{r}

# clean and transform
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(
    c(ration_scripts, ration_period, ration_food), 
    ~ na_if(as.numeric(.x), -99)
    )) %>% mutate(across(
    c(ration_scripts, ration_period, ration_food),
    ~ factor(.x, levels = 1:4,
             labels = c("Frequently", "Sometimes", "Rarely", "Never"),
             ordered = TRUE)
  ))

```

### D. FIRE controls

fire_rare: racial problems in the US are rare, isolated situations

- 1 is strongly agree, 4 strongly disagree

fire_privilege: white people have advantages

- 1 is strongly agree, 4 is strongly disagree

fire_angry: I am angry that racism exists

- 1 is strongly agree, 4 is strongly disagree

fire_fear: I am fearful of other races

- 1 is strongly agree, 4 is strongly disagree

```{r}
fire_vars <- c("fire_rare", "fire_privilege", "fire_angry", "fire_fear")


raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(
    all_of(fire_vars),
    ~ as.numeric(na_if(as.character(.x), "-99"))
  ))


# hist(raw_data_cleaned$fire_rare)
# 
# table(raw_data_cleaned$fire_fear)
# 
# raw_data_cleaned <- raw_data_cleaned %>%
#   mutate(fire_rare)
#     ~ case_when(
#       . == 1 ~ 4,          
#       . == 2 ~ 3,
#       . == 3 ~ 2,
#       . == 4 ~ 1,        
#       TRUE ~ NA_real_
#     )
# 
# 
# raw_data_cleaned <- raw_data_cleaned %>%
#   mutate(fire_fear)
# ~ case_when(
#   . == 1 ~ 4,          
#   . == 2 ~ 3,
#   . == 3 ~ 2,
#   . == 4 ~ 1,        
#   TRUE ~ NA_real_
# )
# 
# hist(raw_data_cleaned$fire_fear)


```

Index?

```{r}
fire_vars <- c("fire_rare", "fire_privilege", "fire_angry", "fire_fear")

fire_matrix <- cor(
  sample[fire_vars],
  use = "pairwise.complete.obs"
)

round(fire_matrix, 2)

library(psych)

alpha_result <- psych::alpha(sample[fire_vars])

print(alpha_result)

names(alpha_result)

alpha_result$alpha.drop

alpha_result

```

### E. Race

q51\_ ▢ White (1) ▢ Black or African American (2) ▢ Native American / Alaska Native (3) ▢ Asian / Asian American (4) ▢ Native Hawaiian / Other Pacific Islander (5) ▢ Middle Eastern/ North African (6) ▢ Hispanic/Latino/a/e (7) ▢ My preferred response not listed (8)

-measured as set of binary indicator variables in the survey, rather than as one categorical race variable. - 1 = selected that category - 0 = ? did not select? - NA = c

```{r}

all_race_vars <- c("White", "Black", "Hisp_Latino", "Asian_Am", "Native_Am", "Pac_Island", "MENA", "race_notlisted")

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(all_of(all_race_vars), ~ na_if(as.numeric(.x), -99)))

```

Responses -- noting consistent NAs so NAs are missing

```{r}

raw_data_cleaned %>%
  select(all_of(all_race_vars)) %>%
  pivot_longer(everything(), names_to = "var", values_to = "val") %>%
  count(var, val) %>%
  pivot_wider(names_from = val, values_from = n, values_fill = 0)

```

How many multiracial respondents?

```{r}

#row level counts
race_df <- raw_data_cleaned %>%
  mutate(
    n_races   = rowSums(across(all_of(all_race_vars), ~ .x == 1), na.rm = TRUE),
    n_missing = rowSums(across(all_of(all_race_vars), is.na)),
    race_item_answered = n_races > 0,
    multiracial = case_when(
      !race_item_answered ~ NA,
      n_races >= 2        ~ TRUE,
      n_races == 1        ~ FALSE
    )
  )

count(race_df, n_races)
count(race_df, multiracial)

summary(race_df$n_races)

```

I have 218 multiracial respondents.

#### multiracial var

```{r}

raw_data_cleaned$multiracial <- race_df$multiracial

all_race_vars <- c(
  "White",
  "Black",
  "Hisp_Latino",
  "Asian_Am",
  "Native_Am",
  "Pac_Island",
  "MENA",
  "race_notlisted",
  "multiracial"
)

```

```{r}
# combinations
combo_counts <- raw_data_cleaned %>%
  mutate(.id = row_number()) %>%
  select(.id, all_of(all_race_vars)) %>%
  pivot_longer(-.id, names_to = "race", values_to = "selected") %>%
  filter(selected == 1) %>%                      # drops both 0s and NAs
  mutate(race = factor(race, levels = all_race_vars)) %>%
  arrange(.id, race) %>%
  group_by(.id) %>%
  summarise(
    n_races    = n(),
    race_combo = paste(race, collapse = " + "),
    .groups = "drop"
  ) %>%
  count(n_races, race_combo, name = "n") %>%
  arrange(desc(n))

View(combo_counts)                                # everything

filter(combo_counts, n_races >= 2)          # multiracial combos only
```

```{r}
# pairs
race_mat <- raw_data_cleaned %>%
  select(all_of(all_race_vars)) %>%
  mutate(across(everything(), ~ replace_na(as.integer(.x == 1), 0L))) %>%
  as.matrix()

co_occur <- crossprod(race_mat)   # diagonal = marginal totals
co_occur
```

#### condensed race var

```{r}
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    OtherRace = case_when(
      MENA == 1 | Pac_Island == 1 | race_notlisted == 1 ~ 1,
      MENA == 0 & Pac_Island == 0 & race_notlisted == 0 ~ 0,
      TRUE ~ NA_real_
    )
  )

#table(raw_data_cleaned$OtherRace, useNA = "ifany")

race_vars <- c(
  "White",
  "Black",
  "Hisp_Latino",
  "Asian_Am",
  "Native_Am",
  "OtherRace"
)

```

#### more condensed race var

```{r}

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    AllOthers = case_when(
      OtherRace == 1 | Asian_Am == 1 | Native_Am == 1 ~ 1,
      OtherRace == 0 & Asian_Am == 0 & Native_Am == 0 ~ 0,
      TRUE ~ NA_real_
    )
  )


problematic_race_vars <- c(
  "White",
  "Black",
  "Hisp_Latino",
  "AllOthers"
)
```

### F. Party + Ideology controls

```{r}
#raw_data_cleaned %>% count(q47)

#clean dem 101
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(dem101 = na_if(as.numeric(dem101), -99))

#single PID variable
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    pid = case_when(
      dem101 == 1 ~ "Republican",
      dem101 == 2 ~ "Democrat",
      dem101 == 3 ~ "Independent",
      dem101 == 4 ~ "Other",
      TRUE ~ NA_character_
    ),
    
    pid = factor(
      pid,
      levels = c(
        "Independent",
        "Democrat",
        "Republican",
        "Other"
      )
    )
  )

#other? 76
# raw_data_cleaned %>%
#   count(pid, .drop = FALSE)
# 
# raw_data_cleaned %>%
#   count(pid, .drop = FALSE) %>%
#   mutate(
#     percent = round(100 * n / sum(n), 2)
#   )

# drop other when doing analyses? 
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    
    # Numeric/ordinal version
    pid_3 = case_when(
      dem101 == 1 ~ 1,   # Republican
      dem101 == 3 ~ 2,   # Independent
      dem101 == 2 ~ 3,   # Democrat
      dem101 == 4 ~ NA_real_,
      TRUE ~ NA_real_
    ),
    
    # Categorical version
    pid_3_factor = factor(
      pid_3,
      levels = c(1, 2, 3),
      labels = c("Republican", "Independent", "Democrat")
    ),
    
    # Preserve reason for exclusion
    pid_other = case_when(
      dem101 == 4 ~ 1,
      dem101 %in% c(1, 2, 3) ~ 0,
      TRUE ~ NA_real_
    )
  )
```

```{r}
#clean ideology
#summary(raw_data_cleaned$q46)
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    fiscal_ideology = na_if(as.numeric(fiscal_ideology), -99),
    social_ideology = na_if(as.numeric(social_ideology), -99)
  )

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(
    fiscal_ideology = ifelse(!is.na(fiscal_ideology), 6 - fiscal_ideology, NA),#now 0 = very lib, 5 = very conservative (left to right)
    social_ideology = ifelse(!is.na(social_ideology), 6 - social_ideology, NA)
  )

#summary(raw_data_cleaned$fiscal_ideology)

#self-reported ideo ident
table(raw_data_cleaned$politcs) # already coded so that left to right
# 
# raw_data_cleaned <- raw_data_cleaned %>%
#    rename(
#      self_ideo = politcs
#      )
```

```{r}
pol_tude_vars <- c("fiscal_ideology", "social_ideology", "self_ideo")

pol_tude <- raw_data_cleaned %>%
  select(all_of(pol_tude_vars))

cor(pol_tude,
    use = "pairwise.complete.obs")

print(psych::corr.test(pol_tude))
```

### I. Gender

```{r, gender}
#raw_data_cleaned %>% count(q55)

#trans clean

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(trans = na_if(as.numeric(trans), -99))


#trans indicator 
raw_data_cleaned <- raw_data_cleaned %>% mutate(trans_id = case_when(
  trans == 1 ~ 1,
  trans %in% c(2, 3) ~ 0,
  TRUE ~ NA_real_
)) 
#Conservative approach: counts only self-identified trans as 1 TRUE ~ NA_real_ 


#gender clean
gender_vars <- c("woman", "man", "gqueer", "g_other")



raw_data_cleaned <- raw_data_cleaned %>% mutate(across(all_of(gender_vars), ~ na_if(as.numeric(.x), -99)))


#constructing the gender binary
raw_data_cleaned <- raw_data_cleaned %>%
  mutate(gender = case_when(
    man == 1 ~ 1,  # Selected man
    woman == 1 & (man != 1 | is.na(man)) ~ 0,  # Woman only or with genderqueer/other
    gqueer == 1 | g_other == 1 ~ 0,  # Genderqueer or other, no man selected
    TRUE ~ NA_real_
  ))


```

### J. Online?

```{r}

raw_data_cleaned <- raw_data_cleaned %>% 
  mutate(
    facebook = na_if(as.numeric(facebook), -99),
    twitter = na_if(as.numeric(twitter), -99),
    truthsocial = na_if(as.numeric(truthsocial), -99),
    reddit = na_if(as.numeric(reddit), -99),
    parlor1 = na_if(as.numeric(parlor1), -99),
    parlor2 = na_if(as.numeric(parlor2), -99),
    youtube = na_if(as.numeric(youtube), -99),
    tiktok = na_if(as.numeric(tiktok), -99)
  )

sm_use <- c("facebook", "twitter", "youtube", "truthsocial", "reddit", "parlor1", "parlor2", "tiktok")

table(raw_data_cleaned$youtube)

raw_data_cleaned <- raw_data_cleaned %>%
  mutate(across(
    all_of(sm_use),
    ~ case_when(
      . == 4 ~ 3,
      . == 5 ~ 2,
      . == 6 ~ 1,
      TRUE ~ NA_real_
    )
  ))

sm_use <- c("facebook", "twitter", "youtube", "truthsocial", "reddit", "parlor1", "parlor2", "tiktok")

sm_matrix <- cor(
  sample[sm_use],
  use = "pairwise.complete.obs"
)

round(sm_matrix, 2)

library(psych)

alpha_result <- psych::alpha(sample[sm_use])

print(alpha_result)

names(alpha_result)

alpha_result$alpha.drop

alpha_result


```

Social media average variable – come back to this later if useful for a control.

### K. LAST OF CLEANING - WRITE CSV

```{r}

#nrow(raw_data_cleaned)
write_csv(raw_data_cleaned, "data/cleaned/coded_data__09_18_26.csv")
#write_csv(raw_data_cleaned, "data/cleaned/coded_data_may9.csv")

```

# ARCHIVED

```{r}
#| eval: false
#| include: false

getwd()
impacted <- read.csv("/Users/kristinamensik/Library/Mobile Documents/com~apple~CloudDocs/Documents/Duke/2024-2025/902 Legislators Pre-Ananlysis Plan/Prelim Folder/Prelim 3-23/data/impacted.csv")

view(impacted)

# Count unique combinations
impacted_summary <- impacted %>%
  select(been_inc, know_inc, none_inc, been_vict, know_vict, no_vict) %>%
  group_by(across(everything())) %>%
  summarise(count = n(), .groups = "drop") %>%
  arrange(desc(count))

# View the result
view(impacted_summary)

#only incarcerated
sum(
  impacted$been_inc == 1 &
  impacted$no_vict == 1 &
  impacted$know_inc == 0 &
  impacted$none_inc == 0 &
  impacted$been_vict == 0 &
  impacted$know_vict == 0
)

#incarcerated & Victimized
sum(
  impacted$been_inc == 1 &
  impacted$no_vict == 0 &
  impacted$know_inc == 0 &
  impacted$none_inc == 0 &
  impacted$been_vict == 1 &
  impacted$know_vict == 0
)

#incarcerated & Victimized & Know Inc
sum(
  impacted$been_inc == 1 &
  impacted$no_vict == 0 &
  impacted$know_inc == 1 &
  impacted$none_inc == 0 &
  impacted$been_vict == 1 &
  impacted$know_vict == 0
)

#only know incarcerated
sum(
  impacted$been_inc == 0 &
  impacted$no_vict == 1 &
  impacted$know_inc == 1 &
  impacted$none_inc == 0 &
  impacted$been_vict == 0 &
  impacted$know_vict == 0
)

#only vict
sum(
  impacted$been_inc == 0 &
  impacted$no_vict == 0 &
  impacted$know_inc == 0 &
  impacted$none_inc == 1 &
  impacted$been_vict == 1 &
  impacted$know_vict == 0
)

#only know vict
sum(
  impacted$been_inc == 0 &
  impacted$no_vict == 0 &
  impacted$know_inc == 0 &
  impacted$none_inc == 1 &
  impacted$been_vict == 0 &
  impacted$know_vict == 1
)

#know vict, know incarc, been vict
sum(
  impacted$been_inc == 0 &
  impacted$no_vict == 0 &
  impacted$know_inc == 1 &
  impacted$none_inc == 0 &
  impacted$been_vict == 1 &
  impacted$know_vict == 1
)


```
