---
title: "prospectus_april3_26"
format: html
editor: visual
---

## 

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,       # Hide code by default
  warning = FALSE,
  message = FALSE
)

# Load packages
library(dplyr)
library(tidyr)
library(readr)
library(janitor)   # for clean_names()
library(tibble)
library(stringr)
library(haven)
library(tibble)
library(gt)
library(marginaleffects)
library(mediation)
library(naniar)
library(rsq)
library(patchwork)

```

Please help me work through my necessary steps. First question: are those models you showed relevant for goal 1 or goal 2? I have already created a table comparing means. My sense is that I have both of the goals 1 and goal 2. I'd like to show you the code I am using right now for comparing means - it might be different than you last saw (I fixed variables and have been trying to streamline and make sure I understand what is happening in the code). 

Step 1: Run Bivariate Models. I want to know how 
Dependent Variables:
-help policy index: prisonhelp
-punitive policy index: prisonpen
-help policy items (see below)
-punitive policy items (see below)

These dependent variables are all defined in this: 
policy_vars <- c(
  "prisonhelp", "prisonpen",
  "q23_2_rev", "q23_7_rev", "q25_2_rev", "q25_3_rev", "q25_4_rev", "q25_6_rev",
  "prisonpen_1", "prisonpen_5"

and this: 

policy_map <- tibble::tribble(
  ~var,          ~label,                                              ~section,
  "prisonhelp",  "Help-oriented policy index",                        "Help",
  "q23_2_rev",   "Voting rights after serving felony sentences",      "Help",
  "q23_7_rev",   "Ensuring eligible incarcerated people can vote",    "Help",
  "q25_2_rev",   "Minimum wage for prison labor",                     "Help",
  "q25_3_rev",   "Free calls with family members",                    "Help",
  "q25_4_rev",   "Funding GED and college courses in prisons",        "Help",
  "q25_6_rev",   "Sentencing alternatives for parents of young children", "Help",
  "prisonpen",   "Punitive policy index",                             "Punish",
  "prisonpen_1", "Death penalty for people convicted of murder",      "Punish",
  "prisonpen_5", "Life without parole sentences",                     "Punish"
)

Independent Variables
-impact (0,1)
-incarc (0,1)
-knowincarc (0,1)
-vict (0,1)

comparison_labels <- c(
  impact = "Impacted",
  incarc = "Ever incarcerated",
  knowincarc = "Knows incarcerated",
  vict = "Victim"
)

all people who are 1 in incarc, knowincarc, and vict should be 1 in impact; many people who are 1 in 1 of incarc, knowincarc, vict, may also be 1 in the other 2

These independent variables are all defined in this: 

impact_vars <- c("impact", "incarc", "knowincarc", "vict")

The first thing that I do is I run simple dummy regressions for each IV, on each DV.


```{r}

#Approach A is doing . . .

# function to add significance star
star_if_sig <- function(p) {
  ifelse(!is.na(p) & p < 0.05, "*", "")
}

# function to build one row of the table
make_pctdiff_row <- function(df, var) {
  ni     <- df[[var]][df$impact == 0]
  direct <- df[[var]][df$incarc == 1]
  prox   <- df[[var]][df$knowincarc == 1]
  vict   <- df[[var]][df$vict == 1]

  ni_mean     <- mean(ni, na.rm = TRUE)
  direct_mean <- mean(direct, na.rm = TRUE)
  prox_mean   <- mean(prox, na.rm = TRUE)
  vict_mean   <- mean(vict, na.rm = TRUE)

  direct_pct <- ((direct_mean - ni_mean) / ni_mean) * 100
  prox_pct   <- ((prox_mean   - ni_mean) / ni_mean) * 100
  vict_pct   <- ((vict_mean   - ni_mean) / ni_mean) * 100

  direct_p <- tryCatch(t.test(ni, direct)$p.value, error = function(e) NA)
  prox_p   <- tryCatch(t.test(ni, prox)$p.value,   error = function(e) NA)
  vict_p   <- tryCatch(t.test(ni, vict)$p.value,   error = function(e) NA)

  tibble(
    policy = policy_labels[[var]],
    not_impacted_mean = round(ni_mean, 2),
    directly_impacted = paste0(round(direct_pct, 1), star_if_sig(direct_p)),
    proximally_impacted = paste0(round(prox_pct, 1), star_if_sig(prox_p)),
    victimized = paste0(round(vict_pct, 1), star_if_sig(vict_p))
  )
}

# build final table
means_table <- bind_rows(
  lapply(policy_items, function(v) make_pctdiff_row(raw_data_cleaned, v))
)

means_table

#output LaTeX rows
latex_rows <- means_table |>
  mutate(
    row = paste0(
      policy, " & ",
      not_impacted_mean, " & ",
      directly_impacted, " & ",
      proximally_impacted, " & ",
      victimized, " \\\\"
    )
  ) |>
  pull(row)

cat(paste(latex_rows, collapse = "\n"))


#Later, will do regressions that may look like

m1_help_index <- . . .

lm(prisonhelp ~ impact, data = raw_data_cleaned)
lm(prisonhelp ~ incarc, data = raw_data_cleaned)
lm(prisonhelp ~ knowincarc, data = raw_data_cleaned)
lm(prisonhelp ~ vict, data = raw_data_cleaned)

m1_pen_index <- . . .

lm(prisonhelp ~ impact, data = raw_data_cleaned)
lm(prisonhelp ~ incarc, data = raw_data_cleaned)
lm(prisonhelp ~ knowincarc, data = raw_data_cleaned)
lm(prisonhelp ~ vict, data = raw_data_cleaned)



m3_help <- lm(prisonhelp ~ incarc + knowincarc + vict +
                ses + race_black + race_hispanic + race_other +
                male + rep + dem + ideology +
                deservingness_core + policy_disposition + racial_resentment,
              data = raw_data_cleaned)

m3_pen <- lm(prisonpen ~ incarc + knowincarc + vict +
               ses + race_black + race_hispanic + race_other +
               male + rep + dem + ideology +
               deservingness_core + policy_disposition + racial_resentment,
             data = raw_data_cleaned)
```
